### 2018-01-25 README.Rmd
Initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# caa2018

The goal of caa2018 is to ...

### Citation

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Title of compendium_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2018-01-25 analysis/paper/paper.Rmd
initial

---
title: "Reproducible research in archaeology using R & rrtools"
author:
  - Sophie C. Schmidt
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      number_sections: false
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  fig.path = "../figures/"
)

library(caa2018) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

## Introduction

These are our notes for preparing our slides for our CAA2018 presentation in the session "R as an archaeological tool: current state and directions" (https://github.com/MartinHinz/isaak_conf_sessions/tree/master/caa_2018_tuebingen)

## What are the top archaeology journals by number of papers published?

We need to know this to plan for the data collection for our JCCA paper, it's not central to our CAA presentation, but I've put it here so I know where to find it later (I hope).

Data are from an 'advanced search' of apps.webofknowledge.com using WC=Archaeology on apps.webofknowledge.com.

For 2017 only:

```{r}
library(tidyverse)

wc_archaeology_2017 <- 
  readr::read_tsv(paste0(here::here(), "/analysis/data/raw_data/savedrecs.txt"), 
            quote = "")
```

```{r top-archaeology-journals-by-articles}
# The 10 journals with the most articles are:

top_journals_by_articles <- 
wc_archaeology_2017 %>% 
  group_by(SO) %>% 
  tally(sort = TRUE)

knitr::kable(head(top_journals_by_articles, 10), 
             caption = "The top 10 archaeology journals, by number of papers published in 2017")
```

For 2007-2017:

```{r}
library(glue)

data_files_from_2007_to_2017 <- 
  paste0(here::here(), glue("/analysis/data/raw_data/savedrecs ({1:2}).txt"))
```

```{r fig.height=10}

wc_archaeology_2007_to_2017 <- 
  map(data_files_from_2007_to_2017,
      ~readr::read_tsv(.x, quote = "")) %>% 
  bind_rows()

# most popular journals over time
most_popular_journals_over_time <- 
wc_archaeology_2007_to_2017 %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE) %>% 
  filter(n >= 50) %>% 
  ungroup()

library(ggrepel)
ggplot(most_popular_journals_over_time,
       aes(PY, 
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(most_popular_journals_over_time, 
                  PY == max(PY)),
    aes(label = SO),
    size = 3,
    nudge_x = 15,
    segment.color = NA
  ) +
  xlim(2007, 2019) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of articles per year in Archaeology journals",
          subtitle = "Data from apps.webofknowledge.com,\nusing WC=Archaeology in SCI-EXPANDED & SSCI for 2007-2017")
```

## Citations of R Core in articles on Web of Science

We wanted to to see how R Core is cited in scientific literature in general, and in archaeology in particular.

We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`

For all scientific journals tracked by the WoS:

```{r}

cran_cites_files <- 
    paste0(here::here(), glue("/analysis/data/raw_data/savedrecs ({3:11}).txt"))
```

```{r fig.height=10}
cran_cites_all_areas_all_years <- 
  map(cran_cites_files,
      ~readr::read_tsv(.x, 
                       quote = "",
                       col_types = cols(.default = col_character()))) %>% 
  bind_rows()

# limit to top journals
top_journals_for_cran_cites <- 
  cran_cites_all_areas_all_years %>% 
  group_by(SO) %>% 
  tally(sort = TRUE)

# tally by year
cran_cites_all_areas_top_journals_by_year <- 
  cran_cites_all_areas_all_years %>% 
  filter(SO %in% top_journals_for_cran_cites$SO[1:10]) %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE) %>% 
  ungroup() %>% 
  mutate(PY = as.numeric(PY))

ggplot(cran_cites_all_areas_top_journals_by_year,
       aes(PY,
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(cran_cites_all_areas_top_journals_by_year, 
                  PY == max(PY)),
    aes(label = SO),
    size = 3,
    nudge_x = 1,
    segment.color = NA
  ) +
  xlim(2007, 2018) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of articles per year citing R Core in top 10 WoS journals",
          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
```

For archaeology journals only:

```{r}

cran_cites_in_archy_journals <- 
  cran_cites_all_areas_all_years %>% 
  filter(SO %in% top_journals_by_articles$SO)

cran_cites_in_archy_journals_per_year <- 
  cran_cites_in_archy_journals %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE) %>% 
  ungroup() %>% 
  mutate(PY = as.numeric(PY))

ggplot(cran_cites_in_archy_journals_per_year,
       aes(PY,
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(cran_cites_in_archy_journals_per_year, 
                  PY == max(PY)),
    aes(label = SO),
    size = 3,
    nudge_x = 1,
    segment.color = NA
  ) +
  xlim(2007, 2018) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of Archaeology articles per year citing R Core",
          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
```

## Text mining titles in archaeology journal articles that cite R Core

What are the key words in the titles of these archaeology articles that cite R Core, and how do they differ from other archaeology articles?

```{r eval = FALSE}
library(tidytext)

cran_cites_in_archy_journals %>% 
  select(TI) %>% 
  unnest_tokens(word, TI) %>% 
  anti_join(stop_words) %>%
  count(word, sort = TRUE) %>% 
  head(20)

library(textrank)

# keywords for each title
keyw <- map(cran_cites_in_archy_journals$TI, 
            ~textrank_keywords(unlist(str_split(paste0(.x, collapse = " "), " "))[!unlist(str_split(paste0(.x, collapse = " "), " ")) %in% stop_words$word]))

map(keyw, ~enframe(.x$pagerank$vector) %>% 
  arrange(desc(value)))

# tf-idf to show words most important to each journal
# not very interesting
cran_cites_in_archy_journals %>% 
  filter(SO %in% top_journals_by_articles$SO) %>% 
  select(SO, TI) %>% 
  unnest_tokens(word, TI) %>% 
  anti_join(stop_words) %>%   
  count(SO, word, sort = TRUE) %>% 
  bind_tf_idf(word, SO, n) %>% 
  arrange(desc(tf_idf)) %>% 
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(SO) %>% 
  top_n(5) %>% 
  ungroup %>%
  ggplot(aes(word, tf_idf, fill = SO)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~SO, ncol = 2, scales = "free") +
  coord_flip()
```

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

### 2018-01-26 analysis/paper/paper.Rmd
complete text mining of titles

@@
 library(caa2018) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
+
+library(ggrepel)
+library(tidyverse)
 ```
@@
 ```{r}
-library(tidyverse)
-
 wc_archaeology_2017 <- 
@@
 
-library(ggrepel)
 ggplot(most_popular_journals_over_time,
@@
 
-We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`
+We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`.
+This function first appeared in 2004 https://github.com/wch/r-source/commits/9a5c086855d2e8fb07f670e3315871b2de31a4fd/src/library/base/inst/CITATION
+
+Note that before 2012 the authors were "R Development Core Team", and then it was simplified to "R Core Team" https://github.com/wch/r-source/commit/9a5c086855d2e8fb07f670e3315871b2de31a4fd#diff-80bb8a8b133a6204e9642d76791927e2
+
+This is our exact search string for WoS:
+
+You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM") AND CITED WORK: ("R: A Language and Environment for Statistical Computing") AND LANGUAGE: (English) AND DOCUMENT TYPES: (Article)
+Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.
 
@@
 cran_cites_files <- 
-    paste0(here::here(), glue("/analysis/data/raw_data/savedrecs ({3:11}).txt"))
+    paste0(here::here(), glue("/analysis/data/raw_data/cited_cran/savedrecs ({1:11}).txt"))
 ```
@@
                        col_types = cols(.default = col_character()))) %>% 
-  bind_rows()
+  bind_rows() %>% 
+  mutate(PY = as.numeric(PY)) %>% 
+  filter(PY <= 2017)
 
@@
   tally(sort = TRUE) %>% 
-  ungroup() %>% 
-  mutate(PY = as.numeric(PY))
+  ungroup()
 
@@
                   PY == max(PY)),
-    aes(label = SO),
+    aes(label = str_wrap(SO, 15)),
     size = 3,
+    direction = "both",
     nudge_x = 1,
@@
                   PY == max(PY)),
-    aes(label = SO),
+    aes(label = str_wrap(SO, 10)),
     size = 3,
@@
   ) +
-  xlim(2007, 2018) +
+  xlim(2007, 2019) +
   theme_minimal() +
@@
 
-```{r eval = FALSE}
+```{r}
 library(tidytext)
+# https://www.tidytextmining.com/twitter.html
 
+# These are all archy articles
+# wc_archaeology_2007_to_2017
+
+# These are ones citing R
+# cran_cites_in_archy_journals
+
+# subtract ones citing R from ones that do not
+library(Hmisc)
+wc_archaeology_2007_to_2017_no_R_cites <- 
+  wc_archaeology_2007_to_2017 %>% 
+  filter(TI %nin% cran_cites_in_archy_journals$TI)
+
+# combine
+archy_journal_articles <- 
+  bind_rows(wc_archaeology_2007_to_2017_no_R_cites %>% 
+            mutate(ID = "no_R_cites",
+                   VL = as.character(VL)),
            cran_cites_in_archy_journals %>% 
-  select(TI) %>% 
+            mutate(ID = "has_R_cites"))
+
+library(tidytext)
+library(stringr)
+
+tidy_cites <- 
+archy_journal_articles %>% 
   unnest_tokens(word, TI) %>%
-  anti_join(stop_words) %>%
+  filter(!word %in% stop_words$word,
+         str_detect(word, "[a-z]"))
+
+frequency <- 
+  tidy_cites %>% 
+  group_by(ID) %>% 
   count(word, sort = TRUE) %>% 
-  head(20)
+  left_join(tidy_cites %>% 
+              group_by(ID) %>% 
+              summarise(total = n())) %>%
+  mutate(freq = n/total)
 
-library(textrank)
+frequency <- frequency %>% 
+  select(ID, word, freq) %>% 
+  spread(ID, freq) %>%
+  arrange(no_R_cites, has_R_cites)
 
-# keywords for each title
-keyw <- map(cran_cites_in_archy_journals$TI, 
-            ~textrank_keywords(unlist(str_split(paste0(.x, collapse = " "), " "))[!unlist(str_split(paste0(.x, collapse = " "), " ")) %in% stop_words$word]))
+library(scales)
 
-map(keyw, ~enframe(.x$pagerank$vector) %>% 
-  arrange(desc(value)))
+ggplot(frequency, 
+       aes(no_R_cites, 
+           has_R_cites)) +
+  geom_jitter(alpha = 0.1, 
+              size = 2.5, 
+              width = 0.25, 
+              height = 0.25) +
+  geom_text(aes(label = word), 
+            check_overlap = TRUE, 
+            vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
 
-# tf-idf to show words most important to each journal
-# not very interesting
-cran_cites_in_archy_journals %>% 
-  filter(SO %in% top_journals_by_articles$SO) %>% 
-  select(SO, TI) %>% 
-  unnest_tokens(word, TI) %>% 
-  anti_join(stop_words) %>%   
-  count(SO, word, sort = TRUE) %>% 
-  bind_tf_idf(word, SO, n) %>% 
-  arrange(desc(tf_idf)) %>% 
-  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
-  group_by(SO) %>% 
-  top_n(5) %>% 
-  ungroup %>%
-  ggplot(aes(word, tf_idf, fill = SO)) +
+word_ratios <- tidy_cites %>%
+  count(word, ID) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(ID, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / sum(. + 1))) %>%
+  mutate(logratio = log(has_R_cites / no_R_cites)) %>%
+  arrange(desc(logratio))
+
+word_ratios %>%
+  group_by(logratio < 0) %>%
+  top_n(10, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
   geom_col(show.legend = FALSE) +
-  labs(x = NULL, y = "tf-idf") +
-  facet_wrap(~SO, ncol = 2, scales = "free") +
-  coord_flip()
+  coord_flip() +
+  ylab("log odds ratio (has_R_cites/no_R_cites)") +
+  scale_fill_discrete(name = "", labels = c("has_R_cites", "no_R_cites"))
+
 ```

### 2018-03-06 analysis/paper/paper.Rmd
text and figs for slides 1-7

@@
 library(tidyverse)
+library(glue)
+library(tidytext)
+library(scales)
 ```
 
-## Introduction
+## Narrative to accompany the slides
 
-These are our notes for preparing our slides for our CAA2018 presentation in the session "R as an archaeological tool: current state and directions" (https://github.com/MartinHinz/isaak_conf_sessions/tree/master/caa_2018_tuebingen)
+- Slide 1: Title slide
 
-## What are the top archaeology journals by number of papers published?
+In this presentation we will talk about how archaeologists are changing their data analysis practices to make their research more reproducible and open.
+We will the rrtools package, which is our specific contribution to help with the challenge of making reproducible research in archaeology easier.
 
-We need to know this to plan for the data collection for our JCCA paper, it's not central to our CAA presentation, but I've put it here so I know where to find it later (I hope).
+- Slide 2: Motivation
 
-Data are from an 'advanced search' of apps.webofknowledge.com using WC=Archaeology on apps.webofknowledge.com.
+Our motivation for this topic comes from a profound change that we have been observing in how researchers analyse their data.
 
-For 2017 only:
+Increasingly, we are seeing researchers publishing research that uses a programming language to analyse their data.
+Similarly, we see journal articles accompanied by scripts of programming language so that readers can study the details of the data analysis methods.
 
-```{r}
-wc_archaeology_2017 <- 
-  readr::read_tsv(paste0(here::here(), "/analysis/data/raw_data/savedrecs.txt"), 
-            quote = "")
-```
+This is an important change away from using mouse-driven, point-and-click software for data analysis.
+When archaeologists use software such as SPSS and Past (as you see in our screenshots here), it is very difficult for them to share the details of their analysis.
+When readers cannot see the details of an analysis, it is difficult for them to be confident in the results described in an article.
 
-```{r top-archaeology-journals-by-articles}
-# The 10 journals with the most articles are:
+- Slide 3: Outline
 
-top_journals_by_articles <- 
-wc_archaeology_2017 %>% 
-  group_by(SO) %>% 
-  tally(sort = TRUE)
+In this presentation we will briefly review some data showing change in the use of programming languages in the sciences generally, and in archaeology in particular.
+We will look at citation data from thousands of journal articles that are available in the Web of Science database produced by Clarivate Analytics.
+We focus on the R programming language, because this is by far the most widely used langauge in archaeology.
 
-knitr::kable(head(top_journals_by_articles, 10), 
-             caption = "The top 10 archaeology journals, by number of papers published in 2017")
-```
+Secondly, we will review a few case studies of how R is used by archaeologists, to show the potential that using a programming language has for opening and accelerating archaeological research.
 
-For 2007-2017:
+Finally, we will briefly introduce our R package, rrtools, which is designed to simplify and standardise the process of using R for archaeology, and making research more open and reproducible.
 
-```{r}
-library(glue)
+- Slide 4: Current uses of R [signpost slide only]
 
-data_files_from_2007_to_2017 <- 
-  paste0(here::here(), glue("/analysis/data/raw_data/savedrecs ({1:2}).txt"))
-```
+- Slide 5: Articles in all sciences citing R
 
-```{r fig.height=10}
+```{r all-sciences-citing-r, fig.height=10}
 
-wc_archaeology_2007_to_2017 <- 
-  map(data_files_from_2007_to_2017,
-      ~readr::read_tsv(.x, quote = "")) %>% 
-  bind_rows()
-
-# most popular journals over time
-most_popular_journals_over_time <- 
-wc_archaeology_2007_to_2017 %>% 
-  group_by(SO, PY) %>% 
-  tally(sort = TRUE) %>% 
-  filter(n >= 50) %>% 
-  ungroup()
-
-ggplot(most_popular_journals_over_time,
-       aes(PY, 
-           n,
-           colour = SO)) +
-  geom_line() +
-  geom_text_repel(
-    data = subset(most_popular_journals_over_time, 
-                  PY == max(PY)),
-    aes(label = SO),
-    size = 3,
-    nudge_x = 15,
-    segment.color = NA
-  ) +
-  xlim(2007, 2019) +
-  theme_minimal() +
-  theme(legend.position="none") +
-  ggtitle("Number of articles per year in Archaeology journals",
-          subtitle = "Data from apps.webofknowledge.com,\nusing WC=Archaeology in SCI-EXPANDED & SSCI for 2007-2017")
-```
-
-## Citations of R Core in articles on Web of Science
-
-We wanted to to see how R Core is cited in scientific literature in general, and in archaeology in particular.
-
-We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`.
+# We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`.
 This function first appeared in 2004 https://github.com/wch/r-source/commits/9a5c086855d2e8fb07f670e3315871b2de31a4fd/src/library/base/inst/CITATION
-
-Note that before 2012 the authors were "R Development Core Team", and then it was simplified to "R Core Team" https://github.com/wch/r-source/commit/9a5c086855d2e8fb07f670e3315871b2de31a4fd#diff-80bb8a8b133a6204e9642d76791927e2
-
-This is our exact search string for WoS:
-
-You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM") AND CITED WORK: ("R: A Language and Environment for Statistical Computing") AND LANGUAGE: (English) AND DOCUMENT TYPES: (Article)
-Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.
-
-For all scientific journals tracked by the WoS:
-
-```{r}
+# 
+# Note that before 2012 the authors were "R Development Core Team", and then it was simplified to "R Core Team" https://github.com/wch/r-source/commit/9a5c086855d2e8fb07f670e3315871b2de31a4fd#diff-80bb8a8b133a6204e9642d76791927e2
+# 
+# This is our exact search string for WoS:
+# 
+# You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM") AND CITED WORK: ("R: A Language and Environment for Statistical Computing") AND LANGUAGE: (English) AND DOCUMENT TYPES: (Article)
+# Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.
 
@@
     paste0(here::here(), glue("/analysis/data/raw_data/cited_cran/savedrecs ({1:11}).txt"))
-```
 
-```{r fig.height=10}
 cran_cites_all_areas_all_years <- 
@@
 
+how_many_articles_in_top_journals_for_cran_cites <- 
+  sum(top_journals_for_cran_cites$n)
+
+top_journals_for_cran_cites_articles_per_year <- 
+  read_csv(str_glue('{here::here()}/analysis/data/raw_data/top_journals_for_cran_cites_articles_per_year.csv')) %>% 
+  select(-X1) %>% 
+  gather(variable, value, -year) %>% 
+  separate(variable, 
+           into = str_glue('X{1:4}'), 
+           by = "_") %>% 
+  select(year, X4, value)
+
 # tally by year
@@
   tally(sort = TRUE) %>% 
+  mutate(journalname = str_remove_all(SO, " |-"),
+         PY = as.integer(PY)) %>% 
+  left_join(top_journals_for_cran_cites_articles_per_year,
+            by = c('journalname' = 'X4', 
+                   'PY' = "year")) %>% 
+  mutate(prop = n / value ) %>% 
+  filter(SO != 'JOURNAL OF STATISTICAL SOFTWARE') %>% 
   ungroup()
 
+min_y <- 2007
+max_y <- 2025
 ggplot(cran_cites_all_areas_top_journals_by_year,
        aes(PY,
-           n,
+           prop,
            colour = SO)) +
-  geom_line() +
+  geom_line(size = 2) +
   geom_text_repel(
@@
                   PY == max(PY)),
-    aes(label = str_wrap(SO, 15)),
-    size = 3,
-    direction = "both",
-    nudge_x = 1,
-    segment.color = NA
+    aes(label = str_wrap(SO, 40)),
+    size = 4,
+    nudge_x = 0,
+    hjust = 0,
+    segment.color = NA,
+    direction = "y"
   ) +
-  xlim(2007, 2018) +
+  scale_x_continuous(breaks = min_y:max_y,
+                     labels = c(min_y:2017, rep("", length(2018:max_y))),
+                     limits = c(min_y, max_y)) +
+  scale_y_continuous(labels = scales::percent) +
+  xlab("Publication year") +
+  ylab("Percentage of articles in that journal") +
   theme_minimal() +
   theme(legend.position="none") +
-  ggtitle("Number of articles per year citing R Core in top 10 WoS journals",
+  ggtitle(str_glue('Percentage of articles per year citing R in top 10 WoS journals ({prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")} articles)'),
           subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
+
+ggsave(filename = "all_sciences_citing_R_over_time.png",
+       path = str_glue('{here::here()}/analysis/figures'),
+       height = 5,
+       width = 9)
 ```
 
-For archaeology journals only:
+This plot shows how data analysis is changing in the sciences more broadly towards the use of the R programming language.
+We searched the Web of Science for citations to the R program over the last ten years, and filtered the results to the top ten journals with the highest number of citations to this software.
+The purpose of this filtering step was to find areas of research were R is generally known, so we can see how it's use is changing over time.
+Note that many articles that use R do not cite the R program, so these numbers are underestimates of the true use of R in science.
 
-```{r}
+We got `r how_many_articles_in_top_journals_for_cran_cites` articles in our sample, and we can make two major observations.
+First, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
+We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.
+The scale of the increases is quite impressive, we see many journals in our sample starting from less than 1% of articles citing R and now have around 20% of articles citing R.
+
+So, we can see this is a substantial change in the way researchers are working in many fields. What about archaeology? Do we also see a change like this?
+
+- Slide 6: Archaeology articles citing R
+
+```{r top-archaeology-journals-by-citations-of-r}
+
+# Data are from an 'advanced search' of apps.webofknowledge.com using WC=Archaeology on apps.webofknowledge.com.
+
+# get the top archaeology journals by numbers of articles.
+wc_archaeology_2017 <- 
+  readr::read_tsv(paste0(here::here(), "/analysis/data/raw_data/savedrecs.txt"), 
+            quote = "")
+
+# The 10 journals with the most articles are:
+top_journals_by_articles <- 
+wc_archaeology_2017 %>% 
+  group_by(SO) %>% 
+  tally(sort = TRUE)
+
+# For 2007-2017:
+
+data_files_from_2007_to_2017 <- 
+  paste0(here::here(), glue("/analysis/data/raw_data/savedrecs ({1:2}).txt"))
+
+wc_archaeology_2007_to_2017 <- 
+  map(data_files_from_2007_to_2017,
+      ~readr::read_tsv(.x, quote = "")) %>% 
+  bind_rows()
+
+total_number_of_articles_in_sample <- 
+  nrow(wc_archaeology_2007_to_2017)
+
+# most populararchaeology journals over time, that have some mention of R
+most_popular_journals_over_time <- 
+wc_archaeology_2007_to_2017 %>% 
+  group_by(SO, PY) %>% 
+  tally(sort = TRUE) %>% 
+  filter(!SO %in% c("JOURNAL OF CULTURAL HERITAGE",
+                    "STUDIES IN CONSERVATION",
+                    "INTERSECCIONES EN ANTROPOLOGIA",
+                    "ISRAEL EXPLORATION JOURNAL",
+                    "JOURNAL OF MATERIAL CULTURE",
+                    "TRABAJOS DE PREHISTORIA",
+                    "ARCHEOSCIENCES-REVUE D ARCHEOMETRIE")) %>% 
+  filter(n >= 2) %>% 
+  ungroup()
+
+most_popular_journals_over_time_names <- 
+  unique(most_popular_journals_over_time$SO)
+
+archy_journals_total_number_of_articles <- 
+  read_csv(str_glue('{here::here()}/analysis/data/raw_data/top_archaeology_journals_articles_per_year.csv'))
+
+top_archaeology_journals_articles_per_year <- 
+  archy_journals_total_number_of_articles %>% 
+  gather(variable, value, -journal_name) %>% 
+  mutate(year = as.numeric(variable)) %>% 
+  select(-variable)
 
@@
   cran_cites_all_areas_all_years %>% 
-  filter(SO %in% top_journals_by_articles$SO)
+  filter(SO %in% most_popular_journals_over_time_names)
+
+# what archaeology journals cite R?
+# unique(cran_cites_in_archy_journals$SO)
+
+archy_journals_total_number_of_citing_articles <- 
+  nrow(cran_cites_in_archy_journals)
 
@@
   ungroup() %>% 
-  mutate(PY = as.numeric(PY))
+  mutate(PY = as.numeric(PY))  %>% 
+  left_join(top_archaeology_journals_articles_per_year,
+            by = c("SO" = "journal_name",
+                   "PY" = "year")) %>% 
+  mutate(prop = n / value) %>% 
+  group_by(SO) %>% 
+  mutate(jn = str_glue('{SO} (n = {sum(n)})')) %>% 
+  ungroup()
+
+min_y <- 2007
+max_y <- 2040
+
+library(ggpmisc)
+jas_cites_r <- 
+cran_cites_in_archy_journals_per_year %>% 
+  filter(SO == "JOURNAL OF ARCHAEOLOGICAL SCIENCE") %>% 
+  ggplot(aes(PY,
+           prop)) +
+  geom_point(size = 4) +
+  geom_smooth(method = "lm",
+              se = FALSE) +
+  stat_poly_eq(aes(label =  paste(..eq.label.., 
+                                  ..adj.rr.label..,
+                                  sep = "~~~~")),
+               formula = y~x, 
+               parse = TRUE,
+               size = 3,
+               label.y.npc = 0.7) +
+  stat_fit_glance(label.y.npc = 0.7,
+                  size = 3,
+                  method = "lm", 
+                  method.args = list(formula = y ~ x),
+                  geom = "text",
+                  aes(label = paste("p-value: ",signif(..p.value.., digits = 4)))) +
+  xlab("year") +
+  ylab("Proportion of articles citing R") +
+  ggtitle(expression(paste("Articles citing R in ", italic("Journal of Archaeological Science")))) +
+  theme_bw(base_size = 6)
 
@@
        aes(PY,
-           n,
+           prop,
            colour = SO)) +
-  geom_line() +
+  geom_line(size = 2) +
   geom_text_repel(
@@
                   PY == max(PY)),
-    aes(label = str_wrap(SO, 10)),
-    size = 3,
+    aes(label = jn),
+    direction = "y",
+    size = 5,
     nudge_x = 1,
-    segment.color = NA
-  ) +
-  xlim(2007, 2019) +
+    hjust = 0) +
   theme_minimal() +
-  theme(legend.position="none") +
-  ggtitle("Number of Archaeology articles per year citing R Core",
-          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
+  theme(legend.position="none",
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
+  scale_x_continuous(breaks = min_y:max_y,
+                     labels = c(min_y:2017, rep("", length(2018:max_y))),
+                     limits = c(min_y, max_y)) +
+  scale_y_continuous(labels = scales::percent) +
+  xlab("Publication year") +
+  ylab("Percentage of articles in that journal") +
+  ggtitle(str_glue('Percentage of Archaeology articles per year citing R ({archy_journals_total_number_of_citing_articles} out of {prettyNum(total_number_of_articles_in_sample, ",")} articles in sample)'),
+          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")  +
+  annotation_custom(grob = ggplotGrob(jas_cites_r), 
+                          xmin = 2030, xmax = 2040, 
+                          ymin = 0.10, ymax = 0.17)
+
+ggsave(filename = "all_archaeology_citing_R_over_time.png",
+       path = str_glue('{here::here()}/analysis/figures'),
+       height = 6,
+       width = 10)
 ```
 
-## Text mining titles in archaeology journal articles that cite R Core
+Here are the results for archaeological articles.
+We found `r archy_journals_total_number_of_citing_articles` articles that cite R over the last ten years.
+Note that this is certinaly not complete, since these are just for the journals tracked by the Web of Science database, and we know that archaeologists publish is a great variaty of journals.
 
-What are the key words in the titles of these archaeology articles that cite R Core, and how do they differ from other archaeology articles?
+There are four important details in these plots:
 
-```{r}
-library(tidytext)
+The first detail we see here is that overall, the proportions of articles citing R in archaeology are much smaller than what we see in the ecology journals on the previous slide.
+The absolute numbers of archaeology articles here are are also very low.
+Only the _Journal of Archaeological Science_ has more than 10 articles in our sample.
+We have yet to adopt programming for data analysis in the same way that ecologists have.
+
+The second detail is that we do not see a significant increase over time, unlike the ecology journals that steeply increase over time.
+Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, we do not yet see much of an upward trend in the archaeology journals.
+Indeed, the first citations of R in several archaeology journals only occured in the last two years.
+The plot on the upper right of this slide shows a statistical test of change over time for articles in JAS.
+With a low r-squared value and high p-value for the linear model, we conclude that there is no evidence for an increase in citations over time in this journal.
+
+The fourth detail we see is the distinctive types of journals that have articles citing R.
+There is a clear focus on scientific archaeology, with _Journal of Archaeological Science_ having the highest total number of articles.
+We also see geoarchaeology, lithics, and environmental archaeology represented in the journal names.
+We can investigate this further with statistical analysis of the words in the titles of all the articles in our sample.
+
+- Slide 7: What are archaeologists doing with R?
+
+```{r text-mining-article-titles-archaeology}
 # https://www.tidytextmining.com/twitter.html
@@
 
+# we have a very skewed sample, so let's match 
+cran_cites_in_archy_journals_lots <- 
+  cran_cites_in_archy_journals %>% 
+  sample_n(nrow(wc_archaeology_2007_to_2017_no_R_cites),
+           replace = TRUE)
+
 # combine
@@
 
-library(tidytext)
-library(stringr)
-
 tidy_cites <- 
@@
   unnest_tokens(word, TI) %>%
-  filter(!word %in% stop_words$word,
+  filter(!word %in% c(stop_words$word, "archaeology", 
+                      "archaeological",
+                      "analysis", "study", "site"),
          str_detect(word, "[a-z]"))
@@
 
-library(scales)
-
 ggplot(frequency, 
@@
            has_R_cites)) +
+    geom_abline(color = "red", 
+              size = 2) +
   geom_jitter(alpha = 0.1, 
@@
               height = 0.25) +
-  geom_text(aes(label = word), 
-            check_overlap = TRUE, 
-            vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
+  geom_text_repel(aes(label = word), 
+            size = 5,
+            segment.colour = NA) +
+  scale_x_log10(labels = percent_format(),
+                limits = c(0.001, 0.02)) +
   scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
+
+  theme_minimal(base_size = 16) +
+  xlab("Articles that do not cite R") +
+  ylab("Articles that cite R") +
+  ggtitle("Comparing the frequency of words used in archaeology articles")
+
+ggsave(filename = "archaeology_articles_word_freqs.png",
+       path = str_glue('{here::here()}/analysis/figures'),
+       height = 9,
+       width = 9)
+```
+
+Our results so far show that only a small number of archaeologists are using R for their research.
+The names of the journals hint that most of these people do scientific archaeology, rather than social or humanistic archaeology.
+
+Here we see a comparison of word frequencies of the titles of archaeology journal articles.
+Words near the red line are used with about equal frequencies by articles that cite R and by articles that do not cite R.
+Words far away from the red line are used much more by one group of articles compared to the other.
+Words that appear in this plot are ones that have been used at least once in article titles of both groups.
+
+...
+
+```{r}
 
@@
 ```
+
+## Notes on data analysis for the presentation
+
+These are our notes for preparing our slides for our CAA2018 presentation in the session "R as an archaeological tool: current state and directions" (https://github.com/MartinHinz/isaak_conf_sessions/tree/master/caa_2018_tuebingen)
+
+## What are the top archaeology journals by number of papers published?
+
+We need to know this to plan for the data collection for our JCCA paper, it's not central to our CAA presentation, but I've put it here so I know where to find it later (I hope).
+
+Data are from an 'advanced search' of apps.webofknowledge.com using WC=Archaeology on apps.webofknowledge.com.
+
+## Citations of R Core in articles on Web of Science
+
+We wanted to to see how R Core is cited in scientific literature in general, and in archaeology in particular.
+
+For all scientific journals tracked by the WoS:
+
+For archaeology journals only:
+
+For 2017 only:
+
+```{r}
+library(glue)
+
+min_y <- 2007
+max_y <- 2030
+library(ggrepel)
+ggplot(most_popular_journals_over_time,
+       aes(PY, 
+           n,
+           colour = SO)) +
+  geom_line(size = 2) +
+  geom_text_repel(
+    data = subset(most_popular_journals_over_time, 
+                  PY == max(PY)),
+    aes(label = SO, 
+        y = (n), 
+        x = PY ),
+    size = 5, 
+    direction = "y",
+    hjust = 0,
+    segment.colour = NA
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = min_y:max_y,
+                     labels = c(min_y:2018, rep("", length(2019:max_y))),
+                     limits = c(min_y, max_y)) +
+  xlab("Publication year") +
+  scale_y_continuous(trans='log2') +
+  ylab("Number of articles (log scale)") +
+  theme(legend.position="none") +
+  ggtitle("Number of articles per year in Archaeology journals",
+          subtitle = "Data from apps.webofknowledge.com,\nusing WC=Archaeology in SCI-EXPANDED & SSCI for 2007-2017")
+
+```
+
+## Text mining titles in archaeology journal articles that cite R Core
+
+What are the key words in the titles of these archaeology articles that cite R Core, and how do they differ from other archaeology articles?
 

### 2018-03-06 analysis/data/raw_data/refs_from_ctv_archaeology.R

@@ -1,0 +1,10 @@
+archy_ctv_readme <- readLines("analysis/data/raw_data/ctv-archaeology-README.md")
+
+archy_ctv_readme <- collapse(archy_ctv_readme)
+archy_ctv_readme <- str_remove_all(archy_ctv_readme, "[[:punct:]]")
+archy_ctv_readme_20XX <- str_extract_all(archy_ctv_readme, " 20[[:digit:]]{2} ")
+archy_ctv_readme_20XX <- str_squish(unlist(archy_ctv_readme_20XX))
+archy_ctv_readme_20XX <- as.numeric(archy_ctv_readme_20XX)
+archy_ctv_readme_20XX <- archy_ctv_readme_20XX[archy_ctv_readme_20XX > 2010]
+hist(archy_ctv_readme_20XX)
+stem(archy_ctv_readme_20XX)

### 2018-03-06 analysis/paper/paper.Rmd
up to slide 9

@@
 
-## Narrative to accompany the slides
+# Narrative to accompany the slides
 
-- Slide 1: Title slide
+## Slide 1: Title slide
 
@@
 
-- Slide 2: Motivation
+## Slide 2: Motivation
 
@@
 
-- Slide 3: Outline
+## Slide 3: Outline
 
@@
 
-- Slide 4: Current uses of R [signpost slide only]
+##- Slide 4: Current uses of R [signpost slide only]
 
-- Slide 5: Articles in all sciences citing R
+## Slide 5: Articles in all sciences citing R
 
@@
 
-- Slide 6: Archaeology articles citing R
+## Slide 6: Archaeology articles citing R
 
@@
 
-- Slide 7: What are archaeologists doing with R?
+## Slide 7: What are archaeologists doing with R?
 
@@
 Words that appear in this plot are ones that have been used at least once in article titles of both groups.
+
+This plot supports our initial observations from the journals titles, which is that archaeologists using R tend to be doing more scientific archaeology.
+We see that many words relating to stone artefacts, zooarchaeology, and geoarchaeology are more frequently used by R-using archaeologists.
+On the other hand, articles that do not cite R more often discuss the Neolithic and metal-age periods, and less often reflect a technical specialisation.
+
+## Slide 8: The unrealised potential of R for reproducible research [signpost slide only]
+
+## Slide 9: The unrealised potential of R for reproducible research [signpost slide only]
+
+We would now like to focus on a small sub-set of these papers using R, because they are doing something very interesting that we think all archaeologists using a programming language should be doing.
+This subset of papers are demonstrating the potential of R for open and reproducible research.
 

### 2018-03-06 analysis/data/raw_data/refs_from_ctv_archaeology.R

@@ -1,10 +1,17 @@
 archy_ctv_readme <- readLines("analysis/data/raw_data/ctv-archaeology-README.md")
 
+
+top_archy_journals <-
+str_subset(archy_ctv_readme, "Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science")
+
+archy_ctv_readme <- top_archy_journals
+
 archy_ctv_readme <- collapse(archy_ctv_readme)
 archy_ctv_readme <- str_remove_all(archy_ctv_readme, "[[:punct:]]")
 archy_ctv_readme_20XX <- str_extract_all(archy_ctv_readme, " 20[[:digit:]]{2} ")
 archy_ctv_readme_20XX <- str_squish(unlist(archy_ctv_readme_20XX))
 archy_ctv_readme_20XX <- as.numeric(archy_ctv_readme_20XX)
 archy_ctv_readme_20XX <- archy_ctv_readme_20XX[archy_ctv_readme_20XX > 2010]
+length(archy_ctv_readme_20XX)
 hist(archy_ctv_readme_20XX)
 stem(archy_ctv_readme_20XX)

### 2018-03-06 analysis/paper/paper.Rmd
up to slide 11

@@
       reference_docx: "../templates/template.docx" # Insert path for the DOCX file
-bibliography: references.bib
-csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
-abstract: |
-  Text of abstract
-keywords: |
-  keyword 1; keyword 2; keyword 3
-highlights: |
-  These are the highlights. 
 ---
@@
 
-# Narrative to accompany the slides
+# Narrative to accompany the slides for our talk
+
+These are our notes for preparing our slides for our CAA2018 presentation in the session "R as an archaeological tool: current state and directions" (https://github.com/MartinHinz/isaak_conf_sessions/tree/master/caa_2018_tuebingen)
 
@@
 We got `r how_many_articles_in_top_journals_for_cran_cites` articles in our sample, and we can make two major observations.
+
 First, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
 We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.
-The scale of the increases is quite impressive, we see many journals in our sample starting from less than 1% of articles citing R and now have around 20% of articles citing R.
+
+Second, the scale of the increases is quite impressive. We see many journals in our sample starting from less than 1% of articles citing R and now have around 20% of articles citing R.
 
@@
 
-We would now like to focus on a small sub-set of these papers using R, because they are doing something very interesting that we think all archaeologists using a programming language should be doing.
+We would now like to focus on a small sub-set of these papers using R, because they are doing something very interesting that we think all archaeologists doing data analysis also should be doing.
 This subset of papers are demonstrating the potential of R for open and reproducible research.
 
-...
+In this figure we show a simple example of what we see in this small-subset of papers.
+On the left we see a plain-text document written in R Markdown.
+This means that is has lines of code in amongst lines of normal text.
+We can format the text with simple additions, such as the underscore to indicate italics and the pound symbol to indicate a heading.
+These simple additions are Markdown, and we can also use Markdown for citations and cross-references, though we don't show them in our example here.
+This document format allows us to mix R code and text in the same document.
+Here we see a short snippet of R code in the text (nrow(cars)) as well as a longer block of code below.
+
+When this R Markdown document is rendered, or converted to another format such as PDF or Microsoft Word, we see that the code is replaced by the output it has generated.
+On the right side of this figure we see that the inline code has been replaced by the number 50, and the code block below has been replaced by the plot.
+
+This is how R helps to make research reproducible - we have found that some archaeologists are writing their entire journal articles useing R Markdown, and they share their plain-text document along with their publication.
+This means that the reader can see all of the details of the data analysis, and all of the small decisions made along the way.
+Many of these decisions are too tedious to describe in the methods section of an article, but they can be important for evaluating the robustness of an anlaysis.
+These details are also important to know if we want to extend a published analysis onto a new set of data.
+
+## Slide 10: An increasing number of archaeologists are publishing reproducible research
 
 ```{r}
+archy_ctv_readme <- readLines(str_glue('{here::here()}/analysis/data/raw_data/ctv-archaeology-README.md'))
 
-word_ratios <- tidy_cites %>%
-  count(word, ID) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(ID, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / sum(. + 1))) %>%
-  mutate(logratio = log(has_R_cites / no_R_cites)) %>%
-  arrange(desc(logratio))
+top_archy_journals <-
+ str_subset(archy_ctv_readme, "Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science")
 
-word_ratios %>%
-  group_by(logratio < 0) %>%
-  top_n(10, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (has_R_cites/no_R_cites)") +
-  scale_fill_discrete(name = "", labels = c("has_R_cites", "no_R_cites"))
+top_archy_journal_names <-
+  unlist(str_extract_all(archy_ctv_readme, "Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science"))
 
+top_archy_journal_names <- ifelse(top_archy_journal_names == "J Archaeol Sci",
+                                  "Journal of Archaeological Science",
+                                  top_archy_journal_names)
+
+archy_ctv_readme <- top_archy_journals
+archy_ctv_readme <- str_remove_all(archy_ctv_readme, "[[:punct:]]")
+archy_ctv_readme_20XX <- str_extract(archy_ctv_readme, " 20[[:digit:]]{2} ")
+archy_ctv_readme_20XX <- str_squish(unlist(archy_ctv_readme_20XX))
+archy_ctv_readme_20XX <- as.numeric(archy_ctv_readme_20XX)
+archy_ctv_readme_20XX <- archy_ctv_readme_20XX[archy_ctv_readme_20XX > 2010]
+
+number_of_reproducible_articles <- length(archy_ctv_readme_20XX)
+
+ggplot(data_frame(archy_ctv_readme_20XX,
+                  top_archy_journal_names),
+       aes(archy_ctv_readme_20XX,
+           fill = str_wrap(top_archy_journal_names, 15))) +
+  geom_bar() +
+  theme_minimal(base_size = 16) +
+  scale_x_continuous(limits = c(min(archy_ctv_readme_20XX),
+                                max(archy_ctv_readme_20XX)+0.5),
+                     breaks = seq(min(archy_ctv_readme_20XX),
+                                  max(archy_ctv_readme_20XX)+0.5, 1)) +
+  ggtitle(str_glue('Articles in archaeology journals\nusing R for reproducible research (n = {length(archy_ctv_readme_20XX)})')) +
+  scale_fill_discrete(name = "Journal") +
+  theme(legend.text=element_text(size=10),
+        legend.key.size = unit(2, 'lines'),
+        legend.position = "bottom"
+        ) +
+  xlab("Year of publication") +
+  ylab("Number of articles")
+
+ggsave(filename = "archaeology_articles_r_reproducible.png",
+       path = str_glue('{here::here()}/analysis/figures'),
+       height = 9,
+       width = 9)
 ```
 
-## Notes on data analysis for the presentation
+In our sample of archaeology articles, we identified `r number_of_reproducible_articles` articles that use R to make their work reproducible.
+These articles include R script files and data files with the articles, either in the supplementary materials or in a trustworthy data repository.
+This plot shows that although the number is small, it is increasing, suggesting the beginning of a new approach to how archaeologists share their research.
 
-These are our notes for preparing our slides for our CAA2018 presentation in the session "R as an archaeological tool: current state and directions" (https://github.com/MartinHinz/isaak_conf_sessions/tree/master/caa_2018_tuebingen)
+These `r number_of_reproducible_articles` articles in the subset of articles that use R to make their work reproducible are distinct from the other articles the cite R because most of the articles that cite R lack several features.
+These include:
 
-## What are the top archaeology journals by number of papers published?
+- making the data files available with their publication
+- making the R script files available with the publication
+- using R Markdown to add narrative text with the R code to give explanation and make it easier for readers to understand the analysis
 
-We need to know this to plan for the data collection for our JCCA paper, it's not central to our CAA presentation, but I've put it here so I know where to find it later (I hope).
+Articles that have these three characteristics are part of a new paradigm in scientific communication, and with these `r number_of_reproducible_articles` articles, we have identified the beginning of this paradigm in archaeology.
 
-Data are from an 'advanced search' of apps.webofknowledge.com using WC=Archaeology on apps.webofknowledge.com.
+## Slide 11: Our new package, rrtools, and how it helps
 
-## Citations of R Core in articles on Web of Science
-
-We wanted to to see how R Core is cited in scientific literature in general, and in archaeology in particular.
-
-For all scientific journals tracked by the WoS:
-
-For archaeology journals only:
-
-For 2017 only:
+...
 
@@
 
-## Text mining titles in archaeology journal articles that cite R Core
-
-What are the key words in the titles of these archaeology articles that cite R Core, and how do they differ from other archaeology articles?
-
 # Acknowledgements

### 2018-03-06 analysis/data/raw_data/refs_from_ctv_archaeology.R

@@ -1,17 +1,0 @@
-archy_ctv_readme <- readLines("analysis/data/raw_data/ctv-archaeology-README.md")
-
-
-top_archy_journals <-
-str_subset(archy_ctv_readme, "Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science")
-
-archy_ctv_readme <- top_archy_journals
-
-archy_ctv_readme <- collapse(archy_ctv_readme)
-archy_ctv_readme <- str_remove_all(archy_ctv_readme, "[[:punct:]]")
-archy_ctv_readme_20XX <- str_extract_all(archy_ctv_readme, " 20[[:digit:]]{2} ")
-archy_ctv_readme_20XX <- str_squish(unlist(archy_ctv_readme_20XX))
-archy_ctv_readme_20XX <- as.numeric(archy_ctv_readme_20XX)
-archy_ctv_readme_20XX <- archy_ctv_readme_20XX[archy_ctv_readme_20XX > 2010]
-length(archy_ctv_readme_20XX)
-hist(archy_ctv_readme_20XX)
-stem(archy_ctv_readme_20XX)

### 2018-03-07 analysis/paper/paper.Rmd
up to slide 13

@@
 
-Articles that have these three characteristics are part of a new paradigm in scientific communication, and with these `r number_of_reproducible_articles` articles, we have identified the beginning of this paradigm in archaeology.
+Articles that have these three characteristics are part of a new paradigm in scientific communication, and with these `r number_of_reproducible_articles` articles, we believe we have identified the beginning of this paradigm in archaeology.
+We think it will not be hard for archaeologists that already use R to make the transition to producing reproducible research by sharing their code and data.
 
@@
 
-...
+As we studied the papers in our sample, we looked for common approaches to making and organising the code and data files.
+We found that there were few common standards used by archaeologists in making their work reproducible.
+There was a great variety of information presented in the code and data files.
+Some papers include code that will generate every figure and numer in the text, while other papers only include code to produce one or two key figures in the paper.
 
-```{r}
-library(glue)
+Looking further afield, we also reviewed current efforts towards improving the reproducibility of research in other disciplines.
+In reviewing the literature we found several useful guides in diverse fields such as ecology and fluid dynamics.
+These guides described several best practices that are not yet common in archaeology.
 
-min_y <- 2007
-max_y <- 2030
-library(ggrepel)
-ggplot(most_popular_journals_over_time,
-       aes(PY, 
-           n,
-           colour = SO)) +
-  geom_line(size = 2) +
-  geom_text_repel(
-    data = subset(most_popular_journals_over_time, 
-                  PY == max(PY)),
-    aes(label = SO, 
-        y = (n), 
-        x = PY ),
-    size = 5, 
-    direction = "y",
-    hjust = 0,
-    segment.colour = NA
-  ) +
-  theme_minimal() +
-  scale_x_continuous(breaks = min_y:max_y,
-                     labels = c(min_y:2018, rep("", length(2019:max_y))),
-                     limits = c(min_y, max_y)) +
-  xlab("Publication year") +
-  scale_y_continuous(trans='log2') +
-  ylab("Number of articles (log scale)") +
-  theme(legend.position="none") +
-  ggtitle("Number of articles per year in Archaeology journals",
-          subtitle = "Data from apps.webofknowledge.com,\nusing WC=Archaeology in SCI-EXPANDED & SSCI for 2007-2017")
+So, after looking at the reproducible articles in archaeology, and the best practices in other disciplines, we wanted to make a simple toolkit that will help archaeologists to write reproducible research as easily as possible.
+This toolkit is an R package called rrtools.
+This package contains a small number of functions that automate many common analysis setup tasks.
 
-```
+On this slide we show the typical use of the key functions in rrtools that a researcher would use to quickly set up a compendium suitable for writing reproducible research in R.
+Each of these steps reflects a best practice that we've noted in our review of the literature:
+
+- use_compendium:  This function creates a new directory for the research project, and creates in R package within this directory.
+We use the R package structure as the standard for our directory because it is widely recognised by R users.
+This means that people know where to look for code, where to look for data, and where to look for dependancies and other information.
+- usethis::use_mit_license: This function creates a copy of the MIT software lisence to the compendium.
+This is important for two reasons.
+First, because it tells the reader that the author is happy for their code to be reused by other people.
+And second, it tells the reader that the author does not take responsibility for any problems that the reader might have when they use the code.
+This is important for setting the expectations about the relationship between the author and the reader, regarding the use of code.
+- devtools::use_github:  This will initiate the use of the Git version control system.
+This is the current state-of-the-art for version control of any kind of plain text file, including text, code, data, and images.
+This will also create a repository on the GitHub website.
+Using version control is important because it allows the author to time-travel back to earlier states of their analysis, if they take a dead-end path.
+It is also important for making collaboration smooth because many people can work on one set of files without losing track of what the most recent version is.
+Plus it is nice to have a remote back-up of our work in case of an emergency.
+The GitHub repository can be kept private if desired, until the work is published.
+- use_readme_rmd:  This will create a simple document that is usually the first thing a reader sees when they browse the code files.
+This document is important because it helps to describe to the reader what to expect in the compendium, and it gives details of how to cite the compendium.
+- use_analysis:  This function creates a set of folders according to the best practices for organising a typical reasearch project.
+We recognise that this file structure wont be perfect for every project.
+But we know that it will suit many projects quite well, and it will at least make the user think about how to organise their files logically, and according to best practises.
+On the next slide we can see the folders that this fucntion creates:
+
+## Slide 12: The best-practice directory structure from rrtools
+
+Here is the file structure the rrtools gives us. We wont discuss it in detail, except to point out two important principles that it enforces:
+
+- Keep the data separate from the methods. This means keep the code in a separate directory to the data    
+- Keep the raw data seprate from data that is created or derived during the analysis.
+
+These two fundamental principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium.
+
+## Slide 13: Advanced reproducibility with rrtools
+
+The last three functions that rrtools gives us are advanced topics that we only summarise briefly here.  ...
 

### 2018-03-07 analysis/paper/paper.Rmd
Update paper.Rmd
adding a few sentences
- is what I added to "who is 'we'" actually correct?
- if there is no trend in an increase of using R to be seen, how plausible is it, that there is a paradigm shift if a few people start sharing code? I'm just unsure whether we can be this definite with these small numbers...  Though I'd like it, if there was one!

@@
 In this presentation we will talk about how archaeologists are changing their data analysis practices to make their research more reproducible and open.
-We will the rrtools package, which is our specific contribution to help with the challenge of making reproducible research in archaeology easier.
+We will introduce the rrtools package, which is our specific contribution to help with the challenge of making reproducible research in archaeology easier.
+"Our" meaning in this case a group of people whose names you can see here and / or who took part in thethe 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel under the guidance of Ben Marwick.
 
@@
 When readers cannot see the details of an analysis, it is difficult for them to be confident in the results described in an article.
+Also, for other researchers it may be near impossible to implement the described analysis in their own work as they lack the details on how to come to the desired conclusions.
 
@@
 First, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
+This is especially significant for archaeologists, as ecology and biology are subjects from which they quite often adapt methods for analysis, as for example in point-pattern-analysis and statistical modelling.
 We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.
@@
 There is a clear focus on scientific archaeology, with _Journal of Archaeological Science_ having the highest total number of articles.
+Archaeology as a subject has always been situated between a humanistic and a scientific tradition, but studies using statistical methods lean towards the sciences, which we mean with the term "scientific archaeology".
 We also see geoarchaeology, lithics, and environmental archaeology represented in the journal names.

### 2018-03-07 analysis/paper/paper.Rmd
all slides and text basically done

--- 2018-03-07 analysis/paper/paper.Rmd
@@
 
-```{r, setup, echo = FALSE}
+```{r, setup, echo = FALSE, message = FALSE, warning = FALSE}
 knitr::opts_chunk$set(
@@
   comment = "#>",
+  fig.height=5,
+  fig.width=5,
   fig.path = "../figures/"
@@
 library(scales)
+library(ggpmisc)
+library(Hmisc)
 ```
@@
 
-## Slide 1: Title slide
+## Slide: Title slide
 
@@
 
-## Slide 2: Motivation
+## Slide: Motivation
 
@@
 
-## Slide 3: Outline
+## Slide: Outline
 
@@
 
-Secondly, we will review a few case studies of how R is used by archaeologists, to show the potential that using a programming language has for opening and accelerating archaeological research.
+Secondly, we will review how R is used by archaeologists, to show the potential that using a programming language has for opening and accelerating archaeological research.
 
@@
 
-##- Slide 4: Current uses of R [signpost slide only]
+## Slide: Current uses of R [signpost slide only]
 
-## Slide 5: Articles in all sciences citing R
+## Slide: Articles in all sciences citing R
 
-```{r all-sciences-citing-r, fig.height=10}
+```{r all-sciences-citing-r}
 
@@
   theme_minimal() +
-  theme(legend.position="none") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
   ggtitle(str_glue('Percentage of articles per year citing R in top 10 WoS journals ({prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")} articles)'),
@@
 
-First, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
+**First**, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
 We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.
 
-Second, the scale of the increases is quite impressive. We see many journals in our sample starting from less than 1% of articles citing R and now have around 20% of articles citing R.
+**Second**, the scale of the increases is quite impressive. We see many journals in our sample starting from less than 1% of articles citing R and now have around 20% of articles citing R.
 
@@
 
-## Slide 6: Archaeology articles citing R
+## Slide: Archaeology articles citing R
 
@@
 
-library(ggpmisc)
 jas_cites_r <- 
@@
 
-There are four important details in these plots:
+There are three important details in these plots:
 
-The first detail we see here is that overall, the proportions of articles citing R in archaeology are much smaller than what we see in the ecology journals on the previous slide.
+The **first** detail we see here is that overall, the proportions of articles citing R in archaeology are much smaller than what we see in the ecology journals on the previous slide.
 The absolute numbers of archaeology articles here are are also very low.
@@
 
-The second detail is that we do not see a significant increase over time, unlike the ecology journals that steeply increase over time.
+The **second** detail is that we do not see a significant increase over time, unlike the ecology journals that steeply increase over time.
 Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, we do not yet see much of an upward trend in the archaeology journals.
@@
 
-The fourth detail we see is the distinctive types of journals that have articles citing R.
+The **third** detail we see is the distinctive types of journals that have articles citing R.
 There is a clear focus on scientific archaeology, with _Journal of Archaeological Science_ having the highest total number of articles.
@@
 
-## Slide 7: What are archaeologists doing with R?
+## Slide: What are archaeologists doing with R?
 
@@
 # subtract ones citing R from ones that do not
-library(Hmisc)
+
 wc_archaeology_2007_to_2017_no_R_cites <- 
@@
 
-## Slide 8: The unrealised potential of R for reproducible research [signpost slide only]
+## Slide: The unrealised potential of R for reproducible research [signpost slide only]
 
-## Slide 9: The unrealised potential of R for reproducible research [signpost slide only]
+## Slide 9: The unrealised potential of R for reproducible research
+
+![](../figures/Marwick et al Fig 3 screenshot from the RStudio program.png)
 
@@
 
-## Slide 10: An increasing number of archaeologists are publishing reproducible research
+## Slide: Real-world example from Barton et al. 2018
+
+![](../figures/example-barton-et-al.png)
+
+Here is a real-world example R Markdown in use, from an archaeology article published earlier this year.
+
+## Slide: An increasing number of archaeologists are publishing reproducible research
 
@@
 We think it will not be hard for archaeologists that already use R to make the transition to producing reproducible research by sharing their code and data.
+If this can be achieved, then we can realise some of the potential that R has for making archaeology more open and reproducible.
 
-## Slide 11: Our new package, rrtools, and how it helps
+## Slide: Our new package, rrtools, and how it helps
+
+![](../figures/rrtools-steps-carbon.png)
 
@@
 
-- use_compendium:  This function creates a new directory for the research project, and creates in R package within this directory.
+- **use_compendium:**  This function creates a new directory for the research project, and creates in R package within this directory.
 We use the R package structure as the standard for our directory because it is widely recognised by R users.
 This means that people know where to look for code, where to look for data, and where to look for dependancies and other information.
-- usethis::use_mit_license: This function creates a copy of the MIT software lisence to the compendium.
+- **usethis::use_mit_license:** This function creates a copy of the MIT software lisence to the compendium.
 This is important for two reasons.
@@
 This is important for setting the expectations about the relationship between the author and the reader, regarding the use of code.
-- devtools::use_github:  This will initiate the use of the Git version control system.
+- **devtools::use_github:**  This will initiate the use of the Git version control system.
 This is the current state-of-the-art for version control of any kind of plain text file, including text, code, data, and images.
@@
 The GitHub repository can be kept private if desired, until the work is published.
-- use_readme_rmd:  This will create a simple document that is usually the first thing a reader sees when they browse the code files.
+- **use_readme_rmd:**  This will create a simple document that is usually the first thing a reader sees when they browse the code files.
 This document is important because it helps to describe to the reader what to expect in the compendium, and it gives details of how to cite the compendium.
-- use_analysis:  This function creates a set of folders according to the best practices for organising a typical reasearch project.
+- **use_analysis:**  This function creates a set of folders according to the best practices for organising a typical reasearch project.
 We recognise that this file structure wont be perfect for every project.
 But we know that it will suit many projects quite well, and it will at least make the user think about how to organise their files logically, and according to best practises.
-On the next slide we can see the folders that this fucntion creates:
+On the next slide we can see the folders that this function creates:
 
-## Slide 12: The best-practice directory structure from rrtools
+## Slide: The best-practice directory structure from rrtools
 
-Here is the file structure the rrtools gives us. We wont discuss it in detail, except to point out two important principles that it enforces:
+![](../figures/rrtools-analysis-dir.png)
+
+Here is the file structure the rrtools gives us. There isn't time here to discuss it in detail, except to point out two important principles that it enforces:
 
@@
 
-## Slide 13: Advanced reproducibility with rrtools
+## Slide: Advanced reproducibility with rrtools
 
-The last three functions that rrtools gives us are advanced topics that we only summarise briefly here.  ...
+![](../figures/rrtools-steps-carbon.png)
+
+The last three functions that rrtools gives us are advanced topics that we summarise only very briefly here.
+
+- **use_dockerfile:** This function generates a configuration file that allows another user to recreate the computational environment of the original author.
+This is important for complex projects that have many pieces of software that might be difficult to install correctly.
+Sometimes when a piece of software is updates, its behaviour can change, which can alter the results of an analysis.
+This function helps to protect again this problem.
+- **use_travis:** This function will create a configuration file so that our compendium can be automatically checked each time we update it.
+This is especially useful when working with collaborators because it helps to identify problems in the analysis before they become too big.
+- **use_testthat:** This creates a template for adding tests for any custom functions used in the analysis.
+This is important so we can ensure that the functions perform as expected while we work on the anlysis.
+
+These three functions are very widely used in software engineering to give quality control to a package.
+With our rrtools package we have brought these tools into a research context where they can help researchers to make their analyses easy to work on and easy for others reproduce.
+
+## Slide: Summary
+
+To summarise:
+
+- We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in the sciences generally.
+We saw that ecological and evolutionary sciences have really embraced R in their publications.
+
+- We have also shown that archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields.
+We've seen that the use of R is most common in more scientific aspects of archaeology.
+
+- We also identified a small sub-set of archaeology papers that use R to make their work fully reproducible.
+There is a lot of potential for other archaeologists to make their work reproducible, and we think this would be a good thing for archaeology generally.
+
+- We have presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
+We hope this will help to realise some of the potential to make archaeological research more open and reproducible.
+
+Thank you.
 
 # Acknowledgements
+
+The rrtools project was developed during the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Exc264 Topoi, CRC1266, and ISAAKiel.
+
+Thanks to Daniel Knitter, Patrick Kennedy, Nils Müller-Scheeßel, Martin Hinz, Clemens Schmid, Ricarda Braun, Wojciech Francuzik for their contributions to the package.
 

### 2018-03-09 analysis/paper/paper.Rmd

@@
 
-**First**, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
+First, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
+This is especially significant for archaeologists, as ecology and biology are subjects from which they quite often adapt methods for analysis, as for example in point-pattern-analysis and statistical modelling.
 We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.

### 2018-03-09 analysis/paper/paper.Rmd
update slides, drop 'advanced repro', move ackno to front

@@
 We will introduce the rrtools package, which is our specific contribution to help with the challenge of making reproducible research in archaeology easier.
-"Our" meaning in this case a group of people whose names you can see here and / or who took part in thethe 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel under the guidance of Ben Marwick.
 
-## Slide: Motivation
+## Acknowledgements
+
+OUr project comes from the group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
+
+## Slide: Motivation & Outline
+
+ _Motivation_
 
@@
 
-## Slide: Outline
+_Outline_
 
@@
 
-## Slide: Advanced reproducibility with rrtools
-
-![](../figures/rrtools-steps-carbon.png)
-
-The last three functions that rrtools gives us are advanced topics that we summarise only very briefly here.
-
-- **use_dockerfile:** This function generates a configuration file that allows another user to recreate the computational environment of the original author.
-This is important for complex projects that have many pieces of software that might be difficult to install correctly.
-Sometimes when a piece of software is updates, its behaviour can change, which can alter the results of an analysis.
-This function helps to protect again this problem.
-- **use_travis:** This function will create a configuration file so that our compendium can be automatically checked each time we update it.
-This is especially useful when working with collaborators because it helps to identify problems in the analysis before they become too big.
-- **use_testthat:** This creates a template for adding tests for any custom functions used in the analysis.
-This is important so we can ensure that the functions perform as expected while we work on the anlysis.
-
-These three functions are very widely used in software engineering to give quality control to a package.
-With our rrtools package we have brought these tools into a research context where they can help researchers to make their analyses easy to work on and easy for others reproduce.
+We have some more advanced functions that we will discuss more at another time.
 
@@
 
-# Acknowledgements
+---
 
-The rrtools project was developed during the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Exc264 Topoi, CRC1266, and ISAAKiel.
+<!--
 
-Thanks to Daniel Knitter, Patrick Kennedy, Nils Müller-Scheeßel, Martin Hinz, Clemens Schmid, Ricarda Braun, Wojciech Francuzik for their contributions to the package.
+These are spare slides that might help with Q&A, but not for the main talk
+
+## Slide: Advanced reproducibility with rrtools
+
+![](../figures/rrtools-steps-carbon.png)
+
+The last three functions that rrtools gives us are advanced topics that we summarise only very briefly here.
+
+- **use_dockerfile:** This function generates a configuration file that allows another user to recreate the computational environment of the original author.
+This is important for complex projects that have many pieces of software that might be difficult to install correctly.
+Sometimes when a piece of software is updates, its behaviour can change, which can alter the results of an analysis.
+This function helps to protect again this problem.
+- **use_travis:** This function will create a configuration file so that our compendium can be automatically checked each time we update it.
+This is especially useful when working with collaborators because it helps to identify problems in the analysis before they become too big.
+- **use_testthat:** This creates a template for adding tests for any custom functions used in the analysis.
+This is important so we can ensure that the functions perform as expected while we work on the anlysis.
+
+These three functions are very widely used in software engineering to give quality control to a package.
+With our rrtools package we have brought these tools into a research context where they can help researchers to make their analyses easy to work on and easy for others reproduce.
+
+-->
 

### 2018-03-10 analysis/paper/paper.Rmd
first try at compressing the text a bit.

@@
 
-OUr project comes from the group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
+Our project comes from the group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
 
@@
 
-Our motivation for this topic comes from a profound change that we have been observing in how researchers analyse their data.
+Our motivation for this topic arises from a profound change that we have been observing in how researchers analyse their data.
 
@@
 ## Slide: Current uses of R [signpost slide only]
+So, let's start with the current uses of R.
 
@@
   ) +
+   annotate("text", x=2019, y=-0.01, label= "n = 50380") +
    scale_x_continuous(breaks = min_y:max_y,
@@
 The purpose of this filtering step was to find areas of research were R is generally known, so we can see how it's use is changing over time.
-Note that many articles that use R do not cite the R program, so these numbers are underestimates of the true use of R in science.
+Note that many articles that use R do not cite the R program, so these numbers underestimate the true use of R in science.
 
-We got `r how_many_articles_in_top_journals_for_cran_cites` articles in our sample, and we can make two major observations.
+We can make two major observations:
 
-First, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
-This is especially significant for archaeologists, as ecology and biology are subjects from which they quite often adapt methods for analysis, as for example in point-pattern-analysis and statistical modelling.
+**First**, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
+This is especially significant for archaeologists, as ecology and biology are subjects from which we quite often adapt methods for analysis, as for example in point-pattern-analysis and statistical modelling.
 We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.
 
-**Second**, the scale of the increases is quite impressive. We see many journals in our sample starting from less than 1% of articles citing R and now have around 20% of articles citing R.
+**Second**, the scale of the increases is quite impressive: Starting only 2012 some journals now have 20% of articles citing R.
 
@@
 We found `r archy_journals_total_number_of_citing_articles` articles that cite R over the last ten years.
-Note that this is certinaly not complete, since these are just for the journals tracked by the Web of Science database, and we know that archaeologists publish is a great variaty of journals.
+Note that this is certainly not complete, since these are just for the journals tracked by the Web of Science database, and we know that archaeologists publish is a great variety of journals.
 
@@
 
-The **first** detail we see here is that overall, the proportions of articles citing R in archaeology are much smaller than what we see in the ecology journals on the previous slide.
-The absolute numbers of archaeology articles here are are also very low.
+The **first** detail we see here is that overall, the proportions and the absolute number of articles citing R in archaeology are much smaller than what we see in the ecology journals on the previous slide.
 Only the _Journal of Archaeological Science_ has more than 10 articles in our sample.
-We have yet to adopt programming for data analysis in the same way that ecologists have.
+It's obvious, we have yet to adopt programming for data analysis in the same way that ecologists have.
 
-The **second** detail is that we do not see a significant increase over time, unlike the ecology journals that steeply increase over time.
-Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, we do not yet see much of an upward trend in the archaeology journals.
+The **second** detail is that we do not see a significant increase over time.
+Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, but we do not yet see much of an upward trend in the archaeology journals in contrast to e.g.
+ecology journals.
 Indeed, the first citations of R in several archaeology journals only occured in the last two years.
@@
 ## Slide: The unrealised potential of R for reproducible research [signpost slide only]
+After this overview how R is being used at the moment, we want to focus on the possibilities of making research reproducible using R.
 
@@
 
-We would now like to focus on a small sub-set of these papers using R, because they are doing something very interesting that we think all archaeologists doing data analysis also should be doing.
-This subset of papers are demonstrating the potential of R for open and reproducible research.
+We would now like to focus on those archaeological papers using R, that are demonstrating the potential of R for open and reproducible research.
+These papers are using R Markdown for their research, which is shown in this figure.
 
-In this figure we show a simple example of what we see in this small-subset of papers.
 On the left we see a plain-text document written in R Markdown.
 This means that is has lines of code in amongst lines of normal text.
-We can format the text with simple additions, such as the underscore to indicate italics and the pound symbol to indicate a heading.
-These simple additions are Markdown, and we can also use Markdown for citations and cross-references, though we don't show them in our example here.
-This document format allows us to mix R code and text in the same document.
-Here we see a short snippet of R code in the text (nrow(cars)) as well as a longer block of code below.
+We can format the text with simple additions, such as the underscore to indicate italics and the hash symbol to indicate a heading.
+These simple additions are Markdown, a very easy markup language with plain text formatting syntax.
+It is designed so that it can be converted to HTML and many other formats.
+**R Markdown** allows us to mix R code and text in the same document.
+Here we see a short snippet of R code right inside the text (nrow(cars)) as well as a longer block of code below.
 
@@
 
-This is how R helps to make research reproducible - we have found that some archaeologists are writing their entire journal articles useing R Markdown, and they share their plain-text document along with their publication.
+We have found that some archaeologists are writing their entire journal articles useing R Markdown, and they share their plain-text R Markdown document along with their publication.
 This means that the reader can see all of the details of the data analysis, and all of the small decisions made along the way.
-Many of these decisions are too tedious to describe in the methods section of an article, but they can be important for evaluating the robustness of an anlaysis.
-These details are also important to know if we want to extend a published analysis onto a new set of data.
+This way the analysis is easily reproducible for others.
 
@@
 
-Here is a real-world example R Markdown in use, from an archaeology article published earlier this year.
+Here is a real-world example R Markdown in use, from an archaeology article published in 2017.
 
@@
 
-In our sample of archaeology articles, we identified `r number_of_reproducible_articles` articles that use R to make their work reproducible.
-These articles include R script files and data files with the articles, either in the supplementary materials or in a trustworthy data repository.
-This plot shows that although the number is small, it is increasing, suggesting the beginning of a new approach to how archaeologists share their research.
+In our sample of R citing archaeology articles, we identified `r number_of_reproducible_articles` articles that include R script- and data files, either in the supplementary materials or in a trustworthy data repository.
+This plot shows that although this number is small, it is increasing, suggesting the beginning of a new approach to how archaeologists share their research.
 
-These `r number_of_reproducible_articles` articles in the subset of articles that use R to make their work reproducible are distinct from the other articles the cite R because most of the articles that cite R lack several features.
-These include:
+The way in which these `r number_of_reproducible_articles` articles make their work reproducible is by:
 
@@
 
-Articles that have these three characteristics are part of a new paradigm in scientific communication, and with these `r number_of_reproducible_articles` articles, we believe we have identified the beginning of this paradigm in archaeology.
+Articles that have these three characteristics are part of a new paradigm in scientific communication, and with these `r number_of_reproducible_articles` articles, we believe we may have identified the beginning of this behaviour in archaeology.
 We think it will not be hard for archaeologists that already use R to make the transition to producing reproducible research by sharing their code and data.
@@
 ## Slide: Our new package, rrtools, and how it helps
+
+So, we developed a package for R, which assists researchers in taking this step.
 
@@
 There was a great variety of information presented in the code and data files.
-Some papers include code that will generate every figure and numer in the text, while other papers only include code to produce one or two key figures in the paper.
+Some papers include code that will generate every figure and number in the text, while other papers only include code to produce one or two key figures in the paper.
 
@@
 
-We have some more advanced functions that we will discuss more at another time.
+Included in the package are some more advanced functions that we will discuss more at another time.
 

### 2018-03-10 analysis/paper/paper.Rmd
eliminated a few typos, replaced doubled words and superfluous sentence.

@@
 
-In this presentation we will talk about how archaeologists are changing their data analysis practices to make their research more reproducible and open.
+*Reproducible research in archaeology using R & rrtools*: In this talk, we will show  how archaeologists are changing their data analysis practices to make their research more reproducible and open.
 We will introduce the rrtools package, which is our specific contribution to help with the challenge of making reproducible research in archaeology easier.
@@
 
-Increasingly, we are seeing researchers publishing research that uses a programming language to analyse their data.
+Increasingly, we are seeing researchers publishing research that uses a programming language to calculate their models.
 Similarly, we see journal articles accompanied by scripts of programming language so that readers can study the details of the data analysis methods.
@@
 This is an important change away from using mouse-driven, point-and-click software for data analysis.
-When archaeologists use software such as SPSS and Past (as you see in our screenshots here), it is very difficult for them to share the details of their analysis.
+When archaeologists use software such as SPSS and Past (as you see in our screenshots here), it is very difficult for them to share the details of their calculations.
 When readers cannot see the details of an analysis, it is difficult for them to be confident in the results described in an article.
-Also, for other researchers it may be near impossible to implement the described analysis in their own work as they lack the details on how to come to the desired conclusions.
+Also, for other researchers it may be near impossible to implement the described methods in their own work as they lack the details on how to come to the desired conclusions.
 
@@
 We will look at citation data from thousands of journal articles that are available in the Web of Science database produced by Clarivate Analytics.
-We focus on the R programming language, because this is by far the most widely used langauge in archaeology.
+We focus on the R programming language, because this is by far the most widely used language in archaeology.
 
@@
 We searched the Web of Science for citations to the R program over the last ten years, and filtered the results to the top ten journals with the highest number of citations to this software.
-The purpose of this filtering step was to find areas of research were R is generally known, so we can see how it's use is changing over time.
+The purpose of this filtering step was to find areas of research were R is generally known, so we can see how its use is changing over time.
 Note that many articles that use R do not cite the R program, so these numbers underestimate the true use of R in science.
@@
 The **second** detail is that we do not see a significant increase over time.
-Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, but we do not yet see much of an upward trend in the archaeology journals in contrast to e.g.
-ecology journals.
-Indeed, the first citations of R in several archaeology journals only occured in the last two years.
+Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, but we do not yet see much of an upward trend in the archaeology journals in contrast to the ecology journals.
+Indeed, in several archaeology journals the first citations of R only occured in the last two years.
 The plot on the upper right of this slide shows a statistical test of change over time for articles in JAS.
@@
 
-Our results so far show that only a small number of archaeologists are using R for their research.
-The names of the journals hint that most of these people do scientific archaeology, rather than social or humanistic archaeology.
-
 Here we see a comparison of word frequencies of the titles of archaeology journal articles.
-Words near the red line are used with about equal frequencies by articles that cite R and by articles that do not cite R.
+Words near the red line are used with about equal frequencies by papers that cite R and by papers that do not cite R.
 Words far away from the red line are used much more by one group of articles compared to the other.
@@
 
-This plot supports our initial observations from the journals titles, which is that archaeologists using R tend to be doing more scientific archaeology.
+This plot supports our initial observations from the journals' titles, which is that archaeologists using R tend to be doing more scientific archaeology.
 We see that many words relating to stone artefacts, zooarchaeology, and geoarchaeology are more frequently used by R-using archaeologists.
@@
 It is designed so that it can be converted to HTML and many other formats.
-**R Markdown** allows us to mix R code and text in the same document.
+**R** Markdown allows us to mix R code and text in the same document.
 Here we see a short snippet of R code right inside the text (nrow(cars)) as well as a longer block of code below.
 
-When this R Markdown document is rendered, or converted to another format such as PDF or Microsoft Word, we see that the code is replaced by the output it has generated.
+When this R Markdown document is rendered, or converted to another format such as PDF or Microsoft Word, we see that the code is replaced by the output it generates.
 On the right side of this figure we see that the inline code has been replaced by the number 50, and the code block below has been replaced by the plot.
@@
 
-So, we developed a package for R, which assists researchers in taking this step.
-
-![](../figures/rrtools-steps-carbon.png)
-
 As we studied the papers in our sample, we looked for common approaches to making and organising the code and data files.
@@
 This package contains a small number of functions that automate many common analysis setup tasks.
+
+![](../figures/rrtools-steps-carbon.png)
 
@@
 But we know that it will suit many projects quite well, and it will at least make the user think about how to organise their files logically, and according to best practises.
-On the next slide we can see the folders that this function creates:
 
@@
 
-Here is the file structure the rrtools gives us. There isn't time here to discuss it in detail, except to point out two important principles that it enforces:
+Here is the file structure the rrtools gives us. For now, we want to point out two important principles that it enforces:
 
 - Keep the data separate from the methods. This means keep the code in a separate directory to the data    
-- Keep the raw data seprate from data that is created or derived during the analysis.
+- Keep the raw data seprate from data that is created or derived during the analysis. This will help you in *never* overwriting your raw data.
 

### 2018-03-10 analysis/paper/paper.Rmd
some more deletions and typo corrections.

@@
 In this presentation we will briefly review some data showing change in the use of programming languages in the sciences generally, and in archaeology in particular.
-We will look at citation data from thousands of journal articles that are available in the Web of Science database produced by Clarivate Analytics.
 We focus on the R programming language, because this is by far the most widely used language in archaeology.
@@
 This plot shows how data analysis is changing in the sciences more broadly towards the use of the R programming language.
-We searched the Web of Science for citations to the R program over the last ten years, and filtered the results to the top ten journals with the highest number of citations to this software.
+We searched the Web of Science, produced by Clarivate Analytics, for citations to the R program over the last ten years, and filtered the results to the top ten journals with the highest number of citations to this software.
 The purpose of this filtering step was to find areas of research were R is generally known, so we can see how its use is changing over time.
@@
 **First**, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
-This is especially significant for archaeologists, as ecology and biology are subjects from which we quite often adapt methods for analysis, as for example in point-pattern-analysis and statistical modelling.
+This is especially significant for us archaeologists, as ecology and biology are subjects from which we quite often adapt methods for analysis, as for example in point-pattern-analysis and statistical modelling.
 We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.
@@
 We found `r archy_journals_total_number_of_citing_articles` articles that cite R over the last ten years.
-Note that this is certainly not complete, since these are just for the journals tracked by the Web of Science database, and we know that archaeologists publish is a great variety of journals.
+Note that this is certainly not complete, since these are just the journals tracked by the Web of Science database, and we know that archaeologists publish in a great variety of journals.
 
@@
 The **second** detail is that we do not see a significant increase over time.
-Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, but we do not yet see much of an upward trend in the archaeology journals in contrast to the ecology journals.
-Indeed, in several archaeology journals the first citations of R only occured in the last two years.
+Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, but we do not yet see much of an upward trend in the archaeology journals.
+Indeed, in several of our journals the first citations of R only occured in the last two years.
 The plot on the upper right of this slide shows a statistical test of change over time for articles in JAS.
@@
 Words far away from the red line are used much more by one group of articles compared to the other.
-Words that appear in this plot are ones that have been used at least once in article titles of both groups.
 
-This plot supports our initial observations from the journals' titles, which is that archaeologists using R tend to be doing more scientific archaeology.
+This plot supports our initial observations: archaeologists using R tend to be doing more scientific archaeology.
 We see that many words relating to stone artefacts, zooarchaeology, and geoarchaeology are more frequently used by R-using archaeologists.
@@
 On this slide we show the typical use of the key functions in rrtools that a researcher would use to quickly set up a compendium suitable for writing reproducible research in R.
-Each of these steps reflects a best practice that we've noted in our review of the literature:
+Each of these steps reflects a best practice that we've noted in our review of the literature.
+We will focus now on the first five:
 
-- **use_compendium:**  This function creates a new directory for the research project, and creates in R package within this directory.
+- **use_compendium:**  This function creates a new directory for the research project, and creates an R package within this directory.
 We use the R package structure as the standard for our directory because it is widely recognised by R users.
@@
 - Keep the data separate from the methods. This means keep the code in a separate directory to the data    
-- Keep the raw data seprate from data that is created or derived during the analysis. This will help you in *never* overwriting your raw data.
+- Keep the raw data seprate from data that is created or derived during the analysis.
 
 These two fundamental principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium.
-
-Included in the package are some more advanced functions that we will discuss more at another time.
 

### 2018-03-10 analysis/paper/paper.Rmd
adding a motivational point about the greatness of reproducibility.
deleted my own added code line again
this " n= " in the plot wasn't needed, as it was already in the heading...

@@
 Also, for other researchers it may be near impossible to implement the described methods in their own work as they lack the details on how to come to the desired conclusions.
+Thus, using scripting language offers the chance of accelerating archaeological research by providing a way to retracing and checking published analyses as well as explaining methods more efficiently.
+These are the main advantages of creating reproducible research, which we would like to encourage.
 
@@
   ) +
-   annotate("text", x=2019, y=-0.01, label= "n = 50380") +
    scale_x_continuous(breaks = min_y:max_y,

### 2018-03-12 analysis/paper/paper.Rmd
make figures show at proper aspect ratio and size

@@
   comment = "#>",
-  fig.height=5,
-  fig.width=5,
   fig.path = "../figures/"
@@
 max_y <- 2025
+all_sciences_citing_R_over_time <- 
 ggplot(cran_cites_all_areas_top_journals_by_year,
@@
 
-ggsave(filename = "all_sciences_citing_R_over_time.png",
+ggsave(plot = all_sciences_citing_R_over_time, 
+       filename = "all_sciences_citing_R_over_time.png",
        path = str_glue('{here::here()}/analysis/figures'),
@@
        width = 9)
+
+knitr::include_graphics(str_glue('{here::here()}/analysis/figures/all_sciences_citing_R_over_time.png'))
 ```
@@
 
+all_archaeology_citing_R_over_time <- 
 ggplot(cran_cites_in_archy_journals_per_year,
@@
 
-ggsave(filename = "all_archaeology_citing_R_over_time.png",
+ggsave(plot = all_archaeology_citing_R_over_time, 
+       filename = "all_archaeology_citing_R_over_time.png",
        path = str_glue('{here::here()}/analysis/figures'),
@@
        width = 10)
+
+knitr::include_graphics(str_glue('{here::here()}/analysis/figures/all_archaeology_citing_R_over_time.png'))
 ```
@@
 
-# we have a very skewed sample, so let's match 
-cran_cites_in_archy_journals_lots <- 
-  cran_cites_in_archy_journals %>% 
-  sample_n(nrow(wc_archaeology_2007_to_2017_no_R_cites),
-           replace = TRUE)
-
 # combine
@@
 
+archaeology_articles_word_freqs <- 
 ggplot(frequency, 
@@
   ylab("Articles that cite R") +
-  ggtitle("Comparing the frequency of words used in archaeology articles")
+  ggtitle(str_glue("Comparing the frequency of words used in titles \nof archaeology articles ({prettyNum(nrow(frequency), ',')} words in {prettyNum(nrow(archy_journal_articles), ',')} articles)"))
 
-ggsave(filename = "archaeology_articles_word_freqs.png",
+ggsave(plot = archaeology_articles_word_freqs, 
+       filename = "archaeology_articles_word_freqs.png",
        path = str_glue('{here::here()}/analysis/figures'),
@@
        width = 9)
+
+knitr::include_graphics(str_glue('{here::here()}/analysis/figures/archaeology_articles_word_freqs.png'))
 ```
@@
 
+archaeology_articles_r_reproducible <- 
 ggplot(data_frame(archy_ctv_readme_20XX,
@@
 
-ggsave(filename = "archaeology_articles_r_reproducible.png",
+ggsave(plot = archaeology_articles_r_reproducible, 
+       filename = "archaeology_articles_r_reproducible.png",
        path = str_glue('{here::here()}/analysis/figures'),
@@
        width = 9)
+
+knitr::include_graphics( str_glue('{here::here()}/analysis/figures/archaeology_articles_r_reproducible.png'))
 ```

### 2018-09-03 analysis/paper/jcaa-paper.Rmd
working on the text

--- 2018-03-12 analysis/paper/paper.Rmd
@@
 ---
-title: "Reproducible research in archaeology using R & rrtools"
+title: "Tool-driven Revolutions in Archaeological Science"
 author:
@@
       reference_docx: "../templates/template.docx" # Insert path for the DOCX file
+abstract: "There is an argument in philosophy of science that revolutions in science are either ideas-driven or tool-driven.
+We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
+One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on pointing-and-clicking with a mouse in commercial software, to a new set of tools based on writing scripts in free and open source programming languages.
+We review bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
+We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches."
 ---
-
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
 
@@
 
-library(caa2018) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
+# library(caa2018) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
 
@@
 
-# Narrative to accompany the slides for our talk
+# Introduction
 
-These are our notes for preparing our slides for our CAA2018 presentation in the session "R as an archaeological tool: current state and directions" (https://github.com/MartinHinz/isaak_conf_sessions/tree/master/caa_2018_tuebingen)
+In this paper we investigate recent changes in the way archaeology is being done, especially the computational aspects.
 
-## Slide: Title slide
+[need some more in here]
 
-*Reproducible research in archaeology using R & rrtools*: In this talk, we will show  how archaeologists are changing their data analysis practices to make their research more reproducible and open.
-We will introduce the rrtools package, which is our specific contribution to help with the challenge of making reproducible research in archaeology easier.
+# Revolutions in Science: Ideas-driven or Tool-driven?
+
+There is an argument among philosophers of science that revolutions in science are either ideas-driven or tool-driven.
+One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] concept of a paradigm shift.
+In brief, this model proposes that the history of science consists of long periods of 'normal science' punctuated by short episodes of revolution.
+Normal science was described by Kuhn as routine day-to-day research focused on 'puzzle solving.' Revolutions in science occur when anomalous observations culminate in changes that replace one paradigm with another.
+
+... Galison
+
+# Tools Emerging from the Reproducibility Crisis
+
+We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
+One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on pointing-and-clicking with a mouse in commercial software, to a new set of tools based on writing scripts in free and open source programming languages...
+
+# Tool-adoption in archaeology and elsewhere
+
+We review bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
+We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches.
+
+# Discussion
+
+rrtools... huskydown
+
+special role of archaeology here, since we are concerned about past behaviours, we have a special interest in research leaving  a visible trace of our work
+
+# Conclusion
 
@@
 Our project comes from the group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
-
-## Slide: Motivation & Outline
-
- _Motivation_
-
-Our motivation for this topic arises from a profound change that we have been observing in how researchers analyse their data.
-
-Increasingly, we are seeing researchers publishing research that uses a programming language to calculate their models.
-Similarly, we see journal articles accompanied by scripts of programming language so that readers can study the details of the data analysis methods.
-
-This is an important change away from using mouse-driven, point-and-click software for data analysis.
-When archaeologists use software such as SPSS and Past (as you see in our screenshots here), it is very difficult for them to share the details of their calculations.
-When readers cannot see the details of an analysis, it is difficult for them to be confident in the results described in an article.
-Also, for other researchers it may be near impossible to implement the described methods in their own work as they lack the details on how to come to the desired conclusions.
-Thus, using scripting language offers the chance of accelerating archaeological research by providing a way to retracing and checking published analyses as well as explaining methods more efficiently.
-These are the main advantages of creating reproducible research, which we would like to encourage.
-
-_Outline_
-
-In this presentation we will briefly review some data showing change in the use of programming languages in the sciences generally, and in archaeology in particular.
-We focus on the R programming language, because this is by far the most widely used language in archaeology.
-
-Secondly, we will review how R is used by archaeologists, to show the potential that using a programming language has for opening and accelerating archaeological research.
-
-Finally, we will briefly introduce our R package, rrtools, which is designed to simplify and standardise the process of using R for archaeology, and making research more open and reproducible.
-
-## Slide: Current uses of R [signpost slide only]
-So, let's start with the current uses of R.
 

### 2018-09-04 analysis/paper/jcaa-paper.Rmd
more text

@@
 There is an argument among philosophers of science that revolutions in science are either ideas-driven or tool-driven.
-One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] concept of a paradigm shift.
-In brief, this model proposes that the history of science consists of long periods of 'normal science' punctuated by short episodes of revolution.
-Normal science was described by Kuhn as routine day-to-day research focused on 'puzzle solving.' Revolutions in science occur when anomalous observations culminate in changes that replace one paradigm with another.
+One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] effort to describe the history of science by proposing paradigm shifts resulting from revolutionary change in communities of researchers.
+In brief, this model proposes that the history of science consists of long periods of tradition-bound 'normal science' punctuated by short episodes of revolution.
+Normal science was described by Kuhn as stagnant, routine day-to-day research focused on 'puzzle solving'.
+What counts as suitable puzzles, and acceptable solutions to these puzzles, is governed by the norms and procedures of the prevailing paradigm.
+Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes that replace one paradigm with another in a conversion or speciation process where the community adopts a new way of doing science.
+There has been substantial debate about how to draw a clear distinction between so-called normal and revolutionary science.
+[@watkins1970against; @toulmin1970does; @Casadevalle00158_16].
+A notable criticism is that Kuhn's normal science is unscientific, a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
+This is at odds with a view of science that it is in a continuous state of revolution, with researchers simultaneously employing several styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
+In this view, based on the post-World War II era of science, radical discontinuities are rare, and major developments instead emerge from the division and recombination of already mature fields [@andersen2013second].
+Given the rarity of radical discontinuities, some biologists have argued that change in biology is better characterised by new ideas (e.g.
+Mendelian heredity, Darwinian evolution, and molecular genetics) replacing not a former paradigm but a conceptual vacuum [@wilkins1996are].
 
-... Galison
+In archaeology we have seen claims for paradigm shifts in several areas.
+The most wide-ranging application is Clark's [-@Clark_1993] comparison of North American archaeology to European archaeology during the 1970s-90s.
+Clark, similar to Binford and Sabloff [-@binford1982paradigms], identified the North American archaeological paradigm as rooted in anthropology, founded on post-positivist biases, and distinctive with it critical-realist ontology, a modified-objectivist epistemology, and an experimental-manipulative methodology.
+In contrast, the European archaeological paradigm is described a kind of history, mostly in a strict empiricist tradition, and typified by a realist ontology, an inductivist epistemology, and by an observational methodology.
+Clark's analysis describes a multi-paradigm situation where multiple research communities (i.e.
+North American and European) simutaneously operate in different paradigms (which Kuhn noted as a sign of a immature discipline).
+His account is not a complete Kuhnian analysis because it lacks a revolutionary change event where one paradigm is replaced by another, although he hints that processual archaeology may be considered a radical discontinuity relative to the culture-historical approaches that preceeded it [@Clark_1993, pp.
+206].
+
+A more thorough treatment of the claim that processual archaeology was a paradigm shift is provided by @Meltzer_1979.
+Meltzer reviews literature arguing for a revolutionary change in archaeology during the 1960s and 1970s, and considers these claims in light of classic Kuhnian revolutionary events such as the replacement of the Ptolemaic system by the Copernican system, and Newtonian dynamics by the new Einsteinian dynamics.
+He finds that the changes that occured in Anglophone archaeology after the 1960s were incremental, and primarily of method, rather than a widespread replacement of one ontological structure by another one, incommensurable or incompatible with the former.
+Trigger [-@Trigger_2006, pp.
+538] came to a similar conclusion at the end of his broad survey of the history of archaeology, that changes throughout archaeology have been mostly additive with only partial replacement.
+
+We see less formal mentions of paradigm shifts in archaeology in reference to processual to post-processual archaeology [@Härke_2002], as well as more localised and specific shifts in archaeological practice and thinking.
+For example, @Snodgrass_2002 claims that among Classical Archaeologists, an increase in research on previously neglected periods odf antiquity, such as the Greek Early Iron Age, constitutes a paradigm shift.
+@McAnany_Rowe_2015 propose that the appearance of community-based participatory models of research among some communities of archaeologists is a paradigm shift, though they concede that it is more transformational rather than revolutionary.
+@Fuller_2010 has used the term to describe a shift in thinking about the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to new concept of the emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
+Most of these paradigms are quite remote in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's paradigm shift approach, both as a tool to understand the history of science and as a guide to changing it, which is Kuhn's multiple definitions of a paradigm.
+
+Masterman [-@masterman1970does] has documented 21 different uses by Kuhn, which she organises into three groups.
+First is the metaphysical notion of a set of beliefs, and this is the sense that dominates most of the critical writing on Kuhn's model.
+Second is the sociological sense of scientific habits, the universially recognised scientific achievements that are the foundations of day-to-day normal science.
+These include specific attention-grabbing successes that provide a touchstone for the coordination of future research.
+Third is the concrete sense of an actual textbook, instrumentation or toolkit.
+This recognition of the multivalent nature of paradigms is important because it broadens the locus for scientific revolution to include sociological practices of science, such as might be produced by changes in norms of publication and peer review, and to include the physical infrastructure of doing science, such as the technology for collecting and analysing data, and for communication and collaboration among researchers.
+
+We can see a compelling exploration of the concrete sense of a paradigm in the work of @Galison_1997, who has offered a tool-driven view of change in research communities.
+Galison analysed of the role of tools in twentieth century particle physics, starting with hand-crafted cloud chambers and bubble chambers and ending with digital counters of and computers.
+He identifies a profound change in physics when analog devices producing pictures were superseded by digital devices producing numerical data.
+At the heart of this change is a shift from an intuitive approach, stimulated by visual and pictorial model-building to an approach based on logic, calculation, and demonstration.
+The changes are not purely at the points of data collection and analysis, but extend to the social and economic organisation of science, with new categories of physicists emerging, people who are not quite experimenters and not quite theorists.
+Dyson [-@Dyson_2000; -@dyson2012science] has further explored this tool-driven approach, linking the Galilean revolution to the use of the telescope in astronomy, and the Crick-Watson revolution to the use of X-ray diffraction.
+
+Galison's emphasis on tool-driven change, as a complement to Kuhnian concept-driven change, has much potential for archaeology.
+The centrality of material culture in Galison's view raises the question about whether there is scope for archaeologists to make productive contributions to understanding change in the history science through an archaeological analysis of scientific instrumentation [cf.
+@Schiffer_2013].
+Furthermore, Galison's analysis of the image/logic contrast in the history of physics invites a similar analysis of the history of archaeology Can we identify archaeological traditions focused on data collection and analysis using images and image-making devices [cf.
+@molyneaux2013cultural], in contrast to digital devices?
+One candidate for this might be the shift from building relative chronologies based on seriation of visually distinctive artefacts to absolute chronologies based on radiometric dating using computer-controlled instruments.
+A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to generate change in archaeology?
 

### 2018-09-05 analysis/paper/jcaa-paper.Rmd
more words

@@
 One candidate for this might be the shift from building relative chronologies based on seriation of visually distinctive artefacts to absolute chronologies based on radiometric dating using computer-controlled instruments.
-A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to generate change in archaeology?
+A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to predict or generate change in archaeology?
 
-# Tools Emerging from the Reproducibility Crisis
+# Bibliometric analysis of tool-driven change
+
+The question of predicting or generating change in scientific practice via adoption of new tools is especially intruiging, and in this section we explore how a change in ecology might be relevant for understanding the future of archaeology.
+@Touchon_McCoy_2016 investigated evaluated changes in statistics used by ecologists as a potential area of tool-driven change.
+They searched nearly 20,000 published articles in ecology published between 1990 and 2013.
+They found that there has been a rise in complex and computationally intensive statistical techniques such as mixed effects models and Bayesian statistics and a decline in reliance on approaches such as ANOVA or t-tests.
+Crucially, they found that ecologists have shifted away from software tools such as SAS and SPSS to the open source program R.
+
+@Touchon_McCoy_2016 identify four factors relating to technological change that might explain the changes they observed in the use of statistics in ecology.
+First, they note that automated data loggers, GPS trackers, remote sensing, and crowd sourcing have greatly increased rate at which ecologists collect data.
+Second, increases in desktop and cluster computing power has made complex analytical processes faster and more convienent to compute.
+Third, the developme of free, open source and easily extensible software such as R allows new methods to spread quickly via online fora and social media
+
+there has been a rise in powerful desktop and supercomputing available to anyone in a university setting.
+Third, and partly as a result of the increase in computing power available, there has been an explosion in new analytical techniques (e.g., Pinheiro and Bates 2000, Gelman and Hill 2006, McCarthy 2007, Bolker 2008, Legendre and Legendre 2012).
+The rise of powerful, freely available, open source analytical software like R (R Development Core Team 2013) and associated wiki's, blogs, and listservs where users consult with specialists and find code written by ecologists for ecologists allows the powerful combination of new methods and large data sets to be harnessed by us all.
+This process has democratized statistical inference, giving the average ecologist the power to conduct analyses previously reserved only for statisticians.
+Although the analyses that one can conduct in R are often achievable in a program like SAS, the community of ecologist statisticians using R has proliferated rapidly and may be driving change in techniques we document here (although it is admittedly impossible to tease apart cause and effect in this case).
+This community of users, who share code and advice about analyses, has enabled a great number of ecologists to conduct advanced analyses on their data.
+Indeed, the flexible nature of statistical programming is what allows those so inclined to “roll their own” statistics.
+Furthermore, the fact that R and other open source software is free has empowered a global community of users, including those in developing nations that may not have access to commercial costly programs.
+The fourth and final factor driving the change in statistical practices is that a number of seminal books and papers have fundamentally changed the way many people think about data analyses.
+To give just one example, Burnham and Anderson's (2002) book on model selection and inference has been cited nearly 30,000 times since its publication in 2002 according to Google Scholar (website accessed 8 December 2015).
 

### 2018-09-05 analysis/paper/paper.Rmd
oold stuff

@@
 output: 
-    bookdown::html_document2:
-      number_sections: false
+    bookdown::word_document2:
       fig_caption: yes
-      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
+      reference_docx: "../templates/template.docx" 
 ---
@@
   message = FALSE,
-  echo = FALSE,
+  echo = TRUE,
   cache = TRUE,

### 2018-09-07 analysis/paper/jcaa-paper.Rmd

@@
 
-In this paper we investigate recent changes in the way archaeology is being done, especially the computational aspects.
+In this paper we investigate recent changes in the way computational archaeology is being done.
+Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discurses[@clark_1993; @härke_2002], but there has been
+an argument in philosophy of science whether revolutions in science are either ideas-driven or tool-driven [@galison_1997; @dyson_2000].
+In this light we explore how many scientific disciplines have started to embrace methods to improve the reproducibility of their research and whether this change is part of a paradigm-shift.
+We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages and review bibliometric evidence for this change in the sciences generally and in archaeology in particular.
+We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and test if the adoption of these new tools may be considered a revolution or some other process.
 

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
no cache so I can build.
just put the first explanatory sentence between the headings

@@
       reference_docx: "../templates/template.docx" # Insert path for the DOCX file
+bibliography: references.bib
 abstract: "There is an argument in philosophy of science that revolutions in science are either ideas-driven or tool-driven.
@@
   echo = FALSE,
-  cache = TRUE,
+  cache = FALSE,
   comment = "#>",
@@
 Clark's analysis describes a multi-paradigm situation where multiple research communities (i.e.
-North American and European) simutaneously operate in different paradigms (which Kuhn noted as a sign of a immature discipline).
+North American and European) simutaneously operate in different paradigms (which Kuhn noted as a sign of a developing or immature discipline).
 His account is not a complete Kuhnian analysis because it lacks a revolutionary change event where one paradigm is replaced by another, although he hints that processual archaeology may be considered a radical discontinuity relative to the culture-historical approaches that preceeded it [@Clark_1993, pp.
@@
 # Bibliometric analysis of tool-driven change
+The question of predicting or generating change in scientific practice via adoption of new tools is especially intruiging, and in this section we explore how a change in ecology might be relevant for understanding the future of archaeology.
 
-The question of predicting or generating change in scientific practice via adoption of new tools is especially intruiging, and in this section we explore how a change in ecology might be relevant for understanding the future of archaeology.
+## Looking back on a tool-driven change in Ecology
+
 @Touchon_McCoy_2016 investigated evaluated changes in statistics used by ecologists as a potential area of tool-driven change.
@@
 Second, increases in desktop and cluster computing power has made complex analytical processes faster and more convienent to compute.
-Third, the developme of free, open source and easily extensible software such as R allows new methods to spread quickly via online fora and social media
+Third, the development of free, open source and easily extensible software for data analysis and visualisation, such as R, allows new methods to spread quickly via online fora and social media.
+Fourth, the publication of several books and papers that have strongly influenced the way many ecologists think about data analysis, such as @burnham2003model's book on model selection and inference.
 
-there has been a rise in powerful desktop and supercomputing available to anyone in a university setting.
-Third, and partly as a result of the increase in computing power available, there has been an explosion in new analytical techniques (e.g., Pinheiro and Bates 2000, Gelman and Hill 2006, McCarthy 2007, Bolker 2008, Legendre and Legendre 2012).
-The rise of powerful, freely available, open source analytical software like R (R Development Core Team 2013) and associated wiki's, blogs, and listservs where users consult with specialists and find code written by ecologists for ecologists allows the powerful combination of new methods and large data sets to be harnessed by us all.
-This process has democratized statistical inference, giving the average ecologist the power to conduct analyses previously reserved only for statisticians.
-Although the analyses that one can conduct in R are often achievable in a program like SAS, the community of ecologist statisticians using R has proliferated rapidly and may be driving change in techniques we document here (although it is admittedly impossible to tease apart cause and effect in this case).
-This community of users, who share code and advice about analyses, has enabled a great number of ecologists to conduct advanced analyses on their data.
-Indeed, the flexible nature of statistical programming is what allows those so inclined to “roll their own” statistics.
-Furthermore, the fact that R and other open source software is free has empowered a global community of users, including those in developing nations that may not have access to commercial costly programs.
-The fourth and final factor driving the change in statistical practices is that a number of seminal books and papers have fundamentally changed the way many people think about data analyses.
-To give just one example, Burnham and Anderson's (2002) book on model selection and inference has been cited nearly 30,000 times since its publication in 2002 according to Google Scholar (website accessed 8 December 2015).
+While all of these four factors could be considered relevant to the tooling of ecology and archaeology, we focus here on the third one because of its relevance to archaeology.
+Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct, with only a small overlap.
+What they share in common is working with the field-collected data on a computer to prepare it for publication.
+To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.
 
-We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
-One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on pointing-and-clicking with a mouse in commercial software, to a new set of tools based on writing scripts in free and open source programming languages...
-
-# Tool-adoption in archaeology and elsewhere
-
-We review bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
-We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches.
-
-# Discussion
-
-rrtools... huskydown
-
-special role of archaeology here, since we are concerned about past behaviours, we have a special interest in research leaving  a visible trace of our work
-
-# Conclusion
-
-## Acknowledgements
-
-Our project comes from the group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
-
-## Slide: Articles in all sciences citing R
-
-```{r all-sciences-citing-r}
+```{r ecology-citing-r}
 
@@
 # This is our exact search string for WoS:
-# 
-# You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM") AND CITED WORK: ("R: A Language and Environment for Statistical Computing") AND LANGUAGE: (English) AND DOCUMENT TYPES: (Article)
-# Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.
+
+# Results: 38,040
+# (from Web of Science Core Collection)
+# You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R # CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM")
+# Refined by: DOCUMENT TYPES: ( ARTICLE ) AND LANGUAGES: ( ENGLISH ) AND RESEARCH AREAS: ( ENVIRONMENTAL SCIENCES ECOLOGY )
+# Timespan: 1900-2018. Indexes: SCI-EXPANDED, SSCI, A&HCI, # ESCI.
+
+# field tags
+# TS= Topic 
+# TI= Title 
+# AU= Author [Index] 
+# AI= Author Identifiers 
+# GP= Group Author [Index] 
+# ED= Editor 
+# SO= Publication Name [Index] 
+# DO= DOI 
+# PY= Year Published 
+# AD= Address 
+# OG= Organization-Enhanced [Index] 
+# OO= Organization 
+# SG= Suborganization 
+# SA= Street Address 
+# CI= City 
+# PS= Province/State 
+# CU= Country/Region 
+# ZP= Zip/Postal Code 
+# FO= Funding Agency 
+# FG= Grant Number 
+# FT= Funding Text 
+# SU= Research Area 
+# WC= Web of Science Category 
+# IS= ISSN/ISBN 
+# UT= Accession Number 
+# PMID= PubMed ID
 
 cran_cites_files <- 
-    paste0(here::here(), glue("/analysis/data/raw_data/cited_cran/savedrecs ({1:11}).txt"))
+    list.files(here::here("/analysis/data/raw_data/cited_cran/"),
+               full.names = TRUE)
 
@@
 
+articles_shown_in_plot <- 
+  cran_cites_all_areas_top_journals_by_year %>% 
+  filter(!is.na(prop)) %>% 
+  summarise(sum_n = sum(n)) %>% 
+  pull(sum_n)
+
+# nice labels with n for plot
+cran_cites_all_areas_top_journals_by_year <- 
+cran_cites_all_areas_top_journals_by_year %>% 
+  group_by(SO) %>% 
+  summarise(total_n_articles = sum(n)) %>% 
+  mutate(SO_n = as.character(glue('{SO} (n = {total_n_articles})'))) %>% 
+  right_join(cran_cites_all_areas_top_journals_by_year,
+             by = "SO")
+```
+
+We used the 'Cited Reference Search' function provided by the Web of Science online scientific citation indexing service to identify journal articles that cite R.
+Although R has been available since the late 1990s [@Thieme_2018], a recommended format for citing the software did not appear until 2004, with the author given as "R Development Core Team".
+This recommended format for citing the software changed slightly in 2012 when the author was updated to "R Core Team".
+We searched the Web of Science using '"R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM"' in the CITED AUTHOR field.
+We sorted the results by frequency of citation, and selected the first 1000 items (the maximum allowed by the Web of Science service).
+These 1000 items represent variations on the recommended format for citing R, and found in the reference lists of `r prettyNum(how_many_articles_in_top_journals_for_cran_cites, ",")` English-languge articles indexed by the Web of Science in the research area of Environmental Sciences Ecology.
+We then downloaded the bibliographic data for all of these articles.
+
+Figure \@ref(fig:ecology-citing-r-plot) shows the percentage of articles citing R in each of several of the top ecology journals (as defined by how often their articles are cited).
+We restrict the start of the observation period to 2007 for convivnence.
+The plot shows that the percentage of articles citing R has increased from less than ten percent in all journals in the late 2000s, to more than 30% in _Ecosphere_, _Ecology and Evolution_, and _Molecular Ecology_ after 2012.
+We would not call this a Kuhnian paradigm shift, but it does show a substantial change in the tools of the discipline, supporting claims for a Galisonian tool-driven change in ecology similar to the changes described by @Touchon_McCoy_2016.
+
+```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Percentage of articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}).
+Data from Web of Science for 2007-2017')}
 min_y <- 2007
@@
                   PY == max(PY)),
-    aes(label = str_wrap(SO, 40)),
+    aes(label = str_wrap(SO_n, 40)),
     size = 4,
@@
   xlab("Publication year") +
-  ylab("Percentage of articles in that journal") +
+  ylab("Percentage of articles in that journal in that year") +
   theme_minimal() +
   theme(legend.position="none",
-        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
-  ggtitle(str_glue('Percentage of articles per year citing R in top 10 WoS journals ({prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")} articles)'),
-          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
 
@@
 
-This plot shows how data analysis is changing in the sciences more broadly towards the use of the R programming language.
-We searched the Web of Science, produced by Clarivate Analytics, for citations to the R program over the last ten years, and filtered the results to the top ten journals with the highest number of citations to this software.
-The purpose of this filtering step was to find areas of research were R is generally known, so we can see how its use is changing over time.
-Note that many articles that use R do not cite the R program, so these numbers underestimate the true use of R in science.
-
-We can make two major observations:
-
-**First**, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences.
-This is especially significant for us archaeologists, as ecology and biology are subjects from which we quite often adapt methods for analysis, as for example in point-pattern-analysis and statistical modelling.
-We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.
-
-**Second**, the scale of the increases is quite impressive: Starting only 2012 some journals now have 20% of articles citing R.
-
-So, we can see this is a substantial change in the way researchers are working in many fields. What about archaeology? Do we also see a change like this?
-
-## Slide: Archaeology articles citing R
+## Looking forward to a tool-driven change in Archaeology
 
@@
 
-# get the top archaeology journals by numbers of articles.
-wc_archaeology_2017 <- 
-  readr::read_tsv(paste0(here::here(), "/analysis/data/raw_data/savedrecs.txt"), 
-            quote = "")
+# Results: 44,047
+# (from Web of Science Core Collection)
+# You searched for: (WC=Archaeology) AND LANGUAGE: (English) AND DOCUMENT TYPES: # (Article)
+# Timespan: 1900-2018. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI
 
-# The 10 journals with the most articles are:
-top_journals_by_articles <- 
-wc_archaeology_2017 %>% 
+#- get journals with the most number of articles
+
+archy_files <- 
+    list.files(here::here("/analysis/data/raw_data/wos_archaeology/"),
+               full.names = TRUE)
+
+archy_all_years <- 
+  map(archy_files,
+      ~readr::read_tsv(.x, 
+                       quote = "",
+                       col_types = cols(.default = col_character()))) %>% 
+  bind_rows() %>% 
+  mutate(PY = as.numeric(PY)) %>% 
+  filter(PY <= 2017)
+
+# limit to top journals
+top_journals_for_archy<- 
+  archy_all_years %>% 
   group_by(SO) %>% 
@@
 
-# For 2007-2017:
+# n articles in top journals per year
+top_journals_for_archy_per_year <- 
+  archy_all_years %>% 
+  group_by(SO, PY) %>% 
+  tally() %>% 
+  arrange(PY)
 
-data_files_from_2007_to_2017 <- 
-  paste0(here::here(), glue("/analysis/data/raw_data/savedrecs ({1:2}).txt"))
+how_many_articles_in_top_archy_journals <- 
+  sum(top_journals_for_archy$n)
 
-wc_archaeology_2007_to_2017 <- 
-  map(data_files_from_2007_to_2017,
-      ~readr::read_tsv(.x, quote = "")) %>% 
-  bind_rows()
+#-- get articles citing R that are archaeology
 
-total_number_of_articles_in_sample <- 
-  nrow(wc_archaeology_2007_to_2017)
+# Results: 137
+# (from Web of Science Core Collection)
+# You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM")
+# Refined by: WEB OF SCIENCE CATEGORIES: ( ANTHROPOLOGY ) AND WEB OF SCIENCE CATEGORIES: ( ARCHAEOLOGY )
+# Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI
 
-# most populararchaeology journals over time, that have some mention of R
-most_popular_journals_over_time <- 
-wc_archaeology_2007_to_2017 %>% 
+archy_cran_files <- 
+      list.files(here::here("/analysis/data/raw_data/cited_cran_archaeology/"),
+               full.names = TRUE)
+
+archy_cran_all_years <- 
+  map(archy_cran_files,
+      ~readr::read_tsv(.x, 
+                       quote = "",
+                       col_types = cols(.default = col_character()))) %>% 
+  bind_rows() %>% 
+  mutate(PY = as.numeric(PY)) %>% 
+  filter(PY <= 2017)
+
+# tally by year
+archy_cran_cites_all_areas_top_journals_by_year <- 
+  archy_cran_all_years %>% 
+  filter(SO %in% top_journals_for_archy$SO) %>% 
   group_by(SO, PY) %>% 
   tally(sort = TRUE)  %>% 
-  filter(!SO %in% c("JOURNAL OF CULTURAL HERITAGE",
-                    "STUDIES IN CONSERVATION",
-                    "INTERSECCIONES EN ANTROPOLOGIA",
-                    "ISRAEL EXPLORATION JOURNAL",
-                    "JOURNAL OF MATERIAL CULTURE",
-                    "TRABAJOS DE PREHISTORIA",
-                    "ARCHEOSCIENCES-REVUE D ARCHEOMETRIE")) %>% 
-  filter(n >= 2) %>% 
+  left_join(top_journals_for_archy_per_year,
+            by = c('SO', 'PY')) %>% 
+   mutate(prop = n.x / n.y) %>% 
   ungroup()
 
-most_popular_journals_over_time_names <- 
-  unique(most_popular_journals_over_time$SO)
+total_number_archaeology_articles_citing_r <- sum(archy_cran_cites_all_areas_top_journals_by_year$n.x)
 
-archy_journals_total_number_of_articles <- 
-  read_csv(str_glue('{here::here()}/analysis/data/raw_data/top_archaeology_journals_articles_per_year.csv'))
+# check for slight differences in journal names
+# sort(unique(archy_cran_cites_all_areas_top_journals_by_year$SO))
 
-top_archaeology_journals_articles_per_year <- 
-  archy_journals_total_number_of_articles %>% 
-  gather(variable, value, -journal_name) %>% 
-  mutate(year = as.numeric(variable)) %>% 
-  select(-variable)
-
-cran_cites_in_archy_journals <- 
-  cran_cites_all_areas_all_years %>% 
-  filter(SO %in% most_popular_journals_over_time_names)
-
-# what archaeology journals cite R?
-# unique(cran_cites_in_archy_journals$SO)
-
-archy_journals_total_number_of_citing_articles <- 
-  nrow(cran_cites_in_archy_journals)
-
-cran_cites_in_archy_journals_per_year <- 
-  cran_cites_in_archy_journals %>% 
-  group_by(SO, PY) %>% 
-  tally(sort = TRUE) %>% 
-  ungroup() %>% 
-  mutate(PY = as.numeric(PY))  %>% 
-  left_join(top_archaeology_journals_articles_per_year,
-            by = c("SO" = "journal_name",
-                   "PY" = "year")) %>% 
-  mutate(prop = n / value) %>% 
+# nice labels with n for plot
+archy_cran_cites_all_areas_top_journals_by_year <- 
+archy_cran_cites_all_areas_top_journals_by_year %>% 
   group_by(SO) %>% 
-  mutate(jn = str_glue('{SO} (n = {sum(n)})')) %>% 
-  ungroup()
+  summarise(total_n_articles = sum(n.x)) %>% 
+  mutate(SO_n = as.character(glue('{SO} (n = {total_n_articles})'))) %>% 
+  right_join(archy_cran_cites_all_areas_top_journals_by_year,
+             by = "SO")
 
@@
 
+all_archaeology_citing_R_over_time <- 
+ggplot(archy_cran_cites_all_areas_top_journals_by_year,
+       aes(PY,
+           prop,
+           colour = SO)) +
+  geom_line(size = 2) +
+  geom_point() +
+  geom_text_repel(
+    data = archy_cran_cites_all_areas_top_journals_by_year %>% 
+      group_by(SO) %>% 
+      slice(which.max(PY)) , 
+    aes(label = SO_n),
+    direction = "y",
+    size = 5,
+    nudge_x = 1,
+    hjust = 0) +
+  theme_minimal() +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle = 90, 
+                                   hjust = 1, 
+                                   vjust = 0.5)) +
+  scale_x_continuous(breaks = min_y:max_y,
+                     labels = c(min_y:2017, 
+                                rep("",
+                                    length(2018:max_y))),
+                     limits = c(min_y, max_y)) +
+  scale_y_continuous(labels = scales::percent,
+                     limits = c(0, 0.2)) +
+  xlab("Publication year") +
+  ylab("Percentage of articles in that journal")
+
+ggsave(plot = all_archaeology_citing_R_over_time, 
+       filename = "all_archaeology_citing_R_over_time.png",
+       path = str_glue('{here::here()}/analysis/figures'),
+       height = 7,
+       width = 11)
+```
+
+```{r j-arch-sci-articles-citing-R}
+# look at the trend in JAS, where there are the most articles
+
+how_many_articles_in_jas <-  
+  archy_cran_cites_all_areas_top_journals_by_year$total_n_articles[1]
+
 jas_cites_r <- 
-cran_cites_in_archy_journals_per_year %>% 
+archy_cran_cites_all_areas_top_journals_by_year %>% 
   filter(SO == "JOURNAL OF ARCHAEOLOGICAL SCIENCE") %>% 
@@
                   geom = "text",
-                  aes(label = paste("p-value: ",signif(..p.value.., digits = 4)))) +
+                  aes(label = paste("p-value: ",
+                                    signif(..p.value.., 
+                                           digits = 4)))) +
   xlab("year") +
   ylab("Proportion of articles citing R") +
-  ggtitle(expression(paste("Articles citing R in ", italic("Journal of Archaeological Science")))) +
+  ggtitle(expression(paste("Articles citing R in ", 
+                           italic("Journal of Archaeological Science")))) +
   theme_bw(base_size = 6)
+```
 
-all_archaeology_citing_R_over_time <- 
-ggplot(cran_cites_in_archy_journals_per_year,
-       aes(PY,
-           prop,
-           colour = SO)) +
-  geom_line(size = 2) +
-  geom_text_repel(
-    data = subset(cran_cites_in_archy_journals_per_year, 
-                  PY == max(PY)),
-    aes(label = jn),
-    direction = "y",
-    size = 5,
-    nudge_x = 1,
-    hjust = 0) +
-  theme_minimal() +
-  theme(legend.position="none",
-        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
-  scale_x_continuous(breaks = min_y:max_y,
-                     labels = c(min_y:2017, rep("", length(2018:max_y))),
-                     limits = c(min_y, max_y)) +
-  scale_y_continuous(labels = scales::percent) +
-  xlab("Publication year") +
-  ylab("Percentage of articles in that journal") +
-  ggtitle(str_glue('Percentage of Archaeology articles per year citing R ({archy_journals_total_number_of_citing_articles} out of {prettyNum(total_number_of_articles_in_sample, ",")} articles in sample)'),
-          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")  +
+We conducted a similar bibliometric analysis for archaeology journals indexed by the Web of Science service.
+We used the 'Cited Reference Search' to find articles citing R in the same way as above, and then refined the results to keep only those articles published during 2007-2017 that are included in the Web of Science category 'Archaeology'.
+This resulted in `r how_many_articles_in_top_archy_journals` articles, of which `r total_number_archaeology_articles_citing_r` cite R.
+Figure \@ref(fig:rchaeology-journals-with-subplot) shows the temporal trend of citations of R in archaeological articles.
+There are three interesting details revealed by this figure.
+
+```{r archaeology-journals-with-subplot, fig.cap=str_glue('Percentage of Archaeology articles per year citing R ({total_number_archaeology_articles_citing_r} out of {prettyNum(how_many_articles_in_top_archy_journals, ",")} articles in sample.
+Data from Web of Science for 2007-2017')}
+
+# combine main plot and subplot
+all_archaeology_citing_R_over_time_and_subplot <- 
+all_archaeology_citing_R_over_time  +
   annotation_custom(grob = ggplotGrob(jas_cites_r), 
@@
 
-ggsave(plot = all_archaeology_citing_R_over_time, 
-       filename = "all_archaeology_citing_R_over_time.png",
+ggsave(plot = all_archaeology_citing_R_over_time_and_subplot, 
+       filename = "all_archaeology_citing_R_over_time_and_subplot.png",
        path = str_glue('{here::here()}/analysis/figures'),
-       height = 6,
-       width = 10)
+       height = 9,
+       width = 12)
 
-knitr::include_graphics(str_glue('{here::here()}/analysis/figures/all_archaeology_citing_R_over_time.png'))
+knitr::include_graphics(str_glue('{here::here()}/analysis/figures/all_archaeology_citing_R_over_time_and_subplot.png'))
 ```
 
-Here are the results for archaeological articles.
-We found `r archy_journals_total_number_of_citing_articles` articles that cite R over the last ten years.
-Note that this is certainly not complete, since these are just the journals tracked by the Web of Science database, and we know that archaeologists publish in a great variety of journals.
+The first detail we see here is that overall, the proportions and the absolute number of articles citing R in archaeology are much smaller than what we see in the ecology journals in Figure \@ref(fig:ecology-citing-r).
+Only the _Journal of Archaeological Science_ has more than 50 articles in our sample.
+We might conclude from this observation that archaeologists have yet to adopt programming for data analysis and visualisation in the same way that ecologists have.
 
-There are three important details in these plots:
+The second detail is that most archaeology journals do not show any increase in the percentage of article citing R over time .
+The ecology journals shown in Figure \@ref(fig:ecology-citing-r) show clear upward trends of increasing proportions of articles citing R over time, but we do not see any clear trends in the archaeology articles considered together.
+Indeed, in several of the archaeological journals in our sample, the first citations of R only occured in the last two years.
+The sample sizes for the archaeology journals are too small to confidently infer any trend of over time, with the exception of the _Journal of Archaeological Science_ with `r how_many_articles_in_jas` articles.
+The inset plot in Figure \@ref(fig:rchaeology-journals-with-subplot) shows a statistical test of change over time for articles in the _Journal of Archaeological Science_.
+With a moderate r-squared value and low p-value for the linear model, we can conclude that there is some evidence for an increase in citations of R over time in this journal.
+This may indicate an increasing use of R among archaeologists, especially those working at the intersection of the natural and physical sciences.
+This may hint at the start of a trend similar to what we see in the ecology journals of a widespread
 
-The **first** detail we see here is that overall, the proportions and the absolute number of articles citing R in archaeology are much smaller than what we see in the ecology journals on the previous slide.
-Only the _Journal of Archaeological Science_ has more than 10 articles in our sample.
-It's obvious, we have yet to adopt programming for data analysis in the same way that ecologists have.
+Following on from this observation about the increasing popularity of R in then _Journal of Archaeological Science_, the third detail we see in these results is the distinctive types of journals that have articles citing R.
+There is a clear focus on journals that publich scientific archaeology, especially those that focus on specialised empirical analysis.
+In addition to the general archaeology journals, we also see _Lithic Technology_, _Anthropozoologica_, and _International Journal of Osteology_ in the journal names, indicating that we find R used by specialists in stone artefacts and faunal remains.
+To further investigate the topics that archaeologists are useing R to study, we conducted a statistical analysis of the words in the titles of all the articles in our sample.
 
-The **second** detail is that we do not see a significant increase over time.
-Both archaeology and the other sciences we saw on the previous slide show an increase in citing R at around 2013-2014, but we do not yet see much of an upward trend in the archaeology journals.
-Indeed, in several of our journals the first citations of R only occured in the last two years.
-The plot on the upper right of this slide shows a statistical test of change over time for articles in JAS.
-With a low r-squared value and high p-value for the linear model, we conclude that there is no evidence for an increase in citations over time in this journal.
-
-The **third** detail we see is the distinctive types of journals that have articles citing R.
-There is a clear focus on scientific archaeology, with _Journal of Archaeological Science_ having the highest total number of articles.
-Archaeology as a subject has always been situated between a humanistic and a scientific tradition, but studies using statistical methods lean towards the sciences, which we mean with the term "scientific archaeology".
-We also see geoarchaeology, lithics, and environmental archaeology represented in the journal names.
-We can investigate this further with statistical analysis of the words in the titles of all the articles in our sample.
-
-## Slide: What are archaeologists doing with R?
+## What are R-using archaeologists writing about?
 
 ```{r text-mining-article-titles-archaeology}
+
 # https://www.tidytextmining.com/twitter.html
@@
 wc_archaeology_2007_to_2017_no_R_cites <- 
-  wc_archaeology_2007_to_2017 %>% 
-  filter(TI %nin% cran_cites_in_archy_journals$TI)
+  archy_all_years %>% 
+  filter(TI %nin% archy_cran_all_years$TI)
 
@@
                    VL = as.character(VL)),
-           cran_cites_in_archy_journals %>% 
+           archy_cran_all_years %>% 
             mutate(ID = "has_R_cites"))
@@
                       "archaeological",
-                      "analysis", "study", "site"),
+                      "analysis", "study", "site", "period", "past", "patterns", "investigation", "analyses", "late", "region", "based", "central", "north", "south", "east", "west", "western", "eastern", "ad", "pre", "sites", "evidence", "prehistoric", "remains", "change", "de", "identification", "northern", "southern", "valley", "implications", "valley", "history", "historical", "modern", "context", "stable", "middle", "approach", "san", "artifacts", "palaeolithic"),
          str_detect(word, "[a-z]"))
@@
 
-frequency <- frequency %>% 
+frequency <- 
+  frequency %>% 
   select(ID, word, freq) %>% 
@@
   arrange(no_R_cites, has_R_cites)
+
+```
+
+We computed a comparison of word frequences in journal article titles to get a better understanding of what topics archaeologists are writing about when they cite R.
+First we separated the archaeology articles into two groups, those that cite R, and those that do not.
+Second, we filter to keep only words that occur in titles in both groups, and remove stop words that are very common and uninformative (e.g.
+'the', 'archaeological', 'study').
+This resulted in `r prettyNum(nrow(frequency), ',')` words in `r prettyNum(nrow(archy_journal_articles))`.
+Third, for each word found in all the titles of the journal articles in each group, we computed its proportion of the total number of words in all the titles in each group.
+Figure \@ref(fig:text-mining-article-titles-archaeology) shows the results of this analysis.
+Words near the red line are used with about equal frequencies by papers that cite R and by papers that do not cite R.
+Words far away from the red line are used much more by one group of articles compared to the other.
+
+```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles \nof archaeology articles ({prettyNum(nrow(frequency), ',')} words in {prettyNum(nrow(archy_journal_articles), ',')} articles)")}
 
@@
     geom_abline(color = "red", 
-              size = 2) +
-  geom_jitter(alpha = 0.1, 
+              size = 1) +
+  geom_point(alpha = 0.1, 
               size = 2.5, 
@@
   geom_text_repel(aes(label = word), 
-            size = 5,
-            segment.colour = NA) +
+            size = 3,
+            segment.colour = "grey80") +
   scale_x_log10(labels = percent_format(),
-                limits = c(0.001, 0.02)) +
-  scale_y_log10(labels = percent_format()) +
-
+                limits = c(0.00075, 0.015)) +
+  scale_y_log10(labels = percent_format(), 
+                limits = c(0.00075, 0.015)) +
   theme_minimal(base_size = 16) +
@@
   ylab("Articles that cite R") +
-  ggtitle(str_glue("Comparing the frequency of words used in titles \nof archaeology articles ({prettyNum(nrow(frequency), ',')} words in {prettyNum(nrow(archy_journal_articles), ',')} articles)"))
+  coord_equal()
 
@@
 
-Here we see a comparison of word frequencies of the titles of archaeology journal articles.
-Words near the red line are used with about equal frequencies by papers that cite R and by papers that do not cite R.
-Words far away from the red line are used much more by one group of articles compared to the other.
+Figure \@ref(fig:text-mining-article-titles-archaeology-plot) supports our initial observations that archaeologists using R tend to be doing archaeology that involves the natural and physical sciences.
+Articles that cite R more often have titles that include terms such as 'lithic', 'experimental', and 'isotope'.
+On the other hand, articles that do not cite R more often discuss the Roman and Bronze Age periods, and less often reflect a technical specialisation, if there is one, it seems to ceramics.
 
-This plot supports our initial observations: archaeologists using R tend to be doing more scientific archaeology.
-We see that many words relating to stone artefacts, zooarchaeology, and geoarchaeology are more frequently used by R-using archaeologists.
-On the other hand, articles that do not cite R more often discuss the Neolithic and metal-age periods, and less often reflect a technical specialisation.
+## From using code in research to sharing code with publications
 
-## Slide: The unrealised potential of R for reproducible research [signpost slide only]
-After this overview how R is being used at the moment, we want to focus on the possibilities of making research reproducible using R.
+```{r archaeology-reproducible-articles}
 
-## Slide 9: The unrealised potential of R for reproducible research
-
-![](../figures/Marwick et al Fig 3 screenshot from the RStudio program.png)
-
-We would now like to focus on those archaeological papers using R, that are demonstrating the potential of R for open and reproducible research.
-These papers are using R Markdown for their research, which is shown in this figure.
-
-On the left we see a plain-text document written in R Markdown.
-This means that is has lines of code in amongst lines of normal text.
-We can format the text with simple additions, such as the underscore to indicate italics and the hash symbol to indicate a heading.
-These simple additions are Markdown, a very easy markup language with plain text formatting syntax.
-It is designed so that it can be converted to HTML and many other formats.
-**R** Markdown allows us to mix R code and text in the same document.
-Here we see a short snippet of R code right inside the text (nrow(cars)) as well as a longer block of code below.
-
-When this R Markdown document is rendered, or converted to another format such as PDF or Microsoft Word, we see that the code is replaced by the output it generates.
-On the right side of this figure we see that the inline code has been replaced by the number 50, and the code block below has been replaced by the plot.
-
-We have found that some archaeologists are writing their entire journal articles useing R Markdown, and they share their plain-text R Markdown document along with their publication.
-This means that the reader can see all of the details of the data analysis, and all of the small decisions made along the way.
-This way the analysis is easily reproducible for others.
-
-## Slide: Real-world example from Barton et al. 2018
-
-![](../figures/example-barton-et-al.png)
-
-Here is a real-world example R Markdown in use, from an archaeology article published in 2017.
-
-## Slide: An increasing number of archaeologists are publishing reproducible research
-
-```{r}
 archy_ctv_readme <- readLines(str_glue('{here::here()}/analysis/data/raw_data/ctv-archaeology-README.md'))
 
+journals <- c("Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science|PLOS|Journal of Human Evolution|Asian Perspectives|Scientific Reports|PLoS ONE|The Holocene|Archaeology in Oceania|Quaternary International|Internet Archaeology|Quaternary Science Reviews|Quaternary International|Journal of Archaeological Science: Reports|Open Quaternary|Evolution and Human Behavior|PaleoAnthropology")
+
 top_archy_journals <-
- str_subset(archy_ctv_readme, "Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science")
+ str_subset(archy_ctv_readme, journals)
 
 top_archy_journal_names <-
-  unlist(str_extract_all(archy_ctv_readme, "Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science"))
+  unlist(str_extract_all(archy_ctv_readme, journals))
 
@@
                                   "Journal of Archaeological Science",
-                                  top_archy_journal_names)
+                                  ifelse(top_archy_journal_names == "PLOS",
+                                  "PLoS ONE",
+                                  top_archy_journal_names))
 
@@
 
+data_to_plot <- data_frame(year = archy_ctv_readme_20XX,
+                  journal_name = top_archy_journal_names) %>% 
+  filter(!is.na(year), 
+         year <= 2017) %>% 
+  mutate(year = factor(year))
+```
+
+One notable practice that we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors make publicly available the code and data that they used to generate the results presented in their article.
+We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository.
+Figure \@ref(fig:archaeology-reproducible-articles) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.
+
+These archaeology articles that include publicly available R code and data are part of a new approach to scientific communication that has also been observed in other fields.
+Although we are not aware of attempts to quantify this change in the same way we have for archaeology, we have observed a number of recent publications that describe and recommend code sharing in statistics [@Baumer2014], genome biology [@markowetz2015five], computational biology [@Sandve2013], biostatistics [@Peng2009], computer science [@peng2011reproducible; @Mitchell2012], applied mathematics [@LeVeque2012], speech science [@Abari2012], political science [@king1995replication; Dafoe2014], and the social sciences generally [@Miguel2014].
+As part of this growing interest in using code for research we also see manifestos aimed at researchers doing any kind of quantitative work [@nosek2015promoting; @Ince2012; @Barnes2010] and articles recommending using R in undergraduate education [@Bray2014; @Eglen2009].
+Much of this literature is concerned with identifying and solving problems of irreproducibility in research.
+This comes from a renewed interest in basic principles of the science method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community.
+Furthermore, an article should sufficiently describe the presented research results such that a colleague can fully understand the results and how they were obtained.
+Many of these authors trace problems of irreproducibility to the increased complextity of computer-based analyses (but only a limited space to describe them in a journal article), and the use of mouse-driven computer programs where the researcher's analytical decisions are not recorded during the research process, and thus cannot be shared with other researchers.
+In our view, using R (or similar open source programming language) to conduct research might not constitute a tool-driven revolution, but publicaly sharing the code used for research, as a solution to problems of irreproducibility could enable revolutionary change, since openly shared code could speed the adoption of code throughout a research community.
+
+```{r archaeology-reproducible-articles-plot,  fig.cap=str_glue('Articles in archaeology journals\nusing R for reproducible research (n = {number_of_reproducible_articles})')}
+
 archaeology_articles_r_reproducible <- 
-ggplot(data_frame(archy_ctv_readme_20XX,
-                  top_archy_journal_names),
-       aes(archy_ctv_readme_20XX,
-           fill = str_wrap(top_archy_journal_names, 15))) +
-  geom_bar() +
-  theme_minimal(base_size = 16) +
-  scale_x_continuous(limits = c(min(archy_ctv_readme_20XX),
-                                max(archy_ctv_readme_20XX)+0.5),
-                     breaks = seq(min(archy_ctv_readme_20XX),
-                                  max(archy_ctv_readme_20XX)+0.5, 1)) +
-  ggtitle(str_glue('Articles in archaeology journals\nusing R for reproducible research (n = {length(archy_ctv_readme_20XX)})')) +
+ggplot(data_to_plot,
+       aes(year,
+           fill = journal_name)) +
+  geom_bar(position = "stack") +
+  scale_y_continuous(breaks = 1:10) +
   scale_fill_discrete(name = "Journal")  +
-  theme(legend.text=element_text(size=10),
-        legend.key.size = unit(2, 'lines'),
-        legend.position = "bottom"
-        ) +
   xlab("Year of publication") +
-  ylab("Number of articles")
+  ylab("Number of articles") +
+  theme_minimal(base_size = 18) +
+  theme(legend.text=element_text(size=16))
 
@@
        height = 9,
-       width = 9)
-
+       width = 16)
 knitr::include_graphics( str_glue('{here::here()}/analysis/figures/archaeology_articles_r_reproducible.png'))
@@
 
-In our sample of R citing archaeology articles, we identified `r number_of_reproducible_articles` articles that include R script- and data files, either in the supplementary materials or in a trustworthy data repository.
-This plot shows that although this number is small, it is increasing, suggesting the beginning of a new approach to how archaeologists share their research.
+A pragamtic issue relating to wide-spread changes in scientific practice is what the incentives are for members of the research community that might not be motivated by abstract basic principles of the science method.
+If code sharing has some concrete career advantages then we might expect it to be more likely to be part of a tool-drive revolution.
+@Vandewalle2012 analysed citation rates for 645 articles published in _IEEE Transactions on Image Processing_ during 2004–2006.
+He found statistically significant differences between the median number of cita?tions for papers that have code available and papers that do not have code available.
+In Figure \@ref(fig:citation-advantage) we conducted a similar anlysis for archaeology articles.
+We see that archaeological articles that cite R are consistently more highly cited themselves than articles that do not cite R.
 
-The way in which these `r number_of_reproducible_articles` articles make their work reproducible is by:
+```{r citation-advantage, fig.cap=str_glue('Median citation rates per year for archaeology articles that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")})')}
+# archy papers that use code are more often cited?
+# is TC total citations?
 
-- making the data files available with their publication
-- making the R script files available with the publication
-- using R Markdown to add narrative text with the R code to give explanation and make it easier for readers to understand the analysis
+archy_journal_articles_tally <- 
+archy_journal_articles %>% 
+  group_by(ID) %>% 
+  tally() %>% 
+  pull(n)
 
-Articles that have these three characteristics are part of a new paradigm in scientific communication, and with these `r number_of_reproducible_articles` articles, we believe we may have identified the beginning of this behaviour in archaeology.
-We think it will not be hard for archaeologists that already use R to make the transition to producing reproducible research by sharing their code and data.
-If this can be achieved, then we can realise some of the potential that R has for making archaeology more open and reproducible.
+archy_journal_articles %>% 
+  mutate(TC = as.numeric(TC),
+         ID = ifelse(ID == "has_R_cites",
+                     "Cites R", "Doesn't cite R")) %>% 
+  group_by(ID, PY) %>% 
+  summarise(median = median(TC)) %>% 
+  filter(PY >= 2010) %>% 
+  ggplot(aes(ID,
+             median)) +
+  geom_col() +
+  facet_wrap( ~ PY) +
+  theme_minimal() +
+  ylab("Median number of citations") +
+  xlab("")
+```
 
-## Slide: Our new package, rrtools, and how it helps
+# Discussion
 
-As we studied the papers in our sample, we looked for common approaches to making and organising the code and data files.
-We found that there were few common standards used by archaeologists in making their work reproducible.
-There was a great variety of information presented in the code and data files.
-Some papers include code that will generate every figure and number in the text, while other papers only include code to produce one or two key figures in the paper.
+The use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R.
+To make it easier for archaeologists and researchers in other fields to use R for research and publication, we describe the R package rrtools ('reproducible research tools').
+This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort.
+The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research.
+First, we keep the data separate from the methods.
+This means keep the code in a separate directory to the data.
+Second, we, keep the raw data seprate from data that is created or derived during the analysis.These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.
 
-Looking further afield, we also reviewed current efforts towards improving the reproducibility of research in other disciplines.
-In reviewing the literature we found several useful guides in diverse fields such as ecology and fluid dynamics.
-These guides described several best practices that are not yet common in archaeology.
+The rrtools package contains several functions to enable a researcher to quickly set up a compendium suitable for writing reproducible research in R.
+Each of these steps reflects a best practice that has been previously articulated in other feilds.
+We will focus here on the first five, as these are most relevant to most researchers:
 
-So, after looking at the reproducible articles in archaeology, and the best practices in other disciplines, we wanted to make a simple toolkit that will help archaeologists to write reproducible research as easily as possible.
-This toolkit is an R package called rrtools.
-This package contains a small number of functions that automate many common analysis setup tasks.
-
-![](../figures/rrtools-steps-carbon.png)
-
-On this slide we show the typical use of the key functions in rrtools that a researcher would use to quickly set up a compendium suitable for writing reproducible research in R.
-Each of these steps reflects a best practice that we've noted in our review of the literature.
-We will focus now on the first five:
-
-- **use_compendium:**  This function creates a new directory for the research project, and creates an R package within this directory.
+- **rrtools::use_compendium:**  This function creates a new directory for the research project, and creates an R package within this directory.
 We use the R package structure as the standard for our directory because it is widely recognised by R users.
@@
 This is important for setting the expectations about the relationship between the author and the reader, regarding the use of code.
-- **devtools::use_github:**  This will initiate the use of the Git version control system.
+- **usethis::use_github:**  This will initiate the use of the Git version control system.
 This is the current state-of-the-art for version control of any kind of plain text file, including text, code, data, and images.
@@
 The GitHub repository can be kept private if desired, until the work is published.
-- **use_readme_rmd:**  This will create a simple document that is usually the first thing a reader sees when they browse the code files.
+- **rrtools::use_readme_rmd:**  This will create a simple document that is usually the first thing a reader sees when they browse the code files.
 This document is important because it helps to describe to the reader what to expect in the compendium, and it gives details of how to cite the compendium.
-- **use_analysis:**  This function creates a set of folders according to the best practices for organising a typical reasearch project.
+- **rrtools::use_analysis:**  This function creates a set of folders according to the best practices for organising a typical reasearch project.
 We recognise that this file structure wont be perfect for every project.
@@
 
-## Slide: The best-practice directory structure from rrtools
+# Conclusion
 
-![](../figures/rrtools-analysis-dir.png)
-
-Here is the file structure the rrtools gives us. For now, we want to point out two important principles that it enforces:
-
-- Keep the data separate from the methods. This means keep the code in a separate directory to the data    
-- Keep the raw data seprate from data that is created or derived during the analysis.
-
-These two fundamental principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium.
-
-## Slide: Summary
-
-To summarise:
-
-- We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in the sciences generally.
-We saw that ecological and evolutionary sciences have really embraced R in their publications.
-
-- We have also shown that archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields.
+We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in ecology  We saw that ecological and evolutionary sciences have strongly embraced R in their publications.
+We have also shown that archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields.
 We've seen that the use of R is most common in more scientific aspects of archaeology.
-
-- We also identified a small sub-set of archaeology papers that use R to make their work fully reproducible.
+We also identified a small sub-set of archaeology papers that use R to make their work fully reproducible.
 There is a lot of potential for other archaeologists to make their work reproducible, and we think this would be a good thing for archaeology generally.
-
-- We have presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
+We have presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
 We hope this will help to realise some of the potential to make archaeological research more open and reproducible.
 
-Thank you.
+## Acknowledgements
 
----
-
-<!--
-
-These are spare slides that might help with Q&A, but not for the main talk
-
-## Slide: Advanced reproducibility with rrtools
-
-![](../figures/rrtools-steps-carbon.png)
-
-The last three functions that rrtools gives us are advanced topics that we summarise only very briefly here.
-
-- **use_dockerfile:** This function generates a configuration file that allows another user to recreate the computational environment of the original author.
-This is important for complex projects that have many pieces of software that might be difficult to install correctly.
-Sometimes when a piece of software is updates, its behaviour can change, which can alter the results of an analysis.
-This function helps to protect again this problem.
-- **use_travis:** This function will create a configuration file so that our compendium can be automatically checked each time we update it.
-This is especially useful when working with collaborators because it helps to identify problems in the analysis before they become too big.
-- **use_testthat:** This creates a template for adding tests for any custom functions used in the analysis.
-This is important so we can ensure that the functions perform as expected while we work on the anlysis.
-
-These three functions are very widely used in software engineering to give quality control to a package.
-With our rrtools package we have brought these tools into a research context where they can help researchers to make their analyses easy to work on and easy for others reproduce.
-
--->
+Our project comes from the group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
 

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
one sentence why R
typo.

@@
 What they share in common is working with the field-collected data on a computer to prepare it for publication.
+R is the most common open source and free ware data analysis tool available and therefore a good proxy for the adoption of open source scientific programming languages tools in a discipline.
 To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.
@@
 Figure \@ref(fig:ecology-citing-r-plot) shows the percentage of articles citing R in each of several of the top ecology journals (as defined by how often their articles are cited).
-We restrict the start of the observation period to 2007 for convivnence.
+We restrict the start of the observation period to 2007 for convenience.
 The plot shows that the percentage of articles citing R has increased from less than ten percent in all journals in the late 2000s, to more than 30% in _Ecosphere_, _Ecology and Evolution_, and _Molecular Ecology_ after 2012.
@@
 
-We conducted a similar bibliometric analysis for archaeology journals indexed by the Web of Science service.
+To compare with ecology we conducted a similar bibliometric analysis for archaeology journals indexed by the Web of Science service.
 We used the 'Cited Reference Search' to find articles citing R in the same way as above, and then refined the results to keep only those articles published during 2007-2017 that are included in the Web of Science category 'Archaeology'.

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
correct end of the sentence?

@@
 What they share in common is working with the field-collected data on a computer to prepare it for publication.
-R is the most common open source and free ware data analysis tool available and therefore a good proxy for the adoption of open source scientific programming languages tools in a discipline.
+R is the most common open source and free ware data analysis tool available and therefore a good proxy for the adoption of open source scientific programming languages in a discipline.
 To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.
@@
 This may indicate an increasing use of R among archaeologists, especially those working at the intersection of the natural and physical sciences.
-This may hint at the start of a trend similar to what we see in the ecology journals of a widespread
+This may hint at the start of a trend similar to what we see in the ecology journals of a widespread adoption of R.
 

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
typos

@@
 
-Following on from this observation about the increasing popularity of R in then _Journal of Archaeological Science_, the third detail we see in these results is the distinctive types of journals that have articles citing R.
-There is a clear focus on journals that publich scientific archaeology, especially those that focus on specialised empirical analysis.
+Following on from this observation about the increasing popularity of R in the _Journal of Archaeological Science_, the third detail we see in these results is the distinctive types of journals that have articles citing R.
+There is a clear focus on journals that publish scientific archaeology, especially those that focus on specialised empirical analysis.
 In addition to the general archaeology journals, we also see _Lithic Technology_, _Anthropozoologica_, and _International Journal of Osteology_ in the journal names, indicating that we find R used by specialists in stone artefacts and faunal remains.
-To further investigate the topics that archaeologists are useing R to study, we conducted a statistical analysis of the words in the titles of all the articles in our sample.
+To further investigate the topics that archaeologists are using R to study, we conducted a statistical analysis of the words in the titles of all the articles in our sample.
 
@@
 First we separated the archaeology articles into two groups, those that cite R, and those that do not.
-Second, we filter to keep only words that occur in titles in both groups, and remove stop words that are very common and uninformative (e.g.
+Second, we filtered to keep only words that occur in titles in both groups, and removed stop words that are very common and uninformative (e.g.
 'the', 'archaeological', 'study').
@@
 Articles that cite R more often have titles that include terms such as 'lithic', 'experimental', and 'isotope'.
-On the other hand, articles that do not cite R more often discuss the Roman and Bronze Age periods, and less often reflect a technical specialisation, if there is one, it seems to ceramics.
+On the other hand, articles that do not cite R more often discuss the Roman and Bronze Age periods, and less often reflect a technical specialisation, if there is one, it seems to be in ceramics.
 

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
added a sentence why sharing code is worth looking at
typo corrections

@@
 
+@Touchon_McCoy_2016 had identified the use of scripting languages as being a factor for easily sharing methods and code, thereby increasing the rate of development of statistical methods.
 One notable practice that we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors make publicly available the code and data that they used to generate the results presented in their article.
@@
 Much of this literature is concerned with identifying and solving problems of irreproducibility in research.
-This comes from a renewed interest in basic principles of the science method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community.
+This comes from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community.
 Furthermore, an article should sufficiently describe the presented research results such that a colleague can fully understand the results and how they were obtained.
 Many of these authors trace problems of irreproducibility to the increased complextity of computer-based analyses (but only a limited space to describe them in a journal article), and the use of mouse-driven computer programs where the researcher's analytical decisions are not recorded during the research process, and thus cannot be shared with other researchers.
-In our view, using R (or similar open source programming language) to conduct research might not constitute a tool-driven revolution, but publicaly sharing the code used for research, as a solution to problems of irreproducibility could enable revolutionary change, since openly shared code could speed the adoption of code throughout a research community.
+In our view, using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolution, but publicaly sharing the code used for research, as a solution to problems of irreproducibility could enable revolutionary change, since openly shared code could speed the adoption of code throughout a research community.
 

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
added two sentences about what truly would be revolutionary: no black box, evaluation of results, equality in access to research community

@@
 Many of these authors trace problems of irreproducibility to the increased complextity of computer-based analyses (but only a limited space to describe them in a journal article), and the use of mouse-driven computer programs where the researcher's analytical decisions are not recorded during the research process, and thus cannot be shared with other researchers.
-In our view, using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolution, but publicaly sharing the code used for research, as a solution to problems of irreproducibility could enable revolutionary change, since openly shared code could speed the adoption of code throughout a research community.
+In our view, using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolution, but publicaly sharing the code used for research, as a solution to problems of irreproducibility could enable revolutionary change: Openly shared code could speed the widespread adoption of the true scientific method, since it frees science from the black box, enabling researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and therefore truly evaluate the published results.
+As open source languages such as R (or Python...) are free as are the publishing platforms (such as zenodo, github or similar), they also enable not-affluent researchers world wide to contribute equally to the research community.
 
@@
 @Vandewalle2012 analysed citation rates for 645 articles published in _IEEE Transactions on Image Processing_ during 2004–2006.
-He found statistically significant differences between the median number of cita?tions for papers that have code available and papers that do not have code available.
+He found statistically significant differences between the median number of citations for papers that have code available and papers that do not have code available.
 In Figure \@ref(fig:citation-advantage) we conducted a similar anlysis for archaeology articles.

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
added to discussion what kind of revolution we are talking about

@@
 
-The use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R.
+The use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be cited more often.
+This also is an indicator for the increase in popularity of R.
+
+The use of R or any other scripting languages enables a paradigm-shift, as it offers the possibility to share one's code.
+The sharing of code has been identified in other disciplines as an amplifier of methodological change [@touchon_and_mccoy_2016].
+It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
+Thus, the sharing of code may be at the root of a tool-driven paradigm-change as desribed in the work of @Galison_1997.
+
+Speaking with Kuhn's view on paradigm-change, a change in the view point of researchers is necessary.
+We propose that a change of point of view of the archaeological researcher is needed: She or he, who works with computational methods, needs to view him/herself as a scientist, who needs to adhere to the scientific method, i.
+e.
+making the research reproducible for others.
+
+We acknowledge, this is not an easy task.
+Learning a scripting language is time-intensive and writing code so that one is comfortable in sharing it even more so.
+[some more info from your case study maybe?]
+
 To make it easier for archaeologists and researchers in other fields to use R for research and publication, we describe the R package rrtools ('reproducible research tools').

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
conclusion: added a few things.

@@
 
-The use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be cited more often.
+The use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often.
 This also is an indicator for the increase in popularity of R.
@@
 
+Using rrtools or its addin for Rstudio [cite Clemens here] enables the researcher to easily set up a digital work place that enables reproducible research.
+[do we name people here, who've already used it?]
+
 # Conclusion
 
+Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [citation needed], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@galison_1997], who focusees on the material culture of a science and its tools.
+A change in the paradigms of a discipline may also only occur, if tools to implement this change are available.
+The change we are aiming for is the move to a reproducible computational archaeology, that adheres to the scientific method.
+The tool which we have been discussing here is R, as an example for scientific scripting languages.
+
 We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in ecology  We saw that ecological and evolutionary sciences have strongly embraced R in their publications.
-We have also shown that archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields.
-We've seen that the use of R is most common in more scientific aspects of archaeology.
-We also identified a small sub-set of archaeology papers that use R to make their work fully reproducible.
-There is a lot of potential for other archaeologists to make their work reproducible, and we think this would be a good thing for archaeology generally.
-We have presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
-We hope this will help to realise some of the potential to make archaeological research more open and reproducible.
+Archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields, most commonly in more scientific aspects of archaeology.
+We also identified a small sub-set of archaeology papers that use R to make their work fully reproducible and could identify an advantage of citing R for researchers as it leads to being referenced more often.
+
+There is a lot of potential for other archaeologists to make their work reproducible, and we think this is a much needed paradigm-shift in computational archaeology.
+We therefore presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
+We hope this will help to realise some of the potential to make archaeological research more open and reproducible and in this way encourage researchers to adhere to the scientific method.
 

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
begin of elaborate discussion

@@
 
-The use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often.
+In light of the discussion in the philosophy of science, whether paradigm-shifts are tool- or idea-driven, we offer the hypothesis that a tool-driven change of paradigms is underway in the biological sciences and, more subtly, in archaeology.
+The tool which drives this change is the use of scripting languages in computational approaches of these disciplines.
+We could show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often.
 This also is an indicator for the increase in popularity of R.
@@
 It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
-Thus, the sharing of code may be at the root of a tool-driven paradigm-change as desribed in the work of @Galison_1997.
+Thus, the sharing of code may be at the root of a tool-driven paradigm-change as desribed in the work of @Galison [-@Galison_1997].
+This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@meltzer_1979, @trigger_2006]).
+Rather, this enables the archaologist to return to the tenet of the scientific method which demands the reproducibility of results.
+We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages.
+There is a longstanding tradition of publishing catalogues, lists and maps showing the data used for analyses.
+In the digital age, though, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses.
+This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results.
+Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which complicates the reproducibility of anlayses.
+As we could show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with biology (zooarchaeology) or focuses on scientific analyses.
 
-Speaking with Kuhn's view on paradigm-change, a change in the view point of researchers is necessary.
-We propose that a change of point of view of the archaeological researcher is needed: She or he, who works with computational methods, needs to view him/herself as a scientist, who needs to adhere to the scientific method, i.
+Speaking with Kuhn's view on paradigm-change, a change in the view point of researchers seems to be happening in these areas: She or he, who works with computational methods, needs to view him/herself as a scientist, who needs to adhere to the scientific method, i.
 e.

### 2018-09-09 analysis/paper/jcaa-paper.Rmd

@@
 This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results.
-Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which complicates the reproducibility of anlayses.
+Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which complicates the reproducibility of analyses.
 As we could show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with biology (zooarchaeology) or focuses on scientific analyses.
 
-Speaking with Kuhn's view on paradigm-change, a change in the view point of researchers seems to be happening in these areas: She or he, who works with computational methods, needs to view him/herself as a scientist, who needs to adhere to the scientific method, i.
-e.
-making the research reproducible for others.
+We do not believe, that speaking with Kuhn's view on paradigm-change, a change in the view point of researchers seems to be happening: Data-and method sharing behaviour has been around in archaeology before.
+What we can see, though, is that the need for methods to share research in a way that is reproducible for others has been acknowledged and that archaeologists have looked for tools to make this possible.
+Using  scripting languages and data repositories is a solution, that enjoys increasing popularity.
 
-We acknowledge, this is not an easy task.
-Learning a scripting language is time-intensive and writing code so that one is comfortable in sharing it even more so.
+We acknowledge, this is not an easy task. Learning a scripting language is time-intensive and writing code so that one is comfortable in sharing it even more so.
+
 [some more info from your case study maybe?]
@@
 
-There is a lot of potential for other archaeologists to make their work reproducible, and we think this is a much needed paradigm-shift in computational archaeology.
+There is a lot of potential for other archaeologists to make their work reproducible, and we think this is a much needed paradigm-shift in computational archaeology - linking the digital and computational times to  older practices of sharing data and methods.
 We therefore presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
-We hope this will help to realise some of the potential to make archaeological research more open and reproducible and in this way encourage researchers to adhere to the scientific method.
+We hope this will help to realise some of the potential to make computational archaeological research more open and reproducible and in this way encourage researchers to adhere to the scientific method.
 

### 2018-09-10 analysis/paper/jcaa-paper.Rmd

@@
 In this paper we investigate recent changes in the way computational archaeology is being done.
-Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discurses[@clark_1993; @härke_2002], but there has been
-an argument in philosophy of science whether revolutions in science are either ideas-driven or tool-driven [@galison_1997; @dyson_2000].
+Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discurses[@clark_1993; @härke_2002], but there has been an argument in philosophy of science whether revolutions in science are either ideas-driven or tool-driven [@galison_1997; @dyson_2000].
 In this light we explore how many scientific disciplines have started to embrace methods to improve the reproducibility of their research and whether this change is part of a paradigm-shift.
@@
 # Bibliometric analysis of tool-driven change
-The question of predicting or generating change in scientific practice via adoption of new tools is especially intruiging, and in this section we explore how a change in ecology might be relevant for understanding the future of archaeology.
+
+The question of predicting or generating change in scientific practice via adoption of new tools is especially intruiging, and in this section we explore how a change that has been observed in ecology might be relevant for understanding or even directing the future of archaeology.
 
@@
 
-We computed a comparison of word frequences in journal article titles to get a better understanding of what topics archaeologists are writing about when they cite R.
+We computed a comparison of word frequences in journal article titles in the Web of Science data to get a better understanding of what topics archaeologists are writing about when they cite R.
 First we separated the archaeology articles into two groups, those that cite R, and those that do not.
@@
 We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository.
+Our contiunuously updated list of articles in this sample is available online here: https://github.com/benmarwick/ctv-archaeology.
 Figure \@ref(fig:archaeology-reproducible-articles) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.
@@
 Many of these authors trace problems of irreproducibility to the increased complextity of computer-based analyses (but only a limited space to describe them in a journal article), and the use of mouse-driven computer programs where the researcher's analytical decisions are not recorded during the research process, and thus cannot be shared with other researchers.
-In our view, using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolution, but publicaly sharing the code used for research, as a solution to problems of irreproducibility could enable revolutionary change: Openly shared code could speed the widespread adoption of the true scientific method, since it frees science from the black box, enabling researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and therefore truly evaluate the published results.
-As open source languages such as R (or Python...) are free as are the publishing platforms (such as zenodo, github or similar), they also enable not-affluent researchers world wide to contribute equally to the research community.
+
+In our view, merely using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolutionb by itself, but publicaly sharing the code used for research, as a solution to problems of irreproducibility, could enable revolutionary change.
+Openly shared code could speed the widespread adoption of the reproducibility are a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and therefore properly evaluate, extend and reuse the published results.
+As open source languages such as R (or Python...) are free to use, and most trustworthy repositories are also free (such as zenodo, osf, figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.
 
@@
 
-A pragamtic issue relating to wide-spread changes in scientific practice is what the incentives are for members of the research community that might not be motivated by abstract basic principles of the science method.
+A pragamtic issue relating to wide-spread changes in scientific practice is what the incentives are for members of the research community.
+This is especially a concern for researchers that might not be motivated by a desire to adhere to the abstract basic principles of what it means to practice science with integrity.
 If code sharing has some concrete career advantages then we might expect it to be more likely to be part of a tool-drive revolution.
@@
 He found statistically significant differences between the median number of citations for papers that have code available and papers that do not have code available.
-In Figure \@ref(fig:citation-advantage) we conducted a similar anlysis for archaeology articles.
+In Figure \@ref(fig:citation-advantage) we conducted a similar anlysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
+TC stands for 'Web of Science Core Collection Times Cited Count'.
+This Times Cited count displays the total number of times a published paper was cited by other papers within the Web of Science Core Collection.
+Unlike Google Scholar, this does not include citations on blogs, pre-prints and other places outside of journal articles, so the citation counts from Web of Science are lower than those on Google Scholar.
 We see that archaeological articles that cite R are consistently more highly cited themselves than articles that do not cite R.
@@
 The tool which drives this change is the use of scripting languages in computational approaches of these disciplines.
-We could show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often.
+We  show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visiblity and influence.
 This also is an indicator for the increase in popularity of R.
 
-The use of R or any other scripting languages enables a paradigm-shift, as it offers the possibility to share one's code.
-The sharing of code has been identified in other disciplines as an amplifier of methodological change [@touchon_and_mccoy_2016].
+The use of R or any other scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to share one's code.
+The sharing of code has been identified in other disciplines as an amplifier of change in methods[@touchon_and_mccoy_2016].
 It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
-Thus, the sharing of code may be at the root of a tool-driven paradigm-change as desribed in the work of @Galison [-@Galison_1997].
+Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with what we see desribed in the work of @Galison [-@Galison_1997].
 This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@meltzer_1979, @trigger_2006]).
-Rather, this enables the archaologist to return to the tenet of the scientific method which demands the reproducibility of results.
+
+Rather, this change enables archaologists to more faithfully adhere to the fundamental tenet of the scientific process which demands the reproducibility of results.
 We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages.
-There is a longstanding tradition of publishing catalogues, lists and maps showing the data used for analyses.
+But instead that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice.
+By using and sharing code we can undo this drift, and return to a robust scrietific practice for archaeology.
+
+There is a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses.
 In the digital age, though, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses.
 This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results.
-Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which complicates the reproducibility of analyses.
-As we could show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with biology (zooarchaeology) or focuses on scientific analyses.
+Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses.
+As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences.
 
-We do not believe, that speaking with Kuhn's view on paradigm-change, a change in the view point of researchers seems to be happening: Data-and method sharing behaviour has been around in archaeology before.
+We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before and do not present incommensurate ways of thinking about the subject.
 What we can see, though, is that the need for methods to share research in a way that is reproducible for others has been acknowledged and that archaeologists have looked for tools to make this possible.
-Using  scripting languages and data repositories is a solution, that enjoys increasing popularity.
+Using scripting languages and data repositories to share the scripts is a solution that we have demonstrated enjoys increasing popularity in the archaeological research community.
 
-We acknowledge, this is not an easy task. Learning a scripting language is time-intensive and writing code so that one is comfortable in sharing it even more so.
+## Making it easier
 
-[some more info from your case study maybe?]
+We acknowledge that learning to use code, if one has not recieved formal instruction, is not an easy task.
+Similarly, sharing code in public takes a little time to ensure that it is fit for others to read.
+Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error, rather than following a model or widely-used approach.
+This suggests to us that there is some scope to improve the ease and efficiency of code-use in archaeology by studying the exsiting practice and providing a model that others can easily adapt to their own research projects.
 
-To make it easier for archaeologists and researchers in other fields to use R for research and publication, we describe the R package rrtools ('reproducible research tools').
+The result of our analysis of existing practices among archaeologists and other researchers, and our study of best practices in scientific computing (described in more detail in @marwick2018packaging) is the R package rrtools ('reproducible research tools', https://github.com/benmarwick/rrtools).
+The goal of rrtools ist o make it easier for archaeologists and researchers in other fields to use R for research and publication.
 This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort.
@@
 
-The rrtools package contains several functions to enable a researcher to quickly set up a compendium suitable for writing reproducible research in R.
+The rrtools package contains several functions for the R user to enable a researcher to quickly set up a compendium suitable for writing reproducible research in R.
 Each of these steps reflects a best practice that has been previously articulated in other feilds.
@@
 
-Using rrtools or its addin for Rstudio [cite Clemens here] enables the researcher to easily set up a digital work place that enables reproducible research.
-[do we name people here, who've already used it?]
-
 # Conclusion
 
-Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [citation needed], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@galison_1997], who focusees on the material culture of a science and its tools.
+Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [citation needed], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@galison_1997], who focuses on the material culture of a science and its tools.
 A change in the paradigms of a discipline may also only occur, if tools to implement this change are available.
@@
 
-We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in ecology  We saw that ecological and evolutionary sciences have strongly embraced R in their publications.
+We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in ecology.
+We saw that ecological and evolutionary sciences have strongly embraced R in their publications.
 Archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields, most commonly in more scientific aspects of archaeology.
@@
 
-There is a lot of potential for other archaeologists to make their work reproducible, and we think this is a much needed paradigm-shift in computational archaeology - linking the digital and computational times to  older practices of sharing data and methods.
+There is great potential for archaeologists to make their work reproducible using R and related tools, and we think this is a much needed shift in computational archaeology - linking the digital and computational times to  older practices of sharing data and methods.
 We therefore presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
-We hope this will help to realise some of the potential to make computational archaeological research more open and reproducible and in this way encourage researchers to adhere to the scientific method.
+We hope this will help to realise some of the potential to make computational archaeological research more open and reproducible speed the transfer of new results and methods throughout the research community without barriers due to resource availabilty, and in this way encourage researchers to adhere to the scientific process.
 

### 2018-09-10 analysis/paper/jcaa-paper.Rmd

@@
 In this paper we investigate recent changes in the way computational archaeology is being done.
-Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discurses[@clark_1993; @härke_2002], but there has been an argument in philosophy of science whether revolutions in science are either ideas-driven or tool-driven [@galison_1997; @dyson_2000].
+Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discurses[@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science are either ideas-driven or tool-driven [@Galison_1997; @Dyson_2000].
 In this light we explore how many scientific disciplines have started to embrace methods to improve the reproducibility of their research and whether this change is part of a paradigm-shift.
@@
 The use of R or any other scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to share one's code.
-The sharing of code has been identified in other disciplines as an amplifier of change in methods[@touchon_and_mccoy_2016].
+The sharing of code has been identified in other disciplines as an amplifier of change in methods[@Touchon_McCoy_2016].
 It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
-Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with what we see desribed in the work of @Galison [-@Galison_1997].
-This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@meltzer_1979, @trigger_2006]).
+Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with what we see desribed in the work of @Galison_1997 [-@Galison_1997].
+This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@Meltzer_1979, @Trigger_2006]).
 
@@
 
-Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [citation needed], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@galison_1997], who focuses on the material culture of a science and its tools.
+Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [citation needed], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@Galison_1997], who focuses on the material culture of a science and its tools.
 A change in the paradigms of a discipline may also only occur, if tools to implement this change are available.

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
add Stodden et al - reference
typo

@@
 As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences.
+These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013].
 
@@
 The result of our analysis of existing practices among archaeologists and other researchers, and our study of best practices in scientific computing (described in more detail in @marwick2018packaging) is the R package rrtools ('reproducible research tools', https://github.com/benmarwick/rrtools).
-The goal of rrtools ist o make it easier for archaeologists and researchers in other fields to use R for research and publication.
+The goal of rrtools ist to make it easier for archaeologists and researchers in other fields to use R for research and publication.
 This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort.

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
added citation of Hugget 2015 + Hugget and Huvalis 2018 + Chrysanthi, Papadopolous, Murietta-Flores 2012.

@@
 The changes are not purely at the points of data collection and analysis, but extend to the social and economic organisation of science, with new categories of physicists emerging, people who are not quite experimenters and not quite theorists.
+(Add footnote: A similar question, if a bit more theoretical, has been raised by @Huvila_2018 for archaeologists, discussing what questions on archaeological practices and knowledge work should be answered to avoid a technologically determisitic approach to computational archaeology.)
 Dyson [-@Dyson_2000; -@dyson2012science] has further explored this tool-driven approach, linking the Galilean revolution to the use of the telescope in astronomy, and the Crick-Watson revolution to the use of X-ray diffraction.
@@
 We acknowledge that learning to use code, if one has not recieved formal instruction, is not an easy task.
-Similarly, sharing code in public takes a little time to ensure that it is fit for others to read.
+As @Chrysanthi_2012 illustrate, it requires a certain familiarity with the computational tools to be able to use them not just on a technological basis, but creatively and in full knowledge of their restrictions and ambiguities.
+The awareness and the appropriate handling of shortcomings and contexts of the data used in an analysis is also a time- and thought-consuming effort, which needs to stand at the beginning of each archaeological research (and which may be exacerbated by using multiple data sets from varying sources -- see e.
+g.
+@Hugget_2015).
+We will not focus on data sharing here, though, but on the sharing of code.
+
+As with data, sharing code in public takes a little time to ensure that it is fit for others to read.
 Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error, rather than following a model or widely-used approach.

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
I may not like docx, but I think its probably a better format to submit.

@@
 output: 
-    bookdown::html_document2:
-      number_sections: false
+    bookdown::word_document2:
       fig_caption: yes

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
resorting Huvila_2018
maybe this is a better place?

--- 2018-09-10 analysis/paper/jcaa-paper.Rmd
@@
 The changes are not purely at the points of data collection and analysis, but extend to the social and economic organisation of science, with new categories of physicists emerging, people who are not quite experimenters and not quite theorists.
-(Add footnote: A similar question, if a bit more theoretical, has been raised by @Huvila_2018 for archaeologists, discussing what questions on archaeological practices and knowledge work should be answered to avoid a technologically determisitic approach to computational archaeology.)
 Dyson [-@Dyson_2000; -@dyson2012science] has further explored this tool-driven approach, linking the Galilean revolution to the use of the telescope in astronomy, and the Crick-Watson revolution to the use of X-ray diffraction.
@@
 
+As @Galison_1997 describes a shift in the practices of physicists, so does archaeological practice change in the digital age.
+What exactly this entails, is diverse and has been discussed more thoroughly by @Huvila_2018 who ventilate what questions on archaeological practices and knowledge work should be answered to avoid a technologically determisitic approach to computational archaeology.
+
 We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before and do not present incommensurate ways of thinking about the subject.

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
added Stupler_2017 to references
reordered part of discussion so we don't jump from data to code back to data.

--- 2018-09-10 analysis/paper/jcaa-paper.Rmd
@@
 We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages.
-But instead that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice.
-By using and sharing code we can undo this drift, and return to a robust scrietific practice for archaeology.
+There is, for example, a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses.
+Archaeological research has always relied on the reuse of data [@Strupler_2017].
 
-There is a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses.
-In the digital age, though, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses.
+Nonetheless we feel that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice.
+In the digital age, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses.
 This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results.
 Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses.
+By using and sharing code we can undo this drift, and return to a robust scrietific practice for archaeology.
 As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences.
 These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013].
+Also in archaeology, more and more funding agencies push for open science methods and transparency [@Stupler_2017].
 

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
yeah. right. let's simply "talk about"

@@
 As @Galison_1997 describes a shift in the practices of physicists, so does archaeological practice change in the digital age.
-What exactly this entails, is diverse and has been discussed more thoroughly by @Huvila_2018 who ventilate what questions on archaeological practices and knowledge work should be answered to avoid a technologically determisitic approach to computational archaeology.
+What exactly this entails, is diverse and has been discussed more thoroughly by @Huvila_2018 who talk about what questions on archaeological practices and knowledge work should be answered to avoid a technologically determisitic approach to computational archaeology.
 

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
rewrote intro to be a bit more precise in what we do in the paper

@@
 
-In this paper we investigate recent changes in the way computational archaeology is being done.
-Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discurses[@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science are either ideas-driven or tool-driven [@Galison_1997; @Dyson_2000].
-In this light we explore how many scientific disciplines have started to embrace methods to improve the reproducibility of their research and whether this change is part of a paradigm-shift.
-We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages and review bibliometric evidence for this change in the sciences generally and in archaeology in particular.
-We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and test if the adoption of these new tools may be considered a revolution or some other process.
+In this paper we investigate recent developments which facilitate computational archaeology to return to principles of the scientific method and whether this might be the beginning of a paradigm-shift.
+Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discourses [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might not just be ideas-driven but also tool-driven [@Galison_1997; @Dyson_2000].
+In this light we explore how Ecology and Archaeology have started to embrace methods which improve the reproducibility of their research.
+We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code.
+A review of bibliometric evidence showcases a strong development towards the use of R for Ecology and hints at the beginning of a similar development in archaeology.
+We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven paradigm-change in archaeology.
+Acknowledging this is not a simple process, we offer a R-based tool to facilitates the task of creating a compendium which enables other researchers to reproduce the published results.
 
-[need some more in here]
+Code and text for this paper are available at: https://github.com/benmarwick/caa2018-Tubingen
 
@@
 
-Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [citation needed], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@Galison_1997], who focuses on the material culture of a science and its tools.
+Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [@Kuhn_1962], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@Galison_1997], who focuses on the material culture of a science and its tools.
 A change in the paradigms of a discipline may also only occur, if tools to implement this change are available.

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
typos

@@
 We saw that ecological and evolutionary sciences have strongly embraced R in their publications.
-Archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields, most commonly in more scientific aspects of archaeology.
+Archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields, and most commonly in more scientific aspects of archaeology.
 We also identified a small sub-set of archaeology papers that use R to make their work fully reproducible and could identify an advantage of citing R for researchers as it leads to being referenced more often.
@@
 We therefore presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
-We hope this will help to realise some of the potential to make computational archaeological research more open and reproducible speed the transfer of new results and methods throughout the research community without barriers due to resource availabilty, and in this way encourage researchers to adhere to the scientific process.
+We hope this will help to realise some of the potential to make computational archaeological research more open and reproducible, speed the transfer of new results and methods throughout the research community without barriers due to resource availabilty, and also in this way encourage researchers to adhere to the scientific process.
 

### 2018-09-11 analysis/paper/jcaa-paper.Rmd
corrected citation
paradigm to just change
Yeah, we should add some commentary in the end why exactly we think so.

@@
 A review of bibliometric evidence showcases a strong development towards the use of R for Ecology and hints at the beginning of a similar development in archaeology.
-We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven paradigm-change in archaeology.
+We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven change in archaeology.
 Acknowledging this is not a simple process, we offer a R-based tool to facilitates the task of creating a compendium which enables other researchers to reproduce the published results.
@@
 These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013].
-Also in archaeology, more and more funding agencies push for open science methods and transparency [@Stupler_2017].
+Also in archaeology, more and more funding agencies push for open science methods and transparency [@Strupler_2017].
 

### 2018-11-16 analysis/paper/jcaa-paper.Rmd
add something on gieryn

@@
 As @Galison_1997 describes a shift in the practices of physicists, so does archaeological practice change in the digital age.
-What exactly this entails, is diverse and has been discussed more thoroughly by @Huvila_2018 who talk about what questions on archaeological practices and knowledge work should be answered to avoid a technologically determisitic approach to computational archaeology.
+@Huvila_2018 have explored the use of digital tools on archaeological practices to understand the boundaries of what counts as archaeological, what is  merely related to archaeology and what should be excluded from archaeology.
+Reflecting on @Huvila_2018, our view is the use use of code is unquestionably archaeology, since it is just one of many tools that archaeologists are using for converting raw data into summaries, visualisations, and insights.
+
+These concerns about how digital tools affect what we consider is and isn't archaeology intersect with the work of @gieryn1999cultural on boundary construction.
+In his cultural cartography of science, Gieryn examines boundaries between science generally and varieties of less authoritative non-science.
+He argues that the boundaries are drawn and redrawn contiuously in inflexible, historically changing and sometimes ambiguous ways.
+Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims by competitors.
+Boundary work involves attributions of selected characteristics to science in order to distinguish it from non-scientific competitors.
+The increasing voluntary use and sharing of code by researchers, and demands for it by readers and peer reviewers is a form of active boundary work that is shifting the boundary between scientific and non- or less-scientific archaeology.
+We argue that the use of code as a research tool, and sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work.
+The reason is because using code, and making available for others to inspect, enhancing the credibility of knowledge claims.
 

### 2018-11-16 analysis/paper/jcaa-paper.Rmd
update the paper for deadline

@@
 
-Figure \@ref(fig:ecology-citing-r-plot) shows the percentage of articles citing R in each of several of the top ecology journals (as defined by how often their articles are cited).
+Figure \@ref(fig:ecology-articles-citing-r-plot) shows the percentage of articles citing R in each of several of the top ecology journals (as defined by how often their articles are cited).
 We restrict the start of the observation period to 2007 for convenience.
@@
 This resulted in `r how_many_articles_in_top_archy_journals` articles, of which `r total_number_archaeology_articles_citing_r` cite R.
-Figure \@ref(fig:rchaeology-journals-with-subplot) shows the temporal trend of citations of R in archaeological articles.
+Figure \@ref(fig:archaeology-journals-with-subplot) shows the temporal trend of citations of R in archaeological articles.
 There are three interesting details revealed by this figure.
@@
 
-The first detail we see here is that overall, the proportions and the absolute number of articles citing R in archaeology are much smaller than what we see in the ecology journals in Figure \@ref(fig:ecology-citing-r).
+The first detail we see here is that overall, the proportions and the absolute number of articles citing R in archaeology are much smaller than what we see in the ecology journals in Figure \@ref(fig:ecology-articles-citing-r-plot).
 Only the _Journal of Archaeological Science_ has more than 50 articles in our sample.
@@
 The sample sizes for the archaeology journals are too small to confidently infer any trend of over time, with the exception of the _Journal of Archaeological Science_ with `r how_many_articles_in_jas` articles.
-The inset plot in Figure \@ref(fig:rchaeology-journals-with-subplot) shows a statistical test of change over time for articles in the _Journal of Archaeological Science_.
+The inset plot in Figure \@ref(fig:archaeology-journals-with-subplot) shows a statistical test of change over time for articles in the _Journal of Archaeological Science_.
 With a moderate r-squared value and low p-value for the linear model, we can conclude that there is some evidence for an increase in citations of R over time in this journal.
@@
 Third, for each word found in all the titles of the journal articles in each group, we computed its proportion of the total number of words in all the titles in each group.
-Figure \@ref(fig:text-mining-article-titles-archaeology) shows the results of this analysis.
+Figure \@ref(fig:text-mining-article-titles-archaeology-plot) shows the results of this analysis.
 Words near the red line are used with about equal frequencies by papers that cite R and by papers that do not cite R.
@@
 
-```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles \nof archaeology articles ({prettyNum(nrow(frequency), ',')} words in {prettyNum(nrow(archy_journal_articles), ',')} articles)")}
+```{r text-mining-article-titles-archaeology-plot, fig.cap=paste0("Comparing the frequency of words used in titles \nof archaeology articles (", prettyNum(nrow(frequency)), " words in ", prettyNum(nrow(archy_journal_articles)), " articles)")}
 
@@
 Our contiunuously updated list of articles in this sample is available online here: https://github.com/benmarwick/ctv-archaeology.
-Figure \@ref(fig:archaeology-reproducible-articles) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.
+Figure \@ref(fig:archaeology-reproducible-articles-plot) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.
 
@@
 Openly shared code could speed the widespread adoption of the reproducibility are a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and therefore properly evaluate, extend and reuse the published results.
-As open source languages such as R (or Python...) are free to use, and most trustworthy repositories are also free (such as zenodo, osf, figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.
+As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as zenodo, osf, figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.
 
-```{r archaeology-reproducible-articles-plot,  fig.cap=str_glue('Articles in archaeology journals\nusing R for reproducible research (n = {number_of_reproducible_articles})')}
+```{r archaeology-reproducible-articles-plot, fig.cap=paste0('Articles in archaeology journals\nusing R for reproducible research (n = ', number_of_reproducible_articles,')')}
 
@@
 The use of R or any other scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to share one's code.
-The sharing of code has been identified in other disciplines as an amplifier of change in methods[@Touchon_McCoy_2016].
+The sharing of code has been identified in other disciplines as an amplifier of change in methods [@Touchon_McCoy_2016].
 It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
 Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with what we see desribed in the work of @Galison_1997 [-@Galison_1997].
-This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@Meltzer_1979, @Trigger_2006]).
+This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@Meltzer_1979; @Trigger_2006]).
 
@@
 
-We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before and do not present incommensurate ways of thinking about the subject.
+We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before now and do not present incommensurate ways of thinking about the subject in the way that a classical paradigm-change should.
 What we can see, though, is that the need for methods to share research in a way that is reproducible for others has been acknowledged and that archaeologists have looked for tools to make this possible.
 Using scripting languages and data repositories to share the scripts is a solution that we have demonstrated enjoys increasing popularity in the archaeological research community.
+It has also expanded the 'trading zone' [@Galison_1997] where archaeologists collaborate with specialists in other fields to develop a shared language to get things done.
+In using scripting languages archaeologists are working at a trading zone with computer scientists and computational reserchers in other fields to coordinate how tools from one domain can be useful in another.
+We can also see archaeologists in similar trading zones when using GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions.
+Archaeologists' embrace of, or turn towards, these methods could also be argued to be tool-driven revolutions.
+However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is universal method with which we can do all, or be central to, of those previously-mentioned approaches.
+That is why we believe archaeologists should pay attention to scripted workflows regardless of their usual toolkit and methodological interests.
 
@@
 The tool which we have been discussing here is R, as an example for scientific scripting languages.
-
 We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in ecology.

### 2018-11-16 analysis/paper/jcaa-paper.Rmd
"inflexible" -> flexible was meant. I think.

@@
 We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven change in archaeology.
-Acknowledging this is not a simple process, we offer a R-based tool to facilitates the task of creating a compendium which enables other researchers to reproduce the published results.
+Acknowledging this is not a simple process, we offer a R-based tool to facilitate the task of creating a compendium which enables other researchers to reproduce the published results.
 
@@
 @Schiffer_2013].
-Furthermore, Galison's analysis of the image/logic contrast in the history of physics invites a similar analysis of the history of archaeology Can we identify archaeological traditions focused on data collection and analysis using images and image-making devices [cf.
+Furthermore, Galison's analysis of the image/logic contrast in the history of physics invites a similar analysis of the history of archaeology.
+Can we identify archaeological traditions focused on data collection and analysis using images and image-making devices [cf.
 @molyneaux2013cultural], in contrast to digital devices?
@@
 
-The second detail is that most archaeology journals do not show any increase in the percentage of article citing R over time .
+The second detail is that most archaeology journals do not show any increase in the percentage of article citing R over time.
 The ecology journals shown in Figure \@ref(fig:ecology-citing-r) show clear upward trends of increasing proportions of articles citing R over time, but we do not see any clear trends in the archaeology articles considered together.
@@
 These archaeology articles that include publicly available R code and data are part of a new approach to scientific communication that has also been observed in other fields.
-Although we are not aware of attempts to quantify this change in the same way we have for archaeology, we have observed a number of recent publications that describe and recommend code sharing in statistics [@Baumer2014], genome biology [@markowetz2015five], computational biology [@Sandve2013], biostatistics [@Peng2009], computer science [@peng2011reproducible; @Mitchell2012], applied mathematics [@LeVeque2012], speech science [@Abari2012], political science [@king1995replication; Dafoe2014], and the social sciences generally [@Miguel2014].
+Although we are not aware of attempts to quantify this change in the same way we have for archaeology, we have observed a number of recent publications that describe and recommend code sharing in statistics [@Baumer2014], genome biology [@markowetz2015five], computational biology [@Sandve2013], biostatistics [@Peng2009], computer science [@peng2011reproducible; @Mitchell2012], applied mathematics [@LeVeque2012], speech science [@Abari2012], political science [@king1995replication; @Dafoe2014], and the social sciences generally [@Miguel2014].
 As part of this growing interest in using code for research we also see manifestos aimed at researchers doing any kind of quantitative work [@nosek2015promoting; @Ince2012; @Barnes2010] and articles recommending using R in undergraduate education [@Bray2014; @Eglen2009].
@@
 
-A pragamtic issue relating to wide-spread changes in scientific practice is what the incentives are for members of the research community.
+A pragmatic issue relating to wide-spread changes in scientific practice is what the incentives are for members of the research community.
 This is especially a concern for researchers that might not be motivated by a desire to adhere to the abstract basic principles of what it means to practice science with integrity.

### 2018-11-16 analysis/paper/jcaa-paper.Rmd
some grammar knots cleared (I hope)

@@
 In his cultural cartography of science, Gieryn examines boundaries between science generally and varieties of less authoritative non-science.
-He argues that the boundaries are drawn and redrawn contiuously in inflexible, historically changing and sometimes ambiguous ways.
+He argues that the boundaries are drawn and redrawn contiuously in flexible, historically changing and sometimes ambiguous ways.
 Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims by competitors.
@@
 The increasing voluntary use and sharing of code by researchers, and demands for it by readers and peer reviewers is a form of active boundary work that is shifting the boundary between scientific and non- or less-scientific archaeology.
-We argue that the use of code as a research tool, and sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work.
-The reason is because using code, and making available for others to inspect, enhancing the credibility of knowledge claims.
+The authors, as members of this group, argue that the use of code as a research tool, and the sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work.
+The reason for this is that using code, and making it available for others to inspect, enhances the credibility of knowledge claims.
 
-We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before now and do not present incommensurate ways of thinking about the subject in the way that a classical paradigm-change should.
+We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before now and does not present incommensurate ways of thinking about the subject in the way that a classical paradigm-change should.
 What we can see, though, is that the need for methods to share research in a way that is reproducible for others has been acknowledged and that archaeologists have looked for tools to make this possible.
@@
 Archaeologists' embrace of, or turn towards, these methods could also be argued to be tool-driven revolutions.
-However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is universal method with which we can do all, or be central to, of those previously-mentioned approaches.
+However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is an universal method with which we can do all, or which is central to any, of those previously-mentioned approaches.
 That is why we believe archaeologists should pay attention to scripted workflows regardless of their usual toolkit and methodological interests.

### 2018-11-16 analysis/paper/jcaa-paper.Rmd
gave reason why ecology is a good comparison for archaeology and added citation.

@@
 The question of predicting or generating change in scientific practice via adoption of new tools is especially intruiging, and in this section we explore how a change that has been observed in ecology might be relevant for understanding or even directing the future of archaeology.
+Biological disciplines have a long tradition of influencing archaeological thought, starting with Oscar Montelius, whose typological method was inspired by Darwin's theory of evolution [@Montelius_1899, 237].
+Also nowadays quite a number of spatial statistical algorithms used in archaeology are derived from Ecology [@keron_use_2015, 7].
 

### 2018-11-20 analysis/paper/jcaa-paper.Rmd

@@
 
+On another level, we'd like to point to the _technicity_ of tools, in our case coding.
+Aud Sissel Hoel and Iris van der Tuin are drawing on Ernst Cassirer's „Form und Technik“ [-@cassirer_form_1930] and Gilbert Simondon's „Du monde d'existence des objets techniques" [-@simondon_mode_1958] to define technicity as a term to describe that technology, i.
+e.
+tools and machines respectively, interact with the human and with nature [@hoel_ontological_2013, 188].
+Technology is said to own an instrumental kind of logos (Cassierer, see @hoel_ontological_2013, 194) and in analogy to language it exercises a certain measure of agency in mediating between human, nature and technology.
+The use of a scripting language might exemplify this quite nicely, as it is not just a technological tool, but quite rightly named "language" as well.
+
+For our paper the concept of technicity is interesting because in contrast to the concept of technological determinism, which puts technology as a force against the human, van der Tuin and Hoel (as well as Cassirer and Simondon) enfold technology into „the human“ without letting it lose its "foreignness" and further develop the notions of „human“ and „nature“ [@hoel_ontological_2013, 190].
+Any change in tools therefore will influence how human researchers interact with their objects of interest.
+Interestingly it is exactly at the point inbetween continously evolving subjects and objects, where van der Tuin and Hoel place technology and the creation of knowledge [@hoel_ontological_2013, 199].
+The "productive back-and-forth" between all parties may challenge established notions of culture, but also offers a concept of formation through differing, called "performative correspondence" [@hoel_ontological_2013, 201].
+
+Another point they make is, that the potential of a certain technology cannot be known before its coming-into-being as it will reflect back on its creation ("concretization" in Simondon's terms) as well as interact with other technologies (developed or yet-to-be developed) [@hoel_ontological_2013, 194].
+Seeing the influence the introduction of scientific scripting languages had on ecology we therefore can only expect similar trends to evolve in archaeology, but not predict these.
+
 We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before now and does not present incommensurate ways of thinking about the subject in the way that a classical paradigm-change should.

### 2018-11-21 analysis/paper/jcaa-paper.Rmd
typos

@@
 
-# Bibliometric analysis of tool-driven change
+# Bibliometric analysis of a tool-driven change
 
@@
 
-In our view, merely using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolutionb by itself, but publicaly sharing the code used for research, as a solution to problems of irreproducibility, could enable revolutionary change.
-Openly shared code could speed the widespread adoption of the reproducibility are a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and therefore properly evaluate, extend and reuse the published results.
+In our view, merely using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolution by itself, but publically sharing the code used for research, as a solution to problems of irreproducibility, could enable revolutionary change.
+Openly shared code could speed the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and therefore properly evaluate, extend and reuse the published results.
 As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as zenodo, osf, figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.
@@
 On another level, we'd like to point to the _technicity_ of tools, in our case coding.
-Aud Sissel Hoel and Iris van der Tuin are drawing on Ernst Cassirer's „Form und Technik“ [-@cassirer_form_1930] and Gilbert Simondon's „Du monde d'existence des objets techniques" [-@simondon_mode_1958] to define technicity as a term to describe that technology, i.
+Aud Sissel Hoel and Iris van der Tuin are drawing on Ernst Cassirer's „Form und Technik“ [@cassirer_form_1930] and Gilbert Simondon's "Du monde d'existence des objets techniques" [@simondon_mode_1958] to define technicity as a term to describe that technology, i.
 e.

### 2018-12-07 analysis/paper/jcaa-paper.Rmd
removed percentage calculations (Jeremys critique)

@@
                    'PY' = "year")) %>% 
-  mutate(prop = n / value ) %>% 
+  mutate(prop = n ) %>% # dropped / value to not have percentages
   filter(SO != 'JOURNAL OF STATISTICAL SOFTWARE') %>% 
@@
 
-```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Percentage of articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}).
+```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}).
 Data from Web of Science for 2007-2017')}
@@
                      limits = c(min_y, max_y)) +
-  scale_y_continuous(labels = scales::percent) +
+#  scale_y_continuous(labels = scales::percent) +
   xlab("Publication year") +
-  ylab("Percentage of articles in that journal in that year") +
+  ylab("Number of Articles in that journal in that year") +
   theme_minimal() +
@@
             by = c('SO', 'PY')) %>% 
-   mutate(prop = n.x / n.y) %>% 
+   mutate(prop = n.x) %>% # dropped /n.y to not have percentage
   ungroup()
@@
              by = "SO")
+
+# filter so not to have the "just one article in R"-journals clutter up the graph
+
+archy_cran_cites_all_areas_top_journals_by_year <- 
+archy_cran_cites_all_areas_top_journals_by_year %>%
+    filter(total_n_articles > 1)
 
@@
                      limits = c(min_y, max_y)) +
-  scale_y_continuous(labels = scales::percent,
-                     limits = c(0, 0.2)) +
+  #scale_y_continuous(labels = scales::percent,
+   #                  limits = c(0, 0.2)) +
   xlab("Publication year") +
-  ylab("Percentage of articles in that journal")
+  ylab("Number of articles in that journal")
 
@@
   xlab("year") +
-  ylab("Proportion of articles citing R") +
+  ylab("Number of articles citing R") +
   ggtitle(expression(paste("Articles citing R in ", 
@@
                           xmin = 2030, xmax = 2040, 
-                          ymin = 0.10, ymax = 0.17)
+                          ymin = 12, ymax = 17) # changed position because of numbers now instead of percentage
 

### 2018-12-07 analysis/paper/jcaa-paper.Rmd
try to make plot-legend in "articles per year citing R in top ecology journals" readable by extending the x-Axis.
still a line break in preoceedings of the royal society. -.-

@@
 min_y <- 2007
-max_y <- 2025
+max_y <- 2050
 all_sciences_citing_R_over_time <- 
@@
   xlab("Publication year") +
-  ylab("Number of Articles in that journal in that year") +
+  ylab("Number of Articles in journal per year") +
   theme_minimal() +
@@
 min_y <- 2007
-max_y <- 2040
+max_y <- 2050
 
@@
   xlab("Publication year") +
-  ylab("Number of articles in that journal")
+  ylab("Number of Articles in journal per year")
 
@@
   annotation_custom(grob = ggplotGrob(jas_cites_r), 
-                          xmin = 2030, xmax = 2040, 
+                          xmin = 2030, xmax = 2045, 
                           ymin = 12, ymax = 17) # changed position because of numbers now instead of percentage

### 2018-12-07 analysis/paper/jcaa-paper.Rmd
haha! I now know str_wrap and the labels don't overlap anymore.

@@
                   PY == max(PY)),
-    aes(label = str_wrap(SO_n, 40)),
+    aes(label = str_wrap(SO_n, 80)), # is this the line break problem solved?
     size = 4,
@@
    scale_x_continuous(breaks = min_y:max_y,
-                     labels = c(min_y:2017, rep("", length(2018:max_y))),
+                     labels = c(min_y:2017rep, rep("", length(2018:max_y))),
                      limits = c(min_y, max_y)) +
@@
       group_by(SO) %>% 
-      slice(which.max(PY)) , 
+      slice(which.max(PY)), 
     aes(label = SO_n),
@@
                      limits = c(min_y, max_y)) +
+  scale_y_continuous(trans = "log2") + # to be able to see smaller numbers better
     #scale_y_continuous(labels = scales::percent,
@@
   annotation_custom(grob = ggplotGrob(jas_cites_r), 
-                          xmin = 2030, xmax = 2045, 
-                          ymin = 12, ymax = 17) # changed position because of numbers now instead of percentage
+                          xmin = 2030, xmax = 2050, 
+                          ymin = 1.2, ymax = 2.6) # changed position because of numbers now instead of percentage + logarithmic scale in bigger plot
 

### 2018-12-07 analysis/paper/jcaa-paper.Rmd
played around with colours.

--- 2018-12-07 analysis/paper/jcaa-paper.Rmd
@@
 min_y <- 2007
-max_y <- 2050
+max_y <- 2040
 all_sciences_citing_R_over_time <- 
@@
   geom_line(size = 2) +
+  scale_colour_brewer(palette = "Paired") + 
   geom_text_repel(
@@
 min_y <- 2007
-max_y <- 2050
+max_y <- 2035
+
+library(RColorBrewer)
 
@@
   geom_line(size = 2) +
+  scale_colour_brewer(palette = "Paired") + 
   geom_point() +
@@
   annotation_custom(grob = ggplotGrob(jas_cites_r), 
-                          xmin = 2030, xmax = 2045, 
-                          ymin = 12, ymax = 17) # changed position because of numbers now instead of percentage
+                          xmin = 2025, xmax = 2035, 
+                          ymin = 2.9, ymax = 4) # changed position because of numbers now instead of percentage + logarithmic scale in bigger plot
 

### 2018-12-13 analysis/paper/jcaa-paper.Rmd
addesd jcaa.csl
but does not seem to be correct. in their description they say they want full first names, not just first letter. and dots seem to be missing.
pity. i don't know anything about csl...

--- 2018-12-07 analysis/paper/jcaa-paper.Rmd
@@
 bibliography: references.bib
+csl: journal-of-computer-applications-in-archaeology.csl
 abstract: "There is an argument in philosophy of science that revolutions in science are either ideas-driven or tool-driven.
@@
                    'PY' = "year")) %>% 
-  mutate(prop = n ) %>% # dropped / value to not have percentages
+  mutate(prop = n / value ) %>% 
   filter(SO != 'JOURNAL OF STATISTICAL SOFTWARE') %>% 
@@
 
-```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}).
+```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Number of articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}).
 Data from Web of Science for 2007-2017')}
 min_y <- 2007
-max_y <- 2050
+max_y <- 2025
 all_sciences_citing_R_over_time <- 
@@
                      limits = c(min_y, max_y)) +
-#  scale_y_continuous(labels = scales::percent) +
+  scale_y_continuous(labels = scales::percent) +
   xlab("Publication year") +
-  ylab("Number of Articles in journal per year") +
+  ylab("Percentage of articles in that journal in that year") +
   theme_minimal() +
@@
             by = c('SO', 'PY')) %>% 
-   mutate(prop = n.x) %>% # dropped /n.y to not have percentage
+   mutate(prop = n.x / n.y) %>% 
   ungroup()
@@
 
-# filter so not to have the "just one article in R"-journals clutter up the graph
-
-archy_cran_cites_all_areas_top_journals_by_year <- 
-archy_cran_cites_all_areas_top_journals_by_year %>%
-    filter(total_n_articles > 1)
-
 min_y <- 2007
-max_y <- 2050
+max_y <- 2040
 
@@
                      limits = c(min_y, max_y)) +
-  #scale_y_continuous(labels = scales::percent,
-   #                  limits = c(0, 0.2)) +
+  scale_y_continuous(labels = scales::percent,
+                     limits = c(0, 0.2)) +
   xlab("Publication year") +
-  ylab("Number of Articles in journal per year")
+  ylab("Percentage of articles in that journal")
 
@@
   xlab("year") +
-  ylab("Number of articles citing R") +
+  ylab("Proportion of articles citing R") +
   ggtitle(expression(paste("Articles citing R in ", 
@@
   annotation_custom(grob = ggplotGrob(jas_cites_r), 
-                          xmin = 2030, xmax = 2045, 
-                          ymin = 12, ymax = 17) # changed position because of numbers now instead of percentage
+                          xmin = 2030, xmax = 2040, 
+                          ymin = 0.10, ymax = 0.17)
 

### 2018-12-18 analysis/paper/jcaa-paper.Rmd
forgotten word

@@
 
-```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Number of articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}).
+```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Percentage of articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}).
 Data from Web of Science for 2007-2017')}
@@
 
-```{r text-mining-article-titles-archaeology-plot, fig.cap=paste0("Comparing the frequency of words used in titles \nof archaeology articles (", prettyNum(nrow(frequency)), " words in ", prettyNum(nrow(archy_journal_articles)), " articles)")}
+```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles \nof archaeology articles (", prettyNum(nrow(frequency)), " words in ", prettyNum(nrow(archy_journal_articles)), " articles)")}
 
@@
 
-```{r archaeology-reproducible-articles-plot, fig.cap=paste0('Articles in archaeology journals\nusing R for reproducible research (n = ', number_of_reproducible_articles,')')}
+```{r archaeology-reproducible-articles-plot, fig.cap=str_glue('Articles in archaeology journals\nusing R for reproducible research (n = ', number_of_reproducible_articles,')')}
 

### 2018-12-18 analysis/paper/jcaa-paper.Rmd
Added quote of author guidelines about reproducibility.

@@
 'the', 'archaeological', 'study').
-This resulted in `r prettyNum(nrow(frequency), ',')` words in `r prettyNum(nrow(archy_journal_articles))`.
+This resulted in `r prettyNum(nrow(frequency), ',')` words in `r prettyNum(nrow(archy_journal_articles))` articles.
 Third, for each word found in all the titles of the journal articles in each group, we computed its proportion of the total number of words in all the titles in each group.
@@
 Figure \@ref(fig:archaeology-reproducible-articles-plot) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.
+The author guidelines of this very journal confirm this trend: "In the interests of open scholarship and the reproducibility of results, JCAA strongly encourages all authors to deposit material relating to their publication in a preservation repository"[@authorguidelines_JCAA].
 

### 2018-12-18 analysis/paper/jcaa-paper.Rmd

@@
 Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with what we see desribed in the work of @Galison_1997 [-@Galison_1997].
-This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@Meltzer_1979; @Trigger_2006]).
+This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@Meltzer_1979; @Trigger_2006].
 

### 2018-12-19 analysis/paper/jcaa-paper.Rmd
adding reference to paper by Fitzpatrick

@@
 The tool which drives this change is the use of scripting languages in computational approaches of these disciplines.
-We  show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visiblity and influence.
+We  show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visiblity and influence (for some critical aspects on open scholarship see Huggett in this volume).
 This also is an indicator for the increase in popularity of R.
@@
 
-As with data, sharing code in public takes a little time to ensure that it is fit for others to read.
+As with data (see Fitzpatrick in this volume), sharing code in public takes a little time to ensure that it is fit for others to read.
 Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error, rather than following a model or widely-used approach.

### 2018-12-25 analysis/paper/scraps.R
add scraps

@@ -1,0 +1,72 @@
+
+the easy topics seem to me to be: 
+- [ ] respond to Heather and Jeremy with something about R as a case study for scripting languages in general
+- [ ] Angus says to clarify our position, 
+- [ ] Jeremy wants more on rrtools and why it's needed
+
+
+Things to add
+-[x] "boundary work" by Gyrien at the beginning of the paper 
+-[x] other tool-driven changes in archaeoology: GIS, isotopes, remote sensing, DNA, networks
+
+Yes! That topic of boundary work by Gieryn that you mention is perfect! What a good find! I believe it may also related to Galison's concept of 'trading zones'. Perhaps we could say something like the increasing use and availability of code, and demands for it by readers and peer reviews is active boundary work that is shifting the boundary of scientific and non-scientific archaeology. And we can follow up with some observation about how archaeologists are increasing found in trading zones with computer scientists and software engineers as they try to find effective and efficient methods to use code and made it available in research compendia. Yes, I think that paper on boundary work is very suitable for our paper!  I will work it into the intro section, and again to wrap up at the end. Thanks again for the tip!
+
+#-----------------
+Our paper for the JCAA:
+
+I've read an article about "boundary work" by Gyrien (attached - it seemed to be quite widely cited) and it seems to discuss how we differentiate "science" from "non-science" and that there are not just theoretical aspects of this discussion ("reasoning and observation" by  Comte; "falsifiability" by Popper, ...), but a practical aspect: Scientists decide on a daily basis, what they consider to be scientific. This process of demarcation is something we actually try to do as well: We say, only papers that also submit code and data are scientific. 
+Yeah, and then the paper goes on about different sociological theories of ideology and what ideologies of science have been developed. It concludes that the boundaries of science are ambigious and defined for specific goals of authority expansion, monopolization and protection of autonomy.
+
+Sounds like something we can add to the beginning of our paper, where we talk about our aims.
+
+#-----------------
+
+You already mention C14-dating, I thought of GIS, networks, ABMs, isotope analysis, remote sensing and DNA analysis.
[…]

### 2018-12-27 analysis/paper/jcaa-paper.Rmd
white space
minor edits, what's wrong with some of the fig  captions?

@@
 
-In this paper we investigate recent developments which facilitate computational archaeology to return to principles of the scientific method and whether this might be the beginning of a paradigm-shift.
-Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discourses [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might not just be ideas-driven but also tool-driven [@Galison_1997; @Dyson_2000].
-In this light we explore how Ecology and Archaeology have started to embrace methods which improve the reproducibility of their research.
+In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method and whether this might be the beginning of a broader paradigm-shift.
+Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might  be not only ideas-driven, but also tool-driven [@Galison_1997; @Dyson_2000].
+In this light we explore how the disciplines of Ecology and Archaeology have started to embrace tools that improve the reproducibility of their research.
 We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code.
@@
 Acknowledging this is not a simple process, we offer a R-based tool to facilitate the task of creating a compendium which enables other researchers to reproduce the published results.
-
-Code and text for this paper are available at: https://github.com/benmarwick/caa2018-Tubingen
 
@@
 
+# Methods
+
 We used the 'Cited Reference Search' function provided by the Web of Science online scientific citation indexing service to identify journal articles that cite R.
@@
 We then downloaded the bibliographic data for all of these articles.
+
+## Reproducibility and open source materials
+
+To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in @Marwick2017, we include the entire R code used for all the analysis and visualizations contained in this paper at http://doi.org/10.17605/OSF.IO/RHVN5.
+Also in this version-controlled research compendium are the raw data for all the results reported here.
+All of the figures, tables and statistical test results presented here can be independently reproduced with the code and data in this repository.
+In our online materials our code is released under the MIT licence, our data as CC-0, and our figures as CC-BY, to enable maximum re-use [for more details, see @Marwick2017].
+
+# Results
 
@@
 
-```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles \nof archaeology articles (", prettyNum(nrow(frequency)), " words in ", prettyNum(nrow(archy_journal_articles)), " articles)")}
+```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles \nof archaeology articles (", prettyNum(nrow(frequency), big.mark = ","), " words in ", prettyNum(nrow(archy_journal_articles), big.mark = ","), " articles)")}
 

### 2018-12-28 analysis/paper/jcaa-paper.Rmd
minor edits, more examples of revolutions and paradigm shifts in archaeology

@@
 Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might  be not only ideas-driven, but also tool-driven [@Galison_1997; @Dyson_2000].
-In this light we explore how the disciplines of Ecology and Archaeology have started to embrace tools that improve the reproducibility of their research.
+In this light we explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of their research.
 We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code.
-A review of bibliometric evidence showcases a strong development towards the use of R for Ecology and hints at the beginning of a similar development in archaeology.
+A review of bibliometric evidence showcases a strong development towards the use of R for ecology and hints at the beginning of a similar development in archaeology.
 We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven change in archaeology.
-Acknowledging this is not a simple process, we offer a R-based tool to facilitate the task of creating a compendium which enables other researchers to reproduce the published results.
+Acknowledging that this is not a simple process, we offer an R-based tool to ease the task of creating a compendium which enables other researchers to reproduce the published results.
 
@@
 
-There is an argument among philosophers of science that revolutions in science are either ideas-driven or tool-driven.
+Philosophers of science disagree if revolutions in science are ideas-driven or tool-driven.
 One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] effort to describe the history of science by proposing paradigm shifts resulting from revolutionary change in communities of researchers.
@@
 What counts as suitable puzzles, and acceptable solutions to these puzzles, is governed by the norms and procedures of the prevailing paradigm.
-Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes that replace one paradigm with another in a conversion or speciation process where the community adopts a new way of doing science.
-There has been substantial debate about how to draw a clear distinction between so-called normal and revolutionary science.
+Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes replacing one paradigm with another in a conversion or speciation process where the community adopts a new way of doing science.
+There has been much debate about how to draw a distinction between so-called normal and revolutionary science.
 [@watkins1970against; @toulmin1970does; @Casadevalle00158_16].
 A notable criticism is that Kuhn's normal science is unscientific, a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
-This is at odds with a view of science that it is in a continuous state of revolution, with researchers simultaneously employing several styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
+This is at odds with a view of science as a continuous state of revolution, with researchers simultaneously employing many styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
 In this view, based on the post-World War II era of science, radical discontinuities are rare, and major developments instead emerge from the division and recombination of already mature fields [@andersen2013second].
@@
 The most wide-ranging application is Clark's [-@Clark_1993] comparison of North American archaeology to European archaeology during the 1970s-90s.
-Clark, similar to Binford and Sabloff [-@binford1982paradigms], identified the North American archaeological paradigm as rooted in anthropology, founded on post-positivist biases, and distinctive with it critical-realist ontology, a modified-objectivist epistemology, and an experimental-manipulative methodology.
-In contrast, the European archaeological paradigm is described a kind of history, mostly in a strict empiricist tradition, and typified by a realist ontology, an inductivist epistemology, and by an observational methodology.
-Clark's analysis describes a multi-paradigm situation where multiple research communities (i.e.
-North American and European) simutaneously operate in different paradigms (which Kuhn noted as a sign of a developing or immature discipline).
-His account is not a complete Kuhnian analysis because it lacks a revolutionary change event where one paradigm is replaced by another, although he hints that processual archaeology may be considered a radical discontinuity relative to the culture-historical approaches that preceeded it [@Clark_1993, pp.
+Clark, similar to Binford and Sabloff [-@binford1982paradigms], identified the North American archaeological paradigm as rooted in anthropology, founded on post-positivist biases, and distinctive with its critical-realist ontology, a modified-objectivist epistemology, and an experimental-manipulative method.
+In contrast, the European archaeological paradigm is described a kind of history, mostly in a strict empiricist tradition, and typified by a realist ontology, an inductivist epistemology, and by an observational method.
+Clark's analysis describes a multi-paradigm situation where many research communities (i.e.
+North American and European) simultaneously operate in different paradigms (which Kuhn noted as a sign of a developing or immature discipline).
+Clark's account is not a complete Kuhnian analysis because it lacks a revolutionary change event where one paradigm is replaced by another, although he hints that processual archaeology may be considered a radical discontinuity relative to the culture-historical approaches that preceded it [@Clark_1993, pp.
 206].
@@
 Meltzer reviews literature arguing for a revolutionary change in archaeology during the 1960s and 1970s, and considers these claims in light of classic Kuhnian revolutionary events such as the replacement of the Ptolemaic system by the Copernican system, and Newtonian dynamics by the new Einsteinian dynamics.
-He finds that the changes that occured in Anglophone archaeology after the 1960s were incremental, and primarily of method, rather than a widespread replacement of one ontological structure by another one, incommensurable or incompatible with the former.
+He finds that the changes occuring in Anglophone archaeology after the 1960s were incremental, mostly of method, rather than a widespread replacement of one ontological structure by another one, incommensurable or incompatible with the former.
 Trigger [-@Trigger_2006, pp.
@@
 
-We see less formal mentions of paradigm shifts in archaeology in reference to processual to post-processual archaeology [@Härke_2002], as well as more localised and specific shifts in archaeological practice and thinking.
-For example, @Snodgrass_2002 claims that among Classical Archaeologists, an increase in research on previously neglected periods odf antiquity, such as the Greek Early Iron Age, constitutes a paradigm shift.
+We also see mentions of paradigm shifts in archaeology in reference to processual to post-processual archaeology [@Härke_2002], as well as more localised and specific shifts in archaeological practice and thinking.
+For example, @Snodgrass_2002 claims that among Classical Archaeologists, an increase in research on previously neglected periods of antiquity, such as the Greek Early Iron Age, constitutes a paradigm shift.
 @McAnany_Rowe_2015 propose that the appearance of community-based participatory models of research among some communities of archaeologists is a paradigm shift, though they concede that it is more transformational rather than revolutionary.
+@harris2012interfacing has similarly argued that community crowdsourced geographic knowledge (or volunteered geographic information) could be paradigm shift for archaeological communities.
+Archaeologists using LIDAR (light detection and ranging) for remote geospatial imaging of cultural landscapes have claimed the application of this technology to archaeology is a paradigm shift [@chase2012geospatial @howey2016detecting].
 @Fuller_2010 has used the term to describe a shift in thinking about the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to new concept of the emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
-Most of these paradigms are quite remote in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's paradigm shift approach, both as a tool to understand the history of science and as a guide to changing it, which is Kuhn's multiple definitions of a paradigm.
+Most of these paradigms differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's paradigm shift approach, both as a tool to understand the history of science and as a guide to changing it, which is the multiple definitions of a paradigm in Kuhn's writings.
 
-Masterman [-@masterman1970does] has documented 21 different uses by Kuhn, which she organises into three groups.
-First is the metaphysical notion of a set of beliefs, and this is the sense that dominates most of the critical writing on Kuhn's model.
-Second is the sociological sense of scientific habits, the universially recognised scientific achievements that are the foundations of day-to-day normal science.
-These include specific attention-grabbing successes that provide a touchstone for the coordination of future research.
-Third is the concrete sense of an actual textbook, instrumentation or toolkit.
-This recognition of the multivalent nature of paradigms is important because it broadens the locus for scientific revolution to include sociological practices of science, such as might be produced by changes in norms of publication and peer review, and to include the physical infrastructure of doing science, such as the technology for collecting and analysing data, and for communication and collaboration among researchers.
+Masterman [-@masterman1970does] has documented 21 different uses of 'paradigm' by Kuhn, which she organises into three groups.
+First is the metaphysical notion of a set of beliefs, and this is the sense dominating most of the commentary on Kuhn's model.
+Second is the sociological sense of scientific habits, the universally recognized scientific achievements that are the foundations of day-to-day normal science.
+These include specific attention-grabbing successes that give a touchstone for the coordination of future research.
+Third is the concrete sense of an actual textbook, instrumentation or toolkit in wide use.
+This recognition of the multivalent nature of paradigms is important because it broadens the locus for scientific revolution to include sociological practices of science, such as might be produced by changes in norms of publication and peer review, and to include the physical infrastructure -- tools -- of doing science, such as the technology for collecting and analysing data, and for communication and collaboration among researchers.
 
-We can see a compelling exploration of the concrete sense of a paradigm in the work of @Galison_1997, who has offered a tool-driven view of change in research communities.
-Galison analysed of the role of tools in twentieth century particle physics, starting with hand-crafted cloud chambers and bubble chambers and ending with digital counters of and computers.
+We see a compelling exploration of the concrete sense of a paradigm in the work of @Galison_1997, who has offered a tool-driven view of change in research communities.
+Galison analysed of the role of tools in twentieth century particle physics, starting with hand-crafted cloud chambers and bubble chambers and ending with digital counters, particle accelerators, and computers.
 He identifies a profound change in physics when analog devices producing pictures were superseded by digital devices producing numerical data.
 At the heart of this change is a shift from an intuitive approach, stimulated by visual and pictorial model-building to an approach based on logic, calculation, and demonstration.
-The changes are not purely at the points of data collection and analysis, but extend to the social and economic organisation of science, with new categories of physicists emerging, people who are not quite experimenters and not quite theorists.
-Dyson [-@Dyson_2000; -@dyson2012science] has further explored this tool-driven approach, linking the Galilean revolution to the use of the telescope in astronomy, and the Crick-Watson revolution to the use of X-ray diffraction.
+The changes are not purely at the points of data collection and analysis, but extend to the social and economic organisation of science, with new categories of physicists emerging, people who are not entirely experimenters and not entirely theorists.
+Dyson [-@Dyson_2000; -@dyson2012science] has further explored this tool-driven approach, tracing the origin of the Galilean revolution to appearance of the telescope in astronomy, and the origin of the Crick-Watson revolution to the use of X-ray diffraction.
 
-Galison's emphasis on tool-driven change, as a complement to Kuhnian concept-driven change, has much potential for archaeology.
+We propose that Galison's emphasis on tool-driven change, as a complement to Kuhnian concept-driven change, has much potential for archaeology.
 The centrality of material culture in Galison's view raises the question about whether there is scope for archaeologists to make productive contributions to understanding change in the history science through an archaeological analysis of scientific instrumentation [cf.
@@
 @molyneaux2013cultural], in contrast to digital devices?
-One candidate for this might be the shift from building relative chronologies based on seriation of visually distinctive artefacts to absolute chronologies based on radiometric dating using computer-controlled instruments.
+One candidate for this might be the shift from building relative chronologies based on seriation using typologies of visually distinctive artefacts to absolute chronologies based on radiometric dating using computer-controlled instruments.
+Three revolutions have been attributed to radiocarbon dating and associated technologies [@bayliss2009rolling], and @chase2012geospatial have claimed the introduction of radiocarbon dating was a paradigm shift in archaeology.
 A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to predict or generate change in archaeology?
@@
 
-The question of predicting or generating change in scientific practice via adoption of new tools is especially intruiging, and in this section we explore how a change that has been observed in ecology might be relevant for understanding or even directing the future of archaeology.
+The question of predicting or generating change in scientific practice via adoption of new tools is especially intriguing, and in this section we explore how a change that has been observed in ecology might be relevant for understanding or even directing the future of archaeology.
 Biological disciplines have a long tradition of influencing archaeological thought, starting with Oscar Montelius, whose typological method was inspired by Darwin's theory of evolution [@Montelius_1899, 237].
@@
 It has also expanded the 'trading zone' [@Galison_1997] where archaeologists collaborate with specialists in other fields to develop a shared language to get things done.
-In using scripting languages archaeologists are working at a trading zone with computer scientists and computational reserchers in other fields to coordinate how tools from one domain can be useful in another.
+In using scripting languages archaeologists are working at a trading zone with computer scientists and computational researchers in other fields to coordinate how tools from one domain can be useful in another.
 We can also see archaeologists in similar trading zones when using GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions.
@@
 The tool which we have been discussing here is R, as an example for scientific scripting languages.
+
 We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in ecology.

### 2018-12-29 analysis/paper/jcaa-paper.Rmd
minor edits and adding citations

@@
 One candidate for this might be the shift from building relative chronologies based on seriation using typologies of visually distinctive artefacts to absolute chronologies based on radiometric dating using computer-controlled instruments.
-Three revolutions have been attributed to radiocarbon dating and associated technologies [@bayliss2009rolling], and @chase2012geospatial have claimed the introduction of radiocarbon dating was a paradigm shift in archaeology.
+Three revolutions in archaeology have been attributed to radiocarbon dating and associated technologies [@bayliss2009rolling].
+@chase2012geospatial have claimed the introduction of radiocarbon dating was a paradigm shift in archaeology, but we are not aware of a Galisonian analysis of this change.
 A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to predict or generate change in archaeology?
@@
 
-The question of predicting or generating change in scientific practice via adoption of new tools is especially intriguing, and in this section we explore how a change that has been observed in ecology might be relevant for understanding or even directing the future of archaeology.
+The question of predicting or generating change in scientific practice via adoption of new tools is intriguing, and in this section we explore how a change observed in ecology might be relevant for understanding or even directing the future of archaeology.
 Biological disciplines have a long tradition of influencing archaeological thought, starting with Oscar Montelius, whose typological method was inspired by Darwin's theory of evolution [@Montelius_1899, 237].
-Also nowadays quite a number of spatial statistical algorithms used in archaeology are derived from Ecology [@keron_use_2015, 7].
+This tradition continues, for example, many spatial statistics in current use in archaeology are derived from Ecology [@keron_use_2015, 7].
 
@@
 
-@Touchon_McCoy_2016 investigated evaluated changes in statistics used by ecologists as a potential area of tool-driven change.
+@Touchon_McCoy_2016 investigated evaluated changes in statistical methods used by ecologists as a potential area of tool-driven change.
 They searched nearly 20,000 published articles in ecology published between 1990 and 2013.
@@
 First, they note that automated data loggers, GPS trackers, remote sensing, and crowd sourcing have greatly increased rate at which ecologists collect data.
-Second, increases in desktop and cluster computing power has made complex analytical processes faster and more convienent to compute.
+Second, increases in desktop and cluster computing power has made complex analytical processes faster and more convenient to compute.
 Third, the development of free, open source and easily extensible software for data analysis and visualisation, such as R, allows new methods to spread quickly via online fora and social media.
@@
 
-While all of these four factors could be considered relevant to the tooling of ecology and archaeology, we focus here on the third one because of its relevance to archaeology.
+We focus here on the third factor, free and open source software, because it is the most generic factor and so the most relevant beyond ecology.
 Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct, with only a small overlap.
 What they share in common is working with the field-collected data on a computer to prepare it for publication.
-R is the most common open source and free ware data analysis tool available and therefore a good proxy for the adoption of open source scientific programming languages in a discipline.
+R is a widely used free and open source data analysis tool in man research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018] and therefore a good proxy for the adoption of open source scientific programming languages in a discipline.
 To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.
 
-```{r ecology-citing-r}
+```{r ecology-citing-r, echo=FALSE}
 
@@
 
-We used the 'Cited Reference Search' function provided by the Web of Science online scientific citation indexing service to identify journal articles that cite R.
+We used the 'Cited Reference Search' function provided by the Web of Science online scientific citation indexing service to find journal articles citing R.
 Although R has been available since the late 1990s [@Thieme_2018], a recommended format for citing the software did not appear until 2004, with the author given as "R Development Core Team".
 This recommended format for citing the software changed slightly in 2012 when the author was updated to "R Core Team".
-We searched the Web of Science using '"R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM"' in the CITED AUTHOR field.
-We sorted the results by frequency of citation, and selected the first 1000 items (the maximum allowed by the Web of Science service).
-These 1000 items represent variations on the recommended format for citing R, and found in the reference lists of `r prettyNum(how_many_articles_in_top_journals_for_cran_cites, ",")` English-languge articles indexed by the Web of Science in the research area of Environmental Sciences Ecology.
-We then downloaded the bibliographic data for all of these articles.
+We searched the Web of Science using '"R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM"' in the CITED AUTHOR field of the Web of Science database.
+We sorted the results by frequency of citations, and selected the first 1000 items (the maximum allowed by the Web of Science service).
+These 1000 items represent variations on the recommended format for citing R.
+We found citations of R in the reference lists of `r prettyNum(how_many_articles_in_top_journals_for_cran_cites, ",")` English-languge articles indexed by the Web of Science in the research area of 'Environmental Sciences Ecology'.
+We then downloaded the bibliographic data and reference lists for each of these articles.
 
@@
 Figure \@ref(fig:ecology-articles-citing-r-plot) shows the percentage of articles citing R in each of several of the top ecology journals (as defined by how often their articles are cited).
-We restrict the start of the observation period to 2007 for convenience.
+We restrict the start of the observation period to 2007 for convenience so we have a ten year study period.
 The plot shows that the percentage of articles citing R has increased from less than ten percent in all journals in the late 2000s, to more than 30% in _Ecosphere_, _Ecology and Evolution_, and _Molecular Ecology_ after 2012.
-We would not call this a Kuhnian paradigm shift, but it does show a substantial change in the tools of the discipline, supporting claims for a Galisonian tool-driven change in ecology similar to the changes described by @Touchon_McCoy_2016.
+We might not call this a Kuhnian paradigm shift, but it does show a substantial change in the tools of the discipline, supporting claims for a Galisonian tool-driven change in ecology similar to the changes described by @Touchon_McCoy_2016.
 
@@
 Only the _Journal of Archaeological Science_ has more than 50 articles in our sample.
-We might conclude from this observation that archaeologists have yet to adopt programming for data analysis and visualisation in the same way that ecologists have.
+We conclude from this observation that archaeologists have yet to adopt programming for data analysis and visualisation in the same way ecologists have.
 
-The second detail is that most archaeology journals do not show any increase in the percentage of article citing R over time.
-The ecology journals shown in Figure \@ref(fig:ecology-citing-r) show clear upward trends of increasing proportions of articles citing R over time, but we do not see any clear trends in the archaeology articles considered together.
-Indeed, in several of the archaeological journals in our sample, the first citations of R only occured in the last two years.
+The second detail is that most archaeology journals do not show any strong increase in the percentage of article citing R over time.
+The ecology journals shown in Figure \@ref(fig:ecology-citing-r) show strong upward trends of increasing proportions of articles citing R over time, but we do not see any unambiguous trends in the archaeology articles when considered together.
+Indeed, in several of the other archaeological journals in our sample, the first citations of R only occurred in the last two years.
 The sample sizes for the archaeology journals are too small to confidently infer any trend of over time, with the exception of the _Journal of Archaeological Science_ with `r how_many_articles_in_jas` articles.
-The inset plot in Figure \@ref(fig:archaeology-journals-with-subplot) shows a statistical test of change over time for articles in the _Journal of Archaeological Science_.
-With a moderate r-squared value and low p-value for the linear model, we can conclude that there is some evidence for an increase in citations of R over time in this journal.
-This may indicate an increasing use of R among archaeologists, especially those working at the intersection of the natural and physical sciences.
-This may hint at the start of a trend similar to what we see in the ecology journals of a widespread adoption of R.
+The inset plot in Figure \@ref(fig:archaeology-journals-with-subplot) shows a statistical test of change over time for articles citing R in the _Journal of Archaeological Science_.
+With a moderate r-squared value and low p-value for the linear model, we conclude there is some evidence for an increase in citations of R over time in this journal.
+This may show an increasing use of R among archaeologists, especially those working at the intersection of archaeology and the natural sciences.
+This may hint at the start of a trend like what we see in the ecology journals of a widespread adoption of R.
 
 Following on from this observation about the increasing popularity of R in the _Journal of Archaeological Science_, the third detail we see in these results is the distinctive types of journals that have articles citing R.
-There is a clear focus on journals that publish scientific archaeology, especially those that focus on specialised empirical analysis.
+There is a focus on journals publishing scientific archaeology, especially those that focus on specialised empirical analysis.
 In addition to the general archaeology journals, we also see _Lithic Technology_, _Anthropozoologica_, and _International Journal of Osteology_ in the journal names, indicating that we find R used by specialists in stone artefacts and faunal remains.
-To further investigate the topics that archaeologists are using R to study, we conducted a statistical analysis of the words in the titles of all the articles in our sample.
+For more fine-grained insights into the topics that archaeologists are using R to study, we conducted a statistical analysis of the words in the titles of all the articles in our sample.
 
@@
 
-We computed a comparison of word frequences in journal article titles in the Web of Science data to get a better understanding of what topics archaeologists are writing about when they cite R.
+We computed a comparison of word frequencies in journal article titles in the Web of Science data to get a better understanding of what topics archaeologists are writing about when they cite R.
 First we separated the archaeology articles into two groups, those that cite R, and those that do not.
-Second, we filtered to keep only words that occur in titles in both groups, and removed stop words that are very common and uninformative (e.g.
-'the', 'archaeological', 'study').
+Second, we filtered to keep only words that occur in titles in both groups, and removed very common and uninformative words (e.g.
+'the', 'archaeological', 'study', etc.).
 This resulted in `r prettyNum(nrow(frequency), ',')` words in `r prettyNum(nrow(archy_journal_articles))` articles.
@@
 Figure \@ref(fig:text-mining-article-titles-archaeology-plot) shows the results of this analysis.
-Words near the red line are used with about equal frequencies by papers that cite R and by papers that do not cite R.
+Words near the red line are used with about equal frequencies by papers citing R and by papers not citing R.
 Words far away from the red line are used much more by one group of articles compared to the other.
@@
 
-Figure \@ref(fig:text-mining-article-titles-archaeology-plot) supports our initial observations that archaeologists using R tend to be doing archaeology that involves the natural and physical sciences.
+Figure \@ref(fig:text-mining-article-titles-archaeology-plot) supports our observation that archaeologists using R tend to be doing archaeology that involves the natural and physical sciences.
 Articles that cite R more often have titles that include terms such as 'lithic', 'experimental', and 'isotope'.
-On the other hand, articles that do not cite R more often discuss the Roman and Bronze Age periods, and less often reflect a technical specialisation, if there is one, it seems to be in ceramics.
+On the other hand, articles that do not cite R more often discuss the Roman and Bronze Age periods, and less often reflect a technical specialisation.
+If there is one that is characteristic of non-citing R articles, it seems to be in ceramics.
 
@@
 
-@Touchon_McCoy_2016 had identified the use of scripting languages as being a factor for easily sharing methods and code, thereby increasing the rate of development of statistical methods.
-One notable practice that we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors make publicly available the code and data that they used to generate the results presented in their article.
-We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository.
-Our contiunuously updated list of articles in this sample is available online here: https://github.com/benmarwick/ctv-archaeology.
+@Touchon_McCoy_2016 identified the use of scripting languages as being a factor for easily sharing methods and code, thereby increasing the rate of development of statistical methods.
+One notable practice we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors make publicly available the code and data that they used to generate the results presented in their article.
+We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository such as Zenodo, Open Science Framework, Figshare, etc.
+Our continuously updated list of articles in this sample is available online here: https://github.com/benmarwick/ctv-archaeology.
 Figure \@ref(fig:archaeology-reproducible-articles-plot) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.
-The author guidelines of this very journal confirm this trend: "In the interests of open scholarship and the reproducibility of results, JCAA strongly encourages all authors to deposit material relating to their publication in a preservation repository"[@authorguidelines_JCAA].
+The author guidelines of this journal confirm this trend: "In the interests of open scholarship and the reproducibility of results, JCAA strongly encourages all authors to deposit material relating to their publication in a preservation repository" [@authorguidelines_JCAA].
 
 These archaeology articles that include publicly available R code and data are part of a new approach to scientific communication that has also been observed in other fields.
+For example, the journal *Nature Neuroscience* requires authors 'to make the code that supports the generation of key figures in their manuscript available for review' [@Editors_2017].
 Although we are not aware of attempts to quantify this change in the same way we have for archaeology, we have observed a number of recent publications that describe and recommend code sharing in statistics [@Baumer2014], genome biology [@markowetz2015five], computational biology [@Sandve2013], biostatistics [@Peng2009], computer science [@peng2011reproducible; @Mitchell2012], applied mathematics [@LeVeque2012], speech science [@Abari2012], political science [@king1995replication; @Dafoe2014], and the social sciences generally [@Miguel2014].
@@
 Much of this literature is concerned with identifying and solving problems of irreproducibility in research.
-This comes from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community.
-Furthermore, an article should sufficiently describe the presented research results such that a colleague can fully understand the results and how they were obtained.
-Many of these authors trace problems of irreproducibility to the increased complextity of computer-based analyses (but only a limited space to describe them in a journal article), and the use of mouse-driven computer programs where the researcher's analytical decisions are not recorded during the research process, and thus cannot be shared with other researchers.
+Furthermore, we see some journals requiring code as a requirement of for publication, for example *Nature Neuroscience* requires authors 'to make the code that supports the generation of key figures in their manuscript available for review' [@Editors_2017].
+These actions comes from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community [@Stark_2018].
+Furthermore, that an article should sufficiently describe the presented research results such that a colleague can fully understand the results and how they were obtained.
+Many of these authors trace problems of irreproducibility to the increased complexity of computer-based analyses, combined with the limited space to describe them in a journal article, and mouse-driven computer programs where the researcher's analytical decisions are not recorded during the research process, and thus cannot be shared with other researchers.
 
-In our view, merely using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolution by itself, but publically sharing the code used for research, as a solution to problems of irreproducibility, could enable revolutionary change.
-Openly shared code could speed the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and therefore properly evaluate, extend and reuse the published results.
-As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as zenodo, osf, figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.
+In our view, merely using R (or similar open source programming languages) to conduct research would not constitute a tool-driven revolution by itself, but publicly sharing the code used for research, as a solution to problems of irreproducibility, is more likely to enable revolutionary change.
+Openly shared code speeds the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
+As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as Zenodo, Open Science Framework, Figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.
+Using and sharing open source code is thus an important action for reducing inequality in the archaeological research community.
+This is because it enables participation by researchers who are unable to buy software licenses or access face-to-face training.
 
@@
 
-A pragmatic issue relating to wide-spread changes in scientific practice is what the incentives are for members of the research community.
-This is especially a concern for researchers that might not be motivated by a desire to adhere to the abstract basic principles of what it means to practice science with integrity.
-If code sharing has some concrete career advantages then we might expect it to be more likely to be part of a tool-drive revolution.
+## Citation effects of sharing code
+
+A pragmatic issue relating to widespread changes in scientific practice is what the incentives are for members of the research community to change.
+This is a concern for researchers that might not be motivated by a desire to adhere to the abstract basic principles of what it means to practice science with integrity.
+If code sharing has concrete career advantages then we might expect it to be more likely to be part of a tool-drive revolution.
 @Vandewalle2012 analysed citation rates for 645 articles published in _IEEE Transactions on Image Processing_ during 2004–2006.
 He found statistically significant differences between the median number of citations for papers that have code available and papers that do not have code available.
-In Figure \@ref(fig:citation-advantage) we conducted a similar anlysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
+In Figure \@ref(fig:citation-advantage) we conducted a similar analysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
 TC stands for 'Web of Science Core Collection Times Cited Count'.
@@
 The tool which drives this change is the use of scripting languages in computational approaches of these disciplines.
-We  show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visiblity and influence (for some critical aspects on open scholarship see Huggett in this volume).
+We  show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visibility and influence (for some critical aspects on open scholarship see Huggett in this volume).
 This also is an indicator for the increase in popularity of R.
@@
 It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
-Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with what we see desribed in the work of @Galison_1997 [-@Galison_1997].
+Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with what we see described in the work of @Galison_1997 [-@Galison_1997].
 This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@Meltzer_1979; @Trigger_2006].
 
-Rather, this change enables archaologists to more faithfully adhere to the fundamental tenet of the scientific process which demands the reproducibility of results.
+Rather, this change enables archaeologists to more faithfully adhere to the fundamental tenet of the scientific process which demands the reproducibility of results.
 We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages.

### 2019-01-01 analysis/paper/jcaa-paper.Rmd
conclusion done, more refs added

@@
 
-In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method and whether this might be the beginning of a broader paradigm-shift.
-Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might  be not only ideas-driven, but also tool-driven [@Galison_1997; @Dyson_2000].
+In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method and whether this might be the beginning of a broader change.
+Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are ideas-driven or tool-driven [@Galison_1997; @Dyson_2000].
 In this light we explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of their research.
-We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code.
-A review of bibliometric evidence showcases a strong development towards the use of R for ecology and hints at the beginning of a similar development in archaeology.
-We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven change in archaeology.
+We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of these scripts.
+Bibliometric evidence shows a strong increase in the use of R among ecologists and the beginning of a similar development in archaeology.
+We evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be a tool-driven change in archaeology.
 Acknowledging that this is not a simple process, we offer an R-based tool to ease the task of creating a compendium which enables other researchers to reproduce the published results.
@@
 
-We also see mentions of paradigm shifts in archaeology in reference to processual to post-processual archaeology [@Härke_2002], as well as more localised and specific shifts in archaeological practice and thinking.
+We also see mentions of paradigm shifts in archaeology in reference to processual to post-processual archaeology [@Härke_2002; @Koerner_2018], as well as more localised and specific shifts in archaeological practice and thinking.
 For example, @Snodgrass_2002 claims that among Classical Archaeologists, an increase in research on previously neglected periods of antiquity, such as the Greek Early Iron Age, constitutes a paradigm shift.
@@
 @harris2012interfacing has similarly argued that community crowdsourced geographic knowledge (or volunteered geographic information) could be paradigm shift for archaeological communities.
-Archaeologists using LIDAR (light detection and ranging) for remote geospatial imaging of cultural landscapes have claimed the application of this technology to archaeology is a paradigm shift [@chase2012geospatial @howey2016detecting].
+Archaeologists using LIDAR (light detection and ranging) for remote geospatial imaging of cultural landscapes have claimed the application of this technology to archaeology is a paradigm shift [@chase2012geospatial; @howey2016detecting].
 @Fuller_2010 has used the term to describe a shift in thinking about the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to new concept of the emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
-Most of these paradigms differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's paradigm shift approach, both as a tool to understand the history of science and as a guide to changing it, which is the multiple definitions of a paradigm in Kuhn's writings.
+Most of these paradigms differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, both as a tool to understand the history of science and as a guide to changing it, which is the multiple definitions of a paradigm in Kuhn's writings.
 
@@
 
-We see a compelling exploration of the concrete sense of a paradigm in the work of @Galison_1997, who has offered a tool-driven view of change in research communities.
+We see a compelling exploration of this third concrete sense of a paradigm in the work of @Galison_1997, who has offered a tool-driven view of change in research communities.
 Galison analysed of the role of tools in twentieth century particle physics, starting with hand-crafted cloud chambers and bubble chambers and ending with digital counters, particle accelerators, and computers.
@@
 Biological disciplines have a long tradition of influencing archaeological thought, starting with Oscar Montelius, whose typological method was inspired by Darwin's theory of evolution [@Montelius_1899, 237].
-This tradition continues, for example, many spatial statistics in current use in archaeology are derived from Ecology [@keron_use_2015, 7].
+This tradition continues, for example, many spatial statistics in current use in archaeology are derived from ecology [@keron_use_2015, 7].
 
@@
 We used the 'Cited Reference Search' to find articles citing R in the same way as above, and then refined the results to keep only those articles published during 2007-2017 that are included in the Web of Science category 'Archaeology'.
-This resulted in `r how_many_articles_in_top_archy_journals` articles, of which `r total_number_archaeology_articles_citing_r` cite R.
+This resulted in `r prettyNum(how_many_articles_in_top_archy_journals, ",")` articles, of which `r total_number_archaeology_articles_citing_r` cite R.
 Figure \@ref(fig:archaeology-journals-with-subplot) shows the temporal trend of citations of R in archaeological articles.
@@
 'the', 'archaeological', 'study', etc.).
-This resulted in `r prettyNum(nrow(frequency), ',')` words in `r prettyNum(nrow(archy_journal_articles))` articles.
+This resulted in `r prettyNum(nrow(frequency), ',')` words in `r prettyNum(nrow(archy_journal_articles), ",")` articles.
 Third, for each word found in all the titles of the journal articles in each group, we computed its proportion of the total number of words in all the titles in each group.
@@
 
-```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles \nof archaeology articles (", prettyNum(nrow(frequency), big.mark = ","), " words in ", prettyNum(nrow(archy_journal_articles), big.mark = ","), " articles)")}
+```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles of archaeology articles ({prettyNum(nrow(frequency), ",")} words in {prettyNum(nrow(archy_journal_articles), ",")} articles)")}
 
@@
 Much of this literature is concerned with identifying and solving problems of irreproducibility in research.
-Furthermore, we see some journals requiring code as a requirement of for publication, for example *Nature Neuroscience* requires authors 'to make the code that supports the generation of key figures in their manuscript available for review' [@Editors_2017].
 These actions comes from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community [@Stark_2018].
@@
 
-```{r archaeology-reproducible-articles-plot, fig.cap=str_glue('Articles in archaeology journals\nusing R for reproducible research (n = ', number_of_reproducible_articles,')')}
+```{r archaeology-reproducible-articles-plot, fig.cap=str_glue('Articles in archaeology journals using R for reproducible research (n = {number_of_reproducible_articles})')}
 
@@
 
-```{r citation-advantage, fig.cap=str_glue('Median citation rates per year for archaeology articles that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")})')}
+```{r citation-advantage}
 # archy papers that use code are more often cited?
@@
 
+archy_journal_articles_citations <- 
 archy_journal_articles %>%
@@
   summarise(median = median(TC)) %>%
+  filter(PY >= 2010)
+
+archy_journal_articles_n <- 
+archy_journal_articles %>%
+  mutate(TC = as.numeric(TC),
+         ID = ifelse(ID == "has_R_cites",
+                     "Cites R", "Doesn't cite R")) %>%
   filter(PY >= 2010) %>% 
-  ggplot(aes(ID,
+  select(ID, PY, TC)
+
+# compute t-test
+archy_journal_articles_citations_t_test <- 
+  t.test(  archy_journal_articles_n$TC ~  factor(archy_journal_articles_n$ID))
+
+tt <- t.test(archy_journal_articles_n$TC ~  factor(archy_journal_articles_n$ID))
+tvalue <- tt$statistic %>% formatC(digits = 2, format = "f")
+pvalue <- tt$p.value %>% formatC(digits = 5, format = "f")
+df <- round(tt$parameter, 0)
+
+dd <- archy_journal_articles_n %>% 
+  group_by(ID) %>% 
+  summarise(mean = mean(TC))
+dd_cites_mean <- dd %>% filter(ID == "Cites R") %>%  pull(mean) %>% round(1)
+dd_no_cites_mean <- dd %>% filter(ID == "Doesn't cite R") %>%  pull(mean) %>% round(1) 
+test_explanation <- 
+str_glue("On average, articles citing R \nhave significantly higher numbers\nof citations (m = {dd_cites_mean}) than articles\nthat do not (m =  {dd_no_cites_mean})\nt({df}) = {tvalue}, p =  {pvalue}")
+
+archy_journal_articles_citations_plot <- 
+  ggplot(archy_journal_articles_citations,
+         aes(ID,
              median)) +
@@
   facet_wrap( ~ PY) +
-  theme_minimal() +
+  theme_minimal(base_size = 16) +
   ylab("Median number of citations") +
@@
 
+```{r results='hide'}
+# add text to empty corner
+png(here::here("analysis", "figures", "archy_journal_articles_citations_plot_cap.png"),
+    width = 20, height = 15, units = "cm", res = 300)
+print(archy_journal_articles_citations_plot)
+library(grid)
+grid.text(test_explanation, 
+          x = 0.70, y = 0.25, 
+          gp=gpar(fontsize=11),
+          just = 'left')
+invisible(dev.off())
+```
+
+```{r  fig.cap=str_glue('Median citation rates per year for archaeology articles 2010-2017 that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")})')}
+knitr::include_graphics(here::here("analysis", "figures", "archy_journal_articles_citations_plot_cap.png"))
+```
+
 # Discussion
 
-In light of the discussion in the philosophy of science, whether paradigm-shifts are tool- or idea-driven, we offer the hypothesis that a tool-driven change of paradigms is underway in the biological sciences and, more subtly, in archaeology.
-The tool which drives this change is the use of scripting languages in computational approaches of these disciplines.
-We  show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visibility and influence (for some critical aspects on open scholarship see Huggett in this volume).
-This also is an indicator for the increase in popularity of R.
+Our hypothesis is that a tool-driven change is underway in the biological sciences, and is emerging in archaeology.
+The tool which drives this change is the use of scripting languages in computational work in these disciplines.
+We have shown that the use of R is on the rise among archaeologists using natural science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visibility and influence (see Huggett in this volume for critical discussion of open scholarship).
 
-The use of R or any other scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to share one's code.
+The use of R or similar scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to communicate one's workflow in a highly detailed an efficient way, and easily share this worksflow with others.
 The sharing of code has been identified in other disciplines as an amplifier of change in methods [@Touchon_McCoy_2016].
 It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
-Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with what we see described in the work of @Galison_1997 [-@Galison_1997].
-This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@Meltzer_1979; @Trigger_2006].
-
-Rather, this change enables archaeologists to more faithfully adhere to the fundamental tenet of the scientific process which demands the reproducibility of results.
-We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages.
-There is, for example, a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses.
-Archaeological research has always relied on the reuse of data [@Strupler_2017].
+Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with the patterns described by @Galison_1997 [-@Galison_1997].
+This change is transformational rather than revolutionary, enabling archaeologists to more faithfully adhere to a fundamental tenet of the scientific process which demands the reproducibility of results.
+We do not want to claim archaeologists did not want to produce reproducible research before the advance of scientific scripting languages.
+Indeed, there is a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses.
+Empirical reproducibility has always been prominent in archaeological research as we can see from the value placed on re-examining museum collections, re-excavating old sites, and reusing previously published data [@Strupler_2017].
 
@@
 Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses.
-By using and sharing code we can undo this drift, and return to a robust scrietific practice for archaeology.
-As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences.
-These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013].
-Also in archaeology, more and more funding agencies push for open science methods and transparency [@Strupler_2017].
+By using and sharing code we can undo this drift, and return to a robust scientific practice for archaeology.
+As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural sciences.
+These sciences also show a trend of increased emphasis on reproducibility, with journals changing their data and code availability policies, requiring code and data to be published alongside articles [@Stodden_2013].
+Also in archaeology, more and more funding agencies show interest in open science methods and transparency [@Strupler_2017].
 
-As @Galison_1997 describes a shift in the practices of physicists, so does archaeological practice change in the digital age.
+As @Galison_1997 describes a shift in the practices of physicists, we believe so is archaeological practice changing in the digital age.
 @Huvila_2018 have explored the use of digital tools on archaeological practices to understand the boundaries of what counts as archaeological, what is merely related to archaeology and what should be excluded from archaeology.
 Reflecting on @Huvila_2018, our view is the use use of code is unquestionably archaeology, since it is just one of many tools that archaeologists are using for converting raw data into summaries, visualisations, and insights.
-
-These concerns about how digital tools affect what we consider is and isn't archaeology intersect with the work of @gieryn1999cultural on boundary construction.
+These concerns about how digital tools affect what we consider is and is not archaeology intersect with the work of @gieryn1999cultural on boundary construction.
 In his cultural cartography of science, Gieryn examines boundaries between science generally and varieties of less authoritative non-science.
@@
 Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims by competitors.
-Boundary work involves attributions of selected characteristics to science in order to distinguish it from non-scientific competitors.
+Boundary work involves attributions of select characteristics to science in order to distinguish it from non-scientific competitors.
 The increasing voluntary use and sharing of code by researchers, and demands for it by readers and peer reviewers is a form of active boundary work that is shifting the boundary between scientific and non- or less-scientific archaeology.
-The authors, as members of this group, argue that the use of code as a research tool, and the sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work.
-The reason for this is that using code, and making it available for others to inspect, enhances the credibility of knowledge claims.
+We, as members of this group, argue that the use of code as a research tool, and the sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work.
+The reason for this is that using code, and making it available for others to inspect, enhances the credibility of knowledge claims and the visibility of those claims.
 
-On another level, we'd like to point to the _technicity_ of tools, in our case coding.
-Aud Sissel Hoel and Iris van der Tuin are drawing on Ernst Cassirer's „Form und Technik“ [@cassirer_form_1930] and Gilbert Simondon's "Du monde d'existence des objets techniques" [@simondon_mode_1958] to define technicity as a term to describe that technology, i.
-e.
-tools and machines respectively, interact with the human and with nature [@hoel_ontological_2013, 188].
-Technology is said to own an instrumental kind of logos (Cassierer, see @hoel_ontological_2013, 194) and in analogy to language it exercises a certain measure of agency in mediating between human, nature and technology.
-The use of a scripting language might exemplify this quite nicely, as it is not just a technological tool, but quite rightly named "language" as well.
+A related issue here is the _technicity_ of tools, in our case coding.
+We draw on @cassirer_form_1930 and @simondon_mode_1958 to define technicity as a term to describe the technology, i.e.
+tools and machines respectively, that interact with humans and with nature [@hoel_ontological_2013, 188].
+Technology is said to own an instrumental kind of logos (Cassierer, see @hoel_ontological_2013, 194) and in analogy to language it exercises a certain measure of agency in mediating between humans, nature and technology.
+The emphasis here is on the function of technology, in the forces that it exercises, with a certain measure of autonomy, on other beings as well as in and through the new realities, it brings into being.
+The use of a scripting language might exemplify this, as it is not just a technological tool, but a proper "language" as well.
+Scripts of code are texts that translate data into visualisations and statements about the world, transversing and revitalizing conceptions of the relationship between people and nature, and between people as members of a research community.
 
-For our paper the concept of technicity is interesting because in contrast to the concept of technological determinism, which puts technology as a force against the human, van der Tuin and Hoel (as well as Cassirer and Simondon) enfold technology into „the human“ without letting it lose its "foreignness" and further develop the notions of „human“ and „nature“ [@hoel_ontological_2013, 190].
-Any change in tools therefore will influence how human researchers interact with their objects of interest.
-Interestingly it is exactly at the point inbetween continously evolving subjects and objects, where van der Tuin and Hoel place technology and the creation of knowledge [@hoel_ontological_2013, 199].
+The concept of technicity is relevant here because in contrast to the concept of technological determinism, which puts technology as a force against the human, van der Tuin and Hoel (as well as Cassirer and Simondon) enfold technology into 'the human' without letting it lose its 'foreignness' and further develop the notions of 'human' and 'nature' [@hoel_ontological_2013, 190].
+Any change in tools therefore will influence how human researchers interact with their objects of interest and other people in their research community.
+For example, tools that make the research process more transparent and open should make the research community more open and inclusive to diverse participants and ideas.
+Using code to 'stand for' or represent the research process will increase our awareness of how archaeology itself 'stands for' the human experience in the past, and how it is a representation of the past [cf.
+@pearson1998beginning].
+Similarly, the code has to be successfully executed to generate a useful result, just as claims about the past have to be successfully engaged with archaeological evidence to count as plausible.
+Interestingly it is exactly at the point between continously evolving subjects and objects, where van der Tuin and Hoel place technology and the creation of knowledge [@hoel_ontological_2013, 199].
 The "productive back-and-forth" between all parties may challenge established notions of culture, but also offers a concept of formation through differing, called "performative correspondence" [@hoel_ontological_2013, 201].
+The potential of a certain technology cannot be known before its coming-into-being as it will reflect back on its creation ("concretization" in Simondon's terms) as well as interact with other technologies (developed or yet-to-be developed) [@hoel_ontological_2013, 194].
+Observing how scientific scripting languages have become a concrete tool in ecology leads us to expect predict similar trends in archaeology.
 
-Another point they make is, that the potential of a certain technology cannot be known before its coming-into-being as it will reflect back on its creation ("concretization" in Simondon's terms) as well as interact with other technologies (developed or yet-to-be developed) [@hoel_ontological_2013, 194].
-Seeing the influence the introduction of scientific scripting languages had on ecology we therefore can only expect similar trends to evolve in archaeology, but not predict these.
+We do not claim that a Kuhnian paradigm-change is happening in ecology or archaeology: Data- and method-sharing behaviours have been around in archaeology before now and do not present incommensurate ways of thinking about the subject in the way that a typical reading of Kuhn's paradigm change concept should.
+Indeed, social science disciplines such as archaeology may not even be easily susceptible to classical Kuhnian paradigm shifts.
+This is because high-power tests (i.e.
+statistical tests with high probabilities of rejecting false null hypotheses) capable of distinguishing between superior and inferior paradigms are rare in the social sciences, such as economics [@Ioannidis_Stanley_Doucouliagos_2017], compared to physical science.
+Modelling and case studies described by @Akerlof_Michaillat_2018 indicate that among discplines lacking evidence to routinely strongly discriminate between theories, the chances of getting trapped in an inferior paradigm are high.
+In this situation, true theories may not be adopted at all and the field is at risk of being captured by false paradigms.
 
-We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before now and does not present incommensurate ways of thinking about the subject in the way that a classical paradigm-change should.
-What we can see, though, is that the need for methods to share research in a way that is reproducible for others has been acknowledged and that archaeologists have looked for tools to make this possible.
-Using scripting languages and data repositories to share the scripts is a solution that we have demonstrated enjoys increasing popularity in the archaeological research community.
+Although a discipline-wide paradigm shift might be off the table for archaeology, what we can see, though, is that there is a need for methods to share research in a way that is reproducible for others, and that archaeologists are looking for tools to make this possible.
+Using scripting languages and data repositories to share the scripts is a solution that enjoys increasing popularity in the archaeological research community.
 It has also expanded the 'trading zone' [@Galison_1997] where archaeologists collaborate with specialists in other fields to develop a shared language to get things done.
@@
 We can also see archaeologists in similar trading zones when using GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions.
-Archaeologists' embrace of, or turn towards, these methods could also be argued to be tool-driven revolutions.
-However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is an universal method with which we can do all, or which is central to any, of those previously-mentioned approaches.
+Archaeologists' embrace of, or turn towards, these methods may also be tool-driven revolutions.
+However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is an universal method with which we can do all, or which can be central to any, of those previously-mentioned approaches.
 That is why we believe archaeologists should pay attention to scripted workflows regardless of their usual toolkit and methodological interests.
@@
 As @Chrysanthi_2012 illustrate, it requires a certain familiarity with the computational tools to be able to use them not just on a technological basis, but creatively and in full knowledge of their restrictions and ambiguities.
-The awareness and the appropriate handling of shortcomings and contexts of the data used in an analysis is also a time- and thought-consuming effort, which needs to stand at the beginning of each archaeological research (and which may be exacerbated by using multiple data sets from varying sources -- see e.
-g.
-@Hugget_2015).
-We will not focus on data sharing here, though, but on the sharing of code.
+The awareness and the appropriate handling of shortcomings and contexts of the data used in an analysis is also a time- and thought-consuming effort.
+This efforts is required at the beginning of each archaeological research, and which may be exacerbated by using multiple data sets from varying sources [cf.
+@Hugget_2015].
 
 As with data (see Fitzpatrick in this volume), sharing code in public takes a little time to ensure that it is fit for others to read.
-Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error, rather than following a model or widely-used approach.
-This suggests to us that there is some scope to improve the ease and efficiency of code-use in archaeology by studying the exsiting practice and providing a model that others can easily adapt to their own research projects.
+Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error.
+This situation of guided variation (learning by trial and error) suggests that using code is not yet widespread enough among archaeologists to propagate due to frequency-dependant biases.
+It also suggests that the professional benefits (i.e.
+citation effects) are not yet widely known for code use behaviours to propagate due to content biases, and finally that there are not yet enough highly-visible prestigous researchers using code for this behaviour due to propagate due to model-based biases [@Rendell_et_al_2011; @boyd1988culture; @henrich2007dual].
+This suggests to us that there is unrealised potential to improve the efficiency of using and code in archaeology by converging on some widely-agreed upon conventions that save time and investing effort in communicating the professional benefits of using code.
 
-The result of our analysis of existing practices among archaeologists and other researchers, and our study of best practices in scientific computing (described in more detail in @marwick2018packaging) is the R package rrtools ('reproducible research tools', https://github.com/benmarwick/rrtools).
-The goal of rrtools ist to make it easier for archaeologists and researchers in other fields to use R for research and publication.
+Our contribution to improving the efficiency of using and sharing code is the R package rrtools ('reproducible research tools', https://github.com/benmarwick/rrtools).
+This package is the result of our analysis of existing practices among archaeologists and other researchers, and our study of best practices and current conventions in scientific computing, described in more detail in @marwick2018packaging.
+The goal of rrtools is to make it easier for archaeologists and other researchers to use R for research and publication.
 This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort.
 The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research.
-First, we keep the data separate from the methods.
+For example, we keep the data separate from the methods.
 This means keep the code in a separate directory to the data.
-Second, we, keep the raw data seprate from data that is created or derived during the analysis.These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.
+We keep the raw data (i.e.
+field-collected and instrument output) seprate from data that is created or derived during the analysis.
+These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.
+At the heart of the project template provided by rrtools is an RMarkdown document that is where the report or journal manuscript is written.
+RMarkdown is a document format that allows the author to combine plain text, R code, and automation of all the usual details of scholarly writing such as citations, captions, and cross-references to tables and figures [@xie2018r].
+The unique concept of RMarkdown is that the R code contained in the document will generate the statistical figures and tables when the document is executed (or rendered) into output such as PDF, MS Word or an HTML file.
 
 The rrtools package contains several functions for the R user to enable a researcher to quickly set up a compendium suitable for writing reproducible research in R.
-Each of these steps reflects a best practice that has been previously articulated in other feilds.
+Each of these steps reflects a best practice that has been previously articulated in other fields [e.g.
+@Wilson_et_al_2017].
 We will focus here on the first five, as these are most relevant to most researchers:
 
-- **rrtools::use_compendium:**  This function creates a new directory for the research project, and creates an R package within this directory.
-We use the R package structure as the standard for our directory because it is widely recognised by R users.
+- **use_compendium:**  This function creates a new directory for the research project, and creates an R package within this directory.
+We use the R package structure as the standard for our directory because it is widely recognised by R users [@wickham2015r].
 This means that people know where to look for code, where to look for data, and where to look for dependancies and other information.
-- **usethis::use_mit_license:** This function creates a copy of the MIT software lisence to the compendium.
+- **use_mit_license:** This function creates a copy of the MIT software license to the compendium.
 This is important for two reasons.
@@
 This is important for setting the expectations about the relationship between the author and the reader, regarding the use of code.
-- **usethis::use_github:**  This will initiate the use of the Git version control system.
-This is the current state-of-the-art for version control of any kind of plain text file, including text, code, data, and images.
+- **use_github:**  This will initiate the use of the Git version control system.
+This is the current state-of-the-art for version control of any kind of plain text file, including text, code, data, and images [@jones2013git].
 This will also create a repository on the GitHub website.
@@
 The GitHub repository can be kept private if desired, until the work is published.
-- **rrtools::use_readme_rmd:**  This will create a simple document that is usually the first thing a reader sees when they browse the code files.
-This document is important because it helps to describe to the reader what to expect in the compendium, and it gives details of how to cite the compendium.
-- **rrtools::use_analysis:**  This function creates a set of folders according to the best practices for organising a typical reasearch project.
+- **use_readme_rmd:**  This will create a simple document that is usually the first thing a reader sees when they browse the code files.
+This document is important because it helps to describe to the reader what to expect in the compendium, and it gives details of how to cite the compendium and engage with it if they want to make changes (e.g.
+if a reader finds a mistake that needs to be fixed).
+- **use_analysis:**  This function creates a set of folders according to the best practices for organising a typical research project [@Wilson_et_al_2017; @noble2009quick].
 We recognise that this file structure wont be perfect for every project.
-But we know that it will suit many projects quite well, and it will at least make the user think about how to organise their files logically, and according to best practises.
+But our observations show that it is in wide use, and will suit many projects quite well.
+It will at least make the user think about how to organise their project logically, and to make it easier for others to navigate.
 
@@
 
-Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [@Kuhn_1962], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@Galison_1997], who focuses on the material culture of a science and its tools.
-A change in the paradigms of a discipline may also only occur, if tools to implement this change are available.
-The change we are aiming for is the move to a reproducible computational archaeology, that adheres to the scientific method.
-The tool which we have been discussing here is R, as an example for scientific scripting languages.
+Starting with the discussion of Kuhn's influential work on paradigm shifts in science [@Kuhn_1962], we take inspiration from @Galison_1997 to propose a another explanation of how changes occurs in research communities, emphasising the transformative role of tools.
+The change we focus on is the use of scripts to document and communicate data analysis.
+This change will increase the computational reproducibility of archaeology, helping the discipline adhere more closely to the core values of science.
+The tool with the highest potential to achieve this currently is the R programming language.
+However, we realise that this will change in time, and in the future newer technologies will replace R.
+The key point here is that this transformative tool is any open source scripting language.
 
-We have shown a bibliometric analysis that reveals substantial change in the use of the R programming languages in ecology.
-We saw that ecological and evolutionary sciences have strongly embraced R in their publications.
+We have presented a bibliometric analysis that reveals substantial change in the use of the R programming languages in ecology.
+Citations patterns show that ecological and evolutionary sciences have strongly embraced R in their publications.
 Archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields, and most commonly in more scientific aspects of archaeology.
-We also identified a small sub-set of archaeology papers that use R to make their work fully reproducible and could identify an advantage of citing R for researchers as it leads to being referenced more often.
+We identifed an advantage of citing R for researchers as it leads to being referenced more often.
+We also identified a small, but increasing, set of archaeology papers that use R to make their work fully reproducible.
+In reflecting on these results, we have noted the technicity of coding in how it mediates between humans, nature and technology.
+We've proposed that the increasing adoption of coding among archaeologists is a form of active boundary work, shifting the boundary between scientific and unscientific archaeology, and is developing a trading zone between archaeologists and programming specialists such as computer scientists and software engineers as they search for useful tools.
+Many useful tools have already emerged from this trading zone, including our R package, rrtools, designed to make it easy for researchers to make their research reproducible.
+We hope this will help to realise some of the potential to make archaeological research more open and reproducible, both within the research community and to the public, as well as speeding the transfer of new results and methods throughout the research community without barriers due to access to resources, and also in this way supporting the practice of archaeology that is more faithful to the core values of science.
 
-There is great potential for archaeologists to make their work reproducible using R and related tools, and we think this is a much needed shift in computational archaeology - linking the digital and computational times to  older practices of sharing data and methods.
-We therefore presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible.
-We hope this will help to realise some of the potential to make computational archaeological research more open and reproducible, speed the transfer of new results and methods throughout the research community without barriers due to resource availabilty, and also in this way encourage researchers to adhere to the scientific process.
+We have outlined a positive future for archaeology where research publications are accompanied by code and data, enabling a 'critical self-consciousness' at a community-wide scale, similar to that imagined by @clarke1973archaeology as part of the discipline's 'loss of innocence'.
+A charactertic of this stage of disciplinary maturity is 'a closer understanding of its internal structure' [@clarke1973archaeology, p.
+7], and we can think of no more efficient and intelligible way to communicate the internals of archaeological analyses to other members of the discipline than by encoding the assumptions, logic, and calculations in scripts of an open source programming language.
+If we accept that that is necessary for the future of the discipline, we must confront two implications.
+First is that there is currently a shortage of incentives to motivate use and sharing of code.
+Our observation is that many researchers will not change unless required to by gate-keepers at prestigious journals and funding sources.
+In other research communities we see people in these gate-keeping roles (e.g.
+in over 5,000 journals and professional organizations) effecting change by adopting the Transparency and Openness Promotion Guidelines that establish community standards of code and data availability that help to align scientific ideals with practices and disciplinary norms [@Nosek_et_al_2015].
+Simple forms of signalling, such as badges on papers to indicate that the authors have made code and data available, may also help to shift norms and raise awareness of these desirable behaviours [@Kidwell_et_al_2016].
+
+However, cultural change on an community-wide scale, such as a shift to the wide-spread use of code among archaeologists, is most likely to happen via the slow enculturation of people as they internalize norms during their formative years, and not via contemporaneous exposure to external cultural influences [@Vaisey_Lizardo_2016].
+This means it is unlikely that a professional archaeologist reading this paper, or ones like it, will be motivated to use code if they were not trained to work with code at the start of their career.
+The implication here is that the most effective way to stimulate this change is during the formative years of professional training.
+This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, whereas the horizontal bar is the bredth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
+A gamma-shaped researcher has expert-level depth in archaeology, and is versed and proficient in other domains the provide skills such as analysing data with a programming language.
+A gamma-shaped researcher with their MA or PhD in archaeology may also be conversant in the language and culture of computer science enough to have conversations and collaborate in the trading zone, but does not need to be an expert in computer science to advance that discipline.
 
@@
 # if your Rmd is not in analysis/paper/
-git2r::repository("../..")
+git2r::repository(here::here())
 ```

### 2019-01-01 analysis/paper/jcaa-paper.Rmd
first submission to the JCAA website

@@
 author:
-  - Sophie C. Schmidt
-  - Ben Marwick
+  - author 1
+  - author 2
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on pointing-and-clicking with a mouse in commercial software, to a new set of tools based on writing scripts in free and open source programming languages.
-We review bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
-We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches."
+We present bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
+We identify a citation effect for papers that use code.
+We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of code."
 ---

### 2019-03-14 analysis/paper/jcaa-paper.Rmd

@@
 csl: journal-of-computer-applications-in-archaeology.csl
-abstract: "There is an argument in philosophy of science that revolutions in science are either ideas-driven or tool-driven.
+abstract: "There is an argument in philosophy of science that revolutions in science are either ideas-driven or tools-driven.
 We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
-One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on pointing-and-clicking with a mouse in commercial software, to a new set of tools based on writing scripts in free and open source programming languages.
+One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on point-and-clicking with a mouse in closed source software, to a new set of tools based on writing scripts in programming languages and making them free and open source.
 We present bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
-We identify a citation effect for papers that use code.
-We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of code."
+We identify a citation effect for papers that use program source code.
+We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of such code."
 ---

### 2019-08-05 analysis/paper/jcaa-paper.Rmd
working on reviewer A's comments

@@
 We present bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
-We identify a citation effect for papers that use program source code.
-We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of such code."
+We identify a citation effect for papers that use the open source programming language R.
+We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of R code."
 ---
@@
 In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method and whether this might be the beginning of a broader change.
-Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are ideas-driven or tool-driven [@Galison_1997; @Dyson_2000].
-In this light we explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of their research.
+Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are ideas-driven or tools-driven [@Galison_1997; @Dyson_2000].
+In this light we explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of research.
 We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of these scripts.
+We focus on R because our prior observations indicate it is by far the most common scripting language used by ecologists and archaeologists.
 Bibliometric evidence shows a strong increase in the use of R among ecologists and the beginning of a similar development in archaeology.
@@
 
-Philosophers of science disagree if revolutions in science are ideas-driven or tool-driven.
+Philosophers of science disagree whether revolutions in science are ideas-driven or tool-driven.
 One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] effort to describe the history of science by proposing paradigm shifts resulting from revolutionary change in communities of researchers.
@@
 This is a concern for researchers that might not be motivated by a desire to adhere to the abstract basic principles of what it means to practice science with integrity.
-If code sharing has concrete career advantages then we might expect it to be more likely to be part of a tool-drive revolution.
+If code sharing has concrete career advantages then we might expect it to be more likely to be part of a tool-driven revolution.
 @Vandewalle2012 analysed citation rates for 645 articles published in _IEEE Transactions on Image Processing_ during 2004–2006.
 He found statistically significant differences between the median number of citations for papers that have code available and papers that do not have code available.
+In considering causal relationships, Vandewalle raises the possibility of self-selection bias, where authors include code more often with their best papers.
+He also notes that for some papers, the code was made available by researchers other than the authors, after the publication has become popular.
+
 In Figure \@ref(fig:citation-advantage) we conducted a similar analysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
@@
 We see that archaeological articles that cite R are consistently more highly cited themselves than articles that do not cite R.
+The mechanisms underlying the citation effects of using R remain to be elucidated.
+The articles in our data are mostly not open access (only 20% are open access), so we cannot attribute the citation effect to open access effects, such as easier availability, and additional time in press of the articles [@Kurtz_Eichhorn_Accomazzi_Grant_Demleitner_Henneken_Murray_2005].
+In our data all the article authors are also the code authors, so there is no popularity effect noted by Vandewalle.
+One possibilty is that authors preferentially tend to promote their articles containing code, beacause they want a greater return for the greater effort they have put into including the code with the article, this is equivalent to the self-selection bias postulated by Vandewalle.
+Another possiblity is that papers with code are perceived by readers to be more robust because of their transparency, and thus a more reliable item to cite in support of claims made by other authods.
+Another possible cause is that when a paper is accompanied by code (for example on a GitHub repository) it is more visible than a paper without code, and appears higher in search result rankings.
+All other factors equal, higher visibility leads to more citations because the reseracher will stop searching when theyve found the first paper that serves their purpose.
+Further work is necessary to understand the mechanisms of this citation effect, additional insights might come from comparisions of papers that use another open source language such as Python, and MATLAB, which is not.
 

### 2019-08-06 analysis/paper/jcaa-paper.Rmd
adding information regarding trustworthyness of repositories (rev. A asked for this).

@@
 We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository such as Zenodo, Open Science Framework, Figshare, etc.
+Data repositories should be hosted by a non-commercial share holder, offer a persistent identifier and long term data management strategies to be regarded as "trustworthy"   (see @marwick_standard_2018 for a more thorough investigation of data sharing practices).
 Our continuously updated list of articles in this sample is available online here: https://github.com/benmarwick/ctv-archaeology.
@@
 Using and sharing open source code is thus an important action for reducing inequality in the archaeological research community.
-This is because it enables participation by researchers who are unable to buy software licenses or access face-to-face training.
+It enables participation by researchers who are unable to buy software licenses or access face-to-face training.
 

### 2019-08-06 analysis/paper/jcaa-paper.Rmd
typos noticed by Reviewer A

@@
 @Touchon_McCoy_2016 identify four factors relating to technological change that might explain the changes they observed in the use of statistics in ecology.
-First, they note that automated data loggers, GPS trackers, remote sensing, and crowd sourcing have greatly increased rate at which ecologists collect data.
+First, they note that automated data loggers, GPS trackers, remote sensing, and crowd sourcing have greatly increased the rate at which ecologists collect data.
 Second, increases in desktop and cluster computing power has made complex analytical processes faster and more convenient to compute.
@@
 The awareness and the appropriate handling of shortcomings and contexts of the data used in an analysis is also a time- and thought-consuming effort.
-This efforts is required at the beginning of each archaeological research, and which may be exacerbated by using multiple data sets from varying sources [cf.
+This effort is required at the beginning of each archaeological research, and which may be exacerbated by using multiple data sets from varying sources [cf.
 @Hugget_2015].

### 2019-08-06 analysis/paper/jcaa-paper.Rmd
stash

--- 2019-08-05 analysis/paper/jcaa-paper.Rmd
@@
 We present bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
-We identify a citation effect for papers that use the open source programming language R.
-We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of R code."
+We identify a citation effect for papers that use program source code.
+We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of such code."
 ---
@@
 In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method and whether this might be the beginning of a broader change.
-Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are ideas-driven or tools-driven [@Galison_1997; @Dyson_2000].
-In this light we explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of research.
+Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are ideas-driven or tool-driven [@Galison_1997; @Dyson_2000].
+In this light we explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of their research.
 We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of these scripts.
-We focus on R because our prior observations indicate it is by far the most common scripting language used by ecologists and archaeologists.
 Bibliometric evidence shows a strong increase in the use of R among ecologists and the beginning of a similar development in archaeology.
@@
 
-Philosophers of science disagree whether revolutions in science are ideas-driven or tool-driven.
+Philosophers of science disagree if revolutions in science are ideas-driven or tool-driven.
 One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] effort to describe the history of science by proposing paradigm shifts resulting from revolutionary change in communities of researchers.
@@
 What they share in common is working with the field-collected data on a computer to prepare it for publication.
-R is a widely used free and open source data analysis tool in man research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018] and therefore a good proxy for the adoption of open source scientific programming languages in a discipline.
+R is a widely used free and open source data analysis tool in many research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018] and therefore a good proxy for the adoption of open source scientific programming languages in a discipline.
 To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.
@@
 This is a concern for researchers that might not be motivated by a desire to adhere to the abstract basic principles of what it means to practice science with integrity.
-If code sharing has concrete career advantages then we might expect it to be more likely to be part of a tool-driven revolution.
+If code sharing has concrete career advantages then we might expect it to be more likely to be part of a tool-drive revolution.
 @Vandewalle2012 analysed citation rates for 645 articles published in _IEEE Transactions on Image Processing_ during 2004–2006.
 He found statistically significant differences between the median number of citations for papers that have code available and papers that do not have code available.
-In considering causal relationships, Vandewalle raises the possibility of self-selection bias, where authors include code more often with their best papers.
-He also notes that for some papers, the code was made available by researchers other than the authors, after the publication has become popular.
-
 In Figure \@ref(fig:citation-advantage) we conducted a similar analysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
@@
 We see that archaeological articles that cite R are consistently more highly cited themselves than articles that do not cite R.
-The mechanisms underlying the citation effects of using R remain to be elucidated.
-The articles in our data are mostly not open access (only 20% are open access), so we cannot attribute the citation effect to open access effects, such as easier availability, and additional time in press of the articles [@Kurtz_Eichhorn_Accomazzi_Grant_Demleitner_Henneken_Murray_2005].
-In our data all the article authors are also the code authors, so there is no popularity effect noted by Vandewalle.
-One possibilty is that authors preferentially tend to promote their articles containing code, beacause they want a greater return for the greater effort they have put into including the code with the article, this is equivalent to the self-selection bias postulated by Vandewalle.
-Another possiblity is that papers with code are perceived by readers to be more robust because of their transparency, and thus a more reliable item to cite in support of claims made by other authods.
-Another possible cause is that when a paper is accompanied by code (for example on a GitHub repository) it is more visible than a paper without code, and appears higher in search result rankings.
-All other factors equal, higher visibility leads to more citations because the reseracher will stop searching when theyve found the first paper that serves their purpose.
-Further work is necessary to understand the mechanisms of this citation effect, additional insights might come from comparisions of papers that use another open source language such as Python, and MATLAB, which is not.
 

### 2019-08-06 analysis/data/raw_data/top_journals_for_cran_cites_articles_per_year.R

@@ -1,124 +1,0 @@
-# total number of articles per year
-# copy-paste from WoS year filter
-top_journals_for_cran_cites_articles_per_year <-
-data_frame(year = 2007:2017,
-           total_number_articles_PLOSONE = c(1230,
-                                             2717,
-                                             4404,
-                                             6729,
-                                             13786,
-                                             23456,
-                                             31504,
-                                             31482,
-                                             29807,
-                                             23039,
-                                             21076),
-           total_number_articles_SCIENTIFICREPORTS = c(0,
-                                                       0,
-                                                       0,
-                                                       0,
-                                                       205,
-                                                       804,
-                                                       2554,
-                                                       4027,
-                                                       10947,
-                                                       21046,
-                                                       24789
-                                                       ),
-           total_number_articles_ECOLOGYANDEVOLUTION = c(0,
-                                                         0,
-                                                         0,
-                                                         0,
-                                                         50,
-                                                         262,
-                                                         430,
-                                                         391,
-                                                         497,
-                                                         730,
-                                                         981),
-           total_number_articles_PEERJ = c(0,
-                                           0,
-                                           0,
-                                           0,
-                                           0,
-                                           0,
-                                           232,
-                                           471,
-                                           799,
-                                           1294,
-                                           1370),
-            total_number_articles_FORESTECOLOGYANDMANAGEMENT = c(442,
-                                                                 707,
-                                                                 612,
-                                                                 475,
-                                                                 482,
-                                                                 526,
-                                                                 656,
-                                                                 556,
-                                                                 455,
-                                                                 606,
-                                                                 611),
-            total_number_articles_JOURNALOFSTATISTICALSOFTWARE =
-             c(63,
-               43,
-               42,
-               60,
-               102,
-               86,
-               55,
-               90,
-               89,
-               83,
-               92),
-           total_number_articles_MOLECULARECOLOGY =
-             c(430,
-               440,
-               428,
-               473,
-               427,
-               476,
-               469,
-               468,
-               446,
-               433,
-               500),
-           total_number_articles_PROCEEDINGSOFTHEROYALSOCIETYBBIOLOGICALSCIENCES =
-             c(413,
-               360,
-               540,
-               463,
-               495,
-               640,
-               547,
-               667,
-               636,
-               575,
-               591),
-            total_number_articles_ECOSPHERE = c(0,
-                                                0,
[…]

### 2019-08-06 analysis/paper/jcaa-paper.Rmd

--- 2019-08-06 analysis/paper/jcaa-paper.Rmd
@@
 We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository such as Zenodo, Open Science Framework, Figshare, etc.
-Data repositories should be hosted by a non-commercial share holder, offer a persistent identifier and long term data management strategies to be regarded as "trustworthy"   (see @marwick_standard_2018 for a more thorough investigation of data sharing practices).
 Our continuously updated list of articles in this sample is available online here: https://github.com/benmarwick/ctv-archaeology.
@@
 Using and sharing open source code is thus an important action for reducing inequality in the archaeological research community.
-It enables participation by researchers who are unable to buy software licenses or access face-to-face training.
+This is because it enables participation by researchers who are unable to buy software licenses or access face-to-face training.
 

### 2019-08-06 analysis/paper/jcaa-paper.Rmd
working on reviewer B's edits

--- 2019-08-06 analysis/paper/jcaa-paper.Rmd
@@
 We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
-One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on point-and-clicking with a mouse in closed source software, to a new set of tools based on writing scripts in programming languages and making them free and open source.
-We present bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process.
-We identify a citation effect for papers that use the open source programming language R.
+One of the most profound changes driven by this concern for reproducibility and transparency is from using data analysis tools dependent on point-and-clicking with a mouse in closed source software, to a set of tools based on writing scripts in open source programming languages and making the scripts openly available.
+We present bibliometric evidence for this change in ecolpogy and archaeology to test if the adoption of these new tools may be considered a revolution.
+We identify a citation effect for archaeology articles that use the open source programming language R.
 We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of R code."
@@
 
-In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method and whether this might be the beginning of a broader change.
+In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method.
+We ask whether this might be the beginning of a broader change in the field.
 Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are ideas-driven or tools-driven [@Galison_1997; @Dyson_2000].
-In this light we explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of research.
-We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of these scripts.
+This argument motivated us to explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of research.
+We focus on one set of tools for this change, based on writing data analysis scripts in free and open source programming languages, exemplified by R, and the practice of sharing of these scripts.
 We focus on R because our prior observations indicate it is by far the most common scripting language used by ecologists and archaeologists.
@@
 One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] effort to describe the history of science by proposing paradigm shifts resulting from revolutionary change in communities of researchers.
-In brief, this model proposes that the history of science consists of long periods of tradition-bound 'normal science' punctuated by short episodes of revolution.
+In brief, this model proposes that the history of science consists of long periods of tradition-bound 'normal science' punctuated by short episodes of 'revolution'.
 Normal science was described by Kuhn as stagnant, routine day-to-day research focused on 'puzzle solving'.
@@
 A notable criticism is that Kuhn's normal science is unscientific, a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
-This is at odds with a view of science as a continuous state of revolution, with researchers simultaneously employing many styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
+This is at odds with a view of science as a continuous state of evolution, with researchers simultaneously employing many styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
 In this view, based on the post-World War II era of science, radical discontinuities are rare, and major developments instead emerge from the division and recombination of already mature fields [@andersen2013second].
 Given the rarity of radical discontinuities, some biologists have argued that change in biology is better characterised by new ideas (e.g.
-Mendelian heredity, Darwinian evolution, and molecular genetics) replacing not a former paradigm but a conceptual vacuum [@wilkins1996are].
+Mendelian heredity, Darwinian evolution, and molecular genetics) replacing not a former paradigm, but a conceptual vacuum [@wilkins1996are].
 
 In archaeology we have seen claims for paradigm shifts in several areas.
-The most wide-ranging application is Clark's [-@Clark_1993] comparison of North American archaeology to European archaeology during the 1970s-90s.
-Clark, similar to Binford and Sabloff [-@binford1982paradigms], identified the North American archaeological paradigm as rooted in anthropology, founded on post-positivist biases, and distinctive with its critical-realist ontology, a modified-objectivist epistemology, and an experimental-manipulative method.
+The most wide-ranging application is Clark's [-@Clark_1993] comparison of North American archaeology to European archaeology during the 1970s-1990s.
+Clark, similarly to Binford and Sabloff [-@binford1982paradigms], identified the North American archaeological paradigm as rooted in anthropology, founded on post-positivist biases, and distinctive with its critical-realist ontology, a modified-objectivist epistemology, and an experimental-manipulative method.
 In contrast, the European archaeological paradigm is described a kind of history, mostly in a strict empiricist tradition, and typified by a realist ontology, an inductivist epistemology, and by an observational method.
@@
 
-A more thorough treatment of the claim that processual archaeology was a paradigm shift is provided by @Meltzer_1979.
+A more thorough treatment of the claim that the appearance of processual archaeology was a paradigm shift is provided by @Meltzer_1979.
 Meltzer reviews literature arguing for a revolutionary change in archaeology during the 1960s and 1970s, and considers these claims in light of classic Kuhnian revolutionary events such as the replacement of the Ptolemaic system by the Copernican system, and Newtonian dynamics by the new Einsteinian dynamics.
@@
 
-We also see mentions of paradigm shifts in archaeology in reference to processual to post-processual archaeology [@Härke_2002; @Koerner_2018], as well as more localised and specific shifts in archaeological practice and thinking.
+We also see mentions of paradigm shifts in archaeology in reference to the shift from processual to post-processual archaeology [@Härke_2002; @Koerner_2018], as well as more localised and specific shifts in archaeological practice and thinking.
 For example, @Snodgrass_2002 claims that among Classical Archaeologists, an increase in research on previously neglected periods of antiquity, such as the Greek Early Iron Age, constitutes a paradigm shift.
-@McAnany_Rowe_2015 propose that the appearance of community-based participatory models of research among some communities of archaeologists is a paradigm shift, though they concede that it is more transformational rather than revolutionary.
-@harris2012interfacing has similarly argued that community crowdsourced geographic knowledge (or volunteered geographic information) could be paradigm shift for archaeological communities.
+@McAnany_Rowe_2015 propose that the appearance of community-based participatory models of research among some communities of archaeologists is a paradigm shift, though they concede that it is transformational rather than revolutionary.
+@harris2012interfacing similarly argues that community crowdsourced geographic knowledge (or volunteered geographic information) could be paradigm shift for archaeological communities.
 Archaeologists using LIDAR (light detection and ranging) for remote geospatial imaging of cultural landscapes have claimed the application of this technology to archaeology is a paradigm shift [@chase2012geospatial; @howey2016detecting].
-@Fuller_2010 has used the term to describe a shift in thinking about the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to new concept of the emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
-Most of these paradigms differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, both as a tool to understand the history of science and as a guide to changing it, which is the multiple definitions of a paradigm in Kuhn's writings.
+@Fuller_2010 has used the term to describe a shift in thinking from the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to its emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
+Most of these claims for paradigm shifts differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, both as a tool to understand the history of science and as a guide to changing it, which is the multiple definitions of a paradigm in Kuhn's writings.
 
@@
 
+archy_cran_cites_all_areas_top_journals_by_year_to_plot <- 
+ archy_cran_cites_all_areas_top_journals_by_year # %>% 
+ # filter(SO != "AFRICAN ARCHAEOLOGICAL REVIEW")
+
 all_archaeology_citing_R_over_time <-
-ggplot(archy_cran_cites_all_areas_top_journals_by_year,
+  ggplot(archy_cran_cites_all_areas_top_journals_by_year_to_plot,
        aes(PY,
@@
   geom_text_repel(
-    data = archy_cran_cites_all_areas_top_journals_by_year %>%
+    data = archy_cran_cites_all_areas_top_journals_by_year_to_plot %>%
       group_by(SO) %>%
@@
     direction = "y",
-    size = 5,
+    size = 6,
     nudge_x = 1,
     hjust = 0) +
-  theme_minimal() +
-  theme(legend.position="none",
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(), 
+         axis.line = element_line(colour = "black"),
+         axis.title.x = element_text(hjust = 0.125),
+        legend.position="none",
         axis.text.x = element_text(angle = 90,
@@
                                    vjust = 0.5)) +
-  scale_x_continuous(breaks = min_y:max_y,
-                     labels = c(min_y:2017,
-                                rep("",
-                                    length(2018:max_y))),
+  scale_x_continuous(breaks = min_y:2017,
                      limits = c(min_y, max_y)) +
@@
   xlab("Publication year") +
-  ylab("Percentage of articles in that journal")
+  ylab("Percentage of articles in that journal") +
+  lemon::coord_capped_cart(bottom='right')
 
@@
                parse = TRUE,
-               size = 3,
-               label.y.npc = 0.7) +
-  stat_fit_glance(label.y.npc = 0.7,
-                  size = 3,
-                  method = "lm",
-                  method.args = list(formula = y ~ x),
-                  geom = "text",
-                  aes(label = paste("p-value: ",
-                                    signif(..p.value..,
-                                           digits = 4)))) +
+              label.y = 0.9) +
+ stat_fit_glance(method = "lm", 
+                  label.y = 0.8,
+                  method.args = list(formula = y~x),
+                  aes(label = paste("italic(P)*\"-value = \"*", 
+                                    signif(..p.value.., digits = 4), sep = "")),
+                  parse = TRUE) +
   xlab("year") +
@@
                            italic("Journal of Archaeological Science")))) +
-  theme_bw(base_size = 6)
+  theme_bw(base_size = 10)
 ```
@@
   annotation_custom(grob = ggplotGrob(jas_cites_r),
-                          xmin = 2030, xmax = 2040,
-                          ymin = 0.10, ymax = 0.17)
+                          xmin = 2027, xmax = 2040,
+                          ymin = 0.10, ymax = 0.20)
 
@@
 We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository such as Zenodo, Open Science Framework, Figshare, etc.
-Data repositories should be hosted by a non-commercial share holder, offer a persistent identifier and long term data management strategies to be regarded as "trustworthy"   (see @marwick_standard_2018 for a more thorough investigation of data sharing practices).
+'Trustworthy' data repositories are those hosted by a non-commercial organisation, offering persistent identifiers, and have transparent commmitments to long term data management and backup strategies (see @marwick_standard_2018 for more details).
 Our continuously updated list of articles in this sample is available online here: https://github.com/benmarwick/ctv-archaeology.
@@
 In our view, merely using R (or similar open source programming languages) to conduct research would not constitute a tool-driven revolution by itself, but publicly sharing the code used for research, as a solution to problems of irreproducibility, is more likely to enable revolutionary change.
-Openly shared code speeds the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
+We expect that openly shared code will speed the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
 As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as Zenodo, Open Science Framework, Figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.

### 2019-08-07 analysis/paper/jcaa-paper.Rmd
up to L408 of Reviewer B's edits on the submitted ms.

@@
 @Fuller_2010 has used the term to describe a shift in thinking from the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to its emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
-Most of these claims for paradigm shifts differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, both as a tool to understand the history of science and as a guide to changing it, which is the multiple definitions of a paradigm in Kuhn's writings.
+Most of these claims for paradigm shifts differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, namely the multiple definitions of a paradigm in Kuhn's writings.
 
@@
 
-We propose that Galison's emphasis on tool-driven change, as a complement to Kuhnian concept-driven change, has much potential for archaeology.
-The centrality of material culture in Galison's view raises the question about whether there is scope for archaeologists to make productive contributions to understanding change in the history science through an archaeological analysis of scientific instrumentation [cf.
+We propose that Galison's emphasis on tool-driven change, as a complement to Kuhnian concept-driven change, has potential to enhance archaeology's social and scientific relevance and contributions.
+The centrality of material culture in Galison's view raises the question about whether there is scope for archaeologists to make productive contributions to understanding change in the history of science through an archaeological analysis of scientific instrumentation [cf.
 @Schiffer_2013].
@@
 @chase2012geospatial have claimed the introduction of radiocarbon dating was a paradigm shift in archaeology, but we are not aware of a Galisonian analysis of this change.
-A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to predict or generate change in archaeology?
+A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to predict or generate change in the practice of archaeology?
 
@@
 
-The question of predicting or generating change in scientific practice via adoption of new tools is intriguing, and in this section we explore how a change observed in ecology might be relevant for understanding or even directing the future of archaeology.
+In this section we explore how a change observed in ecology might be relevant for understanding or even directing the future of archaeology.
 Biological disciplines have a long tradition of influencing archaeological thought, starting with Oscar Montelius, whose typological method was inspired by Darwin's theory of evolution [@Montelius_1899, 237].
-This tradition continues, for example, many spatial statistics in current use in archaeology are derived from ecology [@keron_use_2015, 7].
+This tradition continues, for example, many spatial statistics currently used by archaeologists are derived from ecology [@keron_use_2015, 7].
 
@@
 First, they note that automated data loggers, GPS trackers, remote sensing, and crowd sourcing have greatly increased the rate at which ecologists collect data.
-Second, increases in desktop and cluster computing power has made complex analytical processes faster and more convenient to compute.
+Second, increases in desktop and cluster computing power have made complex analytical processes faster and more convenient to compute.
 Third, the development of free, open source and easily extensible software for data analysis and visualisation, such as R, allows new methods to spread quickly via online fora and social media.
-Fourth, the publication of several books and papers that have strongly influenced the way many ecologists think about data analysis, such as @burnham2003model's book on model selection and inference.
+Fourth, the publication of several books and papers that have strongly influenced the way many ecologists think about data analysis, such as Burnham and Anderson's [-@burnham2003model] book on model selection and inference.
 
 We focus here on the third factor, free and open source software, because it is the most generic factor and so the most relevant beyond ecology.
-Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct, with only a small overlap.
-What they share in common is working with the field-collected data on a computer to prepare it for publication.
-R is a widely used free and open source data analysis tool in man research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018] and therefore a good proxy for the adoption of open source scientific programming languages in a discipline.
+Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct.
+An important similarlity for the two disciplines is working with the field-collected data on a computer to prepare it for publication.
+R is a widely used free and open source data analysis tool in many research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018].
+It is the dominant programming language in archaeology and ecology, so it is a good proxy for the adoption of open source scientific programming languages in these disciplines.
 To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.
@@
 These 1000 items represent variations on the recommended format for citing R.
-We found citations of R in the reference lists of `r prettyNum(how_many_articles_in_top_journals_for_cran_cites, ",")` English-languge articles indexed by the Web of Science in the research area of 'Environmental Sciences Ecology'.
+We found citations of R in the reference lists of `r prettyNum(how_many_articles_in_top_journals_for_cran_cites, ",")` English-language articles indexed by the Web of Science in the research area of 'Environmental Sciences Ecology'.
 We then downloaded the bibliographic data and reference lists for each of these articles.
@@
 On the other hand, articles that do not cite R more often discuss the Roman and Bronze Age periods, and less often reflect a technical specialisation.
-If there is one that is characteristic of non-citing R articles, it seems to be in ceramics.
+If there is one technical specialisation that is characteristic of non-citing R articles, it seems to be in ceramics.
 
@@
 @Touchon_McCoy_2016 identified the use of scripting languages as being a factor for easily sharing methods and code, thereby increasing the rate of development of statistical methods.
-One notable practice we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors make publicly available the code and data that they used to generate the results presented in their article.
+One notable practice we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors publicly share the code and data that they used to generate the results presented in their article.
 We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository such as Zenodo, Open Science Framework, Figshare, etc.
@@
 Figure \@ref(fig:archaeology-reproducible-articles-plot) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.
-The author guidelines of this journal confirm this trend: "In the interests of open scholarship and the reproducibility of results, JCAA strongly encourages all authors to deposit material relating to their publication in a preservation repository" [@authorguidelines_JCAA].
+The author guidelines of this journal support this trend: "In the interests of open scholarship and the reproducibility of results, JCAA strongly encourages all authors to deposit material relating to their publication in a preservation repository" [@authorguidelines_JCAA].
 
-These archaeology articles that include publicly available R code and data are part of a new approach to scientific communication that has also been observed in other fields.
+These archaeology articles that include publicly available R code and data are part of a shift in scientific communication that has also been observed in other fields.
 For example, the journal *Nature Neuroscience* requires authors 'to make the code that supports the generation of key figures in their manuscript available for review' [@Editors_2017].
@@
 Much of this literature is concerned with identifying and solving problems of irreproducibility in research.
-These actions comes from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community [@Stark_2018].
+These actions come from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community [@Stark_2018].
 Furthermore, that an article should sufficiently describe the presented research results such that a colleague can fully understand the results and how they were obtained.
@@
 
-In our view, merely using R (or similar open source programming languages) to conduct research would not constitute a tool-driven revolution by itself, but publicly sharing the code used for research, as a solution to problems of irreproducibility, is more likely to enable revolutionary change.
+In our view, merely using R (or similar open source programming languages) to conduct research would not constitute a tool-driven revolution by itself, but publicly sharing the code used for research, as a solution to problems of irreproducibility, is more likely to lead to revolutionary change.
 We expect that openly shared code will speed the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
@@
 
-A pragmatic issue relating to widespread changes in scientific practice is what the incentives are for members of the research community to change.
-This is a concern for researchers that might not be motivated by a desire to adhere to the abstract basic principles of what it means to practice science with integrity.
-If code sharing has concrete career advantages then we might expect it to be more likely to be part of a tool-driven revolution.
+A pragmatic issue for understanding widespread changes in scientific practice is to identify incentives for members of a research community to change.
+If code sharing has concrete career advantages, for example, if is known to result in increased citations, then we might expect it to be more likely to be part of a tool-driven revolution.
 @Vandewalle2012 analysed citation rates for 645 articles published in _IEEE Transactions on Image Processing_ during 2004–2006.
-He found statistically significant differences between the median number of citations for papers that have code available and papers that do not have code available.
+He found statistically significant differences between the median number of citations for papers that have code available and papers that do not.
 In considering causal relationships, Vandewalle raises the possibility of self-selection bias, where authors include code more often with their best papers.
@@
 We see that archaeological articles that cite R are consistently more highly cited themselves than articles that do not cite R.
+
 The mechanisms underlying the citation effects of using R remain to be elucidated.
@@
 In our data all the article authors are also the code authors, so there is no popularity effect noted by Vandewalle.
-One possibilty is that authors preferentially tend to promote their articles containing code, beacause they want a greater return for the greater effort they have put into including the code with the article, this is equivalent to the self-selection bias postulated by Vandewalle.
-Another possiblity is that papers with code are perceived by readers to be more robust because of their transparency, and thus a more reliable item to cite in support of claims made by other authods.
-Another possible cause is that when a paper is accompanied by code (for example on a GitHub repository) it is more visible than a paper without code, and appears higher in search result rankings.
-All other factors equal, higher visibility leads to more citations because the reseracher will stop searching when theyve found the first paper that serves their purpose.
-Further work is necessary to understand the mechanisms of this citation effect, additional insights might come from comparisions of papers that use another open source language such as Python, and MATLAB, which is not.
+One possibilty is that authors preferentially tend to promote their articles containing code, beacause they want a greater return for the greater effort they have put into including the code with the article, this may be equivalent to the self-selection bias postulated by Vandewalle.
+Another possiblity is that papers with code are perceived by readers to be more robust because of their transparency, and thus a more reliable and credible item to cite in support of knowledge claims.
+Another possible cause is that when a paper is accompanied by code (for example on a trustworthy repository) it is more visible than a paper without code, and appears higher in search result rankings.
+All other factors equal, higher visibility leads to more citations because the reseracher will stop searching when they have found the first paper that serves their purpose.
+Further work is necessary to understand the mechanisms of this citation effect.
+Additional insights might come from comparisions of papers that use another open source language such as Python, and MATLAB, which is not an open source language.
 
@@
 The tool which drives this change is the use of scripting languages in computational work in these disciplines.
-We have shown that the use of R is on the rise among archaeologists using natural science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visibility and influence (see Huggett in this volume for critical discussion of open scholarship).
+We have shown that the use of R is on the rise among archaeologists, especially those using natural science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visibility and influence (cf.
+Huggett in this volume for critical discussion of open scholarship).
 
-The use of R or similar scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to communicate one's workflow in a highly detailed an efficient way, and easily share this worksflow with others.
+The use of R or similar scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to communicate one's workflow in a highly detailed and efficient way, and easily share this workflow with others.
 The sharing of code has been identified in other disciplines as an amplifier of change in methods [@Touchon_McCoy_2016].
-It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
+It also enables researchers to not just reproduce the work of others, but to also thoroughly evaluate them.
 Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with the patterns described by @Galison_1997 [-@Galison_1997].
 This change is transformational rather than revolutionary, enabling archaeologists to more faithfully adhere to a fundamental tenet of the scientific process which demands the reproducibility of results.
-We do not want to claim archaeologists did not want to produce reproducible research before the advance of scientific scripting languages.
+We do not claim archaeologists did not want to produce reproducible research before the advance of scientific scripting languages.
 Indeed, there is a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses.
@@
 
-Nonetheless we feel that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice.
-In the digital age, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses.
+Nonetheless, with the introduction of computers there has been a gradual and unintended drift away from reproducibility [@brinckman2019computing; @bailey2016facilitating], and so from the fundamental tenets of scientific practice.
+Many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses.
 This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results.
-Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses.
+Modern archaeology has often been lacking in this aspect, as much research is  undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses.
 By using and sharing code we can undo this drift, and return to a robust scientific practice for archaeology.
-As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural sciences.
-These sciences also show a trend of increased emphasis on reproducibility, with journals changing their data and code availability policies, requiring code and data to be published alongside articles [@Stodden_2013].
-Also in archaeology, more and more funding agencies show interest in open science methods and transparency [@Strupler_2017].
+As we show, the most frequent use of scripting languages takes place mostly in the part of archaeology that is in close contact with natural sciences.
+These sciences also show a trend of increased emphasis on reproducibility, with journals changing their data and code availability policies to require code and data to be peer reviewed and available alongside articles [@Stodden_2013].
+Also in archaeology, funding agencies are increasingly showing interest in open science methods and transparency [@Strupler_2017].
 

### 2019-08-08 analysis/paper/jcaa-paper.Rmd
up to L470 of reviewer B's edits

@@
 
-As @Galison_1997 describes a shift in the practices of physicists, we believe so is archaeological practice changing in the digital age.
+As @Galison_1997 describes a shift in the practices of physicists, we believe archaeological practice is similarly changing.
 @Huvila_2018 have explored the use of digital tools on archaeological practices to understand the boundaries of what counts as archaeological, what is merely related to archaeology and what should be excluded from archaeology.
-Reflecting on @Huvila_2018, our view is the use use of code is unquestionably archaeology, since it is just one of many tools that archaeologists are using for converting raw data into summaries, visualisations, and insights.
+Reflecting on @Huvila_2018, the use of code is unquestionably archaeological, since it is one of many tools that archaeologists are using for converting raw data into summaries, visualisations, and insights.
 These concerns about how digital tools affect what we consider is and is not archaeology intersect with the work of @gieryn1999cultural on boundary construction.
 In his cultural cartography of science, Gieryn examines boundaries between science generally and varieties of less authoritative non-science.
-He argues that the boundaries are drawn and redrawn contiuously in flexible, historically changing and sometimes ambiguous ways.
-Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims by competitors.
+He argues that the boundaries are drawn and redrawn contiuously in flexible and sometimes ambiguous ways.
+Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims made by their competitors in the scientific community.
 Boundary work involves attributions of select characteristics to science in order to distinguish it from non-scientific competitors.
 The increasing voluntary use and sharing of code by researchers, and demands for it by readers and peer reviewers is a form of active boundary work that is shifting the boundary between scientific and non- or less-scientific archaeology.
-We, as members of this group, argue that the use of code as a research tool, and the sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work.
+We, as members of this group, argue that the use of code as a research tool, and the sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less- and un-scientific work.
 The reason for this is that using code, and making it available for others to inspect, enhances the credibility of knowledge claims and the visibility of those claims.
 
-A related issue here is the _technicity_ of tools, in our case coding.
-We draw on @cassirer_form_1930 and @simondon_mode_1958 to define technicity as a term to describe the technology, i.e.
+A related issue here is the _technicity_ of tools, in our case programming languages @cassirer_form_1930 and @simondon_mode_1958 to define technicity as a term to describe the technology, i.e.
 tools and machines respectively, that interact with humans and with nature [@hoel_ontological_2013, 188].
 Technology is said to own an instrumental kind of logos (Cassierer, see @hoel_ontological_2013, 194) and in analogy to language it exercises a certain measure of agency in mediating between humans, nature and technology.
-The emphasis here is on the function of technology, in the forces that it exercises, with a certain measure of autonomy, on other beings as well as in and through the new realities, it brings into being.
 The use of a scripting language might exemplify this, as it is not just a technological tool, but a proper "language" as well.
@@
 
-The concept of technicity is relevant here because in contrast to the concept of technological determinism, which puts technology as a force against the human, van der Tuin and Hoel (as well as Cassirer and Simondon) enfold technology into 'the human' without letting it lose its 'foreignness' and further develop the notions of 'human' and 'nature' [@hoel_ontological_2013, 190].
-Any change in tools therefore will influence how human researchers interact with their objects of interest and other people in their research community.
-For example, tools that make the research process more transparent and open should make the research community more open and inclusive to diverse participants and ideas.
-Using code to 'stand for' or represent the research process will increase our awareness of how archaeology itself 'stands for' the human experience in the past, and how it is a representation of the past [cf.
+In contrast to the concept of technological determinism, which puts technology as a force against the human experience, technicity enfolds technology into 'the human' without letting it lose its 'foreignness' [@hoel_ontological_2013, 190].
+According to this perspective, changes in tools will influence how human researchers interact with their objects of interest and other people in their research community.
+There are three ways is which this can be anticipated for archaeology.
+Firstly, tools that make the research process more transparent and open might be expected to make the research community more open and inclusive to diverse participants and ideas.
+Secondly, using code to 'stand for' or represent the research process will likely increase our awareness of how archaeology itself 'stands for' the human experience in the past, and how it is a representation of the past [cf.
 @pearson1998beginning].
-Similarly, the code has to be successfully executed to generate a useful result, just as claims about the past have to be successfully engaged with archaeological evidence to count as plausible.
-Interestingly it is exactly at the point between continously evolving subjects and objects, where van der Tuin and Hoel place technology and the creation of knowledge [@hoel_ontological_2013, 199].
-The "productive back-and-forth" between all parties may challenge established notions of culture, but also offers a concept of formation through differing, called "performative correspondence" [@hoel_ontological_2013, 201].
-The potential of a certain technology cannot be known before its coming-into-being as it will reflect back on its creation ("concretization" in Simondon's terms) as well as interact with other technologies (developed or yet-to-be developed) [@hoel_ontological_2013, 194].
-Observing how scientific scripting languages have become a concrete tool in ecology leads us to expect predict similar trends in archaeology.
+Thirdly, the code has to be successfully executed to generate a useful result, just as claims about the past have to be successfully engaged with archaeological evidence to count as plausible.
 

### 2019-08-08 analysis/paper/jcaa-paper.Rmd
typos seen by Rev B L 470-517

@@
 
-Although a discipline-wide paradigm shift might be off the table for archaeology, what we can see, though, is that there is a need for methods to share research in a way that is reproducible for others, and that archaeologists are looking for tools to make this possible.
+Although a discipline-wide paradigm shift might be off the table for archaeology, what we can see, is that there is a need for methods to share research in a way that is reproducible for others, and that archaeologists are looking for tools to make this possible.
 Using scripting languages and data repositories to share the scripts is a solution that enjoys increasing popularity in the archaeological research community.
@@
 
-We acknowledge that learning to use code, if one has not recieved formal instruction, is not an easy task.
+We acknowledge that learning to use code, if one has not received formal instruction, is not an easy task.
 As @Chrysanthi_2012 illustrate, it requires a certain familiarity with the computational tools to be able to use them not just on a technological basis, but creatively and in full knowledge of their restrictions and ambiguities.
@@
 As with data (see Fitzpatrick in this volume), sharing code in public takes a little time to ensure that it is fit for others to read.
-Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error.
+Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independently solving problems of how to write and share code through trial and error.
 This situation of guided variation (learning by trial and error) suggests that using code is not yet widespread enough among archaeologists to propagate due to frequency-dependant biases.

### 2019-08-08 analysis/paper/jcaa-paper.Rmd
typos L532-648 seen by rev B

@@
 The goal of rrtools is to make it easier for archaeologists and other researchers to use R for research and publication.
-This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort.
+This package aims to simplify many of the steps needed to create reproducible research, and to guide users to best practices with minimal effort.
 The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research.
@@
 We have outlined a positive future for archaeology where research publications are accompanied by code and data, enabling a 'critical self-consciousness' at a community-wide scale, similar to that imagined by @clarke1973archaeology as part of the discipline's 'loss of innocence'.
-A charactertic of this stage of disciplinary maturity is 'a closer understanding of its internal structure' [@clarke1973archaeology, p.
+A characteristic of this stage of disciplinary maturity is 'a closer understanding of its internal structure' [@clarke1973archaeology, p.
 7], and we can think of no more efficient and intelligible way to communicate the internals of archaeological analyses to other members of the discipline than by encoding the assumptions, logic, and calculations in scripts of an open source programming language.
@@
 The implication here is that the most effective way to stimulate this change is during the formative years of professional training.
-This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, whereas the horizontal bar is the bredth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
+This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, whereas the horizontal bar is the breadth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
 A gamma-shaped researcher has expert-level depth in archaeology, and is versed and proficient in other domains the provide skills such as analysing data with a programming language.

### 2019-08-08 analysis/paper/jcaa-paper.Rmd
up to L620 of reviewer B's edits

--- 2019-08-08 analysis/paper/jcaa-paper.Rmd
@@
 
-We do not claim that a Kuhnian paradigm-change is happening in ecology or archaeology: Data- and method-sharing behaviours have been around in archaeology before now and do not present incommensurate ways of thinking about the subject in the way that a typical reading of Kuhn's paradigm change concept should.
+We do not claim that a Kuhnian paradigm-change is happening in ecology or archaeology: Data- and method-sharing behaviours have been around in archaeology before now and do not present newly incommensurate ways of thinking about the subject, in the way that a typical reading of Kuhn's paradigm change concept should.
 Indeed, social science disciplines such as archaeology may not even be easily susceptible to classical Kuhnian paradigm shifts.
 This is because high-power tests (i.e.
-statistical tests with high probabilities of rejecting false null hypotheses) capable of distinguishing between superior and inferior paradigms are rare in the social sciences, such as economics [@Ioannidis_Stanley_Doucouliagos_2017], compared to physical science.
+statistical tests with high probabilities of rejecting false null hypotheses) capable of distinguishing between superior and inferior paradigms are rare in the social sciences, such as economics [@Ioannidis_Stanley_Doucouliagos_2017], compared to the physical sciences.
 Modelling and case studies described by @Akerlof_Michaillat_2018 indicate that among discplines lacking evidence to routinely strongly discriminate between theories, the chances of getting trapped in an inferior paradigm are high.
@@
 
-Although a discipline-wide paradigm shift might be off the table for archaeology, what we can see, is that there is a need for methods to share research in a way that is reproducible for others, and that archaeologists are looking for tools to make this possible.
-Using scripting languages and data repositories to share the scripts is a solution that enjoys increasing popularity in the archaeological research community.
+Although a discipline-wide paradigm shift might be off the table for archaeology, there is a need for methods to share research in a way that is reproducible for others, and that archaeologists are actively exploring tools to make this possible.
+Our resuts show that scripting languages and data repositories to share the scripts is a solution that enjoys increasing popularity in the archaeological research community [cf.
+@marwick_standard_2018].
 It has also expanded the 'trading zone' [@Galison_1997] where archaeologists collaborate with specialists in other fields to develop a shared language to get things done.
-In using scripting languages archaeologists are working at a trading zone with computer scientists and computational researchers in other fields to coordinate how tools from one domain can be useful in another.
-We can also see archaeologists in similar trading zones when using GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions.
-Archaeologists' embrace of, or turn towards, these methods may also be tool-driven revolutions.
+In using scripting languages, archaeologists are working at a trading zone with computer scientists and computational researchers in other fields to coordinate how tools from one domain can be useful in another.
+Other examples of archaeologists in  trading zones include the use of GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions.
+Archaeologists' embrace of these methods may also count as tool-driven revolutions.
 However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is an universal method with which we can do all, or which can be central to any, of those previously-mentioned approaches.
-That is why we believe archaeologists should pay attention to scripted workflows regardless of their usual toolkit and methodological interests.
+That is why we believe archaeologists should adopt scripted workflows regardless of their usual toolkit and methodological interests.
 
@@
 
-We acknowledge that learning to use code, if one has not received formal instruction, is not an easy task.
+We acknowledge that learning to use code, if one has not recieved formal instruction, is not an easy task.
 As @Chrysanthi_2012 illustrate, it requires a certain familiarity with the computational tools to be able to use them not just on a technological basis, but creatively and in full knowledge of their restrictions and ambiguities.
 The awareness and the appropriate handling of shortcomings and contexts of the data used in an analysis is also a time- and thought-consuming effort.
-This effort is required at the beginning of each archaeological research, and which may be exacerbated by using multiple data sets from varying sources [cf.
+This effort is required at the beginning of each archaeological research project, and which may be exacerbated by using multiple data sets from varying sources [cf.
 @Hugget_2015].
 
-As with data (see Fitzpatrick in this volume), sharing code in public takes a little time to ensure that it is fit for others to read.
+As with data (see Fitzpatrick in this volume), preparing code to make it publicly available takes time to ensure that it is fit for others to read and use.
 Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independently solving problems of how to write and share code through trial and error.
-This situation of guided variation (learning by trial and error) suggests that using code is not yet widespread enough among archaeologists to propagate due to frequency-dependant biases.
+We can use concepts from cultural evolution to understand this situation.
+Learning by trial and error is known as guided variation in a cultural evolutionary framework.
+The dominance of guided variation suggests that using code is not yet widespread enough among archaeologists to propagate due to frequency-dependent biases.
+Frequency-dependent biases occur when people copy the most abundant variant, in this case the most commonly used tool for analysing data, in the population [cf.
+@boyd1988culture].
 It also suggests that the professional benefits (i.e.
-citation effects) are not yet widely known for code use behaviours to propagate due to content biases, and finally that there are not yet enough highly-visible prestigous researchers using code for this behaviour due to propagate due to model-based biases [@Rendell_et_al_2011; @boyd1988culture; @henrich2007dual].
-This suggests to us that there is unrealised potential to improve the efficiency of using and code in archaeology by converging on some widely-agreed upon conventions that save time and investing effort in communicating the professional benefits of using code.
+the citation effects we describe above) are not yet widely known for code use  to propagate due to content biases.
+Content biases result when some aspect of a variant's content, such as positive citation effects, makes it more likely to be adopted.
+Finally there may not yet enough highly-visible, prestigous researchers using code for this behaviour due to propagate due to model-based biases [@Rendell_et_al_2011; @boyd1988culture; @henrich2007dual].
+This bias is based on imitation of highly prestigous, skilled or successful individuals.
+Considering the high variability of code using cultural evolutionary concepts, we conclude that there is unrealised potential to improve the efficiency of using and code in archaeology, firstly by converging on some widely-agreed upon conventions that will save researcher's time, and secondly by investing effort in communicating the professional benefits of using code.
 
@@
 The goal of rrtools is to make it easier for archaeologists and other researchers to use R for research and publication.
-This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort.
+This package aims to simplify many of the steps needed to write reproducible research papers, and to guide users to best practices with minimal effort.
 The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research.
@@
 - **use_mit_license:** This function creates a copy of the MIT software license to the compendium.
-This is important for two reasons.
-First, because it tells the reader that the author is happy for their code to be reused by other people.
+Other licenses are also possible, such as GPL, but the MIT license is preferred beceause it is widely used for research software, and has two important qualities.
+First, it tells the reader that the author is happy for their code to be reused by others, both in academic and commercial contexts.
 And second, it tells the reader that the author does not take responsibility for any problems that the reader might have when they use the code.
@@
 Using version control is important because it allows the author to time-travel back to earlier states of their analysis, if they take a dead-end path.
-It is also important for making collaboration smooth because many people can work on one set of files without losing track of what the most recent version is.
-Plus it is nice to have a remote back-up of our work in case of an emergency.
+It is also makes collaboration smooth because many people can work on one set of files without losing track of the most recent version.
+Plus it serves as a remote back-up of our work in case of an emergency.
 The GitHub repository can be kept private if desired, until the work is published.
@@
 We recognise that this file structure wont be perfect for every project.
-But our observations show that it is in wide use, and will suit many projects quite well.
+But our observations show that it is in wide use, and will suit most projects well.
 It will at least make the user think about how to organise their project logically, and to make it easier for others to navigate.
@@
 
-Starting with the discussion of Kuhn's influential work on paradigm shifts in science [@Kuhn_1962], we take inspiration from @Galison_1997 to propose a another explanation of how changes occurs in research communities, emphasising the transformative role of tools.
+Starting with the discussion of Kuhn's influential work on paradigm shifts in science [@Kuhn_1962], we took inspiration from @Galison_1997 to propose a another explanation of how changes occurs in research communities, emphasising the transformative role of tools.
 The change we focus on is the use of scripts to document and communicate data analysis.
@@
 If we accept that that is necessary for the future of the discipline, we must confront two implications.
-First is that there is currently a shortage of incentives to motivate use and sharing of code.
+First is the shortage of incentives to motivate the use and sharing of code.
 Our observation is that many researchers will not change unless required to by gate-keepers at prestigious journals and funding sources.

### 2019-08-08 analysis/paper/jcaa-paper.Rmd
fix some ?? in the figure cross-refs

@@
 The second detail is that most archaeology journals do not show any strong increase in the percentage of article citing R over time.
-The ecology journals shown in Figure \@ref(fig:ecology-citing-r) show strong upward trends of increasing proportions of articles citing R over time, but we do not see any unambiguous trends in the archaeology articles when considered together.
+The ecology journals shown in Figure \@ref(fig:ecology-articles-citing-r-plot) show strong upward trends of increasing proportions of articles citing R over time, but we do not see any unambiguous trends in the archaeology articles when considered together.
 Indeed, in several of the other archaeological journals in our sample, the first citations of R only occurred in the last two years.
@@
 
-In Figure \@ref(fig:citation-advantage) we conducted a similar analysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
+In Figure \@ref(fig:citation-advantage-plot) we conducted a similar analysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
 TC stands for 'Web of Science Core Collection Times Cited Count'.
@@
 
-```{r  fig.cap=str_glue('Median citation rates per year for archaeology articles 2010-2017 that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")})')}
+```{r  citation-advantage-plot, fig.cap=str_glue('Median citation rates per year for archaeology articles 2010-2017 that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")})')}
 knitr::include_graphics(here::here("analysis", "figures", "archy_journal_articles_citations_plot_cap.png"))

### 2019-08-08 analysis/paper/jcaa-paper.Rmd
finish adding reviewer B's edits

@@
 In other research communities we see people in these gate-keeping roles (e.g.
-in over 5,000 journals and professional organizations) effecting change by adopting the Transparency and Openness Promotion Guidelines that establish community standards of code and data availability that help to align scientific ideals with practices and disciplinary norms [@Nosek_et_al_2015].
+in over 5,000 journals and professional organizations) effecting change by adopting the Transparency and Openness Promotion Guidelines that establish community standards of code and data availability that help to align scientific ideals with disciplinary practices [@Nosek_et_al_2015].
 Simple forms of signalling, such as badges on papers to indicate that the authors have made code and data available, may also help to shift norms and raise awareness of these desirable behaviours [@Kidwell_et_al_2016].
 
-However, cultural change on an community-wide scale, such as a shift to the wide-spread use of code among archaeologists, is most likely to happen via the slow enculturation of people as they internalize norms during their formative years, and not via contemporaneous exposure to external cultural influences [@Vaisey_Lizardo_2016].
-This means it is unlikely that a professional archaeologist reading this paper, or ones like it, will be motivated to use code if they were not trained to work with code at the start of their career.
-The implication here is that the most effective way to stimulate this change is during the formative years of professional training.
-This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, whereas the horizontal bar is the bredth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
-A gamma-shaped researcher has expert-level depth in archaeology, and is versed and proficient in other domains the provide skills such as analysing data with a programming language.
-A gamma-shaped researcher with their MA or PhD in archaeology may also be conversant in the language and culture of computer science enough to have conversations and collaborate in the trading zone, but does not need to be an expert in computer science to advance that discipline.
+However, cultural change on a community-wide scale, such as a shift to the wide-spread use of code among archaeologists, is most likely to happen via the slow enculturation of professionals as they internalize norms during their formative years, and not via contemporaneous exposure to external cultural influences [@Vaisey_Lizardo_2016].
+This means it is unlikely that an established professional archaeologist reading this paper, or ones like it, will be motivated to use code if they were not trained to work with code at the start of their career.
+The implication here is that the most effective way to stimulate this change is during the first few years of professional training.
+This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, whereas the horizontal bar is the breadth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
+A gamma-shaped researcher has expert-level depth in archaeology, and proficient in other domains that provide skills such as analysing data with a programming language.
+A gamma-shaped researcher with their MA or PhD in archaeology may also be conversant enough in the language and culture of computer science to have conversations and collaborate in the trading zones described above, but does not need to be an expert in computer science to advance that discipline.
 
@@
 
-Our project comes from the group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
+The rrtools was developed by a group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
+A version of this paper was previously presented at the 2018 Computer Applications and Quantitative Methods in Archaeology (CAA) international conference at Tubingen, Germany.
+Early drafts were improved at the Cologne writing workshop in September 2018, which was supported by the COST Action Arkwork: Archaeological Practices and Knowledge Work in the Digital Environment.
 

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
typo.

@@
 One of the most profound changes driven by this concern for reproducibility and transparency is from using data analysis tools dependent on point-and-clicking with a mouse in closed source software, to a set of tools based on writing scripts in open source programming languages and making the scripts openly available.
-We present bibliometric evidence for this change in ecolpogy and archaeology to test if the adoption of these new tools may be considered a revolution.
+We present bibliometric evidence for this change in ecology and archaeology to test if the adoption of these new tools may be considered a revolution.
 We identify a citation effect for archaeology articles that use the open source programming language R.

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
two sentences instead of one for readability. and once tools in the section title

@@
 
-# Revolutions in Science: Ideas-driven or Tool-driven?
+# Revolutions in Science: Ideas-driven or Tools-driven?
 
@@
 What counts as suitable puzzles, and acceptable solutions to these puzzles, is governed by the norms and procedures of the prevailing paradigm.
-Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes replacing one paradigm with another in a conversion or speciation process where the community adopts a new way of doing science.
+Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes replacing one paradigm with another.
+In this conversion or speciation process the community adopts a new way of doing science.
 There has been much debate about how to draw a distinction between so-called normal and revolutionary science.

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
save a few words by cutting the description of American and European archaeology paradigms, cause, anyone who knows theory already knows and who doesn't won't get much out of the big words s/he doesn't know the meaning of.

@@
 [@watkins1970against; @toulmin1970does; @Casadevalle00158_16].
-A notable criticism is that Kuhn's normal science is unscientific, a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
+A notable criticism is that Kuhn's normal science is unscientific, as it describes a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
 This is at odds with a view of science as a continuous state of evolution, with researchers simultaneously employing many styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
@@
 The most wide-ranging application is Clark's [-@Clark_1993] comparison of North American archaeology to European archaeology during the 1970s-1990s.
-Clark, similarly to Binford and Sabloff [-@binford1982paradigms], identified the North American archaeological paradigm as rooted in anthropology, founded on post-positivist biases, and distinctive with its critical-realist ontology, a modified-objectivist epistemology, and an experimental-manipulative method.
-In contrast, the European archaeological paradigm is described a kind of history, mostly in a strict empiricist tradition, and typified by a realist ontology, an inductivist epistemology, and by an observational method.
-Clark's analysis describes a multi-paradigm situation where many research communities (i.e.
+His analysis describes a multi-paradigm situation where many research communities (i.e.
 North American and European) simultaneously operate in different paradigms (which Kuhn noted as a sign of a developing or immature discipline).

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
simplify Fullers paradigm shift

@@
 Archaeologists using LIDAR (light detection and ranging) for remote geospatial imaging of cultural landscapes have claimed the application of this technology to archaeology is a paradigm shift [@chase2012geospatial; @howey2016detecting].
-@Fuller_2010 has used the term to describe a shift in thinking from the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to its emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
+@Fuller_2010 has used the term to describe a shift in thinking from the emergence of agriculture as a  'Neolithic Revolution' to a protracted and entangled process, happening several times independently.
 Most of these claims for paradigm shifts differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, namely the multiple definitions of a paradigm in Kuhn's writings.

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
shortening this sentence a bit

@@
 @Fuller_2010 has used the term to describe a shift in thinking from the emergence of agriculture as a  'Neolithic Revolution' to a protracted and entangled process, happening several times independently.
-Most of these claims for paradigm shifts differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, namely the multiple definitions of a paradigm in Kuhn's writings.
+Most of these claims for  Kuhnian paradigm shifts differ in meaning, which highlights an important complication with using Kuhn's concept of a paradigm-shift, namely the multiple definitions of a paradigm in his writings.
 

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
typos. + program not computer I think

@@
 We see a compelling exploration of this third concrete sense of a paradigm in the work of @Galison_1997, who has offered a tool-driven view of change in research communities.
-Galison analysed of the role of tools in twentieth century particle physics, starting with hand-crafted cloud chambers and bubble chambers and ending with digital counters, particle accelerators, and computers.
+Galison analysed the role of tools in twentieth century particle physics, starting with hand-crafted cloud chambers and bubble chambers and ending with digital counters, particle accelerators, and computers.
 He identifies a profound change in physics when analog devices producing pictures were superseded by digital devices producing numerical data.
@@
 Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct.
-An important similarlity for the two disciplines is working with the field-collected data on a computer to prepare it for publication.
+An important similarity for the two disciplines is working with the field-collected data on a computer to prepare it for publication.
 R is a widely used free and open source data analysis tool in many research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018].
@@
 In our view, merely using R (or similar open source programming languages) to conduct research would not constitute a tool-driven revolution by itself, but publicly sharing the code used for research, as a solution to problems of irreproducibility, is more likely to lead to revolutionary change.
-We expect that openly shared code will speed the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
-As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as Zenodo, Open Science Framework, Figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.
+We expect that openly shared code will speed the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven programs are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
+As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as Zenodo, Open Science Framework, Figshare, etc.), then even researchers with limited resources anywhere in the world can contribute equally to the research community by using and sharing code.
 Using and sharing open source code is thus an important action for reducing inequality in the archaeological research community.

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
stash

--- 2019-08-09 analysis/paper/jcaa-paper.Rmd
@@
 We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
-One of the most profound changes driven by this concern for reproducibility and transparency is from using data analysis tools dependent on point-and-clicking with a mouse in closed source software, to a set of tools based on writing scripts in open source programming languages and making the scripts openly available.
-We present bibliometric evidence for this change in ecology and archaeology to test if the adoption of these new tools may be considered a revolution.
-We identify a citation effect for archaeology articles that use the open source programming language R.
+One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on point-and-clicking with a mouse in closed source software, to a new set of tools based on writing scripts in open source programming languages and making them openly available.
+We present bibliometric evidence for this change in ecology and in archaeology to test if the adoption of these new tools may be considered a revolution or some other process.
+We identify a citation effect for papers that use the open source programming language R.
 We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of R code."
@@
 
-# library(caa2018) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
-
 library(ggrepel)
@@
 
-In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method.
-We ask whether this might be the beginning of a broader change in the field.
+In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method and whether this might be the beginning of a broader change.
 Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are ideas-driven or tools-driven [@Galison_1997; @Dyson_2000].
-This argument motivated us to explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of research.
-We focus on one set of tools for this change, based on writing data analysis scripts in free and open source programming languages, exemplified by R, and the practice of sharing of these scripts.
+In this light we explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of research.
+We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of these scripts.
 We focus on R because our prior observations indicate it is by far the most common scripting language used by ecologists and archaeologists.
@@
 
-# Revolutions in Science: Ideas-driven or Tools-driven?
+# Revolutions in Science: Ideas-driven or Tool-driven?
 
@@
 One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] effort to describe the history of science by proposing paradigm shifts resulting from revolutionary change in communities of researchers.
-In brief, this model proposes that the history of science consists of long periods of tradition-bound 'normal science' punctuated by short episodes of 'revolution'.
+In brief, this model proposes that the history of science consists of long periods of tradition-bound 'normal science' punctuated by short episodes of revolution.
 Normal science was described by Kuhn as stagnant, routine day-to-day research focused on 'puzzle solving'.
 What counts as suitable puzzles, and acceptable solutions to these puzzles, is governed by the norms and procedures of the prevailing paradigm.
-Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes replacing one paradigm with another.
-In this conversion or speciation process the community adopts a new way of doing science.
+Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes replacing one paradigm with another in a conversion or speciation process where the community adopts a new way of doing science.
 There has been much debate about how to draw a distinction between so-called normal and revolutionary science.
 [@watkins1970against; @toulmin1970does; @Casadevalle00158_16].
-A notable criticism is that Kuhn's normal science is unscientific, as it describes a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
-This is at odds with a view of science as a continuous state of evolution, with researchers simultaneously employing many styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
+A notable criticism is that Kuhn's normal science is unscientific, a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
+This is at odds with a view of science as a continuous state of revolution, with researchers simultaneously employing many styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
 In this view, based on the post-World War II era of science, radical discontinuities are rare, and major developments instead emerge from the division and recombination of already mature fields [@andersen2013second].
 Given the rarity of radical discontinuities, some biologists have argued that change in biology is better characterised by new ideas (e.g.
-Mendelian heredity, Darwinian evolution, and molecular genetics) replacing not a former paradigm, but a conceptual vacuum [@wilkins1996are].
+Mendelian heredity, Darwinian evolution, and molecular genetics) replacing not a former paradigm but a conceptual vacuum [@wilkins1996are].
 
 In archaeology we have seen claims for paradigm shifts in several areas.
-The most wide-ranging application is Clark's [-@Clark_1993] comparison of North American archaeology to European archaeology during the 1970s-1990s.
-His analysis describes a multi-paradigm situation where many research communities (i.e.
+The most wide-ranging application is Clark's [-@Clark_1993] comparison of North American archaeology to European archaeology during the 1970s-90s.
+Clark, similar to Binford and Sabloff [-@binford1982paradigms], identified the North American archaeological paradigm as rooted in anthropology, founded on post-positivist biases, and distinctive with its critical-realist ontology, a modified-objectivist epistemology, and an experimental-manipulative method.
+In contrast, the European archaeological paradigm is described a kind of history, mostly in a strict empiricist tradition, and typified by a realist ontology, an inductivist epistemology, and by an observational method.
+Clark's analysis describes a multi-paradigm situation where many research communities (i.e.
 North American and European) simultaneously operate in different paradigms (which Kuhn noted as a sign of a developing or immature discipline).
@@
 
-A more thorough treatment of the claim that the appearance of processual archaeology was a paradigm shift is provided by @Meltzer_1979.
+A more thorough treatment of the claim that processual archaeology was a paradigm shift is provided by @Meltzer_1979.
 Meltzer reviews literature arguing for a revolutionary change in archaeology during the 1960s and 1970s, and considers these claims in light of classic Kuhnian revolutionary events such as the replacement of the Ptolemaic system by the Copernican system, and Newtonian dynamics by the new Einsteinian dynamics.
@@
 
-We also see mentions of paradigm shifts in archaeology in reference to the shift from processual to post-processual archaeology [@Härke_2002; @Koerner_2018], as well as more localised and specific shifts in archaeological practice and thinking.
+We also see mentions of paradigm shifts in archaeology in reference to processual to post-processual archaeology [@Härke_2002; @Koerner_2018], as well as more localised and specific shifts in archaeological practice and thinking.
 For example, @Snodgrass_2002 claims that among Classical Archaeologists, an increase in research on previously neglected periods of antiquity, such as the Greek Early Iron Age, constitutes a paradigm shift.
-@McAnany_Rowe_2015 propose that the appearance of community-based participatory models of research among some communities of archaeologists is a paradigm shift, though they concede that it is transformational rather than revolutionary.
-@harris2012interfacing similarly argues that community crowdsourced geographic knowledge (or volunteered geographic information) could be paradigm shift for archaeological communities.
+@McAnany_Rowe_2015 propose that the appearance of community-based participatory models of research among some communities of archaeologists is a paradigm shift, though they concede that it is more transformational rather than revolutionary.
+@harris2012interfacing has similarly argued that community crowdsourced geographic knowledge (or volunteered geographic information) could be paradigm shift for archaeological communities.
 Archaeologists using LIDAR (light detection and ranging) for remote geospatial imaging of cultural landscapes have claimed the application of this technology to archaeology is a paradigm shift [@chase2012geospatial; @howey2016detecting].
-@Fuller_2010 has used the term to describe a shift in thinking from the emergence of agriculture as a  'Neolithic Revolution' to a protracted and entangled process, happening several times independently.
-Most of these claims for  Kuhnian paradigm shifts differ in meaning, which highlights an important complication with using Kuhn's concept of a paradigm-shift, namely the multiple definitions of a paradigm in his writings.
+@Fuller_2010 has used the term to describe a shift in thinking about the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to new concept of the emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
+Most of these paradigms differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, both as a tool to understand the history of science and as a guide to changing it, which is the multiple definitions of a paradigm in Kuhn's writings.
 
@@
 
-We propose that Galison's emphasis on tool-driven change, as a complement to Kuhnian concept-driven change, has potential to enhance archaeology's social and scientific relevance and contributions.
-The centrality of material culture in Galison's view raises the question about whether there is scope for archaeologists to make productive contributions to understanding change in the history of science through an archaeological analysis of scientific instrumentation [cf.
+We propose that Galison's emphasis on tool-driven change, as a complement to Kuhnian concept-driven change, has much potential for archaeology.
+The centrality of material culture in Galison's view raises the question about whether there is scope for archaeologists to make productive contributions to understanding change in the history science through an archaeological analysis of scientific instrumentation [cf.
 @Schiffer_2013].
@@
 @chase2012geospatial have claimed the introduction of radiocarbon dating was a paradigm shift in archaeology, but we are not aware of a Galisonian analysis of this change.
-A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to predict or generate change in the practice of archaeology?
+A third issue is how we can use a tool-driven approach not just as a framework for understanding change in science, but as a method to predict or generate change in archaeology?
 
@@
 
-In this section we explore how a change observed in ecology might be relevant for understanding or even directing the future of archaeology.
+The question of predicting or generating change in scientific practice via adoption of new tools is intriguing, and in this section we explore how a change observed in ecology might be relevant for understanding or even directing the future of archaeology.
 Biological disciplines have a long tradition of influencing archaeological thought, starting with Oscar Montelius, whose typological method was inspired by Darwin's theory of evolution [@Montelius_1899, 237].
-This tradition continues, for example, many spatial statistics currently used by archaeologists are derived from ecology [@keron_use_2015, 7].
+This tradition continues, for example, many spatial statistics in current use in archaeology are derived from ecology [@keron_use_2015, 7].
 
@@
 @Touchon_McCoy_2016 identify four factors relating to technological change that might explain the changes they observed in the use of statistics in ecology.
-First, they note that automated data loggers, GPS trackers, remote sensing, and crowd sourcing have greatly increased the rate at which ecologists collect data.
-Second, increases in desktop and cluster computing power have made complex analytical processes faster and more convenient to compute.
+First, they note that automated data loggers, GPS trackers, remote sensing, and crowd sourcing have greatly increased rate at which ecologists collect data.
+Second, increases in desktop and cluster computing power has made complex analytical processes faster and more convenient to compute.
 Third, the development of free, open source and easily extensible software for data analysis and visualisation, such as R, allows new methods to spread quickly via online fora and social media.
-Fourth, the publication of several books and papers that have strongly influenced the way many ecologists think about data analysis, such as Burnham and Anderson's [-@burnham2003model] book on model selection and inference.
+Fourth, the publication of several books and papers that have strongly influenced the way many ecologists think about data analysis, such as @burnham2003model's book on model selection and inference.
 
 We focus here on the third factor, free and open source software, because it is the most generic factor and so the most relevant beyond ecology.
-Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct.
-An important similarlity for the two disciplines is working with the field-collected data on a computer to prepare it for publication.
-R is a widely used free and open source data analysis tool in many research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018].
-It is the dominant programming language in archaeology and ecology, so it is a good proxy for the adoption of open source scientific programming languages in these disciplines.
+Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct, with only a small overlap.
+What they share in common is working with the field-collected data on a computer to prepare it for publication.
+R is a widely used free and open source data analysis tool in many research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018] and therefore a good proxy for the adoption of open source scientific programming languages in a discipline.
 To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.
@@
 These 1000 items represent variations on the recommended format for citing R.
-We found citations of R in the reference lists of `r prettyNum(how_many_articles_in_top_journals_for_cran_cites, ",")` English-language articles indexed by the Web of Science in the research area of 'Environmental Sciences Ecology'.
+We found citations of R in the reference lists of `r prettyNum(how_many_articles_in_top_journals_for_cran_cites, ",")` English-languge articles indexed by the Web of Science in the research area of 'Environmental Sciences Ecology'.
 We then downloaded the bibliographic data and reference lists for each of these articles.
@@
 
-archy_cran_cites_all_areas_top_journals_by_year_to_plot <- 
- archy_cran_cites_all_areas_top_journals_by_year # %>% 
- # filter(SO != "AFRICAN ARCHAEOLOGICAL REVIEW")
-
 all_archaeology_citing_R_over_time <-
-  ggplot(archy_cran_cites_all_areas_top_journals_by_year_to_plot,
+ggplot(archy_cran_cites_all_areas_top_journals_by_year,
        aes(PY,
@@
   geom_text_repel(
-    data = archy_cran_cites_all_areas_top_journals_by_year_to_plot %>%
+    data = archy_cran_cites_all_areas_top_journals_by_year %>%
       group_by(SO) %>%
@@
     direction = "y",
-    size = 6,
+    size = 5,
     nudge_x = 1,
     hjust = 0) +
-   theme(panel.grid.major = element_blank(), 
-         panel.grid.minor = element_blank(),
-         panel.background = element_blank(), 
-         axis.line = element_line(colour = "black"),
-         axis.title.x = element_text(hjust = 0.125),
-        legend.position="none",
+  theme_minimal() +
+  theme(legend.position="none",
         axis.text.x = element_text(angle = 90,
@@
                                    vjust = 0.5)) +
-  scale_x_continuous(breaks = min_y:2017,
+  scale_x_continuous(breaks = min_y:max_y,
+                     labels = c(min_y:2017,
+                                rep("",
+                                    length(2018:max_y))),
                      limits = c(min_y, max_y)) +
@@
   xlab("Publication year") +
-  ylab("Percentage of articles in that journal") +
-  lemon::coord_capped_cart(bottom='right')
+  ylab("Percentage of articles in that journal")
 
@@
                parse = TRUE,
-              label.y = 0.9) +
- stat_fit_glance(method = "lm", 
-                  label.y = 0.8,
-                  method.args = list(formula = y~x),
-                  aes(label = paste("italic(P)*\"-value = \"*", 
-                                    signif(..p.value.., digits = 4), sep = "")),
-                  parse = TRUE) +
+               size = 3,
+               label.y.npc = 0.7) +
+  stat_fit_glance(label.y.npc = 0.7,
+                  size = 3,
+                  method = "lm",
+                  method.args = list(formula = y ~ x),
+                  geom = "text",
+                  aes(label = paste("p-value: ",
+                                    signif(..p.value..,
+                                           digits = 4)))) +
   xlab("year") +
@@
                            italic("Journal of Archaeological Science")))) +
-  theme_bw(base_size = 10)
+  theme_bw(base_size = 6)
 ```
@@
   annotation_custom(grob = ggplotGrob(jas_cites_r),
-                          xmin = 2027, xmax = 2040,
-                          ymin = 0.10, ymax = 0.20)
+                          xmin = 2030, xmax = 2040,
+                          ymin = 0.10, ymax = 0.17)
 
@@
 The second detail is that most archaeology journals do not show any strong increase in the percentage of article citing R over time.
-The ecology journals shown in Figure \@ref(fig:ecology-articles-citing-r-plot) show strong upward trends of increasing proportions of articles citing R over time, but we do not see any unambiguous trends in the archaeology articles when considered together.
+The ecology journals shown in Figure \@ref(fig:ecology-citing-r) show strong upward trends of increasing proportions of articles citing R over time, but we do not see any unambiguous trends in the archaeology articles when considered together.
 Indeed, in several of the other archaeological journals in our sample, the first citations of R only occurred in the last two years.
@@
 On the other hand, articles that do not cite R more often discuss the Roman and Bronze Age periods, and less often reflect a technical specialisation.
-If there is one technical specialisation that is characteristic of non-citing R articles, it seems to be in ceramics.
+If there is one that is characteristic of non-citing R articles, it seems to be in ceramics.
 
@@
 @Touchon_McCoy_2016 identified the use of scripting languages as being a factor for easily sharing methods and code, thereby increasing the rate of development of statistical methods.
-One notable practice we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors publicly share the code and data that they used to generate the results presented in their article.
+One notable practice we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors make publicly available the code and data that they used to generate the results presented in their article.
 We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository such as Zenodo, Open Science Framework, Figshare, etc.
-'Trustworthy' data repositories are those hosted by a non-commercial organisation, offering persistent identifiers, and have transparent commmitments to long term data management and backup strategies (see @marwick_standard_2018 for more details).
 Our continuously updated list of articles in this sample is available online here: https://github.com/benmarwick/ctv-archaeology.
 Figure \@ref(fig:archaeology-reproducible-articles-plot) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.
-The author guidelines of this journal support this trend: "In the interests of open scholarship and the reproducibility of results, JCAA strongly encourages all authors to deposit material relating to their publication in a preservation repository" [@authorguidelines_JCAA].
+The author guidelines of this journal confirm this trend: "In the interests of open scholarship and the reproducibility of results, JCAA strongly encourages all authors to deposit material relating to their publication in a preservation repository" [@authorguidelines_JCAA].
 
-These archaeology articles that include publicly available R code and data are part of a shift in scientific communication that has also been observed in other fields.
+These archaeology articles that include publicly available R code and data are part of a new approach to scientific communication that has also been observed in other fields.
 For example, the journal *Nature Neuroscience* requires authors 'to make the code that supports the generation of key figures in their manuscript available for review' [@Editors_2017].
@@
 Much of this literature is concerned with identifying and solving problems of irreproducibility in research.
-These actions come from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community [@Stark_2018].
+These actions comes from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community [@Stark_2018].
 Furthermore, that an article should sufficiently describe the presented research results such that a colleague can fully understand the results and how they were obtained.
@@
 
-In our view, merely using R (or similar open source programming languages) to conduct research would not constitute a tool-driven revolution by itself, but publicly sharing the code used for research, as a solution to problems of irreproducibility, is more likely to lead to revolutionary change.
-We expect that openly shared code will speed the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
+In our view, merely using R (or similar open source programming languages) to conduct research would not constitute a tool-driven revolution by itself, but publicly sharing the code used for research, as a solution to problems of irreproducibility, is more likely to enable revolutionary change.
+Openly shared code speeds the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven computers are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
 As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as Zenodo, Open Science Framework, Figshare, etc.), then even researchers with limited resources anywhere in the world can to contribute equally to the research community by using and sharing code.
 Using and sharing open source code is thus an important action for reducing inequality in the archaeological research community.
-It enables participation by researchers who are unable to buy software licenses or access face-to-face training.
+This is because it enables participation by researchers who are unable to buy software licenses or access face-to-face training.
 
@@
 
-A pragmatic issue for understanding widespread changes in scientific practice is to identify incentives for members of a research community to change.
-If code sharing has concrete career advantages, for example, if is known to result in increased citations, then we might expect it to be more likely to be part of a tool-driven revolution.
+A pragmatic issue relating to widespread changes in scientific practice is what the incentives are for members of the research community to change.
+This is a concern for researchers that might not be motivated by a desire to adhere to the abstract basic principles of what it means to practice science with integrity.
+If code sharing has concrete career advantages then we might expect it to be more likely to be part of a tool-driven revolution.
 @Vandewalle2012 analysed citation rates for 645 articles published in _IEEE Transactions on Image Processing_ during 2004–2006.
-He found statistically significant differences between the median number of citations for papers that have code available and papers that do not.
+He found statistically significant differences between the median number of citations for papers that have code available and papers that do not have code available.
 In considering causal relationships, Vandewalle raises the possibility of self-selection bias, where authors include code more often with their best papers.
@@
 
-In Figure \@ref(fig:citation-advantage-plot) we conducted a similar analysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
+In Figure \@ref(fig:citation-advantage) we conducted a similar analysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
 TC stands for 'Web of Science Core Collection Times Cited Count'.
@@
 We see that archaeological articles that cite R are consistently more highly cited themselves than articles that do not cite R.
-
 The mechanisms underlying the citation effects of using R remain to be elucidated.
@@
 In our data all the article authors are also the code authors, so there is no popularity effect noted by Vandewalle.
-One possibilty is that authors preferentially tend to promote their articles containing code, beacause they want a greater return for the greater effort they have put into including the code with the article, this may be equivalent to the self-selection bias postulated by Vandewalle.
-Another possiblity is that papers with code are perceived by readers to be more robust because of their transparency, and thus a more reliable and credible item to cite in support of knowledge claims.
-Another possible cause is that when a paper is accompanied by code (for example on a trustworthy repository) it is more visible than a paper without code, and appears higher in search result rankings.
-All other factors equal, higher visibility leads to more citations because the reseracher will stop searching when they have found the first paper that serves their purpose.
-Further work is necessary to understand the mechanisms of this citation effect.
-Additional insights might come from comparisions of papers that use another open source language such as Python, and MATLAB, which is not an open source language.
+One possibilty is that authors preferentially tend to promote their articles containing code, beacause they want a greater return for the greater effort they have put into including the code with the article, this is equivalent to the self-selection bias postulated by Vandewalle.
+Another possiblity is that papers with code are perceived by readers to be more robust because of their transparency, and thus a more reliable item to cite in support of claims made by other authods.
+Another possible cause is that when a paper is accompanied by code (for example on a GitHub repository) it is more visible than a paper without code, and appears higher in search result rankings.
+All other factors equal, higher visibility leads to more citations because the reseracher will stop searching when theyve found the first paper that serves their purpose.
+Further work is necessary to understand the mechanisms of this citation effect, additional insights might come from comparisions of papers that use another open source language such as Python, and MATLAB, which is not.
 
@@
 
-```{r  citation-advantage-plot, fig.cap=str_glue('Median citation rates per year for archaeology articles 2010-2017 that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")})')}
+```{r  fig.cap=str_glue('Median citation rates per year for archaeology articles 2010-2017 that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")})')}
 knitr::include_graphics(here::here("analysis", "figures", "archy_journal_articles_citations_plot_cap.png"))
@@
 The tool which drives this change is the use of scripting languages in computational work in these disciplines.
-We have shown that the use of R is on the rise among archaeologists, especially those using natural science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visibility and influence (cf.
-Huggett in this volume for critical discussion of open scholarship).
+We have shown that the use of R is on the rise among archaeologists using natural science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visibility and influence (see Huggett in this volume for critical discussion of open scholarship).
 
-The use of R or similar scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to communicate one's workflow in a highly detailed and efficient way, and easily share this workflow with others.
+The use of R or similar scripting languages enables a major shift in the technical and social dynamics of the research community, as it offers the possibility to communicate one's workflow in a highly detailed an efficient way, and easily share this worksflow with others.
 The sharing of code has been identified in other disciplines as an amplifier of change in methods [@Touchon_McCoy_2016].
-It also enables researchers to not just reproduce the work of others, but to also thoroughly evaluate them.
+It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method.
 Thus, the sharing of code may be at the root of an upcoming tool-driven change for archaeology, consistent with the patterns described by @Galison_1997 [-@Galison_1997].
 This change is transformational rather than revolutionary, enabling archaeologists to more faithfully adhere to a fundamental tenet of the scientific process which demands the reproducibility of results.
-We do not claim archaeologists did not want to produce reproducible research before the advance of scientific scripting languages.
+We do not want to claim archaeologists did not want to produce reproducible research before the advance of scientific scripting languages.
 Indeed, there is a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses.
@@
 
-Nonetheless, with the introduction of computers there has been a gradual and unintended drift away from reproducibility [@brinckman2019computing; @bailey2016facilitating], and so from the fundamental tenets of scientific practice.
-Many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses.
+Nonetheless we feel that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice.
+In the digital age, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses.
 This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results.
-Modern archaeology has often been lacking in this aspect, as much research is  undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses.
+Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses.
 By using and sharing code we can undo this drift, and return to a robust scientific practice for archaeology.
-As we show, the most frequent use of scripting languages takes place mostly in the part of archaeology that is in close contact with natural sciences.
-These sciences also show a trend of increased emphasis on reproducibility, with journals changing their data and code availability policies to require code and data to be peer reviewed and available alongside articles [@Stodden_2013].
-Also in archaeology, funding agencies are increasingly showing interest in open science methods and transparency [@Strupler_2017].
+As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural sciences.
+These sciences also show a trend of increased emphasis on reproducibility, with journals changing their data and code availability policies, requiring code and data to be published alongside articles [@Stodden_2013].
+Also in archaeology, more and more funding agencies show interest in open science methods and transparency [@Strupler_2017].
 
-As @Galison_1997 describes a shift in the practices of physicists, we believe archaeological practice is similarly changing.
+As @Galison_1997 describes a shift in the practices of physicists, we believe so is archaeological practice changing in the digital age.
 @Huvila_2018 have explored the use of digital tools on archaeological practices to understand the boundaries of what counts as archaeological, what is merely related to archaeology and what should be excluded from archaeology.
-Reflecting on @Huvila_2018, the use of code is unquestionably archaeological, since it is one of many tools that archaeologists are using for converting raw data into summaries, visualisations, and insights.
+Reflecting on @Huvila_2018, our view is the use use of code is unquestionably archaeology, since it is just one of many tools that archaeologists are using for converting raw data into summaries, visualisations, and insights.
 These concerns about how digital tools affect what we consider is and is not archaeology intersect with the work of @gieryn1999cultural on boundary construction.
 In his cultural cartography of science, Gieryn examines boundaries between science generally and varieties of less authoritative non-science.
-He argues that the boundaries are drawn and redrawn contiuously in flexible and sometimes ambiguous ways.
-Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims made by their competitors in the scientific community.
+He argues that the boundaries are drawn and redrawn contiuously in flexible, historically changing and sometimes ambiguous ways.
+Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims by competitors.
 Boundary work involves attributions of select characteristics to science in order to distinguish it from non-scientific competitors.
 The increasing voluntary use and sharing of code by researchers, and demands for it by readers and peer reviewers is a form of active boundary work that is shifting the boundary between scientific and non- or less-scientific archaeology.
-We, as members of this group, argue that the use of code as a research tool, and the sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less- and un-scientific work.
+We, as members of this group, argue that the use of code as a research tool, and the sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work.
 The reason for this is that using code, and making it available for others to inspect, enhances the credibility of knowledge claims and the visibility of those claims.
 
-A related issue here is the _technicity_ of tools, in our case programming languages @cassirer_form_1930 and @simondon_mode_1958 to define technicity as a term to describe the technology, i.e.
+A related issue here is the _technicity_ of tools, in our case coding.
+We draw on @cassirer_form_1930 and @simondon_mode_1958 to define technicity as a term to describe the technology, i.e.
 tools and machines respectively, that interact with humans and with nature [@hoel_ontological_2013, 188].
 Technology is said to own an instrumental kind of logos (Cassierer, see @hoel_ontological_2013, 194) and in analogy to language it exercises a certain measure of agency in mediating between humans, nature and technology.
+The emphasis here is on the function of technology, in the forces that it exercises, with a certain measure of autonomy, on other beings as well as in and through the new realities, it brings into being.
 The use of a scripting language might exemplify this, as it is not just a technological tool, but a proper "language" as well.
@@
 
-In contrast to the concept of technological determinism, which puts technology as a force against the human experience, technicity enfolds technology into 'the human' without letting it lose its 'foreignness' [@hoel_ontological_2013, 190].
-According to this perspective, changes in tools will influence how human researchers interact with their objects of interest and other people in their research community.
-There are three ways is which this can be anticipated for archaeology.
-Firstly, tools that make the research process more transparent and open might be expected to make the research community more open and inclusive to diverse participants and ideas.
-Secondly, using code to 'stand for' or represent the research process will likely increase our awareness of how archaeology itself 'stands for' the human experience in the past, and how it is a representation of the past [cf.
+The concept of technicity is relevant here because in contrast to the concept of technological determinism, which puts technology as a force against the human, van der Tuin and Hoel (as well as Cassirer and Simondon) enfold technology into 'the human' without letting it lose its 'foreignness' and further develop the notions of 'human' and 'nature' [@hoel_ontological_2013, 190].
+Any change in tools therefore will influence how human researchers interact with their objects of interest and other people in their research community.
+For example, tools that make the research process more transparent and open should make the research community more open and inclusive to diverse participants and ideas.
+Using code to 'stand for' or represent the research process will increase our awareness of how archaeology itself 'stands for' the human experience in the past, and how it is a representation of the past [cf.
 @pearson1998beginning].
-Thirdly, the code has to be successfully executed to generate a useful result, just as claims about the past have to be successfully engaged with archaeological evidence to count as plausible.
+Similarly, the code has to be successfully executed to generate a useful result, just as claims about the past have to be successfully engaged with archaeological evidence to count as plausible.
+Interestingly it is exactly at the point between continously evolving subjects and objects, where van der Tuin and Hoel place technology and the creation of knowledge [@hoel_ontological_2013, 199].
+The "productive back-and-forth" between all parties may challenge established notions of culture, but also offers a concept of formation through differing, called "performative correspondence" [@hoel_ontological_2013, 201].
+The potential of a certain technology cannot be known before its coming-into-being as it will reflect back on its creation ("concretization" in Simondon's terms) as well as interact with other technologies (developed or yet-to-be developed) [@hoel_ontological_2013, 194].
+Observing how scientific scripting languages have become a concrete tool in ecology leads us to expect predict similar trends in archaeology.
 
-We do not claim that a Kuhnian paradigm-change is happening in ecology or archaeology: Data- and method-sharing behaviours have been around in archaeology before now and do not present newly incommensurate ways of thinking about the subject, in the way that a typical reading of Kuhn's paradigm change concept should.
+We do not claim that a Kuhnian paradigm-change is happening in ecology or archaeology: Data- and method-sharing behaviours have been around in archaeology before now and do not present incommensurate ways of thinking about the subject in the way that a typical reading of Kuhn's paradigm change concept should.
 Indeed, social science disciplines such as archaeology may not even be easily susceptible to classical Kuhnian paradigm shifts.
 This is because high-power tests (i.e.
-statistical tests with high probabilities of rejecting false null hypotheses) capable of distinguishing between superior and inferior paradigms are rare in the social sciences, such as economics [@Ioannidis_Stanley_Doucouliagos_2017], compared to the physical sciences.
+statistical tests with high probabilities of rejecting false null hypotheses) capable of distinguishing between superior and inferior paradigms are rare in the social sciences, such as economics [@Ioannidis_Stanley_Doucouliagos_2017], compared to physical science.
 Modelling and case studies described by @Akerlof_Michaillat_2018 indicate that among discplines lacking evidence to routinely strongly discriminate between theories, the chances of getting trapped in an inferior paradigm are high.
@@
 
-Although a discipline-wide paradigm shift might be off the table for archaeology, there is a need for methods to share research in a way that is reproducible for others, and that archaeologists are actively exploring tools to make this possible.
-Our resuts show that scripting languages and data repositories to share the scripts is a solution that enjoys increasing popularity in the archaeological research community [cf.
-@marwick_standard_2018].
+Although a discipline-wide paradigm shift might be off the table for archaeology, what we can see, though, is that there is a need for methods to share research in a way that is reproducible for others, and that archaeologists are looking for tools to make this possible.
+Using scripting languages and data repositories to share the scripts is a solution that enjoys increasing popularity in the archaeological research community.
 It has also expanded the 'trading zone' [@Galison_1997] where archaeologists collaborate with specialists in other fields to develop a shared language to get things done.
-In using scripting languages, archaeologists are working at a trading zone with computer scientists and computational researchers in other fields to coordinate how tools from one domain can be useful in another.
-Other examples of archaeologists in  trading zones include the use of GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions.
-Archaeologists' embrace of these methods may also count as tool-driven revolutions.
+In using scripting languages archaeologists are working at a trading zone with computer scientists and computational researchers in other fields to coordinate how tools from one domain can be useful in another.
+We can also see archaeologists in similar trading zones when using GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions.
+Archaeologists' embrace of, or turn towards, these methods may also be tool-driven revolutions.
 However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is an universal method with which we can do all, or which can be central to any, of those previously-mentioned approaches.
-That is why we believe archaeologists should adopt scripted workflows regardless of their usual toolkit and methodological interests.
+That is why we believe archaeologists should pay attention to scripted workflows regardless of their usual toolkit and methodological interests.
 
@@
 The awareness and the appropriate handling of shortcomings and contexts of the data used in an analysis is also a time- and thought-consuming effort.
-This effort is required at the beginning of each archaeological research project, and which may be exacerbated by using multiple data sets from varying sources [cf.
+This efforts is required at the beginning of each archaeological research, and which may be exacerbated by using multiple data sets from varying sources [cf.
 @Hugget_2015].
 
-As with data (see Fitzpatrick in this volume), preparing code to make it publicly available takes time to ensure that it is fit for others to read and use.
-Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independently solving problems of how to write and share code through trial and error.
-We can use concepts from cultural evolution to understand this situation.
-Learning by trial and error is known as guided variation in a cultural evolutionary framework.
-The dominance of guided variation suggests that using code is not yet widespread enough among archaeologists to propagate due to frequency-dependent biases.
-Frequency-dependent biases occur when people copy the most abundant variant, in this case the most commonly used tool for analysing data, in the population [cf.
-@boyd1988culture].
+As with data (see Fitzpatrick in this volume), sharing code in public takes a little time to ensure that it is fit for others to read.
+Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error.
+This situation of guided variation (learning by trial and error) suggests that using code is not yet widespread enough among archaeologists to propagate due to frequency-dependant biases.
 It also suggests that the professional benefits (i.e.
-the citation effects we describe above) are not yet widely known for code use  to propagate due to content biases.
-Content biases result when some aspect of a variant's content, such as positive citation effects, makes it more likely to be adopted.
-Finally there may not yet enough highly-visible, prestigous researchers using code for this behaviour due to propagate due to model-based biases [@Rendell_et_al_2011; @boyd1988culture; @henrich2007dual].
-This bias is based on imitation of highly prestigous, skilled or successful individuals.
-Considering the high variability of code using cultural evolutionary concepts, we conclude that there is unrealised potential to improve the efficiency of using and code in archaeology, firstly by converging on some widely-agreed upon conventions that will save researcher's time, and secondly by investing effort in communicating the professional benefits of using code.
+citation effects) are not yet widely known for code use behaviours to propagate due to content biases, and finally that there are not yet enough highly-visible prestigous researchers using code for this behaviour due to propagate due to model-based biases [@Rendell_et_al_2011; @boyd1988culture; @henrich2007dual].
+This suggests to us that there is unrealised potential to improve the efficiency of using and code in archaeology by converging on some widely-agreed upon conventions that save time and investing effort in communicating the professional benefits of using code.
 
@@
 The goal of rrtools is to make it easier for archaeologists and other researchers to use R for research and publication.
-This package aims to simplify many of the steps needed to write reproducible research papers, and to guide users to best practices with minimal effort.
+This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort.
 The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research.
@@
 - **use_mit_license:** This function creates a copy of the MIT software license to the compendium.
-Other licenses are also possible, such as GPL, but the MIT license is preferred beceause it is widely used for research software, and has two important qualities.
-First, it tells the reader that the author is happy for their code to be reused by others, both in academic and commercial contexts.
+This is important for two reasons.
+First, because it tells the reader that the author is happy for their code to be reused by other people.
 And second, it tells the reader that the author does not take responsibility for any problems that the reader might have when they use the code.
@@
 Using version control is important because it allows the author to time-travel back to earlier states of their analysis, if they take a dead-end path.
-It is also makes collaboration smooth because many people can work on one set of files without losing track of the most recent version.
-Plus it serves as a remote back-up of our work in case of an emergency.
+It is also important for making collaboration smooth because many people can work on one set of files without losing track of what the most recent version is.
+Plus it is nice to have a remote back-up of our work in case of an emergency.
 The GitHub repository can be kept private if desired, until the work is published.
@@
 We recognise that this file structure wont be perfect for every project.
-But our observations show that it is in wide use, and will suit most projects well.
+But our observations show that it is in wide use, and will suit many projects quite well.
 It will at least make the user think about how to organise their project logically, and to make it easier for others to navigate.
@@
 
-Starting with the discussion of Kuhn's influential work on paradigm shifts in science [@Kuhn_1962], we took inspiration from @Galison_1997 to propose a another explanation of how changes occurs in research communities, emphasising the transformative role of tools.
+Starting with the discussion of Kuhn's influential work on paradigm shifts in science [@Kuhn_1962], we take inspiration from @Galison_1997 to propose a another explanation of how changes occurs in research communities, emphasising the transformative role of tools.
 The change we focus on is the use of scripts to document and communicate data analysis.
@@
 We have outlined a positive future for archaeology where research publications are accompanied by code and data, enabling a 'critical self-consciousness' at a community-wide scale, similar to that imagined by @clarke1973archaeology as part of the discipline's 'loss of innocence'.
-A characteristic of this stage of disciplinary maturity is 'a closer understanding of its internal structure' [@clarke1973archaeology, p.
+A charactertic of this stage of disciplinary maturity is 'a closer understanding of its internal structure' [@clarke1973archaeology, p.
 7], and we can think of no more efficient and intelligible way to communicate the internals of archaeological analyses to other members of the discipline than by encoding the assumptions, logic, and calculations in scripts of an open source programming language.
 If we accept that that is necessary for the future of the discipline, we must confront two implications.
-First is the shortage of incentives to motivate the use and sharing of code.
+First is that there is currently a shortage of incentives to motivate use and sharing of code.
 Our observation is that many researchers will not change unless required to by gate-keepers at prestigious journals and funding sources.
 In other research communities we see people in these gate-keeping roles (e.g.
-in over 5,000 journals and professional organizations) effecting change by adopting the Transparency and Openness Promotion Guidelines that establish community standards of code and data availability that help to align scientific ideals with disciplinary practices [@Nosek_et_al_2015].
+in over 5,000 journals and professional organizations) effecting change by adopting the Transparency and Openness Promotion Guidelines that establish community standards of code and data availability that help to align scientific ideals with practices and disciplinary norms [@Nosek_et_al_2015].
 Simple forms of signalling, such as badges on papers to indicate that the authors have made code and data available, may also help to shift norms and raise awareness of these desirable behaviours [@Kidwell_et_al_2016].
 
-However, cultural change on a community-wide scale, such as a shift to the wide-spread use of code among archaeologists, is most likely to happen via the slow enculturation of professionals as they internalize norms during their formative years, and not via contemporaneous exposure to external cultural influences [@Vaisey_Lizardo_2016].
-This means it is unlikely that an established professional archaeologist reading this paper, or ones like it, will be motivated to use code if they were not trained to work with code at the start of their career.
-The implication here is that the most effective way to stimulate this change is during the first few years of professional training.
-This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, whereas the horizontal bar is the breadth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
-A gamma-shaped researcher has expert-level depth in archaeology, and proficient in other domains that provide skills such as analysing data with a programming language.
-A gamma-shaped researcher with their MA or PhD in archaeology may also be conversant enough in the language and culture of computer science to have conversations and collaborate in the trading zones described above, but does not need to be an expert in computer science to advance that discipline.
+However, cultural change on an community-wide scale, such as a shift to the wide-spread use of code among archaeologists, is most likely to happen via the slow enculturation of people as they internalize norms during their formative years, and not via contemporaneous exposure to external cultural influences [@Vaisey_Lizardo_2016].
+This means it is unlikely that a professional archaeologist reading this paper, or ones like it, will be motivated to use code if they were not trained to work with code at the start of their career.
+The implication here is that the most effective way to stimulate this change is during the formative years of professional training.
+This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, whereas the horizontal bar is the bredth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
+A gamma-shaped researcher has expert-level depth in archaeology, and is versed and proficient in other domains the provide skills such as analysing data with a programming language.
+A gamma-shaped researcher with their MA or PhD in archaeology may also be conversant in the language and culture of computer science enough to have conversations and collaborate in the trading zone, but does not need to be an expert in computer science to advance that discipline.
 
@@
 
-The rrtools was developed by a group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
-A version of this paper was previously presented at the 2018 Computer Applications and Quantitative Methods in Archaeology (CAA) international conference at Tubingen, Germany.
-Early drafts were improved at the Cologne writing workshop in September 2018, which was supported by the COST Action Arkwork: Archaeological Practices and Knowledge Work in the Digital Environment.
+Our project comes from the group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
 

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
resolve merge conflict

--- 2019-08-09 analysis/paper/jcaa-paper.Rmd
@@
 )
-
-# library(caa2018) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
 
@@
 What counts as suitable puzzles, and acceptable solutions to these puzzles, is governed by the norms and procedures of the prevailing paradigm.
-Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes replacing one paradigm with another.
-In this conversion or speciation process the community adopts a new way of doing science.
+Revolutions in science occur as punctuations in the equilibria of normal science, when anomalous observations culminate in changes replacing one paradigm with another in a conversion or speciation process where the community adopts a new way of doing science.
 There has been much debate about how to draw a distinction between so-called normal and revolutionary science.
 [@watkins1970against; @toulmin1970does; @Casadevalle00158_16].
-A notable criticism is that Kuhn's normal science is unscientific, as it describes a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
+A notable criticism is that Kuhn's normal science is unscientific, a situation where critical science had contracted into defensive metaphysics resulting from the domination of a ruling dogma [@popper1970].
 This is at odds with a view of science as a continuous state of evolution, with researchers simultaneously employing many styles of thinking and doing science, and continuously confronting theory with evidence and modifying their ideas based on the outcome.
@@
 The most wide-ranging application is Clark's [-@Clark_1993] comparison of North American archaeology to European archaeology during the 1970s-1990s.
-His analysis describes a multi-paradigm situation where many research communities (i.e.
+Clark, similarly to Binford and Sabloff [-@binford1982paradigms], identified the North American archaeological paradigm as rooted in anthropology, founded on post-positivist biases, and distinctive with its critical-realist ontology, a modified-objectivist epistemology, and an experimental-manipulative method.
+In contrast, the European archaeological paradigm is described a kind of history, mostly in a strict empiricist tradition, and typified by a realist ontology, an inductivist epistemology, and by an observational method.
+Clark's analysis describes a multi-paradigm situation where many research communities (i.e.
 North American and European) simultaneously operate in different paradigms (which Kuhn noted as a sign of a developing or immature discipline).
@@
 Archaeologists using LIDAR (light detection and ranging) for remote geospatial imaging of cultural landscapes have claimed the application of this technology to archaeology is a paradigm shift [@chase2012geospatial; @howey2016detecting].
-@Fuller_2010 has used the term to describe a shift in thinking from the emergence of agriculture as a  'Neolithic Revolution' to a protracted and entangled process, happening several times independently.
-Most of these claims for  Kuhnian paradigm shifts differ in meaning, which highlights an important complication with using Kuhn's concept of a paradigm-shift, namely the multiple definitions of a paradigm in his writings.
+@Fuller_2010 has used the term to describe a shift in thinking from the emergence of agriculture as a parallel, simultaneous event at the start of the Holocene (the Neolithic Revolution), to its emergence as a protracted and entangled process, happening 20 or so times independently, following different pathways.
+Most of these claims for paradigm shifts differ in meaning from Kuhn's classical formulation, and this highlights an important complication with using Kuhn's concept of a paradigm-shift, namely the multiple definitions of a paradigm in Kuhn's writings.
 
@@
 R is a widely used free and open source data analysis tool in many research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018].
-It is the dominant programming language in archaeology and ecology, so it is a good proxy for the adoption of open source scientific programming languages in these disciplines.
+Our observations suggest it is the dominant programming language in archaeology and ecology, so it is a good proxy for the adoption of open source scientific programming languages in these disciplines.
 To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.

### 2019-08-09 analysis/paper/jcaa-paper.Rmd

--- 2019-08-09 analysis/paper/jcaa-paper.Rmd
@@
 csl: journal-of-computer-applications-in-archaeology.csl
-abstract: "There is an argument in philosophy of science that revolutions in science are either ideas-driven or tools-driven.
+abstract: "There is an argument in philosophy of science that revolutions in science are either idea-driven or tool-driven.
 We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
-One of the most profound changes driven by this concern for reproducibility and transparency is from using data analysis tools dependent on point-and-clicking with a mouse in closed source software, to a set of tools based on writing scripts in open source programming languages and making the scripts openly available.
-We present bibliometric evidence for this change in ecology and archaeology to test if the adoption of these new tools may be considered a revolution.
-We identify a citation effect for archaeology articles that use the open source programming language R.
+One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on point-and-clicking with a mouse in closed source software, to a new set of tools based on writing scripts in open source programming languages and making them openly available.
+We present bibliometric evidence for this change in ecology and in archaeology to test if the adoption of these new tools may be considered a revolution or some other process.
+We identify a citation effect for papers that use the open source programming language R.
 We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of R code."
@@
 Archaeology and ecology each have a high diversity of data collection methods, and types of data analyses they conduct.
-An important similarity for the two disciplines is working with the field-collected data on a computer to prepare it for publication.
+An important similarlity for the two disciplines is working with the field-collected data on a computer to prepare it for publication.
 R is a widely used free and open source data analysis tool in many research communities [@Tippmann_2015; @Baker_2017; @Thieme_2018].
-It is the dominant programming language in archaeology and ecology, so it is a good proxy for the adoption of open source scientific programming languages in these disciplines.
+Our observations suggest it is the dominant programming language in archaeology and ecology, so it is a good proxy for the adoption of open source scientific programming languages in these disciplines.
 To investigate changes in the use of R in archaeology and ecology, we obtained reference lists from a sample of scholarly articles in the Web of Science database and examined patterns in the citation of the R program over time.
@@
 
-The rrtools was developed by a group who took part in the 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
+The rrtools package was developed by the 2017 Summer School on Reproducible Research in Landscape Archaeology group at the Freie Universität Berlin (17-21 July).
+This was funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel.
 A version of this paper was previously presented at the 2018 Computer Applications and Quantitative Methods in Archaeology (CAA) international conference at Tubingen, Germany.

### 2019-08-09 analysis/paper/jcaa-paper.Rmd
updating the data analysis

@@
 author:
-  - author 1
-  - author 2
+  - Sophie C. Schmidt
+  - Ben Marwick
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 # Note that before 2012 the authors were "R Development Core Team", and then it was simplified to "R Core Team" https://github.com/wch/r-source/commit/9a5c086855d2e8fb07f670e3315871b2de31a4fd#diff-80bb8a8b133a6204e9642d76791927e2
+#
+# We select all to get 185,121 results, then refine
+
 #

### 2019-08-09 analysis/paper/update.R

@@ -1,0 +1,149 @@
+
+# We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`. This function first appeared in 2004 https://github.com/wch/r-source/commits/9a5c086855d2e8fb07f670e3315871b2de31a4fd/src/library/base/inst/CITATION
+#
+# Note that before 2012 the authors were "R Development Core Team", and then it was simplified to "R Core Team" https://github.com/wch/r-source/commit/9a5c086855d2e8fb07f670e3315871b2de31a4fd#diff-80bb8a8b133a6204e9642d76791927e2
+#
+# We select all to get 185,121 results, then refine...
+
+# Results: 47,462
+# (from Web of Science Core Collection)
+# You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R # CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM")
+# Refined by: WEB OF SCIENCE CATEGORIES: ( ECOLOGY OR ENVIRONMENTAL SCIENCES ) AND LANGUAGES: ( ENGLISH ) AND DOCUMENT TYPES: ( ARTICLE )
+# Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.
+
+# Export... full record... tab-delimited Mac UTF-8
+# 0. 1-500 x
+# 1. 501-1000 x
+# 2. 1001-1500 x
+# 3. 1501-2000 x
+# 4. 2001-2500 x
+# 5. 2501-3000 x
+# 6. 3001-3500 x
+# 7. 3501-4000 x
+# 8. 4001-4500 x
+# 9. 4501-4747 x
+
+# field tags
+# TS= Topic
+# TI= Title
+# AU= Author [Index]
+# AI= Author Identifiers
+# GP= Group Author [Index]
+# ED= Editor
+# SO= Publication Name [Index]
+# DO= DOI
+# PY= Year Published
+# AD= Address
+# OG= Organization-Enhanced [Index]
+# OO= Organization
+# SG= Suborganization
+# SA= Street Address
+# CI= City
+# PS= Province/State
+# CU= Country/Region
+# ZP= Zip/Postal Code
+# FO= Funding Agency
+# FG= Grant Number
+# FT= Funding Text
+# SU= Research Area
+# WC= Web of Science Category
+# IS= ISSN/ISBN
[…]

### 2019-08-10 analysis/paper/jcaa-paper.Rmd
reanalysis complete, plots polished up a bit

@@
 
-# We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`.
-This function first appeared in 2004 https://github.com/wch/r-source/commits/9a5c086855d2e8fb07f670e3315871b2de31a4fd/src/library/base/inst/CITATION
+# We did a 'Cited Reference Search' in WoS for author:
+
+# "R DEV COR TEAM" OR "R # CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM"
+
+# which is the output from  `citation()`. 
+# This function first appeared in 2004 https://github.com/wch/r-source/commits/9a5c086855d2e8fb07f670e3315871b2de31a4fd/src/library/base/inst/CITATION
 #
@@
 #
-# We select all to get 185,121 results, then refine
-
-#
-# This is our exact search string for WoS:
-
-# Results: 38,040
+# we get  5,317 cited references, which are varients on how R is cited, we sort by 'citing articles', most cited at the top
+# then 'select all' to get the first 1000 matches, then 'finish search'
+
+# We select all to get 185,121 results, then refine...
+
+# Results: 47,462
 # (from Web of Science Core Collection)
 # You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R # CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM")
-# Refined by: DOCUMENT TYPES: ( ARTICLE ) AND LANGUAGES: ( ENGLISH ) AND RESEARCH AREAS: ( ENVIRONMENTAL SCIENCES ECOLOGY )
-# Timespan: 1900-2018. Indexes: SCI-EXPANDED, SSCI, A&HCI, # ESCI.
+# Refined by: WEB OF SCIENCE CATEGORIES: ( ECOLOGY OR ENVIRONMENTAL SCIENCES ) AND LANGUAGES: ( ENGLISH ) AND DOCUMENT TYPES: ( ARTICLE )
+# Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.
+
+# Export to fast 5K...  tab-delimited Win
+# 0. 1-5000 x
+# 1. 5001-10000 x
+# 2. 10001-15000 x
+# 3. 15001-20000 x
+# 4. 20001-25000 x
+# 5. 25001-30000 x
+# 6. 30001-35000 x
+# 7. 35001-40000 x
+# 8. 40001-45000 x
+# 9. 45001-47538 x
 
@@
 
-cran_cites_files <-
-    list.files(here::here("/analysis/data/raw_data/cited_cran/"),
+# ecology articles citing R
+ecology_articles_citing_r_files <-
+  list.files(here::here("/analysis/data/raw_data/post-review/ecology-articles-citing-R"),
              full.names = TRUE)
 
-cran_cites_all_areas_all_years <-
-  map(cran_cites_files,
+ecology_articles_citing_r <-
+  map(ecology_articles_citing_r_files,
       ~readr::read_tsv(.x,
@@
   mutate(PY = as.numeric(PY)) %>%
-  filter(PY <= 2017)
+  filter(PY <= 2018)
 
 # limit to top journals
-top_journals_for_cran_cites <-
-  cran_cites_all_areas_all_years %>%
+top_ecology_journals_for_cran_cites <-
+  ecology_articles_citing_r %>%
   group_by(SO) %>%
@@
 how_many_articles_in_top_journals_for_cran_cites <-
-  sum(top_journals_for_cran_cites$n)
-
+  sum(top_ecology_journals_for_cran_cites$n)
+
+# this is a hand-collected spreadsheet of the total number 
+# articles published in these journals. We get these from Web
+# of Science 'publication name' search
 top_journals_for_cran_cites_articles_per_year <-
@@
            by = "_") %>%
-  select(year, X4, value)
+  select(year, X4, value) %>% 
+  mutate(year = as.integer(year)) %>% 
+  rename( "journalname" = "X4", 
+          'PY' = "year")
 
 # tally by year
-cran_cites_all_areas_top_journals_by_year <-
-  cran_cites_all_areas_all_years %>%
-  filter(SO %in% top_journals_for_cran_cites$SO[1:10]) %>%
+cran_cites_all_areas_top_ecol_journals_by_year <-
+  ecology_articles_citing_r %>%
+  filter(SO %in% top_ecology_journals_for_cran_cites$SO[1:10]) %>%
   group_by(SO, PY) %>%
@@
          PY = as.integer(PY)) %>%
-  left_join(top_journals_for_cran_cites_articles_per_year,
-            by = c('journalname' = 'X4',
-                   'PY' = "year")) %>%
+  left_join(top_journals_for_cran_cites_articles_per_year) %>%
   mutate(prop = n / value ) %>%
@@
 articles_shown_in_plot <-
-  cran_cites_all_areas_top_journals_by_year %>%
+  cran_cites_all_areas_top_ecol_journals_by_year %>%
   filter(!is.na(prop)) %>%
@@
 # nice labels with n for plot
-cran_cites_all_areas_top_journals_by_year <-
-cran_cites_all_areas_top_journals_by_year %>%
+cran_cites_all_areas_top_ecol_journals_by_year <-
+  cran_cites_all_areas_top_ecol_journals_by_year %>%
   group_by(SO) %>%
@@
   mutate(SO_n = as.character(glue('{SO} (n = {total_n_articles})'))) %>%
-  right_join(cran_cites_all_areas_top_journals_by_year,
+  right_join(cran_cites_all_areas_top_ecol_journals_by_year,
              by = "SO")
+
 ```
@@
 Figure \@ref(fig:ecology-articles-citing-r-plot) shows the percentage of articles citing R in each of several of the top ecology journals (as defined by how often their articles are cited).
-We restrict the start of the observation period to 2007 for convenience so we have a ten year study period.
+We restrict the start of the observation period to 2008 for convenience so we have a ten year study period.
 The plot shows that the percentage of articles citing R has increased from less than ten percent in all journals in the late 2000s, to more than 30% in _Ecosphere_, _Ecology and Evolution_, and _Molecular Ecology_ after 2012.
@@
 ```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Percentage of articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}).
-Data from Web of Science for 2007-2017')}
-min_y <- 2007
-max_y <- 2025
-all_sciences_citing_R_over_time <-
-ggplot(cran_cites_all_areas_top_journals_by_year,
+Data from Web of Science for 2008-2018')}
+
+cran_cites_all_areas_top_ecol_journals_by_year <- 
+  cran_cites_all_areas_top_ecol_journals_by_year %>% 
+  mutate(SO_n = str_wrap(SO_n, 30))
+
+min_y <- 2008
+max_y <- 2030
+eco_citing_R_over_time <-
+  ggplot(cran_cites_all_areas_top_ecol_journals_by_year,
          aes(PY,
@@
   geom_text_repel(
-    data = subset(cran_cites_all_areas_top_journals_by_year,
-                  PY == max(PY)),
-    aes(label = str_wrap(SO_n, 40)),
-    size = 4,
-    nudge_x = 0,
-    hjust = 0,
-    segment.color = NA,
-    direction = "y"
-  ) +
-   scale_x_continuous(breaks = min_y:max_y,
-                     labels = c(min_y:2017, rep("", length(2018:max_y))),
-                     limits = c(min_y, max_y)) +
-  scale_y_continuous(labels = scales::percent) +
-  xlab("Publication year") +
-  ylab("Percentage of articles in that journal in that year") +
-  theme_minimal() +
-  theme(legend.position="none",
-        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
-
-ggsave(plot = all_sciences_citing_R_over_time,
-       filename = "all_sciences_citing_R_over_time.png",
-       path = str_glue('{here::here()}/analysis/figures'),
-       height = 5,
-       width = 9)
-
-knitr::include_graphics(str_glue('{here::here()}/analysis/figures/all_sciences_citing_R_over_time.png'))
-```
-
-## Looking forward to a tool-driven change in Archaeology
-
-```{r top-archaeology-journals-by-citations-of-r}
-
-# Data are from an 'advanced search' of apps.webofknowledge.com using WC=Archaeology on apps.webofknowledge.com.
-
-# Results: 44,047
-# (from Web of Science Core Collection)
-# You searched for: (WC=Archaeology) AND LANGUAGE: (English) AND DOCUMENT TYPES: # (Article)
-# Timespan: 1900-2018. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI
-
-#- get journals with the most number of articles
-
-archy_files <-
-    list.files(here::here("/analysis/data/raw_data/wos_archaeology/"),
-               full.names = TRUE)
-
-archy_all_years <-
-  map(archy_files,
-      ~readr::read_tsv(.x,
-                       quote = "",
-                       col_types = cols(.default = col_character()))) %>%
-  bind_rows() %>%
-  mutate(PY = as.numeric(PY)) %>%
-  filter(PY <= 2017)
-
-# limit to top journals
-top_journals_for_archy<-
-  archy_all_years %>%
-  group_by(SO) %>%
-  tally(sort = TRUE)
-
-# n articles in top journals per year
-top_journals_for_archy_per_year <-
-  archy_all_years %>%
-  group_by(SO, PY) %>%
-  tally() %>%
-  arrange(PY)
-
-how_many_articles_in_top_archy_journals <-
-  sum(top_journals_for_archy$n)
-
-#-- get articles citing R that are archaeology
-
-# Results: 137
-# (from Web of Science Core Collection)
-# You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM")
-# Refined by: WEB OF SCIENCE CATEGORIES: ( ANTHROPOLOGY ) AND WEB OF SCIENCE CATEGORIES: ( ARCHAEOLOGY )
-# Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI
-
-archy_cran_files <-
-      list.files(here::here("/analysis/data/raw_data/cited_cran_archaeology/"),
-               full.names = TRUE)
-
-archy_cran_all_years <-
-  map(archy_cran_files,
-      ~readr::read_tsv(.x,
-                       quote = "",
-                       col_types = cols(.default = col_character()))) %>%
-  bind_rows() %>%
-  mutate(PY = as.numeric(PY)) %>%
-  filter(PY <= 2017)
-
-# tally by year
-archy_cran_cites_all_areas_top_journals_by_year <-
-  archy_cran_all_years %>%
-  filter(SO %in% top_journals_for_archy$SO) %>%
-  group_by(SO, PY) %>%
-  tally(sort = TRUE)  %>%
-  left_join(top_journals_for_archy_per_year,
-            by = c('SO', 'PY')) %>%
-   mutate(prop = n.x / n.y) %>%
-  ungroup()
-
-total_number_archaeology_articles_citing_r <- sum(archy_cran_cites_all_areas_top_journals_by_year$n.x)
-
-# check for slight differences in journal names
-# sort(unique(archy_cran_cites_all_areas_top_journals_by_year$SO))
-
-# nice labels with n for plot
-archy_cran_cites_all_areas_top_journals_by_year <-
-archy_cran_cites_all_areas_top_journals_by_year %>%
-  group_by(SO) %>%
-  summarise(total_n_articles = sum(n.x)) %>%
-  mutate(SO_n = as.character(glue('{SO} (n = {total_n_articles})'))) %>%
-  right_join(archy_cran_cites_all_areas_top_journals_by_year,
-             by = "SO")
-
-min_y <- 2007
-max_y <- 2040
-
-archy_cran_cites_all_areas_top_journals_by_year_to_plot <- 
- archy_cran_cites_all_areas_top_journals_by_year # %>% 
- # filter(SO != "AFRICAN ARCHAEOLOGICAL REVIEW")
-
-all_archaeology_citing_R_over_time <-
-  ggplot(archy_cran_cites_all_areas_top_journals_by_year_to_plot,
-       aes(PY,
-           prop,
-           colour = SO)) +
-  geom_line(size = 2) +
-  geom_point() +
-  geom_text_repel(
-    data = archy_cran_cites_all_areas_top_journals_by_year_to_plot %>%
-      group_by(SO) %>%
-      slice(which.max(PY)) ,
+    data = cran_cites_all_areas_top_ecol_journals_by_year %>% 
+                  filter(PY == max(PY)),
     aes(label = SO_n),
     direction = "y",
-    size = 6,
+    size = 4,
     nudge_x = 1,
@@
                                    vjust = 0.5)) +
-  scale_x_continuous(breaks = min_y:2017,
+  scale_x_continuous(breaks = min_y:2018,
                      limits = c(min_y, max_y)) +
   scale_y_continuous(labels = scales::percent,
-                     limits = c(0, 0.2)) +
+                     limits = c(0, 0.7)) +
+  xlab("Publication year") +
+  ylab("Percentage of articles in that journal in that year") +
+  lemon::coord_capped_cart(bottom='right')
+
+ggsave(plot = eco_citing_R_over_time,
+       filename = "eco_citing_R_over_time.png",
+       path = str_glue('{here::here()}/analysis/figures'),
+       height = 5,
+       width = 7)
+
+knitr::include_graphics(str_glue('{here::here()}/analysis/figures/eco_citing_R_over_time.png'))
+```
+
+## Looking forward to a tool-driven change in Archaeology
+
+```{r top-archaeology-journals-by-citations-of-r}
+
+# Data are from an 'advanced search' of apps.webofknowledge.com using WC=Archaeology on apps.webofknowledge.com.
+
+# Results: 44,047
+# (from Web of Science Core Collection)
+# You searched for: (WC=Archaeology) AND LANGUAGE: (English) AND DOCUMENT TYPES: # (Article)
+# Timespan: 1900-2018. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI
+
+#- get journals with the most number of articles
+
+archy_files <-
+    list.files(here::here("/analysis/data/raw_data/post-review/archy-articles-in-wos"),
+               full.names = TRUE)
+
+archy_all_years <-
+  map(archy_files,
+      ~readr::read_tsv(.x,
+                       quote = "",
+                       col_types = cols(.default = col_character()))) %>%
+  bind_rows() %>%
+  mutate(PY = as.numeric(PY)) %>%
+  filter(PY <= 2017) %>% 
+  mutate(SO = ifelse(SO == "ASIAN PERSPECTIVES-THE JOURNAL OF ARCHAEOLOGY FOR ASIA AND THE PACIFIC",
+                     "ASIAN PERSPECTIVES", SO)) %>% 
+  mutate(SO = ifelse(SO == "GEOARCHAEOLOGY-AN INTERNATIONAL JOURNAL",
+                     "GEOARCHAEOLOGY", SO)) %>% 
+  mutate(SO = ifelse(SO == "AZANIA-ARCHAEOLOGICAL RESEARCH IN AFRICA" ,
+                     "AZANIA", SO))
+
+# limit to top journals
+top_journals_for_archy<-
+  archy_all_years %>%
+  group_by(SO) %>%
+  tally(sort = TRUE)
+
+# n articles in top journals per year
+top_journals_for_archy_per_year <-
+  archy_all_years %>%
+  group_by(SO, PY) %>%
+  tally() %>%
+  arrange(PY)
+
+how_many_articles_in_top_archy_journals <-
+  sum(top_journals_for_archy$n)
+
+#-- get articles citing R that are archaeology
+
+# We did a 'Cited Reference Search' in WoS for author:
+
+# "R DEV COR TEAM" OR "R # CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM"
+
+# which is the output from  `citation()`. 
+# This function first appeared in 2004 https://github.com/wch/r-source/commits/9a5c086855d2e8fb07f670e3315871b2de31a4fd/src/library/base/inst/CITATION
+#
+# Note that before 2012 the authors were "R Development Core Team", and then it was simplified to "R Core Team" https://github.com/wch/r-source/commit/9a5c086855d2e8fb07f670e3315871b2de31a4fd#diff-80bb8a8b133a6204e9642d76791927e2
+#
+# we get  5,317 cited references, which are varients on how R is cited, we sort by 'citing articles', most cited at the top
+# then 'select all' to get the first 1000 matches, then 'finish search'
+
+# We select all to get 185,121 results, then refine...
+# WOS research area "Archaeology"
+
+# Results: 
+# (from Web of Science Core Collection)
+# You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R # CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM")
+# Refined by: LANGUAGES: ( ENGLISH ) AND DOCUMENT TYPES: ( ARTICLE ) AND RESEARCH AREAS: ( ARCHAEOLOGY )
+# Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.
+
+# Export to fast 5K...  tab-delimited Win
+
+# Results: 262
+
+archy_cran_files <-
+  list.files(here::here("/analysis/data/raw_data/post-review/archy-articles-citing-R"),
+             full.names = TRUE)
+
+archy_cran_all_years <-
+  map(archy_cran_files,
+      ~readr::read_tsv(.x,
+                       quote = "",
+                       col_types = cols(.default = col_character()))) %>%
+  bind_rows() %>%
+  mutate(PY = as.numeric(PY)) %>%
+  filter(PY <= 2018) %>% 
+  mutate(SO = ifelse(SO == "ASIAN PERSPECTIVES-THE JOURNAL OF ARCHAEOLOGY FOR ASIA AND THE PACIFIC",
+                     "ASIAN PERSPECTIVES", SO)) %>% 
+  mutate(SO = ifelse(SO == "GEOARCHAEOLOGY-AN INTERNATIONAL JOURNAL",
+                     "GEOARCHAEOLOGY", SO)) %>% 
+  mutate(SO = ifelse(SO == "AZANIA-ARCHAEOLOGICAL RESEARCH IN AFRICA" ,
+                     "AZANIA", SO))
+
+# tally by year
+archy_cran_cites_all_areas_top_journals_by_year <-
+  archy_cran_all_years %>%
+  filter(SO %in% {archy_cran_all_years %>% 
+           count(SO, sort = TRUE)  %>% 
+           slice(1:25) %>% 
+           pull(SO)}
+         ) %>%
+  group_by(SO, PY) %>%
+  tally(sort = TRUE)  %>%
+  left_join(top_journals_for_archy_per_year,
+            by = c('SO', 'PY')) %>%
+  mutate(prop = n.x / n.y) %>%
+  ungroup() %>% 
+  filter(!is.na(prop))
+
+total_number_archaeology_articles_citing_r <- sum(archy_cran_cites_all_areas_top_journals_by_year$n.x, na.rm = TRUE)
+
+# check for slight differences in journal names
+# sort(unique(archy_cran_cites_all_areas_top_journals_by_year$SO))
+
+# nice labels with n for plot
+archy_cran_cites_all_areas_top_journals_by_year <-
+  archy_cran_cites_all_areas_top_journals_by_year %>%
+  group_by(SO) %>%
+  summarise(total_n_articles = sum(n.x)) %>%
+  mutate(SO_n = as.character(glue('{SO} (n = {total_n_articles})'))) %>%
+  right_join(archy_cran_cites_all_areas_top_journals_by_year,
+             by = "SO")
+
+min_y <- 2008
+max_y <- 2021
+label_text_size = 2
+
+archy_cran_cites_all_areas_top_journals_by_year_to_plot <- 
+  archy_cran_cites_all_areas_top_journals_by_year # %>% 
+# filter(SO != "AFRICAN ARCHAEOLOGICAL REVIEW")
+
+all_archaeology_citing_R_over_time <-
+  ggplot(archy_cran_cites_all_areas_top_journals_by_year_to_plot,
+         aes(PY,
+             prop,
+             colour = SO)) +
+  scale_x_continuous(breaks = min_y:2018,
+                     limits = c(min_y, max_y)) +
+  scale_y_continuous(limits = c(0, 0.2)) +
+  geom_line(size = 2) +
+  geom_point() +
+  geom_text_repel(
+    data = archy_cran_cites_all_areas_top_journals_by_year_to_plot %>%
+      group_by(SO) %>%
+      slice(which.max(PY)) %>% 
+      mutate(SO_n = str_wrap(SO_n, 60)),
+    aes(label = SO_n),
+    hjust = -1,
+    direction = "y",
+    size = label_text_size,
+    nudge_x = 2) +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(), 
+        axis.line = element_line(colour = "black"),
+        axis.title.x = element_text(hjust = 0.5),
+        legend.position="none",
+        axis.text.x = element_text(angle = 90,
+                                   hjust = 1,
+                                   vjust = 0.5)) +
   xlab("Publication year") +
@@
        height = 7,
-       width = 11)
+       width = 7)
 ```
@@
   archy_cran_cites_all_areas_top_journals_by_year$total_n_articles[1]
+
+labels_and_breaks <- 
+  sort(unique(archy_cran_cites_all_areas_top_journals_by_year$PY))
 
@@
   xlab("year") +
-  ylab("Proportion of articles citing R") +
-  ggtitle(expression(paste("Articles citing R in ",
-                           italic("Journal of Archaeological Science")))) +
-  theme_bw(base_size = 10)
+  scale_x_continuous(breaks = labels_and_breaks,
+                     labels = labels_and_breaks) +
+  ylab("Proportion of articles citing R per year") +
+  theme_bw(base_size = 14) +
+  theme(axis.text.x = element_text(angle = 90,
+                                   hjust = 1,
+                                   vjust = 0.5))
+
+ggsave(plot = jas_cites_r,
+       filename = "jas_cites_r.png",
+       path = str_glue('{here::here()}/analysis/figures'),
+       height = 7,
+       width = 11)
 ```
@@
 
-```{r archaeology-journals-with-subplot, fig.cap=str_glue('Percentage of Archaeology articles per year citing R ({total_number_archaeology_articles_citing_r} out of {prettyNum(how_many_articles_in_top_archy_journals, ",")} articles in sample.
-Data from Web of Science for 2007-2017')}
-
+```{r archaeology-journals-with-subplot, fig.cap=str_glue('Proportion of Archaeology articles per year citing R ({total_number_archaeology_articles_citing_r} out of {prettyNum(how_many_articles_in_top_archy_journals, ",")} articles in sample.
+Sub-plot shows articles in the _Journal of Archaeological Science_.
+Data from Web of Science for 2008-2018')}
 # combine main plot and subplot
@@
   annotation_custom(grob = ggplotGrob(jas_cites_r),
-                          xmin = 2027, xmax = 2040,
-                          ymin = 0.10, ymax = 0.20)
-
+                          xmin = 2008, xmax = 2015,
+                          ymin = 0.08, ymax = 0.20)
 ggsave(plot = all_archaeology_citing_R_over_time_and_subplot,
@@
        width = 12)
-
 knitr::include_graphics(str_glue('{here::here()}/analysis/figures/all_archaeology_citing_R_over_time_and_subplot.png'))
@@
 The sample sizes for the archaeology journals are too small to confidently infer any trend of over time, with the exception of the _Journal of Archaeological Science_ with `r how_many_articles_in_jas` articles.
-The inset plot in Figure \@ref(fig:archaeology-journals-with-subplot) shows a statistical test of change over time for articles citing R in the _Journal of Archaeological Science_.
-With a moderate r-squared value and low p-value for the linear model, we conclude there is some evidence for an increase in citations of R over time in this journal.
+Figure \@ref(fig:j-arch-sci-articles-citing-R-plot) shows a statistical test of change over time for articles citing R in the _Journal of Archaeological Science_.
+With a moderate r-squared value and low p-value for the linear model, we conclude there is some evidence for a non-random increase in citations of R over time in this journal.
 This may show an increasing use of R among archaeologists, especially those working at the intersection of archaeology and the natural sciences.
@@
 
-frequency <-
+word_frequency <-
   tidy_cites %>%
@@
 
-frequency <-
-  frequency %>%
+word_frequency <-
+  word_frequency %>%
   select(ID, word, freq) %>%
@@
   arrange(no_R_cites, has_R_cites)
+
+hm_words <- prettyNum(nrow(word_frequency), big.mark  = ",")
+hm_articles <- prettyNum(nrow(archy_journal_articles), big.mark  = ",")
 
@@
 
-```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles of archaeology articles ({prettyNum(nrow(frequency), ",")} words in {prettyNum(nrow(archy_journal_articles), ",")} articles)")}
+```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles of archaeology articles ({hm_words} words in {hm_articles} articles)")}
 
 archaeology_articles_word_freqs <-
-ggplot(frequency,
+ggplot(word_frequency,
        aes(no_R_cites,
@@
 
-archy_ctv_readme <- readLines(str_glue('{here::here()}/analysis/data/raw_data/ctv-archaeology-README.md'))
-
-journals <- c("Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science|PLOS|Journal of Human Evolution|Asian Perspectives|Scientific Reports|PLoS ONE|The Holocene|Archaeology in Oceania|Quaternary International|Internet Archaeology|Quaternary Science Reviews|Quaternary International|Journal of Archaeological Science: Reports|Open Quaternary|Evolution and Human Behavior|PaleoAnthropology")
-
-top_archy_journals <-
- str_subset(archy_ctv_readme, journals)
-
-top_archy_journal_names <-
-  unlist(str_extract_all(archy_ctv_readme, journals))
-
-top_archy_journal_names <- ifelse(top_archy_journal_names == "J Archaeol Sci",
-                                  "Journal of Archaeological Science",
-                                  ifelse(top_archy_journal_names == "PLOS",
-                                  "PLoS ONE",
-                                  top_archy_journal_names))
-
-archy_ctv_readme <- top_archy_journals
+archy_ctv_readme <- readLines(str_glue('{here::here()}/analysis/data/raw_data/post-review/README.md'))
+
+# get just the articles
+archy_ctv_readme_start <- str_which(archy_ctv_readme, " Publications that include R code")
+archy_ctv_readme <- 
+  archy_ctv_readme[archy_ctv_readme_start:length(archy_ctv_readme)]
+
+# get all dates of publication
 archy_ctv_readme <- str_remove_all(archy_ctv_readme, "[[:punct:]]")
@@
 archy_ctv_readme_20XX <- as.numeric(archy_ctv_readme_20XX)
-archy_ctv_readme_20XX <- archy_ctv_readme_20XX[archy_ctv_readme_20XX > 2010]
+archy_ctv_readme_20XX <- archy_ctv_readme_20XX[!is.na(archy_ctv_readme_20XX)]
 
@@
 
-data_to_plot <- data_frame(year = archy_ctv_readme_20XX,
-                  journal_name = top_archy_journal_names) %>%
-  filter(!is.na(year),
-         year <= 2017) %>%
-  mutate(year = factor(year))
+# get a few journal names
+archy_ctv_readme_journals <- str_which(archy_ctv_readme, glue_collapse(archy_ctv_readme_20XX, "|"))
+archy_ctv_readme_journals <- archy_ctv_readme[archy_ctv_readme_journals]
+
+journals <- c("Journal of Archaeological Method and Theory|J Archaeol Sci|Journal of Anthropological Archaeology|Journal of Archaeological Science|PLOS|Journal of Human Evolution|Asian Perspectives|Scientific Reports|PLoS ONE|The Holocene|Archaeology in Oceania|Quaternary International|Internet Archaeology|Quaternary Science Reviews|Quaternary International|Journal of Archaeological Science: Reports|Open Quaternary|Evolution and Human Behavior|PaleoAnthropology|Archaeological and Anthropological Sciences")
+
+top_archy_journals <-
+  tibble( citation = archy_ctv_readme_journals, 
+          year = archy_ctv_readme_20XX) %>% 
+  mutate(journal_name = str_extract(tolower(citation), tolower(journals))) %>% 
+  mutate(journal_name = ifelse(journal_name == "j archaeol sci",
+                                  "journal of archaeological science",
+                                  ifelse(journal_name == "plos",
+                                  "plos one",
+                                  journal_name))) %>% 
+  mutate(journal_name = str_to_title(journal_name)) %>% 
+  mutate(journal_name = str_replace(journal_name, "And", "and")) %>% 
+  mutate(journal_name = str_replace(journal_name, "Of", "of")) %>% 
+  mutate(journal_name = str_replace(journal_name, "In ", "in ")) %>% 
+  mutate(journal_name = str_replace(journal_name, "Plos", "PLOS")) %>% 
+  mutate(journal_name = str_replace_na(journal_name, "Other")) %>% 
+  mutate(year = factor(year)) %>% 
+  add_count(journal_name) %>% 
+  mutate(journal_name = str_glue('{journal_name} (n = {n})'))
 ```
@@
 For example, the journal *Nature Neuroscience* requires authors 'to make the code that supports the generation of key figures in their manuscript available for review' [@Editors_2017].
-Although we are not aware of attempts to quantify this change in the same way we have for archaeology, we have observed a number of recent publications that describe and recommend code sharing in statistics [@Baumer2014], genome biology [@markowetz2015five], computational biology [@Sandve2013], biostatistics [@Peng2009], computer science [@peng2011reproducible; @Mitchell2012], applied mathematics [@LeVeque2012], speech science [@Abari2012], political science [@king1995replication; @Dafoe2014], and the social sciences generally [@Miguel2014].
+Although we are not aware of attempts to quantify this change in the same way we have for archaeology, we have observed a number of recent publications that describe and recommend code sharing in statistics [@Baumer2014], genome biology [@markowetz2015five], computational biology [@Sandve2013], hydrology [@Slater_et_al_2019], biostatistics [@Peng2009], computer science [@peng2011reproducible; @Mitchell2012], applied mathematics [@LeVeque2012], speech science [@Abari2012], political science [@king1995replication; @Dafoe2014], and the social sciences generally [@Miguel2014].
 As part of this growing interest in using code for research we also see manifestos aimed at researchers doing any kind of quantitative work [@nosek2015promoting; @Ince2012; @Barnes2010] and articles recommending using R in undergraduate education [@Bray2014; @Eglen2009].
@@
 
-```{r archaeology-reproducible-articles-plot, fig.cap=str_glue('Articles in archaeology journals using R for reproducible research (n = {number_of_reproducible_articles})')}
+```{r archaeology-reproducible-articles-plot, fig.cap=str_glue('Articles in archaeology journals using R for reproducible research, and making code files openly available to accompany the published article (n = {number_of_reproducible_articles})')}
 
 archaeology_articles_r_reproducible <-
-ggplot(data_to_plot,
+ggplot(top_archy_journals,
        aes(year,
@@
   geom_bar(position = "stack") +
-  scale_y_continuous(breaks = 1:10) +
+  scale_y_continuous(breaks = 1:30) +
   scale_fill_discrete(name = "Journal")  +
@@
   ylab("Number of articles") +
-  theme_minimal(base_size = 18) +
-  theme(legend.text=element_text(size=16))
+  theme_bw(base_size = 14) +
+  theme(legend.text=element_text(size=14),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
+        legend.justification = c(0, 1), 
+        legend.position = c(0.05, 0.98),
+        legend.spacing.y = unit(0.65, 'cm'),
+        legend.key.size = unit(0.65, "cm"))
 
@@
   summarise(median = median(TC)) %>%
-  filter(PY >= 2010)
+  filter(PY >= 2010) %>% 
+  filter(PY < 2018)
 
@@
                      "Cites R", "Doesn't cite R")) %>%
+  select(ID, PY, TC) %>% 
   filter(PY >= 2010) %>% 
-  select(ID, PY, TC)
+  filter(PY < 2018)
 
@@
 test_explanation <- 
-str_glue("On average, articles citing R \nhave significantly higher numbers\nof citations (m = {dd_cites_mean}) than articles\nthat do not (m =  {dd_no_cites_mean})\nt({df}) = {tvalue}, p =  {pvalue}")
+str_glue("On average, articles citing R have higher numbers of citations (m = {dd_cites_mean}) than articles that do not (m =  {dd_no_cites_mean}) t({df}) = {tvalue}, p =  {pvalue}")
 
@@
          aes(ID,
-             median)) +
+             median, 
+             fill = ID)) +
   geom_col() +
+  scale_fill_manual(values  =c("green", "red"),
+                    guide = FALSE) +
   facet_wrap( ~ PY) +
@@
   xlab("")
+
+ggsave(plot = archy_journal_articles_citations_plot,
+       filename = "archy_journal_articles_citations_plot.png",
+       path = str_glue('{here::here()}/analysis/figures'),
+       height = 9,
+       width = 16)
+
 ```
@@
 ```{r results='hide'}
-# add text to empty corner
-png(here::here("analysis", "figures", "archy_journal_articles_citations_plot_cap.png"),
-    width = 20, height = 15, units = "cm", res = 300)
-print(archy_journal_articles_citations_plot)
-library(grid)
-grid.text(test_explanation, 
-          x = 0.70, y = 0.25, 
-          gp=gpar(fontsize=11),
-          just = 'left')
-invisible(dev.off())
-```
-
-```{r  citation-advantage-plot, fig.cap=str_glue('Median citation rates per year for archaeology articles 2010-2017 that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")})')}
-knitr::include_graphics(here::here("analysis", "figures", "archy_journal_articles_citations_plot_cap.png"))
+
+```
+
+```{r  citation-advantage-plot, fig.cap=str_glue('Median citation rates per year for archaeology articles 2010-2017 that cite R (n = {prettyNum(archy_journal_articles_tally[1], ",")}) and articles that do not cite R (n = {prettyNum(archy_journal_articles_tally[2], ",")}).
+{test_explanation}.')}
+knitr::include_graphics(here::here("analysis", "figures", "archy_journal_articles_citations_plot.png"))
 ```

### 2019-08-10 analysis/paper/update.R

@@ -1,149 +1,0 @@
-
-# We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`. This function first appeared in 2004 https://github.com/wch/r-source/commits/9a5c086855d2e8fb07f670e3315871b2de31a4fd/src/library/base/inst/CITATION
-#
-# Note that before 2012 the authors were "R Development Core Team", and then it was simplified to "R Core Team" https://github.com/wch/r-source/commit/9a5c086855d2e8fb07f670e3315871b2de31a4fd#diff-80bb8a8b133a6204e9642d76791927e2
-#
-# We select all to get 185,121 results, then refine...
-
-# Results: 47,462
-# (from Web of Science Core Collection)
-# You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R # CORE" OR "R CORE TEAM" OR "R DEVELOPMENT CORE TEAM")
-# Refined by: WEB OF SCIENCE CATEGORIES: ( ECOLOGY OR ENVIRONMENTAL SCIENCES ) AND LANGUAGES: ( ENGLISH ) AND DOCUMENT TYPES: ( ARTICLE )
-# Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.
-
-# Export... full record... tab-delimited Mac UTF-8
-# 0. 1-500 x
-# 1. 501-1000 x
-# 2. 1001-1500 x
-# 3. 1501-2000 x
-# 4. 2001-2500 x
-# 5. 2501-3000 x
-# 6. 3001-3500 x
-# 7. 3501-4000 x
-# 8. 4001-4500 x
-# 9. 4501-4747 x
-
-# field tags
-# TS= Topic
-# TI= Title
-# AU= Author [Index]
-# AI= Author Identifiers
-# GP= Group Author [Index]
-# ED= Editor
-# SO= Publication Name [Index]
-# DO= DOI
-# PY= Year Published
-# AD= Address
-# OG= Organization-Enhanced [Index]
-# OO= Organization
-# SG= Suborganization
-# SA= Street Address
-# CI= City
-# PS= Province/State
-# CU= Country/Region
-# ZP= Zip/Postal Code
-# FO= Funding Agency
-# FG= Grant Number
-# FT= Funding Text
-# SU= Research Area
-# WC= Web of Science Category
-# IS= ISSN/ISBN
[…]

### 2019-08-12 analysis/paper/jcaa-paper.Rmd
ideas-driven and tools-driven to idea-driven and tool-driven

@@
 We ask whether this might be the beginning of a broader change in the field.
-Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are ideas-driven or tools-driven [@Galison_1997; @Dyson_2000].
+Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there is an argument among philosophers of science whether revolutions in science are idea-driven or tool-driven [@Galison_1997; @Dyson_2000].
 This argument motivated us to explore how the disciplines of ecology and archaeology have started to embrace tools that improve the reproducibility of research.
@@
 
-# Revolutions in Science: Ideas-driven or Tools-driven?
-
-Philosophers of science disagree whether revolutions in science are ideas-driven or tool-driven.
-One of the most widely-known models of ideas-driven change in science is Kuhn's [-@Kuhn_1962] effort to describe the history of science by proposing paradigm shifts resulting from revolutionary change in communities of researchers.
+# Revolutions in Science: idea-driven or tool-driven?
+
+Philosophers of science disagree whether revolutions in science are idea-driven or tool-driven.
+One of the most widely-known models of idea-driven change in science is Kuhn's [-@Kuhn_1962] effort to describe the history of science by proposing paradigm shifts resulting from revolutionary change in communities of researchers.
 In brief, this model proposes that the history of science consists of long periods of tradition-bound 'normal science' punctuated by short episodes of 'revolution'.

### 2019-08-12 analysis/paper/jcaa-paper.Rmd
typos.

@@
 In our data all the article authors are also the code authors, so there is no popularity effect noted by Vandewalle.
-One possibilty is that authors preferentially tend to promote their articles containing code, beacause they want a greater return for the greater effort they have put into including the code with the article, this may be equivalent to the self-selection bias postulated by Vandewalle.
-Another possiblity is that papers with code are perceived by readers to be more robust because of their transparency, and thus a more reliable and credible item to cite in support of knowledge claims.
+One possibility is that authors preferentially tend to promote their articles containing code, because they want a greater return for the greater effort they have put into including the code with the article, this may be equivalent to the self-selection bias postulated by Vandewalle.
+Another possibility is that papers with code are perceived by readers to be more robust because of their transparency, and thus a more reliable and credible item to cite in support of knowledge claims.
 Another possible cause is that when a paper is accompanied by code (for example on a trustworthy repository) it is more visible than a paper without code, and appears higher in search result rankings.
-All other factors equal, higher visibility leads to more citations because the reseracher will stop searching when they have found the first paper that serves their purpose.
+All other factors equal, higher visibility leads to more citations because the researcher will stop searching when they have found the first paper that serves their purpose.
 Further work is necessary to understand the mechanisms of this citation effect.
-Additional insights might come from comparisions of papers that use another open source language such as Python, and MATLAB, which is not an open source language.
+Additional insights might come from comparisons of papers that use another open source language such as Python, and MATLAB, which is not an open source language.
 
@@
 In his cultural cartography of science, Gieryn examines boundaries between science generally and varieties of less authoritative non-science.
-He argues that the boundaries are drawn and redrawn contiuously in flexible and sometimes ambiguous ways.
+He argues that the boundaries are drawn and redrawn continuously in flexible and sometimes ambiguous ways.
 Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims made by their competitors in the scientific community.
@@
 statistical tests with high probabilities of rejecting false null hypotheses) capable of distinguishing between superior and inferior paradigms are rare in the social sciences, such as economics [@Ioannidis_Stanley_Doucouliagos_2017], compared to the physical sciences.
-Modelling and case studies described by @Akerlof_Michaillat_2018 indicate that among discplines lacking evidence to routinely strongly discriminate between theories, the chances of getting trapped in an inferior paradigm are high.
+Modelling and case studies described by @Akerlof_Michaillat_2018 indicate that among disciplines lacking evidence to routinely strongly discriminate between theories, the chances of getting trapped in an inferior paradigm are high.
 In this situation, true theories may not be adopted at all and the field is at risk of being captured by false paradigms.
@@
 Although a discipline-wide paradigm shift might be off the table for archaeology, there is a need for methods to share research in a way that is reproducible for others, and that archaeologists are actively exploring tools to make this possible.
-Our resuts show that scripting languages and data repositories to share the scripts is a solution that enjoys increasing popularity in the archaeological research community [cf.
+Our results show that scripting languages and data repositories to share the scripts is a solution that enjoys increasing popularity in the archaeological research community [cf.
 @marwick_standard_2018].
@@
 We keep the raw data (i.e.
-field-collected and instrument output) seprate from data that is created or derived during the analysis.
+field-collected and instrument output) separate from data that is created or derived during the analysis.
 These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.
@@
 We use the R package structure as the standard for our directory because it is widely recognised by R users [@wickham2015r].
-This means that people know where to look for code, where to look for data, and where to look for dependancies and other information.
+This means that people know where to look for code, where to look for data, and where to look for dependencies and other information.
 - **use_mit_license:** This function creates a copy of the MIT software license to the compendium.
-Other licenses are also possible, such as GPL, but the MIT license is preferred beceause it is widely used for research software, and has two important qualities.
+Other licenses are also possible, such as GPL, but the MIT license is preferred because it is widely used for research software, and has two important qualities.
 First, it tells the reader that the author is happy for their code to be reused by others, both in academic and commercial contexts.
@@
 Archaeologists have more recently taken up R, but to a much lesser degree, so far, than what we see in the other fields, and most commonly in more scientific aspects of archaeology.
-We identifed an advantage of citing R for researchers as it leads to being referenced more often.
+We identified an advantage of citing R for researchers as it leads to being referenced more often.
 We also identified a small, but increasing, set of archaeology papers that use R to make their work fully reproducible.

### 2019-08-12 analysis/paper/jcaa-paper.Rmd

@@
 We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
-One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on point-and-clicking with a mouse in closed source software, to a new set of tools based on writing scripts in open source programming languages and making them openly available.
-We present bibliometric evidence for this change in ecology and in archaeology to test if the adoption of these new tools may be considered a revolution or some other process.
-We identify a citation effect for papers that use the open source programming language R.
+One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on point-and-clicking with a mouse in closed source software, to tools based on writing scripts in open source programming languages and making them openly available.
+We present bibliometric evidence for this change in ecology and in archaeology to test if the adoption of these new tools is revolutionary or transformational.
+We identify a positive citation effect for papers that use the open source programming language R.
 We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of R code."
@@
 Bibliometric evidence shows a strong increase in the use of R among ecologists and the beginning of a similar development in archaeology.
-We evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be a tool-driven change in archaeology.
+We evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might reflect a tool-driven change in archaeology.
 Acknowledging that this is not a simple process, we offer an R-based tool to ease the task of creating a compendium which enables other researchers to reproduce the published results.
@@
                      limits = c(min_y, max_y)) +
-  scale_y_continuous(labels = scales::percent,
-                     limits = c(0, 0.7)) +
+  scale_y_continuous(limits = c(0, 0.7)) +
   xlab("Publication year") +
-  ylab("Percentage of articles in that journal in that year") +
+  ylab("Proportion of articles in that journal in that year") +
   lemon::coord_capped_cart(bottom='right')
@@
         axis.line = element_line(colour = "black"),
-        axis.title.x = element_text(hjust = 0.5),
+        axis.title.x = element_text(hjust = 0.4),
         legend.position="none",
@@
   xlab("Publication year") +
-  ylab("Percentage of articles in that journal") +
+  ylab("Proportion of articles in that journal") +
   lemon::coord_capped_cart(bottom='right')
@@
 To compare with ecology we conducted a similar bibliometric analysis for archaeology journals indexed by the Web of Science service.
-We used the 'Cited Reference Search' to find articles citing R in the same way as above, and then refined the results to keep only those articles published during 2007-2017 that are included in the Web of Science category 'Archaeology'.
+We used the 'Cited Reference Search' to find articles citing R in the same way as above, and then refined the results to keep only those articles published during 2008-2018 that are included in the Web of Science category 'Archaeology'.
 This resulted in `r prettyNum(how_many_articles_in_top_archy_journals, ",")` articles, of which `r total_number_archaeology_articles_citing_r` cite R.
@@
 
-```{r archaeology-journals-with-subplot, fig.cap=str_glue('Proportion of Archaeology articles per year citing R ({total_number_archaeology_articles_citing_r} out of {prettyNum(how_many_articles_in_top_archy_journals, ",")} articles in sample.
-Sub-plot shows articles in the _Journal of Archaeological Science_.
-Data from Web of Science for 2008-2018')}
+```{r archaeology-journals-with-subplot, fig.cap=str_glue('Proportion of Archaeology articles per year citing R (a total of {total_number_archaeology_articles_citing_r} out of {prettyNum(how_many_articles_in_top_archy_journals, ",")} articles in our sample for 2008-2018).
+Sub-plot shows articles published in the _Journal of Archaeological Science_ during 2008-2017.
+Data from Web of Science')}
 # combine main plot and subplot
@@
 The sample sizes for the archaeology journals are too small to confidently infer any trend of over time, with the exception of the _Journal of Archaeological Science_ with `r how_many_articles_in_jas` articles.
-Figure \@ref(fig:j-arch-sci-articles-citing-R-plot) shows a statistical test of change over time for articles citing R in the _Journal of Archaeological Science_.
+Figure \@ref(fig:archaeology-journals-with-subplot) shows a statistical test of change over time for articles citing R in the _Journal of Archaeological Science_.
 With a moderate r-squared value and low p-value for the linear model, we conclude there is some evidence for a non-random increase in citations of R over time in this journal.
@@
 'the', 'archaeological', 'study', etc.).
-This resulted in `r prettyNum(nrow(frequency), ',')` words in `r prettyNum(nrow(archy_journal_articles), ",")` articles.
+This resulted in `r hm_words` words in `r hm_articles` articles.
 Third, for each word found in all the titles of the journal articles in each group, we computed its proportion of the total number of words in all the titles in each group.
@@
 On the other hand, articles that do not cite R more often discuss the Roman and Bronze Age periods, and less often reflect a technical specialisation.
-If there is one technical specialisation that is characteristic of non-citing R articles, it seems to be in ceramics.
+If there is one technical specialisation that is characteristic of articles that do not cite R, it seems to be in ceramics.
 
@@
 
-These archaeology articles that include publicly available R code and data are part of a shift in scientific communication that has also been observed in other fields.
+These archaeology articles that include publicly available R code and data are part of a shift in scientific communication that is also underway in other fields.
 For example, the journal *Nature Neuroscience* requires authors 'to make the code that supports the generation of key figures in their manuscript available for review' [@Editors_2017].
@@
 We expect that openly shared code will speed the widespread adoption of reproducibility as a core tenet of the scientific process, since it frees researchers from the black box that most mouse-driven programs are, and enables researchers to not only rerun the shared analysis, but to gain access to all parameter settings, empowering them to change these and so properly evaluate, extend and reuse the published results.
-As open source languages such as R (and Python) are free to use, and most trustworthy repositories are also free (such as Zenodo, Open Science Framework, Figshare, etc.), then even researchers with limited resources anywhere in the world can contribute equally to the research community by using and sharing code.
+As open source languages such as R (and Python) are free to use, and many trustworthy repositories are also free (such as Zenodo, Open Science Framework, Figshare, etc.), then even researchers with limited resources anywhere in the world can contribute equally to the research community by using and sharing code.
 Using and sharing open source code is thus an important action for reducing inequality in the archaeological research community.
-It enables participation by researchers who are unable to buy software licenses or access face-to-face training.
+It enables participation by researchers who are unable to buy software licenses.
 
@@
 @Vandewalle2012 analysed citation rates for 645 articles published in _IEEE Transactions on Image Processing_ during 2004–2006.
-He found statistically significant differences between the median number of citations for papers that have code available and papers that do not.
+He found non-random differences between the median number of citations for papers that have code available and papers that do not.
 In considering causal relationships, Vandewalle raises the possibility of self-selection bias, where authors include code more often with their best papers.
@@
 
-In Figure \@ref(fig:citation-advantage-plot) we conducted a similar analysis for archaeology articles using the 'TC' field in the data we obtained from the Web of Science service.
-TC stands for 'Web of Science Core Collection Times Cited Count'.
+We conducted a similar analysis for archaeology articles using the Web of Science Core Collection 'Times Cited' ('TC') count (Figure \@ref(fig:citation-advantage-plot)).
 This Times Cited count displays the total number of times a published paper was cited by other papers within the Web of Science Core Collection.
@@
 Further work is necessary to understand the mechanisms of this citation effect.
-Additional insights might come from comparisons of papers that use another open source language such as Python, and MATLAB, which is not an open source language.
+Additional insights might come from comparisons of papers that use another open source language such as Python, or a closed source language like MATLAB.
 
@@
 The tool which drives this change is the use of scripting languages in computational work in these disciplines.
-We have shown that the use of R is on the rise among archaeologists, especially those using natural science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visibility and influence (cf.
+We have shown that the use of R is on the rise among archaeologists, especially those using natural science methods, and has concrete advantages for authors that publish papers that cite R.
+The advantage is that papers using and citing R tend to be referenced more often in other papers, giving the authors more visibility and influence (cf.
 Huggett in this volume for critical discussion of open scholarship).
@@
 
-A related issue here is the _technicity_ of tools, in our case programming languages @cassirer_form_1930 and @simondon_mode_1958 to define technicity as a term to describe the technology, i.e.
+A related issue here is the _technicity_ of tools, in our case programming languages.
+@cassirer_form_1930 and @simondon_mode_1958 to define technicity as a term to describe the technology, i.e.
 tools and machines respectively, that interact with humans and with nature [@hoel_ontological_2013, 188].
-Technology is said to own an instrumental kind of logos (Cassierer, see @hoel_ontological_2013, 194) and in analogy to language it exercises a certain measure of agency in mediating between humans, nature and technology.
+Technology is said to own an instrumental kind of logos (Cassierer, see @hoel_ontological_2013, 194) and, in analogy to language, it exercises a certain measure of agency in mediating between humans, nature and technology.
 The use of a scripting language might exemplify this, as it is not just a technological tool, but a proper "language" as well.
@@
 In contrast to the concept of technological determinism, which puts technology as a force against the human experience, technicity enfolds technology into 'the human' without letting it lose its 'foreignness' [@hoel_ontological_2013, 190].
-According to this perspective, changes in tools will influence how human researchers interact with their objects of interest and other people in their research community.
+Changes in tools will influence how human researchers interact with their objects of interest and other people in their research community.
 There are three ways is which this can be anticipated for archaeology.
-Firstly, tools that make the research process more transparent and open might be expected to make the research community more open and inclusive to diverse participants and ideas.
-Secondly, using code to 'stand for' or represent the research process will likely increase our awareness of how archaeology itself 'stands for' the human experience in the past, and how it is a representation of the past [cf.
+First, tools that make the research process more transparent and open might be expected to make the research community more open and inclusive to diverse participants and ideas.
+Second, using code to 'stand for' or represent the research process will likely increase our awareness of how archaeology itself 'stands for' the human experience in the past, and how it is a representation of the past [cf.
 @pearson1998beginning].
-Thirdly, the code has to be successfully executed to generate a useful result, just as claims about the past have to be successfully engaged with archaeological evidence to count as plausible.
-
-We do not claim that a Kuhnian paradigm-change is happening in ecology or archaeology: Data- and method-sharing behaviours have been around in archaeology before now and do not present newly incommensurate ways of thinking about the subject, in the way that a typical reading of Kuhn's paradigm change concept should.
+Third, the code has to be successfully executed to generate a useful result, just as claims about the past have to be successfully engaged with archaeological evidence to count as plausible.
+
+We do not claim that a Kuhnian paradigm-change is happening in ecology or archaeology: Data- and method-sharing behaviours have been around in archaeology before now [@Lodwick_2019] and do not present newly incommensurate ways of thinking about the subject, in the way that a typical reading of Kuhn's paradigm change concept should.
 Indeed, social science disciplines such as archaeology may not even be easily susceptible to classical Kuhnian paradigm shifts.
@@
 We acknowledge that learning to use code, if one has not received formal instruction, is not an easy task.
-As @Chrysanthi_2012 illustrate, it requires a certain familiarity with the computational tools to be able to use them not just on a technological basis, but creatively and in full knowledge of their restrictions and ambiguities.
+A certain familiarity with the computational tools is required to be able to use them not just on a technological basis, but creatively and in full knowledge of their restrictions and ambiguities [@Chrysanthi_2012].
 The awareness and the appropriate handling of shortcomings and contexts of the data used in an analysis is also a time- and thought-consuming effort.
@@
 Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independently solving problems of how to write and share code through trial and error.
-We can use concepts from cultural evolution to understand this situation.
-Learning by trial and error is known as guided variation in a cultural evolutionary framework.
+We can use concepts from cultural evolution to understand this situation: learning by trial and error is known as guided variation in a cultural evolutionary framework.
 The dominance of guided variation suggests that using code is not yet widespread enough among archaeologists to propagate due to frequency-dependent biases.
@@
 These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.
-At the heart of the project template provided by rrtools is an RMarkdown document that is where the report or journal manuscript is written.
+At the heart of the project template provided by rrtools is an RMarkdown document.
+This is where the report or journal manuscript is written.
 RMarkdown is a document format that allows the author to combine plain text, R code, and automation of all the usual details of scholarly writing such as citations, captions, and cross-references to tables and figures [@xie2018r].
@@
 - **use_mit_license:** This function creates a copy of the MIT software license to the compendium.
-Other licenses are also possible, such as GPL, but the MIT license is preferred because it is widely used for research software, and has two important qualities.
+Other licenses are also possible, such as GPL, but the MIT license is preferred because it is widely used for research software because of its two important qualities.
 First, it tells the reader that the author is happy for their code to be reused by others, both in academic and commercial contexts.
@@
 This will also create a repository on the GitHub website.
-Using version control is important because it allows the author to time-travel back to earlier states of their analysis, if they take a dead-end path.
+Using version control is important because it allows the author to time-travel back to earlier states of their analysis, if they take a dead-end path [@Ram_2013].
 It is also makes collaboration smooth because many people can work on one set of files without losing track of the most recent version.
@@
 The tool with the highest potential to achieve this currently is the R programming language.
-However, we realise that this will change in time, and in the future newer technologies will replace R.
+However, we realise that this will change in time, and in the future newer technologies may replace R.
 The key point here is that this transformative tool is any open source scripting language.
@@
 We identified an advantage of citing R for researchers as it leads to being referenced more often.
-We also identified a small, but increasing, set of archaeology papers that use R to make their work fully reproducible.
+We also identified a small, but increasing, set of archaeology papers that use R to make their work fully transparent and reproducible by others.
 In reflecting on these results, we have noted the technicity of coding in how it mediates between humans, nature and technology.
@@
 Many useful tools have already emerged from this trading zone, including our R package, rrtools, designed to make it easy for researchers to make their research reproducible.
-We hope this will help to realise some of the potential to make archaeological research more open and reproducible, both within the research community and to the public, as well as speeding the transfer of new results and methods throughout the research community without barriers due to access to resources, and also in this way supporting the practice of archaeology that is more faithful to the core values of science.
+We hope this will help to realise some of the potential to make archaeological research more open and reproducible, both within the research community and to the public, as well as speeding the transfer of new results and methods throughout the research community without barriers due to access to resources.
+Tools emerging from the trading zone are important because they support practices of archaeology that are more faithful to the core values of science.
 
@@
 The implication here is that the most effective way to stimulate this change is during the first few years of professional training.
-This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, whereas the horizontal bar is the breadth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
+This means updating our professional training curricula by shifting from a model of creating T-shaped researchers (where the vertical bar on the T represents the depth of specialist skills and expertise in archaeology, and the horizontal bar is the breadth of skills and knowledge of related and intersecting fields) to gamma-shaped researchers [@Fiore_Gartland_2017].
 A gamma-shaped researcher has expert-level depth in archaeology, and proficient in other domains that provide skills such as analysing data with a programming language.
-A gamma-shaped researcher with their MA or PhD in archaeology may also be conversant enough in the language and culture of computer science to have conversations and collaborate in the trading zones described above, but does not need to be an expert in computer science to advance that discipline.
+A gamma-shaped researcher with their MA or PhD in archaeology may also be conversant enough in the language and culture of computer science to have conversations and collaborate in the trading zones described above, but does not need to be an expert in computer science.
 

### 2019-08-12 analysis/paper/jcaa-paper.Rmd
resubmit to JCAA with response to editor and peer reviewers

@@
 max_y <- 2021
-label_text_size = 2
+label_text_size = 4
 
@@
       slice(which.max(PY)) %>% 
-      mutate(SO_n = str_wrap(SO_n, 60)),
+      ungroup() %>% 
+      top_n(5, total_n_articles) %>% 
+      mutate(SO_n = str_wrap(SO_n, 30)),
     aes(label = SO_n),
@@
     size = label_text_size,
-    nudge_x = 2) +
+    nudge_x = 5) +
   theme(panel.grid.major = element_blank(), 
@@
 We used the 'Cited Reference Search' to find articles citing R in the same way as above, and then refined the results to keep only those articles published during 2008-2018 that are included in the Web of Science category 'Archaeology'.
-This resulted in `r prettyNum(how_many_articles_in_top_archy_journals, ",")` articles, of which `r total_number_archaeology_articles_citing_r` cite R.
-Figure \@ref(fig:archaeology-journals-with-subplot) shows the temporal trend of citations of R in archaeological articles.
+This resulted in `r prettyNum(how_many_articles_in_top_archy_journals, ",")` articles, of which `r total_number_archaeology_articles_citing_r` cite R.Figure \@ref(fig:archaeology-journals-with-subplot) shows the temporal trend of citations of R in archaeological articles.
 There are three interesting details revealed by this figure.
@@
 ```{r archaeology-journals-with-subplot, fig.cap=str_glue('Proportion of Archaeology articles per year citing R (a total of {total_number_archaeology_articles_citing_r} out of {prettyNum(how_many_articles_in_top_archy_journals, ",")} articles in our sample for 2008-2018).
+Labels to the right show the top five journals in our sample.
 Sub-plot shows articles published in the _Journal of Archaeological Science_ during 2008-2017.
@@
 
-As with data (see Fitzpatrick in this volume), preparing code to make it publicly available takes time to ensure that it is fit for others to read and use.
+As with data, preparing code to make it publicly available takes time to ensure that it is fit for others to read and use.
 Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independently solving problems of how to write and share code through trial and error.

### 2019-10-15 analysis/paper/jcaa-paper.Rmd
add heading numbering and heading for abstract according to JCAA author guidelines

@@
 csl: journal-of-computer-applications-in-archaeology.csl
-abstract: "There is an argument in philosophy of science that revolutions in science are either idea-driven or tool-driven.
+abstract: "*Abstract* \n There is an argument in philosophy of science that revolutions in science are either idea-driven or tool-driven.
 We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research.
@@
 
-# Introduction
+# 1. Introduction
 
@@
 
-# Revolutions in Science: idea-driven or tool-driven?
+# 2. Revolutions in Science: idea-driven or tool-driven?
 
@@
 
-# Bibliometric analysis of a tool-driven change
+# 3. Bibliometric analysis of a tool-driven change
 
@@
 
-## Looking back on a tool-driven change in Ecology
+## 3.1 Looking back on a tool-driven change in Ecology
 
@@
 
-# Methods
+# 4. Methods
 
@@
 
-## Reproducibility and open source materials
+## 4.1 Reproducibility and open source materials
 
@@
 
-# Results
+# 5. Results
 
@@
 
-## Looking forward to a tool-driven change in Archaeology
+## 5.1 Looking forward to a tool-driven change in Archaeology
 
@@
 
-## What are R-using archaeologists writing about?
+## 5.2 What are R-using archaeologists writing about?
 
@@
 
-## From using code in research to sharing code with publications
+## 5.3 From using code in research to sharing code with publications
 
@@
 
-## Citation effects of sharing code
+## 5.4 Citation effects of sharing code
 
@@
 
-# Discussion
+# 6. Discussion
 
@@
 
-## Making it easier
+## 6.1 Making it easier
 
@@
 
-# Conclusion
+# 7. Conclusion
 

### 2019-10-15 analysis/paper/jcaa-paper.Rmd
Upper cases in headings as specified by jcaa author guidelines
the journal-of-computer-applications-in-archaeology.csl doesn't add dots (.), where it should JCAA guidelines say they want harvard style
testing the harvard-european-archaeolgoy style by zotero

@@
 bibliography: references.bib
-csl: journal-of-computer-applications-in-archaeology.csl
+csl: havard-european-archaeology.csl
 abstract: "*Abstract* \n There is an argument in philosophy of science that revolutions in science are either idea-driven or tool-driven.
@@
 
-# 2. Revolutions in Science: idea-driven or tool-driven?
+# 2. Revolutions in Science: Idea-driven or Tool-driven?
 
@@
 
-# 3. Bibliometric analysis of a tool-driven change
+# 3. Bibliometric Analysis of a Tool-driven Change
 

### 2019-10-15 analysis/paper/jcaa-paper.Rmd
typo
corrected

@@
 bibliography: references.bib
-csl: havard-european-archaeology.csl
+csl: harvard-european-archaeology.csl
 abstract: "*Abstract* \n There is an argument in philosophy of science that revolutions in science are either idea-driven or tool-driven.

### 2019-10-15 analysis/paper/jcaa-paper.Rmd

@@
 We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of R code."
+keywords: reproducibility, philosophy of science, programming language R, computational archaeology, open science
 ---

### 2019-10-15 analysis/paper/jcaa-paper.Rmd

--- 2019-10-15 analysis/paper/jcaa-paper.Rmd
@@
 We discuss how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by the use of R code."
+keywords: philosophy of science, reproducibility, programming language R, computational archaeology, open science
 ---

### 2019-10-16 analysis/paper/jcaa-paper.Rmd
by -> sep in data input & cleaning, update ctv data

@@
            into = str_glue('X{1:4}'),
-           by = "_") %>%
+           sep = "_") %>%
   select(year, X4, value) %>% 

### 2019-10-16 analysis/paper/jcaa-paper.Rmd
much fiddling with csl and bib file

@@
 bibliography: references.bib
-csl: harvard-european-archaeology.csl
+csl: ../templates/harvard-european-archaeology.csl
 abstract: "*Abstract* \n There is an argument in philosophy of science that revolutions in science are either idea-driven or tool-driven.
@@
 Can we identify archaeological traditions focused on data collection and analysis using images and image-making devices [cf.
-@molyneaux2013cultural], in contrast to digital devices?
+@molyneaux2013cultural], in contrast to digital devices [@marwick_2019]?
 One candidate for this might be the shift from building relative chronologies based on seriation using typologies of visually distinctive artefacts to absolute chronologies based on radiometric dating using computer-controlled instruments.
@@
              colour = SO)) +
+   scale_color_viridis_d(option = "D") +
   geom_line(size = 2) +
@@
              colour = SO)) +
+  scale_color_viridis_d(option = "A") +
   scale_x_continuous(breaks = min_y:2018,
@@
   scale_y_continuous(breaks = 1:30) +
-  scale_fill_discrete(name = "Journal")  +
+  scale_fill_viridis_d(name = "Journal")  +
   xlab("Year of publication") +
@@
   geom_col() +
-  scale_fill_manual(values  =c("green", "red"),
+  scale_fill_viridis_d(
                     guide = FALSE) +

### 2020-01-22 analysis/paper/jcaa-paper.Rmd

@@
 archy_cran_cites_all_areas_top_journals_by_year_to_plot <- 
-  archy_cran_cites_all_areas_top_journals_by_year # %>% 
+  archy_cran_cites_all_areas_top_journals_by_year %>%
+  filter(total_n_articles > 5)
+# %>% 
 # filter(SO != "AFRICAN ARCHAEOLOGICAL REVIEW")
@@
              colour = SO)) +
-  scale_color_viridis_d(option = "A") +
+  scale_color_viridis_d(option = "D") +
   scale_x_continuous(breaks = min_y:2018,
